### 2015-09-18 papers/announcement/manuscript.tex
add paper-announcement skeleton

%&pdflatex
\documentclass[pdftex,preprint,3p,times,numbers]{elsarticle}

\journal{Computer Physics Communications}

%\usepackage{moreverb}

\usepackage{hyperref}
\hypersetup{pdfborder={0 0 0}}

\usepackage{float}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
\graphicspath{{Figures/}}

\usepackage[pdftex,usenames]{xcolor}

\usepackage{booktabs}
%\usepackage{colortbl}
\usepackage{multirow}

%\usepackage{amssymb}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{xspace}

\usepackage{xcolor}
\definecolor{Maroon}{cmyk}{0,0.87,0.68,0.32}
\definecolor{RoyalBlue}{cmyk}{1,0.50,0,0}
\definecolor{gray}{cmyk}{0.01,0.01,0.01,0.01}
\usepackage{listings}
\lstdefinelanguage{MyFortran}[95]{Fortran}{morecomment=[l]{\#},morestring=[d]',morekeywords={procedure,non_overridable,class,is}}
\lstdefinestyle{code}{%
  basicstyle=\footnotesize,%
  backgroundcolor=\color{gray},%
  language=MyFortran,%
  captionpos=b,%
  columns=fixed,%
  xleftmargin=10pt,%
  numbers=none,%
  numberstyle={\tiny},%
  keywordstyle=\color{RoyalBlue},%
  stringstyle={\sffamily},%
  texcl=true,%
  upquote=true,%
  commentstyle=\color{Maroon}%
}

\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\vardiamond}{\mathalpha}{extraup}{87}
\definecolor{OMP}{RGB}{255,127,0}
\definecolor{MPI}{RGB}{0,127,255}
\definecolor{HYB}{RGB}{127,0,255}

\DeclareGraphicsExtensions{.pdf,.png,.jpg,.bmp,.mps}
\newcommand{\citeh}[1]{\citeauthor{#1} \citenum{#1}}

\renewcommand{\thesubfigure}{\Alph{subfigure}}

\begin{document}

\begin{frontmatter}

\title{FOODiE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern}

\author[insean]{Zaghi S.\corref{cor1}\fnref{sz}}
\ead{stefano.zaghi@cnr.it}
\fntext[sz]{Ph. D., Aerospace Engineer, Research Scientist, Dept. of Computational Hydrodynamics at CNR-INSEAN.}
\address[insean]{CNR-INSEAN, Istituto Nazionale per Studi ed Esperienze di Architettura Navale, Via di Vallerano 139, Rome, Italy, 00128}
\cortext[cor1]{Corresponding author}

\author[rsmas]{Curcic M.\fnref{mc}}
\ead{milan@orca.rsmas.miami.edu}
\fntext[mc]{Ph.D. Meteorology and Physical Oceanography, Research Scientist, Dept. of Ocean Sciences Rosenstiel School of Marine and Atmospheric Science at University of Miami}
\address[rsmas]{Ocean Sciences Rosenstiel School of Marine and Atmospheric Science, University of Miami, 4600 Rickenbacker Causeway Miami, FL 33149-1098 +1 305.421.4000}

\author[sourcery]{Rouson D.\fnref{dr}}
\ead{damian@sourceryinstitute.org}
\fntext[dr]{Ph.D. Mechanical Engineering, Founder and President Sourcery Institute and Sourcery, Inc.}
\address[sourcery]{Sourcery Institute 482 Michigan Ave., Berkeley, CA 94707}

\author[sourcery]{Beekman I.\fnref{iz}}
\ead{izaak@izaakbeekman.com}
\fntext[iz]{Graduate Research Assistant, Princeton/UMD CCROCCO LAB}

\begin{abstract}
To be written.
\end{abstract}

\begin{keyword}
  Ordinary Differential Equations (ODE) \sep
  Partial Differential Equations (PDE) \sep
  Object Oriented Programming (OOP) \sep
  Abstract Calculus Pattern (ACP) \sep
  Fortran \sep
\end{keyword}

\end{frontmatter}

{\bf PROGRAM SUMMARY}

\begin{small}
\noindent
\emph{Manuscript Title:} FOODiE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern \\
\emph{Authors:} Zaghi, S., Curcic, M., Rouson, D., Beekman, I. \\
\emph{Program title:} FOODiE \\
\emph{Journal Reference:} \\
\emph{Catalogue identifier:} \\
\emph{Licensing provisions:} GNU General Public License (GPL) v3 \\
\emph{Programming language:} Fortran (standard 2008 or newer); developed and tested with GNU gfortran 5.2 or newer \\
\emph{Computer(s) for which the program has been designed:} designed for shared-memory multi-cores workstations and for hybrid distributed/shared-memory supercomputers, but any computer system with a Fortran (2008+) compiler is suited \\
\emph{Operating system(s) for which the program has been designed:} designed for POSIX architecture and tested on GNU/Linux one \\
\emph{RAM required to execute with typical data: bytes:} $[1MB,1GB]\times core$, simulation-dependent \\
\emph{Has the code been vectorised or parallelized?:} the library is not aware of the parallel back-end, it providing a high-level models, but the provided tests suite shows parallel usage by means of MPI library and OpenMP paradigm \\
\emph{Number of processors used:} tested up to 256 \\
\emph{Supplementary material:}    \\
\emph{Keywords:} ODE, PDE, OOP, ACP, Fortran \\
\emph{CPC Library Classification:} 4.3 Differential Equations, 4.10 Interpolation, 12 Gases and Fluids \\
\emph{External routines/libraries used:} \\
\emph{CPC Program Library subprograms used:} \\
\emph{Nature of problem:} \\
Numerical integration of (general) Ordinary Differential Equations system \\
\emph{Solution method:} \\
\emph{Restrictions:} \\
\emph{Unusual features:} \\
\emph{Additional comments:} \\
\emph{Running time:} \\
\emph{References:} \\
% \begin{thebibliography}{0}
% \end{thebibliography}
\end{small}

\section{Introduction}\label{sec:introduction}
\subsection{Background}

\subsection{Related Works}

\subsection{Motivations and Aims}

\subsection{General Description}

\section{Mathematical and Numerical Models}\label{sec:MNmodels}

\section{Application Program Interface}\label{sec:API}

\section{Tests and Examples}\label{sec:Tests}

\section{Ongoing Development Activities}\label{sec:ongoing}

\section{Concluding Remarks and Perspectives}\label{sec:conclusions}

\bibliographystyle{mycpc2}
\bibliography{Reference}

\end{document}

### 2015-09-28 papers/announcement/errors_analysis/errors_analysis.sh
Add Oscillation Errors Analysis
Add Oscillation Errors Analysis

@@ -1,0 +1,33 @@
+#!/bin/bash
+# script automotion for FOODiE's errors analysis of the provided tests
+cd ../../../
+echo "Building tests executables"
+FoBiS.py build -mode error-analysis --build_dir papers/announcement/error_analysis/tests > buils.log
+cd -
+echo "Oscillation test analysis"
+rm -rf results-oscillation
+echo "  Analize leapfrog solver"
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s leapfrog > leapfrog.log
+echo "  Analize Adams-Bashforth solver"
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s adams-bashforth --ss 2 > adams-bashforth-2.log
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s adams-bashforth --ss 3 > adams-bashforth-3.log
+echo "  Analize low storage Runge-Kutta solver"
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s ls-runge-kutta --ss 5 > ls-runge-kutta-5.log
+echo "  Analize TVD/SSP Runge-Kutta solver"
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s tvd-runge-kutta --ss 2 > tvd-runge-kutta-2.log
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s tvd-runge-kutta --ss 3 > tvd-runge-kutta-3.log
+./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s tvd-runge-kutta --ss 5 > tvd-runge-kutta-5.log
+mkdir -p results-oscillation
+mv *log *dat results-oscillation
+cd results-oscillation
+cp ../utilities-oscillation/*lay .
+cp ../utilities/lay_export* .
+lay_export_all_f error_analysis-oscillation-adams-bashforth-2.lay
+lay_export_all_f error_analysis-oscillation-adams-bashforth-3.lay
+lay_export_all_f error_analysis-oscillation-leapfrog.lay
+lay_export_all_f error_analysis-oscillation-ls-runge-kutta-5.lay
[…]

### 2015-09-28 papers/announcement/errors_analysis/utilities-oscillation/error_analysis.py

@@ -1,0 +1,164 @@
+#!/usr/bin/env python
+from __future__ import division
+from __future__ import print_function
+import argparse
+import numpy as np
+import re
+import subprocess
+import sys
+
+
+def syswork(cmd):
+  """
+  Execute system command 'cmd'.
+  """
+  work_error = 0
+  try:
+    work_output = subprocess.check_output(cmd, shell=True, stderr=subprocess.STDOUT)
+  except subprocess.CalledProcessError as err:
+    work_error = err.returncode
+    work_output = err.output
+  if sys.version_info[0] > 2:
+    work_output = work_output.decode("ascii")
+  return [work_error, str(work_output)]
+
+
+def compute_error(exact, numerical):
+  """
+  Compute error of numerical solution once the exact one is provided.
+
+  Parameters
+  ----------
+  exact : numpy array
+    exact solution
+  numerical : numpy array
+    numerical solution
+
+  Returns
+  error_estimate : real
+    Norm L2 of error estimate
+  """
+  error_estimate = 0.
+  for i in range(numerical.shape[0]):
+    error_estimate += (numerical[i][0] - exact[i][0])**2 + (numerical[i][1] - exact[i][1])**2
+  return np.sqrt(error_estimate)
+
+
+def load_oscillation_results(filename):
+  """Load Osciallation test results.
+
+  Parameters
+  ----------
+  filename : str
+    output file name
+
+  Returns
+  -------
+  numerical : list
+    list of numerical results
+  exact : list
+    list of exact results
+  """
+  numerical = []
+  exact = []
+  with open(filename) as foodie_results:
+    zones = re.split('ZONE.*\n', foodie_results.read())
+    sol_steps = zones[1].split('\n')
+    for sol in sol_steps:
+      if sol.strip() != '':
+        variables = re.split(' *', sol.strip())
+        numerical.append([variables[1], variables[2]])
+    sol_steps = zones[2].split('\n')
+    for sol in sol_steps:
+      if sol.strip() != '':
+        variables = re.split(' *', sol.strip())
+        exact.append([variables[1], variables[2]])
+  return numerical, exact
+
+
+def collect_errors(cliargs):
+  """Collect the errors on all simulations performed.
+
+  Parameters
+  ----------
+  cliargs : argpargse parsed arguments
+
+  Returns
+  -------
+  errors : list
+    list of float64 errors
+  """
+  time_steps = cliargs.time_steps
[…]

### 2015-09-28 src/tests/burgers/burgers.f90

@@ -40,15 +40,13 @@ call cli%init(progname    = 'burgers',                                          
               authors     = 'Fortran-FOSS-Programmers',                            &
               license     = 'GNU GPLv3',                                           &
               description = 'Test FOODiE library on Burgers equation integration', &
-              examples    = ["burgers --solver euler --results",                   &
-                             "burgers --solver runge-kutta -r ",                   &
-                             "burgers --solver adams-bashforth",                   &
-                             "burgers --solver all --plots -r "])
-call cli%add(switch='--solver', switch_ab='-s', help='ODE solver used', required=.true., act='store', error=error)
-call cli%add(switch='--results', switch_ab='-r', help='Save results', required=.false., act='store_true', def='.false.', &
-             error=error)
-call cli%add(switch='--plots', switch_ab='-p', help='Save plots of results', required=.false., act='store_true', def='.false.', &
-             error=error)
+              examples    = ["burgers --solver euler --results  ",                 &
+                             "burgers --solver ls-runge-kutta -r",                 &
+                             "burgers --solver adams-bashforth  ",                 &
+                             "burgers --solver all --plots -r   "])
+call cli%add(switch='--solver', switch_ab='-s', help='ODE solver used', required=.true., act='store')
+call cli%add(switch='--results', switch_ab='-r', help='Save results', required=.false., act='store_true', def='.false.')
+call cli%add(switch='--plots', switch_ab='-p', help='Save plots of results', required=.false., act='store_true', def='.false.')
 ! parsing Command Line Interface
 call cli%parse(error=error)
 call cli%get(switch='-s', val=solver, error=error) ; if (error/=0) stop

### 2015-09-28 src/tests/euler-1D-openmp/euler-1D-openmp.f90

@@ -75,12 +75,12 @@ call cli%add(switch='--verbose', help='Verbose output', required=.false., act='s
 call cli%add(switch='--omp_threads', help='Number of OpenMP threads used', required=.false., act='store', def='1', error=error)
 ! parsing Command Line Interface
 call cli%parse(error=error)
+call cli%get(switch='--omp_threads', val=omp_threads, error=error) ; if (error/=0) stop
+call cli%get(switch='--Ni', val=Ni, error=error) ; if (error/=0) stop
 call cli%get(switch='-r', val=results, error=error) ; if (error/=0) stop
 call cli%get(switch='-p', val=plots, error=error) ; if (error/=0) stop
 call cli%get(switch='-t', val=time_serie, error=error) ; if (error/=0) stop
 call cli%get(switch='--verbose', val=verbose, error=error) ; if (error/=0) stop
-call cli%get(switch='--omp_threads', val=omp_threads, error=error) ; if (error/=0) stop
-call cli%get(switch='--Ni', val=Ni, error=error) ; if (error/=0) stop
 #ifdef OPENMP
 call omp_set_dynamic(.false.)
 call omp_set_num_threads(omp_threads)

### 2015-09-28 src/tests/euler-1D-openmp/type_euler-1D-openmp.f90

@@ -85,6 +85,21 @@ type, extends(integrand) :: euler_1D_openmp
   !<$$
   !< where \(N_s\) is the number of initial species composing the gas mixture.
   !<
+  !<#### Numerical grid organization
+  !< The finite volume, Godunov's like approach is employed. The conservative variables (and the primitive ones) are co-located at
+  !< the cell center. The cell and (inter)faces numeration is as follow.
+  !<```
+  !<                cell            (inter)faces
+  !<                 |                   |
+  !<                 v                   v
+  !<     |-------|-------|-.....-|-------|-------|-------|-------|-.....-|-------|-------|-------|-.....-|-------|-------|
+  !<     | 1-Ng  | 2-Ng  | ..... |  -1   |   0   |   1   |  2    | ..... |  Ni   | Ni+1  | Ni+1  | ..... |Ni+Ng-1| Ni+Ng |
+  !<     |-------|-------|-.....-|-------|-------|-------|-------|-.....-|-------|-------|-------|-.....-|-------|-------|
+  !<    0-Ng                             -1      0       1       2      Ni-1     Ni                                    Ni+Ng
+  !<```
+  !< Where *Ni* are the finite volumes (cells) used for discretizing the domain and *Ng* are the ghost cells used for imposing the
+  !< left and right boundary conditions (for a total of *2Ng* cells).
+  !<
   !<#### Primitive variables organization
   !< Primitive variables are organized as an array of reals which the first index means:
   !<

### 2015-09-28 papers/announcement/manuscript.tex

@@
 \usepackage{multirow}
-\graphicspath{{Figures/}}
+\graphicspath{{images/}}
 
@@
 
-%\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage[utf8x]{inputenc}
@@
 
+\clearpage
+
+\subsection{Oscillation equations test}
+
+\begin{equation}
+\begin{matrix}
+U_t = R(U)  \\
+U = \begin{bmatrix}
+x \\
+y
+\end{bmatrix}\;\;\;
+R(U) = \begin{bmatrix}
+-f y \\
+f x
+\end{bmatrix}
+\end{matrix}
+\label{eq:oscillation}
+\end{equation}
+
+where the frequency is chosen as $f=10^4$. The ODE system \ref{eq:oscillation} is completed by the following initial conditions:
+
+\begin{equation}
+\begin{matrix}
+  x(t_0) = 0 \\
+  y(t_0) = 1
+\end{matrix}
+\label{eq:oscillation-ic}
+\end{equation}
+
+where $t_0=0$ is the initial time considered. The exact solution is:
+
+\begin{equation}
+\begin{matrix}
+  x(t_0 + \Delta t) = X_0 cos(f \Delta t) - y_0 sin(f \Delta t) \\
+  y(t_0 + \Delta t) = X_0 sin(f \Delta t) + y_0 cos(f \Delta t)
+\end{matrix}
+\label{eq:oscillation-exact}
+\end{equation}
+
+where $\Delta t$ is an arbitrary time step.
+
+\subsubsection{Errors Analysis}
+
+For the analysis of the accuracy of each solvers implemented into FOODiE library, we have integrated the Oscillation equations \ref{eq:oscillation} with different, decreasing time steps in the range $[5000, 2500, 1250, 625, 320, 100]$.
+
+The error is estimated by the L2 norm of the difference between the exact ($U_e$) and the numerical ($U_{\Delta t}$) solutions for each time step:
+
+\begin{equation}
+  \varepsilon (\Delta t) = || U_e - U_{\Delta t} ||_2 = \sqrt{ \sum_{s=1}^{N_s} { \left(U_e(t_0 + s * \Delta t) - U_{\Delta t}(t_0 + s * \Delta t) \right)^2 }}
+\label{eq:oscillation-error}
+\end{equation}
+
+Using two pairs of subsequent-decreasing time steps solution is possible to estimate the order of accuracy of the solver employed computing the \emph{observed order} of accuracy:
+
+\begin{equation}
+  p = \frac{log10 \left( \frac{\varepsilon (\Delta t_1)}{\varepsilon (\Delta t_2)} \right)}{log10 \left( \frac{\Delta t_1}{\Delta t_2} \right)}
+\label{eq:oscillation-observed-order}
+\end{equation}
+
+where $\frac{\Delta t_1}{\Delta t_2}>1$.
+
+Table \ref{tab:oscillation} summarizes the errors analysis.
+
+\begin{table}[!ht]
+  \centering
+  \caption{Oscillation test errors analysis}\label{tab:oscillation}
+  \resizebox{0.80\textwidth}{!}{%
+  \begin{tabular}{cccc}
+    \toprule
+    {\sc Solver}                             & {\sc Time Step} & {\sc Error}     & {\sc Observed Order of Accuracy} \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{Adams-Bashforth 2 steps}
+                                             & 5000            & 1565.886        &  /                               \\
+                                             & 2500            & 34.608          &  5.500                           \\
+                                             & 1250            & 11.117          &  1.638                           \\
+                                             & 625             & 3.815           &  1.543                           \\
+                                             & 320             & 1.391           &  1.508                           \\
+                                             & 100             & 0.243           &  1.500                           \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{Adams-Bashforth 3 steps}
+                                             & 5000            & 11.491          & /                                \\
+                                             & 2500            & 3.746           & 1.617                            \\
+                                             & 1250            & 1.698           & 1.141                            \\
+                                             & 625             & 0.769           & 1.142                            \\
+                                             & 320             & 0.314           & 1.339                            \\
+                                             & 100             & 0.059           & 1.438                            \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{leapfrog 2 steps}
+                                             & 5000            & 21.456          & /                                \\
+                                             & 2500            & 7.099           & 1.596                            \\
+                                             & 1250            & 2.705           & 1.392                            \\
+                                             & 625             & 2.560           & 0.079                            \\
+                                             & 320             & 2.222           & 0.212                            \\
+                                             & 100             & 1.417           & 0.387                            \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{low storage Runge-Kutta 5 stages}
+                                             & 5000            & 1.715 $10^{-1}$ & /                                \\
+                                             & 2500            & 1.509 $10^{-2}$ & 3.507                            \\
+                                             & 1250            & 1.331 $10^{-3}$ & 3.503                            \\
+                                             & 625             & 1.175 $10^{-4}$ & 3.501                            \\
+                                             & 320             & 1.128 $10^{-5}$ & 3.500                            \\
+                                             & 100             & 1.925 $10^{-7}$ & 3.500                            \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{TVD/SSP Runge-Kutta 2 stages}
+                                             & 5000            & 44.954          & /                                \\
+                                             & 2500            & 12.626          & 1.832                            \\
+                                             & 1250            & 4.287           & 1.558                            \\
+                                             & 625             & 1.506           & 1.510                            \\
+                                             & 320             & 0.551           & 1.502                            \\
+                                             & 100             & 0.096           & 1.500                            \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{TVD/SSP Runge-Kutta 3 stages}
+                                             & 5000            & 3.582           & /                                \\
+                                             & 2500            & 7.350 $10^{-1}$ & 2.285                            \\
+                                             & 1250            & 1.330 $10^{-1}$ & 2.471                            \\
+                                             & 625             & 2.349 $10^{-2}$ & 2.497                            \\
+                                             & 320             & 4.407 $10^{-3}$ & 2.500                            \\
+                                             & 100             & 2.406 $10^{-4}$ & 2.500                            \\
+    \cmidrule{2-4}
+    \multirow{6}{*}{TVD/SSP Runge-Kutta 5 stages}
+                                             & 5000            & 1.976 $10^{-1}$ & /                                \\
+                                             & 2500            & 1.744 $10^{-2}$ & 3.502                            \\
+                                             & 1250            & 1.539 $10^{-3}$ & 3.502                            \\
+                                             & 625             & 1.361 $10^{-4}$ & 3.499                            \\
+                                             & 320             & 1.333 $10^{-5}$ & 3.470                            \\
+                                             & 100             & 7.294 $10^{-7}$ & 2.498                            \\
+    \bottomrule
+  \end{tabular}}
+\end{table}
+
+\paragraph{Adams-Bashforth}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-adams-bashforth-2.png}
+  \caption{Oscillation equations solutions computed by means of Adams-Bashforth 2 steps solver}\label{fig:results-oscillation-adams-bashforth-2}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-adams-bashforth-3.png}
+  \caption{Oscillation equations solutions computed by means of Adams-Bashforth 3 steps solver}\label{fig:results-oscillation-adams-bashforth-3}
+\end{figure}
+
+\paragraph{Leapfrog}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-leapfrog.png}
+  \caption{Oscillation equations solutions computed by means of leapfrog 2 steps solver}\label{fig:results-oscillation-leapfrog}
+\end{figure}
+
+\paragraph{Low Storage Runge-Kutta}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-ls-runge-kutta-5.png}
+  \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta 5 stages solver}\label{fig:results-oscillation-ls-runge-kutta-5}
+\end{figure}
+
+\paragraph{TVD/SSP Runge-Kutta}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-tvd-runge-kutta-2.png}
+  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 2 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-2}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-tvd-runge-kutta-3.png}
+  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 3 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-3}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-tvd-runge-kutta-5.png}
+  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 5 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-5}
+\end{figure}
+
+\clearpage
+
 \section{Ongoing Development Activities}\label{sec:ongoing}

### 2015-09-28 src/oscillation.f90
Travis CI autocommit from travis build 52

@@ -5,7 +5,7 @@ program integrate_oscillation
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use IR_Precision, only : R_P, I_P, FR_P, str
+use IR_Precision, only : R_P, I_P, FR_P, str, strz
 use type_oscillation, only : oscillation
 use Data_Type_Command_Line_Interface, only : Type_Command_Line_Interface
 use foodie, only : adams_bashforth_integrator, &
@@ -20,16 +20,19 @@ use pyplot_module, only :  pyplot
 implicit none
 type(Type_Command_Line_Interface) :: cli                                               !< Command line interface handler.
 type(oscillation)                 :: attractor                                         !< Oscillation field.
-integer,   parameter              :: num_steps=1e4                                     !< Maximum time steps.
 integer,   parameter              :: space_dimension=2                                 !< Space dimensions.
-real(R_P), parameter              :: f=1e-4_R_P                                        !< Frequency.
-real(R_P), parameter              :: dt = 100._R_P                                     !< Time step.
 real(R_P), parameter              :: initial_state(1:space_dimension)=[0._R_P, 1._R_P] !< Initial state.
-real(R_P)                         :: solution(0:space_dimension, 0:num_steps)          !< Solution at each time step.
+real(R_P), parameter              :: f=1e-4_R_P                                        !< Frequency.
+real(R_P)                         :: t_final                                           !< Final time.
+real(R_P)                         :: Dt                                                !< Time step.
+real(R_P), allocatable            :: solution(:,:)                                     !< Solution at each time step.
 integer(I_P)                      :: error                                             !< Error handler.
+integer(I_P)                      :: stages_steps                                      !< Number of stages/steps used.
 character(99)                     :: solver                                            !< Solver used.
 logical                           :: plots                                             !< Flag for activating plots saving.
 logical                           :: results                                           !< Flag for activating results saving.
+character(len=:), allocatable     :: output                                            !< Output files basename.
+character(99)                     :: output_cli                                        !< Output files basename.
[…]

### 2015-09-29 papers/announcement/manuscript.tex
Embed errors analysis into oscillation test source
Embed errors analysis into oscillation test source.

Why:

Avoid Python for higher control on numerical computations involved into
the errors analysis.

This change addresses the need by:

Improve the oscillation.f90 test program.

Side effects:

Nothing. Minor improvements into the manuscript are also provided.

@@
 \usepackage{listings}
-\lstdefinelanguage{MyFortran}[95]{Fortran}{morecomment=[l]{\#},morestring=[d]',morekeywords={procedure,non_overridable,class,is}}
+\lstdefinelanguage{MyFortran}[95]{Fortran}{morecomment=[l]{\#},morestring=[d]',morekeywords={procedure,pass,deferred,non_overridable,generic,class,is}}
 \lstdefinestyle{code}{%
@@
 
+Initial Value Problem (IVP, or Cauchy problem) constitutes a class of mathematical models of paramount relevance, it being applied to the modelling of a wide range of problems.
+Briefly, an IVP is an Ordinary Differential Equations system (ODE) coupled with specified initial values of the unknown state variables, the solution of which are searched at a given time after the initial time considered.
+
+The prototype of IVP can be expressed as:
+\begin{equation}
+  \begin{matrix}
+  U_t = R(t,U) \\
+  U_0 = F(t_0)
+  \end{matrix}
+\label{eq:IVP}
+\end{equation}
+where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residual function and $F$ is the (vectorial) initial conditions function evaluated at the initial time $t_0$.
+In general, the residual function $R$ is a function of the state variable $U$ through which it is a function of time, but it can be also a direct function of time, thus in general $R=R(t,U(t))$ holds.
+
+The problem prototype \ref{eq:IVP} is ubiquitous in the mathematical modelling of physical problems: essentially whenever an \emph{evolutionary} phenomenon is considered the prevision (simulation) of the future solutions involves the solution of an IVP.
+As a matter of facts, many physical problems (fluid dynamics, chemistry, biology, evolutionary-anthropology, etc...) are described by means of an IVP.
+
+It is worth noting that the state vector variables $U$ and its corresponding residual function $U_t = \frac{dU}{dt} = R(t,U)$ are \emph{problem dependent}: the number and the meaning of the state variables as well as the equations governing their evolution (that are embedded into the residual function) are different for the Navier-Stokes conservation laws with respect the Burgers one, as an example.
+Nevertheless, the \emph{solvers} used for the prediction of the Navier-Stokes equations evolution (hereafter the \emph{solution}) are the same that are used for Burgers equations time-integration.
+As a consequence, the solution of the IVP model prototype can be generalized, allowing the application of the same solver to many different problems, thus eliminating the necessity to re-implement the same solver for each different problem.
+
+FOODiE library is designed for solving the generalized IVP \ref{eq:IVP}, it being completely unaware of the actual problem's definition.
+FOODiE library provides a high-level, well-documented, simple Application Program Interface (API) for many well-known ODE integration solvers, its aims being twofold:
+
+\begin{itemize}
+  \item provide a robust set of ODE solvers ready to be applied to a wide range of different problems;
+  \item provide a simple framework for the rapid development of new ODE solvers.
+\end{itemize}
+
+{\color{red} Add citations to IVP, Cauchy, ODE references.}
+
 \subsection{Related Works}
+
+{\color{red} To be written.}
 
@@
 
-\subsection{General Description}
+FOODiE library must:
+
+\begin{itemize}
+  \item be written in modern Fortran (standard 2008 or newer);
+  \item be written by means of Object Oriented Programming (OOP) paradigm;
+  \item be well documented;
+  \item be Tests Driven Developed (TDD);
+  \item be collaboratively developed;
+  \item be free.
+\end{itemize}
+
+FOODiE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has been developed with the aim to satisfy the above specifications.
+The present paper is its first comprehensive presentation.
+
+The Fortran (Formula Translator) programming language is the \emph{de facto} standard into computer science field: it strongly facilitates the effective and efficient translation of (even complex) mathematical and numerical models into an operative software without compromise on computations speed and accuracy.
+Moreover, its simple syntax is suitable for scientific researchers that are interested (and skilled) in the physical aspects of the numerical computations rather than computer technicians.
+Consequently, we develop FOODiE using Fortran language: FOODiE is written by research scientists for research scientists.
+
+One key-point of the FOODiE development is the \emph{problem generalization}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its actual definitions.
+Consequently, we must rely on a \emph{generic} implementation of the solvers.
+To this aim, OOP is very useful: it allows to express IVP \ref{eq:IVP} in a very concise and clear formulation that is really generic.
+In particular, our implementation is based on \emph{Abstract Calculus Pattern} (ACP) concept.
+
+\paragraph{The Abstract Calculus Pattern}
+The abstract calculus pattern provides a simple solution for the connection between the very high-level expression of IVP \ref{eq:IVP} and the eventual concrete (low-level) implementation of the ODE problem being solved.
+ACP essentially constitutes a \emph{contract} based on an Abstract Data Type (ADT): we specify an ADT supporting a certain set of mathematical operators (differential and integral ones) and implement FOODiE solvers only on the basis of this ADT.
+FOODiE clients must formulate the ODE problem under integration defining their own concrete extensions of our ADT (implementing all the deferred operators).
+Such an approach defines the abstract calculus pattern: FOODiE solvers are aware of only the ADT, while FOODiE clients extend the ADT for defining the concrete ODE problem.
+
+Is is worth noting that this ACP emancipates the solvers implementations from any low-level problem-dependent details: the ODE solvers developed with this pattern are extremely concise, clear, maintainable and less errors-prone with respect a low-level (non abstract) pattern.
+Moreover, the FOODiE clients can use solvers being extremely robust: as a matter of facts, FOODiE solvers are expressed in a formulation very close to the mathematical one and are tested on an extremely varying family of problems.
+As shown in the following, such a great flexibility does not compromise the computational efficiency.
+
+{\color{red} Add citations to Fortran standards, OOP, TDD, free software, ACP, ADT, Rouson's book references.}
 
@@
 
+In this section the mathematical and numerical models of the currently implemented FOODiE solvers are described.
+
+\subsection{The explicit forward Euler scheme}
+
+{\color{red} To be written.}
+
+\subsection{The explicit Adams-Bashforth class of schemes}
+
+{\color{red} To be written.}
+
+\subsection{The leapfrog solver}
+
+{\color{red} To be written.}
+
+\subsection{The explicit low storage Runge-Kutta class of schemes}
+
+{\color{red} To be written.}
+
+\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
+
+{\color{red} To be written.}
+
+\clearpage
+
 \section{Application Program Interface}\label{sec:API}
 
+In this section we review the FOODiE API providing a detailed discussion of the implementation choices.
+
+\subsection{The main FOODiE Abstract Data Type: the integrand type}
+Our ACP is based on one main ADT, the \emph{integrand} type, the definition of which is shown in listing \ref{list:type_integrand}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={integrand type definition},label={list:type_integrand}]
+type, abstract :: integrand
+  !< Abstract type for building FOODiE ODE integrators.
+  contains
+    ! public deferred procedures that concrete integrand-field must implement
+    procedure(time_derivative),       pass(self), deferred, public:: t
+    procedure(update_previous_steps), pass(self), deferred, public:: update_previous_steps
+    procedure(previous_step),         pass(self), deferred, public:: previous_step
+    ! operators
+    procedure(symmetric_operator),   pass(lhs), deferred, public:: integrand_multiply_integrand
+    procedure(integrand_op_real),    pass(lhs), deferred, public:: integrand_multiply_real
+    procedure(real_op_integrand),    pass(rhs), deferred, public:: real_multiply_integrand
+    procedure(symmetric_operator),   pass(lhs), deferred, public:: add
+    procedure(symmetric_operator),   pass(lhs), deferred, public:: sub
+    procedure(assignment_integrand), pass(lhs), deferred, public:: assign_integrand
+    ! operators overloading
+    generic, public:: operator(+) => add
+    generic, public:: operator(-) => sub
+    generic, public:: operator(*) => integrand_multiply_integrand, &
+                                     real_multiply_integrand, &
+                                     integrand_multiply_real
+    generic, public:: assignment(=) => assign_integrand
+endtype integrand
+\end{lstlisting}
+
+The \emph{integrand} type does not implement any actual integrand field, it being and abstract type.
+It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODiE solver.
+Listing \ref{list:type_integrand} specifies the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT.
+These procedures are analyzed in the following paragraphs.
+
+\subsubsection{Time derivative procedure, the residual function}
+
+The abstract interface of the time derivative procedure \emph{t} is shown in listing \ref{list:time_derivative}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={time derivative procedure interface},label={list:time_derivative}]
+function time_derivative(self, n, t) result(dState_dt)
+import :: integrand, R_P, I_P
+class(integrand),       intent(IN) :: self      !< Integrand field.
+integer(I_P), optional, intent(IN) :: n         !< Time level.
+real(R_P),    optional, intent(IN) :: t         !< Time.
+class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.
+endfunction time_derivative
+\end{lstlisting}
+
+This procedure-function takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optional, and returns an integrand object.
+The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
+The optional argument \emph{n} indicated the \emph{time level} at which the residual function must be evaluated: this argument is necessary for multi-level (often referred as multi-step) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family.
+The optional argument \emph{t} is the {time} at which the residual function must be computed: it can be omitted in the case the residual function does not depend directly from time.
+
+Commonly, into the concrete implementation of this deferred abstract procedure clients embed the actual ODE equations being solved.
+As an example, for the Burgers equation, that is a Partial Differential Equations (PDE) system involving also a boundary value problem, this procedure embeds the spatial operator that convert the PDE to a system of algebraic ODE.
+As a consequence, the eventual concrete implementation of this procedure can be very complex and errors-prone.
+Nevertheless, the FOODiE solvers are implemented only on the above abstract interface, thus emancipating the solvers implementation from any concrete complexity.
+
+{\color{red} Add citations to Burgers, Adams-Bashfort, leapfrog references.}
+
+\subsubsection{Update previous time steps procedure}
+
+The abstract interface of the \emph{update previous time steps} procedure is shown in listing \ref{list:update_previous_time_steps}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={update previous time steps procedure interface},label={list:update_previous_time_steps}]
+subroutine update_previous_steps(self, filter, weights)
+import :: integrand, R_P
+class(integrand),           intent(INOUT) :: self       !< Integrand field.
+class(integrand), optional, intent(IN)    :: filter     !< Filter field displacement.
+real(R_P),        optional, intent(IN)    :: weights(:) !< Weights for filtering the steps.
+endsubroutine update_previous_steps
+\end{lstlisting}
+
+{\color{red} To be completed.}
+
+\subsubsection{Previous step procedure}
+
+The abstract interface of the \emph{previous step} procedure is shown in listing \ref{list:previous_step}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={previous step procedure interface},label={list:previous_step}]
+function previous_step(self, n) result(previous)
+import :: integrand, I_P
+class(integrand), intent(IN)  :: self     !< Integrand field.
+integer(I_P),     intent(IN)  :: n        !< Time level.
+class(integrand), allocatable :: previous !< Selected previous time integrand field.
+endfunction previous_step
+\end{lstlisting}
+
+{\color{red} To be completed.}
+
+\subsubsection{Symmetric operators procedures}
+
+The abstract interface of \emph{symmetric} procedures is shown in listing \ref{list:symmetric_operator}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={symmetric operator procedure interface},label={list:symmetric_operator}]
+function symmetric_operator(lhs, rhs) result(operator_result)
+import :: integrand
+class(integrand), intent(IN)  :: lhs             !< Left hand side.
+class(integrand), intent(IN)  :: rhs             !< Right hand side.
+class(integrand), allocatable :: operator_result !< Operator result.
+endfunction symmetric_operator
+\end{lstlisting}
+
+{\color{red} To be completed.}
+
+\subsubsection{Integrand/real and real/integrand operators procedures}
+
+The abstract interfaces of \emph{Integrand/real and real/integrand operators} procedures are shown in listing \ref{list:integrand_real_ops}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={Integrand/real and real/integrand operators procedure interfaces},label={list:integrand_real_ops}]
+function integrand_op_real(lhs, rhs) result(operator_result)
+import :: integrand, R_P
+class(integrand), intent(IN)  :: lhs              !< Left hand side.
+real(R_P),        intent(IN)  :: rhs              !< Right hand side.
+class(integrand), allocatable :: operator_result  !< Operator result.
+endfunction integrand_op_real
+
+function real_op_integrand(lhs, rhs) result(operator_result)
+import :: integrand, R_P
+real(R_P),        intent(IN)  :: lhs             !< Left hand side.
+class(integrand), intent(IN)  :: rhs             !< Right hand side.
+class(integrand), allocatable :: operator_result !< Operator result.
+endfunction real_op_integrand
+\end{lstlisting}
+
+{\color{red} To be completed.}
+
+\subsubsection{Integrand assignment procedure}
+
+The abstract interface of \emph{integrand assignment} procedure is shown in listing \ref{list:integrand_assignment}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={integrand assignment procedure interface},label={list:integrand_assignment}]
+subroutine assignment_integrand(lhs, rhs)
+import :: integrand
+class(integrand), intent(INOUT) :: lhs !< Left hand side.
+class(integrand), intent(IN)    :: rhs !< Right hand side.
+endsubroutine assignment_integrand
+\end{lstlisting}
+
+{\color{red} To be completed.}
+
+\clearpage
+
+\subsection{The explicit forward Euler solver}
+
+{\color{red} To be written.}
+
+\subsection{The explicit Adams-Bashforth class of solvers}
+
+{\color{red} To be written.}
+
+\subsection{The leapfrog solver}
+
+{\color{red} To be written.}
+
+\subsection{The explicit low storage Runge-Kutta class of solvers}
+
+{\color{red} To be written.}
+
+\subsection{The explicit TVD/SSP Runge-Kutta class of solvers}
+
+{\color{red} To be written.}
+
 \section{Tests and Examples}\label{sec:Tests}
+
+{\color{red} To be written.}
 
@@
 \subsection{Oscillation equations test}
+
+{\color{red} To be completed.}
 
@@
   \resizebox{0.80\textwidth}{!}{%
-  \begin{tabular}{cccc}
+  \begin{tabular}{cccccc}
     \toprule
-    {\sc Solver}                             & {\sc Time Step} & {\sc Error}     & {\sc Observed Order of Accuracy} \\
-    \cmidrule{2-4}
+    {\sc Solver} & {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+    \cmidrule{2-6}
     \multirow{6}{*}{Adams-Bashforth 2 steps}
-                                             & 5000            & 1565.886        &  /                               \\
-                                             & 2500            & 34.608          &  5.500                           \\
-                                             & 1250            & 11.117          &  1.638                           \\
-                                             & 625             & 3.815           &  1.543                           \\
-                                             & 320             & 1.391           &  1.508                           \\
-                                             & 100             & 0.243           &  1.500                           \\
-    \cmidrule{2-4}
+                 &   5000.0        &  0.109E+04    &  0.112E+04    & /             & /             \\
+                 &   2500.0        &  0.246E+02    &  0.243E+02    & 5.47          & 5.53          \\
+                 &   1250.0        &  0.783E+01    &  0.789E+01    & 1.65          & 1.62          \\
+                 &    625.0        &  0.268E+01    &  0.272E+01    & 1.55          & 1.54          \\
+                 &    320.0        &  0.976E+00    &  0.990E+00    & 1.51          & 1.51          \\
+                 &    100.0        &  0.171E+00    &  0.173E+00    & 1.50          & 1.50          \\
+    \cmidrule{2-6}
     \multirow{6}{*}{Adams-Bashforth 3 steps}
-                                             & 5000            & 11.491          & /                                \\
-                                             & 2500            & 3.746           & 1.617                            \\
-                                             & 1250            & 1.698           & 1.141                            \\
-                                             & 625             & 0.769           & 1.142                            \\
-                                             & 320             & 0.314           & 1.339                            \\
-                                             & 100             & 0.059           & 1.438                            \\
-    \cmidrule{2-4}
+                 &   5000.0        &  0.816E+01    &  0.809E+01     & /            & /             \\
+                 &   2500.0        &  0.266E+01    &  0.264E+01     & 1.62         & 1.62          \\
+                 &   1250.0        &  0.120E+01    &  0.120E+01     & 1.15         & 1.14          \\
+                 &    625.0        &  0.545E+00    &  0.543E+00     & 1.14         & 1.14          \\
+                 &    320.0        &  0.223E+00    &  0.222E+00     & 1.34         & 1.34          \\
+                 &    100.0        &  0.418E-01    &  0.416E-01     & 1.44         & 1.44          \\
+    \cmidrule{2-6}
     \multirow{6}{*}{leapfrog 2 steps}
-                                             & 5000            & 21.456          & /                                \\
-                                             & 2500            & 7.099           & 1.596                            \\
-                                             & 1250            & 2.705           & 1.392                            \\
-                                             & 625             & 2.560           & 0.079                            \\
-                                             & 320             & 2.222           & 0.212                            \\
-                                             & 100             & 1.417           & 0.387                            \\
-    \cmidrule{2-4}
+                 &   5000.0        &  0.151E+02    &  0.152E+02     & /            & /             \\
+                 &   2500.0        &  0.500E+01    &  0.504E+01     & 1.60         & 1.59          \\
+                 &   1250.0        &  0.190E+01    &  0.193E+01     & 1.40         & 1.39          \\
+                 &    625.0        &  0.181E+01    &  0.181E+01     & 0.07         & 0.09          \\
+                 &    320.0        &  0.157E+01    &  0.157E+01     & 0.21         & 0.21          \\
+                 &    100.0        &  0.100E+01    &  0.100E+01     & 0.39         & 0.39          \\
+    \cmidrule{2-6}
     \multirow{6}{*}{low storage Runge-Kutta 5 stages}
-                                             & 5000            & 1.715 $10^{-1}$ & /                                \\
-                                             & 2500            & 1.509 $10^{-2}$ & 3.507                            \\
-                                             & 1250            & 1.331 $10^{-3}$ & 3.503                            \\
-                                             & 625             & 1.175 $10^{-4}$ & 3.501                            \\
-                                             & 320             & 1.128 $10^{-5}$ & 3.500                            \\
-                                             & 100             & 1.925 $10^{-7}$ & 3.500                            \\
-    \cmidrule{2-4}
+                 &   5000.0        &  0.120E+00    &  0.122E+00     & /            & /             \\
+                 &   2500.0        &  0.106E-01    &  0.107E-01     & 3.51         & 3.51          \\
+                 &   1250.0        &  0.935E-03    &  0.947E-03     & 3.50         & 3.50          \\
+                 &    625.0        &  0.826E-04    &  0.836E-04     & 3.50         & 3.50          \\
+                 &    320.0        &  0.793E-05    &  0.803E-05     & 3.50         & 3.50          \\
+                 &    100.0        &  0.135E-06    &  0.137E-06     & 3.50         & 3.50          \\
+    \cmidrule{2-6}
     \multirow{6}{*}{TVD/SSP Runge-Kutta 2 stages}
-                                             & 5000            & 44.954          & /                                \\
-                                             & 2500            & 12.626          & 1.832                            \\
-                                             & 1250            & 4.287           & 1.558                            \\
-                                             & 625             & 1.506           & 1.510                            \\
-                                             & 320             & 0.551           & 1.502                            \\
-                                             & 100             & 0.096           & 1.500                            \\
-    \cmidrule{2-4}
+                 &   5000.0        &  0.316E+02    &  0.319E+02     & /            & /             \\
+                 &   2500.0        &  0.892E+01    &  0.894E+01     & 1.83         & 1.84          \\
+                 &   1250.0        &  0.301E+01    &  0.305E+01     & 1.57         & 1.55          \\
+                 &    625.0        &  0.106E+01    &  0.107E+01     & 1.51         & 1.51          \\
+                 &    320.0        &  0.387E+00    &  0.392E+00     & 1.50         & 1.50          \\
+                 &    100.0        &  0.676E-01    &  0.685E-01     & 1.50         & 1.50          \\
+    \cmidrule{2-6}
     \multirow{6}{*}{TVD/SSP Runge-Kutta 3 stages}
-                                             & 5000            & 3.582           & /                                \\
-                                             & 2500            & 7.350 $10^{-1}$ & 2.285                            \\
-                                             & 1250            & 1.330 $10^{-1}$ & 2.471                            \\
-                                             & 625             & 2.349 $10^{-2}$ & 2.497                            \\
-                                             & 320             & 4.407 $10^{-3}$ & 2.500                            \\
-                                             & 100             & 2.406 $10^{-4}$ & 2.500                            \\
-    \cmidrule{2-4}
+                 &   5000.0        &  0.255E+01    &  0.252E+01     & /            & /             \\
+                 &   2500.0        &  0.523E+00    &  0.516E+00     & 2.28         & 2.29          \\
+                 &   1250.0        &  0.944E-01    &  0.931E-01     & 2.47         & 2.47          \\
+                 &    625.0        &  0.167E-01    &  0.165E-01     & 2.50         & 2.50          \\
+                 &    320.0        &  0.314E-02    &  0.310E-02     & 2.50         & 2.50          \\
+                 &    100.0        &  0.171E-03    &  0.169E-03     & 2.50         & 2.50          \\
+    \cmidrule{2-6}
     \multirow{6}{*}{TVD/SSP Runge-Kutta 5 stages}
-                                             & 5000            & 1.976 $10^{-1}$ & /                                \\
-                                             & 2500            & 1.744 $10^{-2}$ & 3.502                            \\
-                                             & 1250            & 1.539 $10^{-3}$ & 3.502                            \\
-                                             & 625             & 1.361 $10^{-4}$ & 3.499                            \\
-                                             & 320             & 1.333 $10^{-5}$ & 3.470                            \\
-                                             & 100             & 7.294 $10^{-7}$ & 2.498                            \\
+                 &   5000.0        &  0.139E+00    &  0.141E+00     & /            & /             \\
+                 &   2500.0        &  0.122E-01    &  0.124E-01     & 3.50         & 3.50          \\
+                 &   1250.0        &  0.108E-02    &  0.110E-02     & 3.50         & 3.50          \\
+                 &    625.0        &  0.956E-04    &  0.969E-04     & 3.50         & 3.50          \\
+                 &    320.0        &  0.937E-05    &  0.949E-05     & 3.47         & 3.47          \\
+                 &    100.0        &  0.512E-06    &  0.519E-06     & 2.50         & 2.50          \\
     \bottomrule
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-adams-bashforth-2.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
   \caption{Oscillation equations solutions computed by means of Adams-Bashforth 2 steps solver}\label{fig:results-oscillation-adams-bashforth-2}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-adams-bashforth-3.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
   \caption{Oscillation equations solutions computed by means of Adams-Bashforth 3 steps solver}\label{fig:results-oscillation-adams-bashforth-3}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-leapfrog.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
   \caption{Oscillation equations solutions computed by means of leapfrog 2 steps solver}\label{fig:results-oscillation-leapfrog}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-ls-runge-kutta-5.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-5.png}
   \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta 5 stages solver}\label{fig:results-oscillation-ls-runge-kutta-5}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-tvd-runge-kutta-2.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-2.png}
   \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 2 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-2}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-tvd-runge-kutta-3.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-3.png}
   \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 3 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-3}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/error_analysis-oscillation-tvd-runge-kutta-5.png}
+  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-5.png}
   \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 5 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-5}
@@
 
+{\color{red} To be written.}
+
 \section{Concluding Remarks and Perspectives}\label{sec:conclusions}
+
+{\color{red} To be written.}
 

### 2015-09-29 papers/announcement/errors_analysis/errors_analysis.sh

@@ -2,32 +2,30 @@
 # script automotion for FOODiE's errors analysis of the provided tests
 cd ../../../
 echo "Building tests executables"
-FoBiS.py build -mode error-analysis --build_dir papers/announcement/error_analysis/tests > buils.log
+FoBiS.py build -mode errors-analysis --build_dir papers/announcement/errors_analysis/tests > buils.log
 cd -
-echo "Oscillation test analysis"
+echo "Oscillation test errors analysis"
 rm -rf results-oscillation
-echo "  Analize leapfrog solver"
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s leapfrog > leapfrog.log
-echo "  Analize Adams-Bashforth solver"
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s adams-bashforth --ss 2 > adams-bashforth-2.log
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s adams-bashforth --ss 3 > adams-bashforth-3.log
-echo "  Analize low storage Runge-Kutta solver"
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s ls-runge-kutta --ss 5 > ls-runge-kutta-5.log
-echo "  Analize TVD/SSP Runge-Kutta solver"
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s tvd-runge-kutta --ss 2 > tvd-runge-kutta-2.log
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s tvd-runge-kutta --ss 3 > tvd-runge-kutta-3.log
-./utilities-oscillation/error_analysis.py -Dts 5000 2500 1250 625 320 100 -ex -s tvd-runge-kutta --ss 5 > tvd-runge-kutta-5.log
-mkdir -p results-oscillation
-mv *log *dat results-oscillation
+mkdir results-oscillation
 cd results-oscillation
-cp ../utilities-oscillation/*lay .
-cp ../utilities/lay_export* .
-lay_export_all_f error_analysis-oscillation-adams-bashforth-2.lay
-lay_export_all_f error_analysis-oscillation-adams-bashforth-3.lay
-lay_export_all_f error_analysis-oscillation-leapfrog.lay
[…]

### 2015-09-29 papers/announcement/errors_analysis/utilities-oscillation/error_analysis.py

@@ -1,164 +1,0 @@
-#!/usr/bin/env python
-from __future__ import division
-from __future__ import print_function
-import argparse
-import numpy as np
-import re
-import subprocess
-import sys
-
-
-def syswork(cmd):
-  """
-  Execute system command 'cmd'.
-  """
-  work_error = 0
-  try:
-    work_output = subprocess.check_output(cmd, shell=True, stderr=subprocess.STDOUT)
-  except subprocess.CalledProcessError as err:
-    work_error = err.returncode
-    work_output = err.output
-  if sys.version_info[0] > 2:
-    work_output = work_output.decode("ascii")
-  return [work_error, str(work_output)]
-
-
-def compute_error(exact, numerical):
-  """
-  Compute error of numerical solution once the exact one is provided.
-
-  Parameters
-  ----------
-  exact : numpy array
-    exact solution
-  numerical : numpy array
-    numerical solution
-
-  Returns
-  error_estimate : real
-    Norm L2 of error estimate
-  """
-  error_estimate = 0.
-  for i in range(numerical.shape[0]):
-    error_estimate += (numerical[i][0] - exact[i][0])**2 + (numerical[i][1] - exact[i][1])**2
-  return np.sqrt(error_estimate)
-
-
-def load_oscillation_results(filename):
-  """Load Osciallation test results.
-
-  Parameters
-  ----------
-  filename : str
-    output file name
-
-  Returns
-  -------
-  numerical : list
-    list of numerical results
-  exact : list
-    list of exact results
-  """
-  numerical = []
-  exact = []
-  with open(filename) as foodie_results:
-    zones = re.split('ZONE.*\n', foodie_results.read())
-    sol_steps = zones[1].split('\n')
-    for sol in sol_steps:
-      if sol.strip() != '':
-        variables = re.split(' *', sol.strip())
-        numerical.append([variables[1], variables[2]])
-    sol_steps = zones[2].split('\n')
-    for sol in sol_steps:
-      if sol.strip() != '':
-        variables = re.split(' *', sol.strip())
-        exact.append([variables[1], variables[2]])
-  return numerical, exact
-
-
-def collect_errors(cliargs):
-  """Collect the errors on all simulations performed.
-
-  Parameters
-  ----------
-  cliargs : argpargse parsed arguments
-
-  Returns
-  -------
-  errors : list
-    list of float64 errors
-  """
-  time_steps = cliargs.time_steps
[…]

### 2015-10-01 papers/announcement/manuscript.tex

@@
   U_t = R(t,U) \\
-  U_0 = F(t_0)
+  U_0 = U(t_0)
   \end{matrix}
@@
 \end{equation}
-where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residual function and $F$ is the (vectorial) initial conditions function evaluated at the initial time $t_0$.
+where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residual function and $U(t_0)$ is the (vectorial) initial conditions, namely the state variables function evaluated at the initial time $t_0$.
 In general, the residual function $R$ is a function of the state variable $U$ through which it is a function of time, but it can be also a direct function of time, thus in general $R=R(t,U(t))$ holds.
@@
 
+The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODiE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODiE applications is presented; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.
+
 {\color{red} Add citations to Fortran standards, OOP, TDD, free software, ACP, ADT, Rouson's book references.}
@@
 
-In this section the mathematical and numerical models of the currently implemented FOODiE solvers are described.
+In many (most) circumstances, the solution of equation \ref{eq:IVP} cannot be computed in a closed, exact form (even if it exists and is unique) due to the complexity and nature of the residual functions, that is often non linear.
+Consequently, the problem is often solved relying on a numerical approach: the solution of system \ref{eq:IVP} at a time $t^n$, namely $U(t^n)$, is approximated by a subsequent time-marching approximations $U_0=u_0 \rightarrow u_1 \rightarrow u_2 \rightarrow ...
+\rightarrow u_N\approx U(t^n)$ where the relation $u_i \rightarrow u_{i+1}$ implies a \emph{stepping, numerical integration} from the time $t^i$ to time $t^{i+1}$ and $N$ is the total number of numerical time steps necessary to evolve the initial conditions toward the searched solution $U(t^n)$.
+To this aim, many numerical schemes have been devised.
+Notably, the numerical schemes of practical usefulness must posses some necessary proprieties such as \emph{consistency} and \emph{stability} to ensure that the numerical approximation \emph{converges} to the exact solution as the numerical time step tends to zero.
+A detailed discussion of these details is out the scope of the present work and is omitted.
+Here, we briefly recall some classifications necessary to introduce the schemes implemented into the FOODiE library.
+
+A non comprehensive classification of the most widely used schemes could distinguish between \emph{multi-step} versus \emph{one-step} schemes and between \emph{explicit} versus \emph{implicit} schemes.
+
+Essentially, the multi-step schemes have been developed to obtain an accurate approximation of the subsequent numerical steps using the informations contained into the previously computed steps, thus this approach relates the next step approximation to a set of the previously computed steps.
+On the contrary, a one-step scheme evolves the solution toward the next step using only the information coming from the current time approximation.
+In the framework of one-step schemes family an equivalent accurate approximation can be obtained by means of a multi-stage approach as the one due to Runge-Kutta.
+The current version of FOODiE provides schemes belonging to both these families.
+
+The other ODE solvers classification concerns with explicit or implicit nature of the schemes employed.
+Briefly, an explicit scheme computes the next step approximation using the previously computed steps at most to the current time, whereas an implicit scheme uses also the next step approximation (that is the unknown), thus it requires extra computations.
+The implicit approach is of practical use for \emph{stiff} systems where the usage of explicit schemes could require an extremely small time step to evolve in a \emph{stable} way the solution.
+Currently, FOODiE provides only explicit solvers.
+
+FOODiE currently implements the following ODE schemes:
+
+\begin{itemize}
+  \item one-step schemes:
+    \begin{itemize}
+      \item explicit forward Euler scheme, it being $1^{st}$ order accurate;
+      \item explicit Runge-Kutta schemes:
+        \begin{itemize}
+          \item low storage Runge-Kutta schemes:
+            \begin{itemize}
+              \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
+              \end{itemize}
+          \item TVD/SSP Runge-Kutta schemes:
+            \begin{itemize}
+              \item 2-stages, it being $2^{nd}$ order accurate;
+              \item 3-stages, it being $3^{rd}$ order accurate;
+              \item 5-stages, it being $4^{th}$ order accurate;
+              \end{itemize}
+          \end{itemize}
+      \end{itemize}
+  \item multi-step schemes:
+    \begin{itemize}
+      \item Adams-Bashforth schemes:
+        \begin{itemize}
+          \item 2-steps, it being $2^{nd}$ order accurate;
+          \item 3-steps, it being $3^{rd}$ order accurate;
+          \end{itemize}
+      \item Leapfrog schemes:
+        \begin{itemize}
+          \item 2-steps unfiltered, it being $2^{nd}$ order accurate, but mostly \emph{unstable};
+          \item 2-steps Robert-Asselin filtered, it being $1^{st}$ order accurate;
+          \item 2-steps Robert-Asselin-Williams filtered, it being $3^{rd}$ order accurate;
+          \end{itemize}
+      \end{itemize}
+  \end{itemize}
+
+{\color{red} Add citations to all solvers reference papers.}
 
 \subsection{The explicit forward Euler scheme}
+
+{\color{red} To be written.}
+
+\subsection{The explicit low storage Runge-Kutta class of schemes}
+
+{\color{red} To be written.}
+
+\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
 
@@
 \subsection{The leapfrog solver}
-
-{\color{red} To be written.}
-
-\subsection{The explicit low storage Runge-Kutta class of schemes}
-
-{\color{red} To be written.}
-
-\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
 
@@
 
-\section{Ongoing Development Activities}\label{sec:ongoing}
+% \section{Ongoing Development Activities}\label{sec:ongoing}
 
-{\color{red} To be written.}
+% {\color{red} To be written.}
 

### 2015-10-05 papers/announcement/errors_analysis/errors_analysis.sh
Some progress on the manuscript
+ Some progress on the manuscript;
+ add Adams-Bashforth 4 steps scheme;
+ modify the solvers integrate API: change *field* argument name with
*U* name in order to be more consistent with the math formulation of
IVP.

@@ -15,6 +15,7 @@ if [ -x "$tecplot" ] ; then
   ln -fs ../utilities/lay_export* .
   ./lay_export_all_f errors_analysis-oscillation-adams-bashforth-2.lay
   ./lay_export_all_f errors_analysis-oscillation-adams-bashforth-3.lay
+  ./lay_export_all_f errors_analysis-oscillation-adams-bashforth-4.lay
   ./lay_export_all_f errors_analysis-oscillation-leapfrog.lay
   ./lay_export_all_f errors_analysis-oscillation-ls-runge-kutta-5.lay
   ./lay_export_all_f errors_analysis-oscillation-tvd-runge-kutta-2.lay

### 2015-10-05 src/lib/foodie_integrator_low_storage_runge_kutta.f90

@@ -158,14 +158,14 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine destroy
 
-  subroutine integrate(self, field, stage, Dt, t)
+  subroutine integrate(self, U, stage, Dt, t)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with explicit low storage Runge-Kutta scheme.
   !<
   !< @note This method can be used **after** the integrator is created (i.e. the RK coeficients are initialized).
   !---------------------------------------------------------------------------------------------------------------------------------
   class(ls_runge_kutta_integrator), intent(IN)    :: self       !< Actual RK integrator.
-  class(integrand),                 intent(INOUT) :: field      !< Field to be integrated.
+  class(integrand),                 intent(INOUT) :: U          !< Field to be integrated.
   class(integrand),                 intent(INOUT) :: stage(1:2) !< Runge-Kutta registers [1:2].
   real(R_P),                        intent(IN)    :: Dt         !< Time step.
   real(R_P),                        intent(IN)    :: t          !< Time.
@@ -176,13 +176,13 @@ contains
   select type(stage)
   class is(integrand)
     ! computing stages
-    stage(1) = field
-    stage(2) = field*0._R_P
+    stage(1) = U
+    stage(2) = U*0._R_P
     do s=1, self%stages
       stage(2) = stage(2) * self%A(s) + stage(1)%t(t=t + self%C(s) * Dt) * Dt
       stage(1) = stage(1) + stage(2) * self%B(s)
     enddo
-    field = stage(1)
+    U = stage(1)
   endselect
   return
   !---------------------------------------------------------------------------------------------------------------------------------

### 2015-10-05 src/lib/foodie_integrator_tvd_runge_kutta.f90

@@ -190,14 +190,14 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine destroy
 
-  subroutine integrate(self, field, stage, Dt, t)
+  subroutine integrate(self, U, stage, Dt, t)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with explicit TVD (or SSP) Runge-Kutta scheme.
   !<
   !< @note This method can be used **after** the integrator is created (i.e. the RK coeficients are initialized).
   !---------------------------------------------------------------------------------------------------------------------------------
   class(tvd_runge_kutta_integrator), intent(IN)    :: self      !< Actual RK integrator.
-  class(integrand),                  intent(INOUT) :: field     !< Field to be integrated.
+  class(integrand),                  intent(INOUT) :: U         !< Field to be integrated.
   class(integrand),                  intent(INOUT) :: stage(1:) !< Runge-Kutta stages [1:stages].
   real(R_P),                         intent(IN)    :: Dt        !< Time step.
   real(R_P),                         intent(IN)    :: t         !< Time.
@@ -210,7 +210,7 @@ contains
   class is(integrand)
     ! computing stages
     do s=1, self%stages
-      stage(s) = field
+      stage(s) = U
       do ss=1, s - 1
         stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
       enddo
@@ -218,7 +218,7 @@ contains
     enddo
     ! computing new time step
     do s=1, self%stages
-      field = field +  stage(s) * (Dt * self%beta(s))
+      U = U +  stage(s) * (Dt * self%beta(s))
     enddo
   endselect
   return

### 2015-10-05 src/lib/type_integrand.f90

@@ -29,14 +29,13 @@ type, abstract :: integrand
     procedure(symmetric_operator),   pass(lhs),  deferred, public :: add                          !< Integrand + integrand oprator.
     procedure(symmetric_operator),   pass(lhs),  deferred, public :: sub                          !< Integrand - integrand oprator.
     procedure(assignment_integrand), pass(lhs),  deferred, public :: assign_integrand             !< Integrand = integrand.
-    procedure(assignment_real),      pass(lhs),  deferred, public :: assign_real                  !< Integrand = real.
     ! operators overloading
     generic, public :: operator(+) => add                              !< Overloading + operator.
     generic, public :: operator(-) => sub                              !< Overloading - operator.
     generic, public :: operator(*) => integrand_multiply_integrand, &
                                       real_multiply_integrand, &
                                       integrand_multiply_real          !< Overloading * operator.
-    generic, public :: assignment(=) => assign_integrand, assign_real  !< Overloading = assignament.
+    generic, public :: assignment(=) => assign_integrand               !< Overloading = assignament.
 endtype integrand
 
 abstract interface
@@ -117,16 +116,6 @@ abstract interface
   class(integrand), intent(IN)    :: rhs !< Right hand side.
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine assignment_integrand
-
-  subroutine assignment_real(lhs, rhs)
-  !---------------------------------------------------------------------------------------------------------------------------------
-  !< Asymmetric assignment integrand = real.
-  !---------------------------------------------------------------------------------------------------------------------------------
-  import :: integrand, R_P
-  class(integrand), intent(INOUT) :: lhs !< Left hand side.
-  real(R_P),        intent(IN)    :: rhs !< Right hand side.
-  !---------------------------------------------------------------------------------------------------------------------------------
-  endsubroutine assignment_real
 endinterface
 !-----------------------------------------------------------------------------------------------------------------------------------
 endmodule type_integrand

### 2015-10-05 src/tests/burgers/burgers.f90

@@ -141,7 +141,7 @@ contains
   integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(burgers)                    :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer                          :: step                  !< Time steps counter.
   real(R_P)                        :: dt                    !< Time step.
   real(R_P)                        :: t(1:ab_steps)         !< Times.
@@ -154,7 +154,12 @@ contains
   do s=1, ab_steps
     print "(A)", ' AB-'//trim(str(.true.,s))
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
+    select case(s)
+    case(1, 2, 3)
+      call rk_integrator%init(stages=s)
+    case(4)
+      call rk_integrator%init(stages=5)
+    endselect
     call domain%init(initial_state=initial_state, Ni=Ni, h=h, nu=nu, steps=s)
     dt = domain%dt(CFL=CFL)
     t = 0._R_P
@@ -162,14 +167,14 @@ contains
     do while(t(s)<t_final)
       if (s>=step) then
         ! the time steps from 1 to s - 1 must be computed with other scheme...
-        call rk_integrator%integrate(field=domain, stage=rk_stage(1:s), dt=dt, t=t(s))
+        call rk_integrator%integrate(U=domain, stage=rk_stage(1:s), dt=dt, t=t(s))
         if (step>1) then
           t(step) = t(step-1) + dt
         else
           t(step) = dt
         endif
       else
-        call ab_integrator%integrate(field=domain, dt=dt, t=t)
+        call ab_integrator%integrate(U=domain, dt=dt, t=t)
         do ss=1, s-1
           t(ss) = t(ss + 1)
         enddo
@@ -202,7 +207,7 @@ contains
   dt = domain%dt(CFL=CFL)
   t = 0._R_P
   do while(t<t_final)
-    call euler_integrator%integrate(field=domain, dt=dt, t=t)
+    call euler_integrator%integrate(U=domain, dt=dt, t=t)
     t = t + dt
   enddo
[…]

### 2015-10-05 src/tests/euler-1D/euler-1D.f90

@@ -338,7 +338,7 @@ contains
   integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(euler_1D)                   :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer                          :: step                  !< Time steps counter.
   real(R_P)                        :: dt                    !< Time step.
   real(R_P)                        :: t(1:ab_steps)         !< Times.
@@ -355,14 +355,16 @@ contains
     print "(A)", ' AB-'//trim(str(.true.,s))
     title = '1D Euler equations integration, explicit Adams-Bashforth, t='//str(n=t_final)//trim(str(.true., s))//' steps'
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
     select case(s)
     case(1)
       call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, steps=s, ord=1)
+      call rk_integrator%init(stages=s)
     case(2, 3)
       call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, steps=s, ord=3)
-    case(5)
+      call rk_integrator%init(stages=s)
+    case(4, 5)
       call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, steps=s, ord=7)
+      call rk_integrator%init(stages=5)
     endselect
     t = 0._R_P
     call save_time_serie(title=title, filename=output//'-'//trim(str(.true., s))//'-time_serie.dat', t=t(s))
@@ -372,14 +374,14 @@ contains
       dt = domain%dt(Nmax=0, Tmax=t_final, t=t(s), CFL=0.1_R_P*CFL)
       if (s>=step) then
         ! the time steps from 1 to s - 1 must be computed with other scheme...
[…]

### 2015-10-05 src/tests/lorenz/lorenz.f90

@@ -130,7 +130,7 @@ contains
   integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(lorenz)                     :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer(I_P)                     :: s                     !< AB steps counter.
   integer                          :: step                  !< Time steps counter.
   !---------------------------------------------------------------------------------------------------------------------------------
@@ -140,16 +140,21 @@ contains
   do s=1, ab_steps
     print "(A)", ' AB-'//trim(str(.true.,s))
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
+    select case(s)
+    case(1, 2, 3)
+      call rk_integrator%init(stages=s)
+    case(4)
+      call rk_integrator%init(stages=5)
+    endselect
     call attractor%init(initial_state=initial_state, sigma=sigma, rho=rho, beta=beta, steps=s)
     solution(0, 0) = 0._R_P
     solution(1:space_dimension, 0) = attractor%output()
     do step = 1, num_steps
       if (s>=step) then
         ! the time steps from 1 to s - 1 must be computed with other scheme...
-        call rk_integrator%integrate(field=attractor, stage=rk_stage(1:s), dt=dt, t=solution(0, step))
+        call rk_integrator%integrate(U=attractor, stage=rk_stage(1:s), dt=dt, t=solution(0, step))
       else
-        call ab_integrator%integrate(field=attractor, dt=dt, t=solution(0, step-s:step-1))
+        call ab_integrator%integrate(U=attractor, dt=dt, t=solution(0, step-s:step-1))
       endif
       solution(0, step) = step * dt
       solution(1:space_dimension, step) = attractor%output()
@@ -176,7 +181,7 @@ contains
   solution(0, 0) = 0._R_P
   solution(1:space_dimension, 0) = attractor%output()
   do step = 1, num_steps
-    call euler_integrator%integrate(field=attractor, dt=dt, t=solution(0, step))
+    call euler_integrator%integrate(U=attractor, dt=dt, t=solution(0, step))
[…]

### 2015-10-05 src/tests/oscillation/oscillation.f90

@@ -19,7 +19,7 @@ use pyplot_module, only :  pyplot
 !-----------------------------------------------------------------------------------------------------------------------------------
 implicit none
 type(Type_Command_Line_Interface) :: cli                                               !< Command line interface handler.
-type(oscillation)                 :: attractor                                         !< Oscillation field.
+type(oscillation)                 :: oscillator                                        !< Oscillation field.
 integer,   parameter              :: space_dimension=2                                 !< Space dimensions.
 real(R_P), parameter              :: initial_state(1:space_dimension)=[0._R_P, 1._R_P] !< Initial state.
 real(R_P)                         :: f                                                 !< Oscillation frequency.
@@ -190,7 +190,7 @@ contains
   integer, parameter               :: rk_stages=3           !< Runge-Kutta stages number.
   type(oscillation)                :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer                          :: step                  !< Time steps counter.
   integer(I_P)                     :: s                     !< AB steps counter.
   integer(I_P)                     :: steps_range(1:2)      !< Steps used.
@@ -204,21 +204,26 @@ contains
     print "(A)", ' AB-'//trim(str(.true.,s))
     title = 'Oscillation equation integration, explicit Adams-Bashforth, t='//str(n=t_final)//' steps='//trim(str(.true., s))
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
-    call attractor%init(initial_state=initial_state, f=f, steps=s)
+    select case(s)
+    case(1, 2, 3)
+      call rk_integrator%init(stages=s)
+    case(4)
+      call rk_integrator%init(stages=5)
+    endselect
+    call oscillator%init(initial_state=initial_state, f=f, steps=s)
     solution(0, 0) = 0._R_P
-    solution(1:space_dimension, 0) = attractor%output()
+    solution(1:space_dimension, 0) = oscillator%output()
[…]

### 2015-10-05 papers/announcement/manuscript.tex

@@
 
-The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODiE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODiE applications is presented; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.
+The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODiE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODiE applications is presented; section \ref{sec:parallel} provides an analysis of FOODiE performances under parallel frameworks scenario like the OpenMP and MPI paradigms; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.
 
@@
         \begin{itemize}
-          \item low storage Runge-Kutta schemes:
-            \begin{itemize}
-              \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \end{itemize}
           \item TVD/SSP Runge-Kutta schemes:
@@
               \item 5-stages, it being $4^{th}$ order accurate;
+              \end{itemize}
+          \item low storage Runge-Kutta schemes:
+            \begin{itemize}
+              \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
               \end{itemize}
@@
           \item 3-steps, it being $3^{rd}$ order accurate;
+          \item 4-steps, it being $4^{th}$ order accurate;
           \end{itemize}
-      \item Leapfrog schemes:
+        \item Leapfrog schemes\footnote{Leapfrog method is a 2-steps solver being, in general $2^{nd}$ order accurate, but mostly \emph{unstable}.
+It is well suited for periodic/oscillatory problems, thus the error is generally categorized accordingly to the phase and amplitude errors: the unfiltered scheme provides a $2^{nd}$ order error on phase approximation while the amplitude error is null.
+However, the unfiltered scheme is mostly unstable, thus many filters have been devised.
+In general, the filters retain the accuracy on the phase error, while affect the amplitude one.}:
         \begin{itemize}
           \item 2-steps unfiltered, it being $2^{nd}$ order accurate, but mostly \emph{unstable};
-          \item 2-steps Robert-Asselin filtered, it being $1^{st}$ order accurate;
-          \item 2-steps Robert-Asselin-Williams filtered, it being $3^{rd}$ order accurate;
+          \item 2-steps Robert-Asselin filtered, it being $1^{st}$ order accurate (on \emph{amplitude error});
+          \item 2-steps Robert-Asselin-Williams filtered, it being $3^{rd}$ order accurate (on \emph{amplitude error});
           \end{itemize}
@@
 
-{\color{red} To be written.}
+The explicit forward Euler scheme for ODE integration is probably the simplest solver ever devised.
+Considering the system \ref{eq:IVP}, the solution (approximation) of the state vector $U$ at the time $t^{n+1}=t^n+\Delta t$ assuming to known the solution at time $t^n$ is:
+
+\begin{equation}
+  U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]
+\label{eq:solver-euler}
+\end{equation}
+where the solution at the new time step is computed by means of only the current time solution, thus this is an explicit scheme.
+The solution is an approximation of $1^{st}$ order, the local truncation error being $O(Dt^2)$.
+As well known, this scheme has an absolute (linear) stability locus equals to $|1+\Delta t \lambda|\le 1$ where $\lambda$ contains the eigenvalues of the linear (or linearized) Jacobian matrix of the system.
+
+This scheme is Total Variation Diminishing (TVD), thus satisfies the maximum principle (or the equivalent positivity preserving property).
+
+{\color{red} Add citations.}
+
+\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
+
+Runge-Kutta methods belong to the more general multi-stage family of schemes.
+This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme, but without increasing the number of time steps used, as it is done with the multi-step schemes.
+Essentially, the high order of accuracy is obtained by means of \emph{intermediate values} (the stages) of the solution and its derivative are generated and used within a single time step.
+This commonly implies the allocation of some auxiliary memory registers for storing the intermediate stages.
+
+Notably, the multi-stage schemes class has the attractive property to be \emph{self-starting}: the high order accurate solution can be obtained directly from the previous one, without the necessity to compute \emph{before} a certain number of previous steps, as it happens for the multi-step schemes.
+Moreover, one-step multi-stage methods are suitable for adaptively-varying time-step size (that is also possible for multi-step schemes, but at a cost of more complexity) and for discontinuous solutions, namely discontinued solutions happening at a certain time $t*$ (that in a multi-step framework can involve an overall accuracy degradation).
+
+In general, the TVD/SSP Runge-Kutta schemes provided by FOODiE library are written by means of the following algorithm:
+\begin{equation}
+  U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
+\label{eq:RK}
+\end{equation}
+where $N_s$ is the number of Runge-Kutta stages used and $K_s$ is the $s^{th}$ stage defined as:
+\begin{equation}
+  K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right)
+\label{eq:RK-stage}
+\end{equation}
+It is worth noting that the equations \ref{eq:RK} and \ref{eq:RK-stage} contain also implicit schemes.
+A scheme belonging to this family is operative once the coefficients $\alpha,\, \beta,\, \gamma$ are provided.
+We represent these coefficients using the Butcher's table, that for an explicit scheme where $\gamma_1=\alpha_{1,*}=\alpha_{i,i}=0$ has the form reported in table \ref{tab:butcher-table}.
+
+\begin{table}[!ht]
+\caption{Butcher's table for explicit Runge-Kutta schemes}\label{tab:butcher-table}
+\centering
+\resizebox{0.40\textwidth}{!}{%
+\begin{tabular}{c|ccccc}
+ $\gamma_2$     & $\alpha_{2,1}$   &                  &          &                      &              \\
+ $\gamma_3$     & $\alpha_{3,1}$   & $\alpha_{3,2}$   &          &                      &              \\
+ $\vdots$       & $\vdots$         &                  & $\ddots$ &                      &              \\
+ $\gamma_{N_s}$ & $\alpha_{N_s,1}$ & $\alpha_{N_s,2}$ & $\cdots$ & $\alpha_{N_s,N_s-1}$ &              \\
+ \hline
+                & $\beta_1$        & $\beta_2$        & $\cdots$ & $\beta_{N_s-1}$      & $\beta_{N_s}$\\
+\end{tabular}}
+\end{table}
+
+The equations \ref{eq:RK} and \ref{eq:RK-stage} show that Runge-Kutta methods do not require any additional differentiations of the ODE system for achieving high order accuracy, rather they require additional evaluations of the residual function $R$.
+
+The nature of the scheme and the properties of the solutions obtained depend on the number of stages and on the value of the coefficients selected.
+Currently, FOODiE provides 3 Runge-Kutta schemes having TVD or Strong Stability Preserving (SSP) propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.
+
+\begin{table}[!ht]
+  \centering
+  \caption{Butcher's table of 2 stages, $2^{nd}$ order, Runge-Kutta TVD scheme\label{tab:RK-TVD-butcher-2}}
+  \resizebox{0.15\textwidth}{!}{%
+  \begin{tabular}{c|cc}
+   1 & 1 & 0       \\
+   \hline
+       & 1/2 & 1/2   \\
+  \end{tabular}}
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Butcher's table of 3 stages, $3^{rd}$ order, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-3}}
+  \resizebox{0.20\textwidth}{!}{%
+  \begin{tabular}{c|ccc}
+   1   & 1   &     &      \\
+   1/2 & 1/4 & 1/4 &      \\
+   \hline
+       & 1/6 & 1/6 & 2/3  \\
+  \end{tabular}}
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Butcher's table of 5 stages, $4^{th}$ order accurate, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-5}}
+  \resizebox{0.99\textwidth}{!}{%
+  \begin{tabular}{c|ccccc}
+   0.39175222700392 & 0.39175222700392 &                  &                  &                  &                  \\
+   0.58607968896779 & 0.21766909633821 & 0.36841059262959 &                  &                  &                  \\
+   0.47454236302687 & 0.08269208670950 & 0.13995850206999 & 0.25189177424738 &                  &                  \\
+   0.93501063100924 & 0.06796628370320 & 0.11503469844438 & 0.20703489864929 & 0.54497475021237 &                  \\
+   \hline
+                    & 0.14681187618661 & 0.24848290924556 & 0.10425883036650 & 0.27443890091960 & 0.22600748319395 \\
+  \end{tabular}}
+\end{table}
+The absolute stability locus depends on the coefficients selected, however, as a general principle, we can assume that greater is the stages number and wider is the stability locus on equal accuracy orders.
+
+It is worth noting that FODDiE also provides a one-stage TVD Runge-Kutta solver that reverts back to the explicit forward Euler scheme: it can be used, for example, into a Recursive Order Reduction (ROR) framework that automatically checks some properties of the solution and, in case, reduces the order of the Runge-Kutta solver until those properties are obtained.
+
+{\color{red} Add citations.}
 
@@
 
-{\color{red} To be written.}
+As aforementioned, standard Runge-Kutta schemes have the drawback to require $N_S$ auxiliary memory registers to store the necessary stages data.
+In order to make an efficient use of the available limited computer memory, the class of low storage Runge-Kutta scheme was devised.
+Essentially, the standard Runge-Kutta class (under some specific conditions) can be reformulated allowing a more efficient memory management.
+Currently FOODiE provides a class of \emph{2N registers} storage Runge-Kutta schemes, meaning that the storage of all stages requires only 2 registers of memory with a \emph{word} length $N$ (namely the length of the state vector) in contrast to the standard formulation where $N_s$ registers of the same length $N$ are required.
+This is a dramatic improvement of memory efficiency especially for schemes using a high number of stages ($N_s \ge 4$) where the memory necessary is an half with respect the original formulation.
+Unfortunately, not all standard Runge-Kutta schemes can be reformulated as a low storage one.
 
-\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
+Following the Williamson's approach the standard coefficients are reformulated to the coefficients vectors $A$, $B$ and $C$ and the Runge-Kutta algorithm becomes:
 
-{\color{red} To be written.}
+\begin{equation}
+\begin{matrix}
+  K_1 = U\left(t^n\right) \\
+  K_2 = 0 \\
+  \left.\begin{matrix}
+    K_2 = A_s K_2 + \Delta t \cdot R\left(t^n + C_s \Delta t, K_1\right) \\
+    K_1 = K_1 + B_s K_2
+  \end{matrix}\right\} s=1,2,...N_s\\
+  U\left(t^{n+1}\right) = K_1
+  \end{matrix}
+\label{eq:RK-ls}
+\end{equation}
+
+Currently FOODiE provides only one 5 stages, $4^{th}$ order, 2N registers explicit scheme, the coefficients of which are listed in table \ref{tab:RK-ls-5}.
+
+\begin{table}[!ht]
+  \centering
+  \caption{Williamson's table of 5 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-5}}
+  \resizebox{0.40\textwidth}{!}{%
+  \begin{tabular}{ccc}
+       $A$                                    & $B$                                    & $C$                                   \\
+    \hline
+       $ 0                                  $ & $\frac{1432997174477}{9575080441755 }$ & $ 0                                 $ \\
+       $-\frac{567301805773 }{1357537059087}$ & $\frac{5161836677717}{13612068292357}$ & $\frac{1432997174477}{9575080441755}$ \\
+       $-\frac{2404267990393}{2016746695238}$ & $\frac{1720146321549}{2090206949498 }$ & $\frac{2526269341429}{6820363962896}$ \\
+       $-\frac{3550918686646}{2091501179385}$ & $\frac{3134564353537}{4481467310338 }$ & $\frac{2006345519317}{3224310063776}$ \\
+       $-\frac{1275806237668}{842570457699 }$ & $\frac{2277821191437}{14882151754819}$ & $\frac{2802321613138}{2924317926251}$ \\
+  \end{tabular}}
+\end{table}
+
+Similarly to the TVD/SSP Runge-Kutta class, the low storage class also provides a fail-safe one-stage solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
+
+{\color{red} Add citations.}
 
@@
 
-{\color{red} To be written.}
+Adams-Bashforth methods belong to the more general (linear) multi-step family of schemes.
+This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
+Typically only one new residual function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
+
+In general, the Adams-Bashforth schemes provided by FOODiE library are written by means of the following algorithm (for only explicit schemes):
+
+\begin{equation}
+U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}
+\label{eq:AB}
+\end{equation}
+where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.
+
+Currently FOODiE provides 2, 3, and 4 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.
+
+\begin{table}[!ht]
+  \centering
+  \caption{Explicit Adams-Bashforth coefficients\label{tab:AB-coeff}}
+  \resizebox{0.30\textwidth}{!}{%
+  \begin{tabular}{c|cccc}
+    $N_s$ & $b_1$           & $b_2$            & $b_3$            & $b_4$           \\
+    \hline
+    2     & $-\frac{1}{2}$  & $\frac{3}{2}$    &       /          &       /         \\
+    3     & $\frac{5}{12}$  & $-\frac{16}{12}$ & $\frac{23}{12}$  &       /         \\
+    4     & $-\frac{9}{24}$ & $\frac{37}{24}$  & $-\frac{59}{24}$ & $\frac{55}{24}$ \\
+  \end{tabular}}
+\end{table}
+
+Similarly to the Runge-Kutta classes, the Adams-Bashforth class also provides a fail-safe one-step solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
+
+It is worth noting that for $N_s>1$ the Adams-Bashforth class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided.
+To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
+
+{\color{red} Add citations.}
 
@@
 
-{\color{red} To be written.}
+The \emph{leapfrog} scheme belongs to the multi-step family, it being formally a centered second order approximation in time.
+The leapfrog method (in its original formulation) is mostly unstable, however it is well suited for periodic-oscillatory problems providing a null error on the amplitude value and a formal second order error on the phase one, under the satisfaction of the time-step size stable limit.
+Commonly, the leapfrog methods are said to provide a $2 \Delta t$ computational mode that can generate unphysical, unstable solutions.
+As consequence, the original leapfrog scheme is generally \emph{filtered} in order to suppress these computational modes.
+
+The unfiltered leapfrog scheme provided by FOODiE is:
+\begin{equation}
+  U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]
+\label{eq:leapfrog}
+\end{equation}
+
+FOODiE provides, in a \emph{seamless} API, also filtered leapfrog schemes.
+A widely used filter is due to Robert and Asselin, that suppress the computational modes at the cost of accuracy reduction resulting into a $1^{st}$ order error in amplitude value.
+A more accurate filter, able to provide a $3^{rd}$ order error on amplitude, is a modification of the Robert-Asselin filter due to Williams known as Robert-Asselin-Williams (RAW) filter, that filters the approximation of $U\left(t^{n+1}\right)$ and $U\left(t^{n+2}\right)$ by the following scalar coefficient:
+
+\begin{equation}
+  \begin{matrix}
+    U\left(t^{n+1}\right) = U\left(t^{n+1}\right) + \Delta * \alpha     \\
+    U\left(t^{n+2}\right) = U\left(t^{n+2}\right) + \Delta * (\alpha-1) \\
+    where \\
+    \Delta = \frac{\nu}{2}(U^{n} - 2 U^{n+1} + U^{n+2})
+  \end{matrix}
+\label{eq:leapfrog-RAW}
+\end{equation}
+
+The filter coefficients should be taken as $\nu \in [0,1]$ and $\alpha \in [0.5,1]$.
+If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude.
+The default values of the FOODiE provided scheme are $\nu=0.01$ $\alpha=0.5$, but they can be customized at runtime.
+
+{\color{red} Add citations.}
 
@@
 
-\subsection{The main FOODiE Abstract Data Type: the integrand type}
-Our ACP is based on one main ADT, the \emph{integrand} type, the definition of which is shown in listing \ref{list:type_integrand}.
+As aforementioned, the programming language used is the Fortran 2008 standard, that is a minor revision of the previous Fortran 2003 standard.
+Such a new Fortran idioms provide (among other useful features) an almost complete support for OOP, in particular for ADT concept.
+Fortran 2003 has introduced the \emph{abstract derived type}: it is a derived type suitable to serve as \emph{contract} for concrete type-extensions that has not any actual implementations, rather it provides a well-defined set of type bound procedures interfaces, that in Fortran nomenclature are called \emph{deferred} procedures.
+Using such an abstract definition, we can implement algorithms operating on only this abstract type and on \emph{all its concrete extensions}.
+This is the key feature of FOODiE library: all the above described ODE solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax.
+In the meanwhile, client codes must implement their own IVPs extending only one simple abstract type.
+
+In the subsection \ref{subsec:type_integrand} a review of the FOODiE main ADT, the \emph{integrand} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.
+
+It is worth noting that all FOODiE public \emph{entities} (ADT and solvers) must be accessed by the FOODiE module, see listing \ref{list:main_module} for an example on how access to all public FOODiE entities.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={usage example importing all public entities of FOODiE main module},label={list:main_module}]
+use foodie, only: integrand, &
+                  adams_bashforth_integrator, &
+                  euler_explicit_integrator, &
+                  leapfrog_integrator, &
+                  ls_runge_kutta_integrator, &
+                  tvd_runge_kutta_integrator
+! or simply
+use foodie
+\end{lstlisting}
+
+\subsection{The main FOODiE Abstract Data Type: the integrand type}\label{subsec:type_integrand}
+
+The implemented ACP is based on one main ADT, the \emph{integrand} type, the definition of which is shown in listing~\ref{list:type_integrand}.
 
@@
 It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODiE solver.
-Listing \ref{list:type_integrand} specifies the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT.
+
+As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 3 ODE-specific procedures plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects.
 These procedures are analyzed in the following paragraphs.
@@
 
-This procedure-function takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optional, and returns an integrand object.
+This procedure-function takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optionals, and it returns an integrand object.
 The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
-The optional argument \emph{n} indicated the \emph{time level} at which the residual function must be evaluated: this argument is necessary for multi-level (often referred as multi-step) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family.
-The optional argument \emph{t} is the {time} at which the residual function must be computed: it can be omitted in the case the residual function does not depend directly from time.
+The optional argument \emph{n} indicates the \emph{time level} at which the residual function must be evaluated: this argument is necessary for multi-step (often referred as multi-level) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family.
+The optional argument \emph{t} is the \emph{time} at which the residual function must be computed: it can be omitted in the case the residual function does not depend directly on time.
 
@@
 
-{\color{red} To be completed.}
+This procedure is a necessary complement for using multi-step solvers.
+As its name suggests, it updates the previous time steps solutions (that must be stored on the care of the clients code).
+It takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optionals, and it returns nothing it being a subroutine directly modifying the passed argument.
+The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
+The optional argument \emph{filter} indicates a \emph{displacement} for filtering the field while the last optional argument \emph{weights} specifies the weights that must applied to the displacement for each time steps filtered.
+
+Let us consider the multi-step scheme expressed by the equation \ref{eq:AB}.
+Once the solution at time $t^{n+N_s}$ has been computed the previous $N_s-1$ steps must be updated, e.g.
+an unfiltered update can read as:
+\begin{equation}
+U\left(t^{n+s}\right) = U\left(t^{n+s+1}\right)\quad s=1,..,N_s-1
+\label{eq:previous-steps-update}
+\end{equation}
+and a filtered update can read as:
+\begin{equation}
+U\left(t^{n+s}\right) = \Delta * U\left(t^{n+s+1}\right) \omega_s\quad s=1,..,N_s-1
+\label{eq:previous-steps-update-filtered}
+\end{equation}
+where $\Delta$ is the filter displacement and $\omega_s$ the weight of the step $s-th$.
+This interface implies that the previous time steps must be embedded into the client \emph{integrand} extension.
+This procedure is automatically called by each multi-step solver after the new step $t^{n+N_s}$ has been computed, thus alleviating client codes from its explicit call.
+
+In the case the client codes will use only one-step solvers this procedure can be defined as simple \emph{empty} subroutine as shown in listing \ref{list:empty_update_previous_time_steps}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={empty implementation of the update previous time steps procedure},label={list:empty_update_previous_time_steps}]
+subroutine update_previous_steps(self, filter, weights)
+class(my_integrand),        intent(INOUT) :: self       !< Integrand field.
+class(integrand), optional, intent(IN)    :: filter     !< Filter field displacement.
+real(R_P),        optional, intent(IN)    :: weights(:) !< Weights for filtering the steps.
+! do nothing
+return
+endsubroutine update_previous_steps
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
 
@@
 
-{\color{red} To be completed.}
+This procedure is a necessary complement for using multi-step solvers.
+It returns the integrand solution at the specified previous time step (level).
+It takes two arguments and it returns an integrand object.
+The first argument passed as a \emph{type bounded} argument is the current instance of the integrand field object.
+The second argument, named \emph{n}, indicates the \emph{time level} at which the solution must be returned.
+
+This procedure is used in multi-step methods like the leapfrog where there is an explicit dependency on also the solution at previous time steps.
+
+For the case in which multi-step solvers are not used, this procedure can have a concrete implementation that does nothing as shown in listing \ref{list:empty_previous_step}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={empty implementation of the previous step procedure},label={list:empty_previous_step}]
+function previous_step(self, n) result(previous)
+class(my_integrand), intent(IN)  :: self     !< Integrand field.
+integer(I_P),        intent(IN)  :: n        !< Time level.
+class(integrand), allocatable    :: previous !< Selected previous time integrand field.
+! do nothing
+return
+endfunction previous_step
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT. Note that, in this case, the returned result is not allocated.
 
@@
 
-{\color{red} To be completed.}
+This interface defines a class of procedures operating on 2 \emph{integrand} objects, namely it is used for the definition of the operators \emph{multiplication}, \emph{summation} and \emph{subtraction} of integrand objects.
+These operators are used into the above described ODE solvers, for example see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
+The implementation details of such a procedures class are strictly dependent on the concrete extension of the integrand type.
+From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are two integrand object and the returned object is still an integrand one.
 
@@
 
-{\color{red} To be completed.}
+These two interfaces are necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the multiplication of an integrand object for a real number, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
+The implementation details of these procedures are strictly dependent on the concrete extension of the integrand type.
+From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are an integrand object and real number of viceversa and the returned object is still an integrand one.
 
@@
 
-{\color{red} To be completed.}
+The assignment statement is necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the assignment of an integrand object by another one, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
+The implementation details of this assignment is strictly dependent on the concrete extension of the integrand type.
+From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the assignment, and the second argument, the right-hand-side of the assignment, are two integrand objects.
 
@@
 
-\subsection{The explicit forward Euler solver}
+\subsection{The explicit forward Euler solver}\label{subsec:solver_euler}
+
+The explicit forward Euler solver is exposed (by the FOODiE main module that must imported, see listing \ref{list:main_module}) as a single derived type (that is a standard convention for all FOODiE solvers) named \emph{euler\_explicit\_integrator}.
+It provides the type bound procedure (also referred as \emph{method}) \emph{integrate} for integrating in time an \emph{integrand} object, or any of its polymorphic concrete extensions.
+Consequently, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_euler_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit forward Euler integrator},label={list:define_euler_solver}]
+use FOODiE, only: euler_explicit_integrator
+type(euler_explicit_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it can be directly used without any initialization, for example see listing \ref{list:use_euler_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an explicit forward Euler integrator},label={list:use_euler_solver}]
+type(my_integrand) :: my_field
+call integrator%integrate(U=my_field, Dt=0.1)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
+
+The complete implementation of the \emph{integrate} method of the explicit forward Euler solver is reported in listing \ref{list:euler_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of Euler solver},label={list:euler_integrate}]
+subroutine integrate(U, Dt, t)
+class(integrand),    intent(INOUT) :: U  !< Field to be integrated.
+real(R_P),           intent(IN)    :: Dt !< Time step.
+real(R_P), optional, intent(IN)    :: t  !< Time.
+U = U + U%t(t=t) * Dt
+return
+endsubroutine integrate
+\end{lstlisting}
+This method takes three arguments, the first argument is an integrand class, it being the integrand field that must integrated one-step-over in time, the second is the time step used and the third, that is optional, is the current time value that is passed to the residual function for taking into account the cases where the time derivative explicitly depends on time.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+
+As shown from listing \ref{list:euler_integrate} the \emph{integrand} ADT has allowed a very high-level implementation of the Euler scheme.
+For example let us compare the implementation instructions with the Euler's scheme equation \ref{eq:solver-euler}:
+
+$$
+U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right] \rightarrow U = U + U\%t(t=t) * Dt
+$$
+
+As shown, the Fortran implementation is almost equivalent to the rigorous mathematical formulation.
+This aspect directly implies that the implementation of Euler's scheme (as of all other schemes) is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the Euler's solver must be implemented for each specific definition of the integrand $U$.
+
+\subsection{The explicit TVD/SSP Runge-Kutta class of solvers}\label{subsec:solver_tvd_rk}
+
+The TVD/SSP Runge-Kutta class of solvers is exposed as a single derived type named \emph{tvd\_runge\_kutta\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator among the possibilities offered by the class of solvers;
+  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time.
+  \end{itemize}
+
+As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_tvd_rk_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit TVD/SSP Runge-Kutta integrator},label={list:define_tvd_rk_solver}]
+use FOODiE, only: tvd_runge_kutta_integrator
+type(tvd_runge_kutta_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_tvd_rk_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit TVD/SSP Runge-Kutta integrator},label={list:init_tvd_rk_solver}]
+call integrator%init(stages=3)
+\end{lstlisting}
+
+In the listing \ref{list:init_tvd_rk_solver} a 3-stages solver has been initialized.
+As a matter of facts, from the equations \ref{eq:RK} and \ref{eq:RK-stage} a solver belonging to this class is completely defined once the number of stages adopted has been chosen.
+The complete definition of the \emph{tvd\_runge\_kutta\_integrator} type is reported into listing \ref{list:tvd_rk_solver}.
+As shown, the Butcher's coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{tvd\_runge\_kutta\_integrator} type},label={list:tvd_rk_solver}]
+type :: tvd_runge_kutta_integrator
+  integer(I_P)           :: stages=0  ! Number of stages.
+  real(R_P), allocatable :: alph(:,:) ! alpha Butcher's coefficients.
+  real(R_P), allocatable :: beta(:)   ! beta Butcher's coefficients.
+  real(R_P), allocatable :: gamm(:)   ! gamma Butcher's coefficients.
+  contains
+    procedure, pass(self), public :: destroy   ! Destroy the integrator.
+    procedure, pass(self), public :: init      ! Initialize (create) the integrator.
+    procedure, pass(self), public :: integrate ! Integrate integrand field.
+    final                         :: finalize  ! Finalize object.
+endtype tvd_runge_kutta_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can used for integrating an integrand field, as shown in listing \ref{list:use_tvd_rk_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a TVD/SSP Runge-Kutta integrator},label={list:use_tvd_rk_solver}]
+type(my_integrand) :: my_field
+type(my_integrand) :: my_stages(1:3)
+call integrator%integrate(U=my_field, stage=my_stage, Dt=0.1)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
+Listing \ref{list:use_tvd_rk_solver} shows that the memory registers necessary for storing the Runge-Kutta stages must be supplied by the client code.
+
+The complete implementation of the \emph{integrate} method of the explicit TVD/SSP Runge-Kutta class of solvers is reported in listing \ref{list:tvd_rk_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of TVD/SSP Runge-Kutta class},label={list:tvd_rk_integrate}]
+subroutine integrate(self, U, stage, Dt, t)
+class(tvd_runge_kutta_integrator), intent(IN)    :: self      !< Actual RK integrator.
+class(integrand),                  intent(INOUT) :: U         !< Field to be integrated.
+class(integrand),                  intent(INOUT) :: stage(1:) !< Runge-Kutta stages [1:stages].
+real(R_P),                         intent(IN)    :: Dt        !< Time step.
+real(R_P),                         intent(IN)    :: t         !< Time.
+integer(I_P)                                     :: s         !< First stages counter.
+integer(I_P)                                     :: ss        !< Second stages counter.
+select type(stage)
+class is(integrand)
+  do s=1, self%stages
+    stage(s) = U
+    do ss=1, s - 1
+      stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
+    enddo
+    stage(s) = stage(s)%t(t=t + self%gamm(s) * Dt)
+  enddo
+  do s=1, self%stages
+    U = U +  stage(s) * (Dt * self%beta(s))
+  enddo
+endselect
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residual function for taking into account the cases where the time derivative explicitly depends on time.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+
+It is worth noting that the stages memory registers, namely the array \emph{stage}, must be passed as argument because it is defined as a \emph{not-passed} polymorphic argument, thus we are not allowed to define it as an automatic array of the \emph{integrate} method.
+
+As it happens for the Euler's scheme, the \emph{integrand} ADT greatly simplifies the Runge-Kutta algorithm implementation.
+For example let us compare the implementation instructions with the Runge-Kutta equations \ref{eq:RK} and \ref{eq:RK-stage}:
+
+$$
+\begin{matrix}
+  K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right) \\
+  U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
+\end{matrix}
+\rightarrow
+\begin{matrix}
+  stage(s) = stage(s) + stage(ss) * (Dt * self\%alph(s, ss)) \\
+  U = U +  stage(s) * (Dt * self\%beta(s))
+\end{matrix}
+$$
+where the syntax "$(s)$" and "$(ss)$" imply the summation operation. As for the Euler's scheme, the Fortran implementation is almost equivalent to the rigorous mathematical formulation.
+
+\subsection{The explicit low storage Runge-Kutta class of solvers}\label{subsec:solver_ls_rk}
 
@@
 
-\subsection{The explicit Adams-Bashforth class of solvers}
+\subsection{The explicit Adams-Bashforth class of solvers}\label{subsec:solver_ab}
 
@@
 
-\subsection{The leapfrog solver}
-
-{\color{red} To be written.}
-
-\subsection{The explicit low storage Runge-Kutta class of solvers}
-
-{\color{red} To be written.}
-
-\subsection{The explicit TVD/SSP Runge-Kutta class of solvers}
+\subsection{The leapfrog solver}\label{subsec:solver_leapfrog}
 
@@
 
-% \section{Ongoing Development Activities}\label{sec:ongoing}
+\section{Benchmarks on parallel frameworks}\label{sec:parallel}
 
-% {\color{red} To be written.}
+{\color{red} To be written.}
+
+\subsection{OpenMP benchmark}\label{subsec:openmp}
+
+{\color{red} To be written.}
+
+\subsection{MPI benchmark}\label{subsec:mpi}
+
+{\color{red} To be written.}
 

### 2015-10-06 papers/announcement/manuscript.tex
Improve manuscript
Improve manuscript:

+ complete API review and disccusion;
+ almost complete Oscillation test disccusion (must be completed the
results analysis).

@@
 \end{equation}
-where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residual function and $U(t_0)$ is the (vectorial) initial conditions, namely the state variables function evaluated at the initial time $t_0$.
-In general, the residual function $R$ is a function of the state variable $U$ through which it is a function of time, but it can be also a direct function of time, thus in general $R=R(t,U(t))$ holds.
+where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residuals function and $U(t_0)$ is the (vectorial) initial conditions, namely the state variables function evaluated at the initial time $t_0$.
+In general, the residuals function $R$ is a function of the state variable $U$ through which it is a function of time, but it can be also a direct function of time, thus in general $R=R(t,U(t))$ holds.
 
@@
 
-It is worth noting that the state vector variables $U$ and its corresponding residual function $U_t = \frac{dU}{dt} = R(t,U)$ are \emph{problem dependent}: the number and the meaning of the state variables as well as the equations governing their evolution (that are embedded into the residual function) are different for the Navier-Stokes conservation laws with respect the Burgers one, as an example.
+It is worth noting that the state vector variables $U$ and its corresponding residuals function $U_t = \frac{dU}{dt} = R(t,U)$ are \emph{problem dependent}: the number and the meaning of the state variables as well as the equations governing their evolution (that are embedded into the residuals function) are different for the Navier-Stokes conservation laws with respect the Burgers one, as an example.
 Nevertheless, the \emph{solvers} used for the prediction of the Navier-Stokes equations evolution (hereafter the \emph{solution}) are the same that are used for Burgers equations time-integration.
@@
 
-In many (most) circumstances, the solution of equation \ref{eq:IVP} cannot be computed in a closed, exact form (even if it exists and is unique) due to the complexity and nature of the residual functions, that is often non linear.
+In many (most) circumstances, the solution of equation \ref{eq:IVP} cannot be computed in a closed, exact form (even if it exists and is unique) due to the complexity and nature of the residuals functions, that is often non linear.
 Consequently, the problem is often solved relying on a numerical approach: the solution of system \ref{eq:IVP} at a time $t^n$, namely $U(t^n)$, is approximated by a subsequent time-marching approximations $U_0=u_0 \rightarrow u_1 \rightarrow u_2 \rightarrow ...
@@
 
-The equations \ref{eq:RK} and \ref{eq:RK-stage} show that Runge-Kutta methods do not require any additional differentiations of the ODE system for achieving high order accuracy, rather they require additional evaluations of the residual function $R$.
+The equations \ref{eq:RK} and \ref{eq:RK-stage} show that Runge-Kutta methods do not require any additional differentiations of the ODE system for achieving high order accuracy, rather they require additional evaluations of the residuals function $R$.
 
@@
 This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
-Typically only one new residual function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
+Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
 
@@
 
-\subsubsection{Time derivative procedure, the residual function}
+\subsubsection{Time derivative procedure, the residuals function}
 
@@
 The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
-The optional argument \emph{n} indicates the \emph{time level} at which the residual function must be evaluated: this argument is necessary for multi-step (often referred as multi-level) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family.
-The optional argument \emph{t} is the \emph{time} at which the residual function must be computed: it can be omitted in the case the residual function does not depend directly on time.
+The optional argument \emph{n} indicates the \emph{time level} at which the residuals function must be evaluated: this argument is necessary for multi-step (often referred as multi-level) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family.
+The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.
 
@@
 \end{lstlisting}
-This method takes three arguments, the first argument is an integrand class, it being the integrand field that must integrated one-step-over in time, the second is the time step used and the third, that is optional, is the current time value that is passed to the residual function for taking into account the cases where the time derivative explicitly depends on time.
+This method takes three arguments, the first argument is an integrand class, it being the integrand field that must integrated one-step-over in time, the second is the time step used and the third, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
 The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
 
-As shown from listing \ref{list:euler_integrate} the \emph{integrand} ADT has allowed a very high-level implementation of the Euler scheme.
-For example let us compare the implementation instructions with the Euler's scheme equation \ref{eq:solver-euler}:
-
-$$
-U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right] \rightarrow U = U + U\%t(t=t) * Dt
-$$
-
-As shown, the Fortran implementation is almost equivalent to the rigorous mathematical formulation.
-This aspect directly implies that the implementation of Euler's scheme (as of all other schemes) is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the Euler's solver must be implemented for each specific definition of the integrand $U$.
-
 \subsection{The explicit TVD/SSP Runge-Kutta class of solvers}\label{subsec:solver_tvd_rk}
@@
 \begin{itemize}
-  \item \emph{init}: initialize the integrator among the possibilities offered by the class of solvers;
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
   \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
@@
   contains
-    procedure, pass(self), public :: destroy   ! Destroy the integrator.
-    procedure, pass(self), public :: init      ! Initialize (create) the integrator.
-    procedure, pass(self), public :: integrate ! Integrate integrand field.
-    final                         :: finalize  ! Finalize object.
+    procedure, pass(self), public :: destroy
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+    final                         :: finalize
 endtype tvd_runge_kutta_integrator
@@
 
-After the solver has been initialized it can used for integrating an integrand field, as shown in listing \ref{list:use_tvd_rk_solver}.
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_tvd_rk_solver}.
 
@@
 
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of TVD/SSP Runge-Kutta class},label={list:tvd_rk_integrate}]
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit TVD/SSP Runge-Kutta class},label={list:tvd_rk_integrate}]
 subroutine integrate(self, U, stage, Dt, t)
-class(tvd_runge_kutta_integrator), intent(IN)    :: self      !< Actual RK integrator.
-class(integrand),                  intent(INOUT) :: U         !< Field to be integrated.
-class(integrand),                  intent(INOUT) :: stage(1:) !< Runge-Kutta stages [1:stages].
-real(R_P),                         intent(IN)    :: Dt        !< Time step.
-real(R_P),                         intent(IN)    :: t         !< Time.
-integer(I_P)                                     :: s         !< First stages counter.
-integer(I_P)                                     :: ss        !< Second stages counter.
+class(tvd_runge_kutta_integrator), intent(IN)    :: self      ! Actual RK integrator.
+class(integrand),                  intent(INOUT) :: U         ! Field to be integrated.
+class(integrand),                  intent(INOUT) :: stage(1:) ! Runge-Kutta stages [1:stages].
+real(R_P),                         intent(IN)    :: Dt        ! Time step.
+real(R_P),                         intent(IN)    :: t         ! Time.
+integer(I_P)                                     :: s         ! First stages counter.
+integer(I_P)                                     :: ss        ! Second stages counter.
 select type(stage)
@@
 
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residual function for taking into account the cases where the time derivative explicitly depends on time.
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
 The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
@@
 
-As it happens for the Euler's scheme, the \emph{integrand} ADT greatly simplifies the Runge-Kutta algorithm implementation.
-For example let us compare the implementation instructions with the Runge-Kutta equations \ref{eq:RK} and \ref{eq:RK-stage}:
-
-$$
-\begin{matrix}
+\subsection{The explicit low storage Runge-Kutta class of solvers}\label{subsec:solver_ls_rk}
+
+The low storage variant of Runge-Kutta class of solvers is exposed as a single derived type named \emph{ls\_runge\_kutta\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
+  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time.
+  \end{itemize}
+
+As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ls_rk_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit low storage Runge-Kutta integrator},label={list:define_ls_rk_solver}]
+use FOODiE, only: ls_runge_kutta_integrator
+type(ls_runge_kutta_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ls_rk_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit low storage Runge-Kutta integrator},label={list:init_ls_rk_solver}]
+call integrator%init(stages=5)
+\end{lstlisting}
+
+In the listing \ref{list:init_ls_rk_solver} a 5-stages solver has been initialized.
+As a matter of facts, from the equation \ref{eq:RK-ls} a solver belonging to this class is completely defined once the number of stages adopted has been chosen.
+The complete definition of the \emph{ls\_runge\_kutta\_integrator} type is reported into listing \ref{list:ls_rk_solver}.
+As shown, the Williamson's coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{ls\_runge\_kutta\_integrator} type},label={list:ls_rk_solver}]
+type :: ls_runge_kutta_integrator
+  integer(I_P)           :: stages=0 ! Number of stages.
+  real(R_P), allocatable :: A(:)     ! Low storage *A* coefficients.
+  real(R_P), allocatable :: B(:)     ! Low storage *B* coefficients.
+  real(R_P), allocatable :: C(:)     ! Low storage *C* coefficients.
+  contains
+    procedure, pass(self), public :: destroy
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+    final                         :: finalize
+endtype ls_runge_kutta_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ls_rk_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a low storage Runge-Kutta integrator},label={list:use_ls_rk_solver}]
+type(my_integrand) :: my_field
+type(my_integrand) :: my_stages(1:2)
+call integrator%integrate(U=my_field, stage=my_stage, Dt=0.1)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
+Listing \ref{list:use_ls_rk_solver} shows that the memory registers necessary for storing the Runge-Kutta stages must be supplied by the client code, as it happens of the TVD/SSP Runge-Kutta class.
+However, now the registers necessary is always 2, independently on the number of stages used, that in the example considered are 5.
+
+The complete implementation of the \emph{integrate} method of the explicit low storage Runge-Kutta class of solvers is reported in listing \ref{list:ls_rk_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit low storage Runge-Kutta class},label={list:ls_rk_integrate}]
+subroutine integrate(self, U, stage, Dt, t)
+class(ls_runge_kutta_integrator), intent(IN)    :: self       ! Actual RK integrator.
+class(integrand),                 intent(INOUT) :: U          ! Field to be integrated.
+class(integrand),                 intent(INOUT) :: stage(1:2) ! Runge-Kutta registers [1:2].
+real(R_P),                        intent(IN)    :: Dt         ! Time step.
+real(R_P),                        intent(IN)    :: t          ! Time.
+integer(I_P)                                    :: s          ! First stages counter.
+select type(stage)
+class is(integrand)
+  stage(1) = U
+  stage(2) = U*0._R_P
+  do s=1, self%stages
+    stage(2) = stage(2) * self%A(s) + stage(1)%t(t=t + self%C(s) * Dt) * Dt
+    stage(1) = stage(1) + stage(2) * self%B(s)
+  enddo
+  U = stage(1)
+endselect
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+
+It is worth noting that the stages memory registers, namely the array \emph{stage}, must be passed as argument because it is defined as a \emph{not-passed} polymorphic argument, thus we are not allowed to define it as an automatic array of the \emph{integrate} method.
+
+\subsection{The explicit Adams-Bashforth class of solvers}\label{subsec:solver_ab}
+
+The explicit Adams-Bashforth class of solvers is exposed as a single derived type named \emph{adams\_bashforth\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
+  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time.
+  \end{itemize}
+
+As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ab_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Adams-Bashforth integrator},label={list:define_ab_solver}]
+use FOODiE, only: adams_bashforth_integrator
+type(adams_bashforth_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ab_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Adams-Bashforth integrator},label={list:init_ab_solver}]
+call integrator%init(steps=4)
+\end{lstlisting}
+
+In the listing \ref{list:init_ab_solver} a 4-steps solver has been initialized.
+As a matter of facts, from the equation \ref{eq:AB} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen.
+The complete definition of the \emph{adams\_bashforth\_integrator} type is reported into listing \ref{list:ab_solver}.
+As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_bashforth\_integrator} type},label={list:ab_solver}]
+type :: adams_bashforth_integrator
+  private
+  integer(I_P)           :: steps=0 ! Number of time steps.
+  real(R_P), allocatable :: b(:)    ! \(b\) coefficients.
+  contains
+    procedure, pass(self), public :: destroy
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+    final                         :: finalize
+endtype adams_bashforth_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ab_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
+real               :: times(1:4)
+type(my_integrand) :: my_field
+call integrator%integrate(U=my_field, Dt=Dt, t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT and \emph{times} are the time at each 4 steps considered for the current one-step-over integration.
+
+The complete implementation of the \emph{integrate} method of the explicit Adams-Bashforth class of solvers is reported in listing \ref{list:ab_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
+subroutine integrate(self, U, Dt, t)
+class(adams_bashforth_integrator), intent(IN)    :: self ! Actual AB integrator.
+class(integrand),                  intent(INOUT) :: U    ! Field to be integrated.
+real(R_P),                         intent(IN)    :: Dt   ! Time steps.
+real(R_P),                         intent(IN)    :: t(:) ! Times.
+integer(I_P)                                     :: s    ! Steps counter.
+do s=1, self%steps
+  U = U + U%t(n=s, t=t(s)) * (Dt * self%b(s))
+enddo
+call U%update_previous_steps
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes four arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the time step used and the fourth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+
+It is worth noting that there is an explicit call to the method \emph{update\_previous\_steps} for updating the values of the previously computed steps.
+
+\subsection{The leapfrog solver}\label{subsec:solver_leapfrog}
+
+The explicit Leapfrog class of solvers is exposed as a single derived type named \emph{leapfrog\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time.
+  \end{itemize}
+
+As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_lf_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Leapfrog integrator},label={list:define_lf_solver}]
+use FOODiE, only: leapfrog_integrator
+type(leapfrog_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_lf_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Leapfrog integrator},label={list:init_lf_solver}]
+! default coefficients nu=0.01, alpha=0.5
+call integrator%init()
+! custom coefficients
+call integrator%init(nu=0.015, alpha=0.6)
+\end{lstlisting}
+
+The complete definition of the \emph{leapfrog\_integrator} type is reported into listing \ref{list:lf_solver}.
+As shown, the filter coefficients are initialized to zero, suitable values are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{leapfrog\_integrator} type},label={list:lf_solver}]
+type :: leapfrog_integrator
+  private
+  real(R_P) :: nu=0.01_R_P   ! Robert-Asselin filter coefficient.
+  real(R_P) :: alpha=0.5_R_P ! Robert-Asselin-Williams filter coefficient.
+  contains
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+endtype leapfrog_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_lf_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a Leapfrog integrator},label={list:use_lf_solver}]
+real               :: times(1:2)
+type(my_integrand) :: filter_displacement
+type(my_integrand) :: my_field
+call integrator%integrate(U=my_field, filter=filter_displacement, Dt=Dt, t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.
+
+The complete implementation of the \emph{integrate} method of the explicit Leapfrog class of solvers is reported in listing \ref{list:lf_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Leapfrog class},label={list:lf_integrate}]
+subroutine integrate(self, U, filter, Dt, t)
+class(leapfrog_integrator), intent(IN)    :: self   ! LF integrator.
+class(integrand),           intent(INOUT) :: U      ! Field to be integrated.
+class(integrand), optional, intent(INOUT) :: filter ! Filter field displacement.
+real(R_P),                  intent(in)    :: Dt     ! Time step.
+real(R_P),                  intent(IN)    :: t      ! Time.
+U = U%previous_step(n=1) + U%t(n=2, t=t) * (Dt * 2._R_P)
+if (present(filter)) then
+  filter = (U%previous_step(n=1) - U%previous_step(n=2) * 2._R_P + U) * self%nu * 0.5_R_P
+endif
+call U%update_previous_steps(filter=filter, weights=[self%alpha, self%alpha - 1._R_P])
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the optional filter-displacement-register, the fourth is the time step used and the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+It is worth noting that if the filter displacement argument is not passed, the solver reverts back to the standard unfiltered Leapfrog method.
+
+It is worth noting that there is an explicit call to the method \emph{update\_previous\_steps} for updating the values of the previously computed steps.
+
+\subsection{General Remarks}\label{subsec:API-remarks}
+
+Table \ref{tab:foodie-vs-math} presents a comparison of the relevant parts of equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:RK-stage}, \ref{eq:RK-ls}, \ref{eq:AB} and \ref{eq:leapfrog} with the corresponding FOODiE implementations reported in listings \ref{list:euler_integrate}, \ref{list:tvd_rk_integrate}, \ref{list:ls_rk_integrate}, \ref{list:ab_integrate} and \ref{list:lf_integrate}, respectively.
+This comparison proves that the \emph{integrand} ADT has allowed a very high-level implementation syntax.
+The Fortran implementation is almost equivalent to the rigorous mathematical formulation.
+This aspect directly implies that the implementation of a ODE solver into the FOODiE library is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the solvers must be implemented for each specific definition of the integrand type, it being not abstract.
+
+\begin{table}[!ht]
+  \centering
+  \caption{Comparison between rigorous mathematical formulation and FOODiE high-level implementation; the syntax "$(s)$" and "$(ss)$" imply the summation operation}\label{tab:foodie-vs-math}
+  \resizebox{0.99\textwidth}{!}{%
+  \begin{tabular}{c|cc}
+    \toprule
+    {\sc Solver} & {\sc Mathematical formulation} & {\sc FOODiE implementation} \\
+    \hline
+    explicit forward Euler & $U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]$ & $U = U + U\%t(t=t) * Dt$\\
+    \hline
+    TVD/SSP Runge-Kutta & $\begin{matrix}
   K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right) \\
   U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
-\end{matrix}
-\rightarrow
-\begin{matrix}
+\end{matrix}$ & $\begin{matrix}
   stage(s) = stage(s) + stage(ss) * (Dt * self\%alph(s, ss)) \\
   U = U +  stage(s) * (Dt * self\%beta(s))
-\end{matrix}
-$$
-where the syntax "$(s)$" and "$(ss)$" imply the summation operation. As for the Euler's scheme, the Fortran implementation is almost equivalent to the rigorous mathematical formulation.
-
-\subsection{The explicit low storage Runge-Kutta class of solvers}\label{subsec:solver_ls_rk}
+\end{matrix}$\\
+    \hline
+    low storage Runge-Kutta & $\begin{matrix}
+  K_2 = A_s K_2 + \Delta t \cdot R\left(t^n + C_s \Delta t, K_1\right) \\
+  K_1 = K_1 + B_s K_2
+\end{matrix}$ & $\begin{matrix}
+  stage(2) = stage(2) * self\%A(s) + \\ + stage(1)\%t(t=t + self\%C(s) * Dt) * Dt \\
+  stage(1) = stage(1) + stage(2) * self\%B(s)
+\end{matrix}$\\
+    \hline
+  explicit Adams-Bashforth & $\begin{matrix}U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) + \\ +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}\end{matrix}$ & $U = U + U\%t(n=s, t=t(s)) * (Dt * self\%b(s))$\\
+    \hline
+    explicit Leapfrog & $ U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]$ & $U = U\%previous\_step(n=1) + U\%t(n=2, t=t) * (Dt * 2.)$\\
+
+  \end{tabular}}
+\end{table}
+
+\clearpage
+
+\section{Tests and Examples}\label{sec:Tests}
 
@@
 
-\subsection{The explicit Adams-Bashforth class of solvers}\label{subsec:solver_ab}
-
-{\color{red} To be written.}
-
-\subsection{The leapfrog solver}\label{subsec:solver_leapfrog}
-
-{\color{red} To be written.}
-
-\section{Tests and Examples}\label{sec:Tests}
-
-{\color{red} To be written.}
-
-\clearpage
-
 \subsection{Oscillation equations test}
 
-{\color{red} To be completed.}
+Let us consider the \emph{oscillation} problem, it being a simple, yet interesting IVP.
+Briefly, the oscillation problem is a prototype problem of non dissipative, oscillatory phenomena.
+For example, let us consider a pendulum subjected to the Coriolis accelerations without dissipation, the motion equations of which can be described by the ODE system \ref{eq:oscillation}.
 
@@
 \end{equation}
-
 where the frequency is chosen as $f=10^4$. The ODE system \ref{eq:oscillation} is completed by the following initial conditions:
@@
 \end{equation}
-
-where $t_0=0$ is the initial time considered. The exact solution is:
+where $t_0=0$ is the initial time considered.
+
+The IVP constituted by equations \ref{eq:oscillation} and \ref{eq:oscillation-ic} is (apparently) simple and its exact solution is known:
 
@@
 \end{equation}
-
 where $\Delta t$ is an arbitrary time step.
 
-\subsubsection{Errors Analysis}
-
-For the analysis of the accuracy of each solvers implemented into FOODiE library, we have integrated the Oscillation equations \ref{eq:oscillation} with different, decreasing time steps in the range $[5000, 2500, 1250, 625, 320, 100]$.
-
+This problem is only apparently simple.
+As a matter of facts, in a non dissipative oscillatory problem the eventual errors in the amplitude approximation can rapidly drive the subsequent series of approximations to an unphysical solution.
+This is of particular relevance if the solution (that is numerically approximated) constitutes a \emph{prediction} far from the initial conditions, that is the common case in weather forecasting.
+
+\subsubsection{FOODiE aware implementation of an oscillation numerical solver}
+The IVP \ref{eq:oscillation} can be easily solved by means of FOODiE library.
+The first block of a FOODiE aware solution consists to define an \emph{oscillation integrand field} defining a concrete extension of the FOODiE \emph{integrand} type.
+Listing \ref{list:oscillation_type} reports the implementation of such an integrand field that is contained into the tests suite shipped within the FOODiE library.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} type},label={list:oscillation_type}]
+type, extends(integrand) :: oscillation
+  private
+  integer(I_P)                           :: dims=0   ! Space dimensions.
+  real(R_P)                              :: f=0._R_P ! Oscillation frequency (Hz).
+  real(R_P), dimension(:),   allocatable :: U        ! Integrand (state) variables, [1:dims].
+  integer(I_P)                           :: steps=0  ! Number of time steps stored.
+  real(R_P), dimension(:,:), allocatable :: previous ! Previous time steps states, [1:dims,1:steps].
+  contains
+    ! auxiliary methods
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: output
+    ! type\_integrand deferred methods
+    procedure, pass(self), public :: t => dOscillation_dt
+    procedure, pass(self), public :: update_previous_steps
+    procedure, pass(self), public :: previous_step
+    procedure, pass(lhs),  public :: integrand_multiply_integrand => &
+                                     oscillation_multiply_oscillation
+    procedure, pass(lhs),  public :: integrand_multiply_real => oscillation_multiply_real
+    procedure, pass(rhs),  public :: real_multiply_integrand => real_multiply_oscillation
+    procedure, pass(lhs),  public :: add => add_oscillation
+    procedure, pass(lhs),  public :: sub => sub_oscillation
+    procedure, pass(lhs),  public :: assign_integrand => oscillation_assign_oscillation
+    procedure, pass(lhs),  public :: assign_real => oscillation_assign_real
+endtype oscillation
+\end{lstlisting}
+
+The \emph{oscillation} field extends the \emph{integrand} ADT making it a concrete type.
+This derived type is very simple: it has 5 data members for storing the state vector and some auxiliary variables, and it implements all the deferred methods necessary for defining a valid concrete extension of the \emph{integrand} ADT (plus 2 auxiliary methods that are not relevant for our discussion).
+The key point is here constituted by the implementation of the deferred methods: the \emph{integrand} ADT does not impose any structure for the data members, that are consequently free to be customized by the client code.
+In this example the data members have a very simple, clean and concise structure:
+\begin{itemize}
+  \item $dims$ is the number of space dimensions; in the case of equation \ref{eq:oscillation} we have $dims=2$, however the test implementation has been kept more general parametrizing this dimension in order to easily allow future modification of the test-program itself;
+  \item $f$ stores the frequency of the oscillatory problem solved, that is here set to $10^{4}$, but it can be changed at runtime in the test-program;
+  \item $U$ is the state vector corresponding directly to the state vector of equation \ref{eq:oscillation};
+  \item $steps$ and $previous$ are two auxiliary data members used only in case the oscillation field is integrated by means of a multi-step solver (thus they could be omitted in case of one-step solver usage):
+    \begin{itemize}
+      \item $steps$ is the number of previous time steps used for computing the next one-step-over integration with respect the current time;
+      \item $previous$ is an array where previously computed time step solutions are stored.
+      \end{itemize}
+  \end{itemize}
+
+As the listing \ref{list:oscillation_type} shows, the FOODiE implementation strictly corresponds to the mathematical formulation embracing all the relevant mathematical aspects into one derived type, a single \emph{object}.
+Here we not review the implementation of all deferred methods, this being out of the scope of the present work: the interested reader can see the tests suite sources shipped within the FOODiE library.
+However, some of these methods are relevant for our discussion, thus they are reviewed.
+
+\paragraph{dOscillation\_dt, the oscillation residuals function}
+
+Probably, the most important methods for an IVP solver is the residuals function.
+As a matter of facts, the ODE equations are implemented into the residuals function.
+However, the FOODiE ADT strongly alleviates the subtle problems that could arise when the ODE solver is hard-implemented within the specific ODE equations.
+As a matter of facts, the \emph{integrand} ADT specifies the precise interface the residuals function must have: if the client code implements a compliant interface, the FOODiE solvers will work as expected, reducing the possible errors location into the ODE equations, having designed the solvers on the ADT and not on the concrete type.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} residuals function},label={list:oscillation_t}]
+function dOscillation_dt(self, n, t) result(dState_dt)
+class(oscillation),     intent(IN) :: self      ! Oscillation field.
+integer(I_P), optional, intent(IN) :: n         ! Time level.
+real(R_P),    optional, intent(IN) :: t         ! Time.
+class(integrand),  allocatable     :: dState_dt ! Oscillation field time derivative.
+integer(I_P)                       :: dn        ! Time level, dummy variable.
+allocate(oscillation :: dState_dt)
+select type(dState_dt)
+class is(oscillation)
+  dState_dt = self
+  if (self%steps>=2) then ! self\%previous should be used
+    dn = self%steps ; if (present(n)) dn = n
+    dState_dt%U(1) = -self%f * self%previous(2, dn)
+    dState_dt%U(2) =  self%f * self%previous(1, dn)
+  else ! self\%previous should not be used, use directly self\%U
+    dState_dt%U(1) = -self%f * self%U(2)
+    dState_dt%U(2) =  self%f * self%U(1)
+  endif
+endselect
+return
+endfunction dOscillation_dt
+\end{lstlisting}
+
+Listing \ref{list:oscillation_t} reports the implementation of the oscillation residuals function: it is very clear and concise.
+Moreover, comparing this listing with the equation \ref{eq:oscillation} the close correspondence between the mathematical formulation and Fortran implementation is evident.
+
+\paragraph{add method, an example of oscillation symmetric operator}
+
+As a prototype of the operators overloading let us consider the \emph{add} operator, it being a prototype of symmetric operators, the implementation of which is presented in listing \ref{list:oscillation_add}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} add operator},label={list:oscillation_add}]
+function add_oscillation(lhs, rhs) result(opr)
+class(oscillation), intent(IN) :: lhs ! Left hand side.
+class(integrand),   intent(IN) :: rhs ! Right hand side.
+class(integrand), allocatable  :: opr ! Operator result.
+allocate(oscillation :: opr)
+select type(opr)
+class is(oscillation)
+  opr = lhs
+  select type(rhs)
+  class is (oscillation)
+    opr%U = lhs%U + rhs%U
+  endselect
+endselect
+return
+endfunction add_Oscillation
+\end{lstlisting}
+It is very simple and clear: firstly all the auxiliary data are copied into the operator result, then the state vector of the result is populated with the addiction between the state vectors of the left-hand-side and right-hand-side.
+This is very intuitive from the mathematical point of view and it helps to reduce implementation errors.
+Similar implementations are possible for all the other operators necessary to define a valid \emph{intregrand} ADT concrete extension.
+
+\paragraph{assignment of an oscillation object}
+
+The assignment overloading of the \emph{oscillation} type is the last key-method that enforces the conciseness of the FOODiE aware implementation.
+Listing \ref{list:oscillation_assign} reports the implementation of the assignment overloading.
+Essentially, to all the data members of the left-hand-side are assigned the values of the corresponding right-hand-side.
+Notably, for the assignment of the state vector and of the previous time steps solution array we take advantage of the automatic re-allocation of the left-hand-side variables when they are not allocated or allocated differently from the right-hand-side, that is a Fortran 2003 feature.
+In spite its simplicity, the assignment overloading is a key-method enabling the usage of FOODiE solver: effectively, the assignment between two \emph{integrand} ADT variables is ubiquitous into the solvers implementations, see equation \ref{eq:RK} for example.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} assignment},label={list:oscillation_assign}]
+subroutine oscillation_assign_oscillation(lhs, rhs)
+class(oscillation), intent(INOUT) :: lhs ! Left hand side.
+class(integrand),   intent(IN)    :: rhs ! Right hand side.
+select type(rhs)
+class is (oscillation)
+  lhs%dims = rhs%dims
+  lhs%steps = rhs%steps
+  if (allocated(rhs%U)) lhs%U = rhs%U
+  if (allocated(rhs%previous)) lhs%previous = rhs%previous
+  lhs%f = rhs%f
+endselect
+return
+endsubroutine oscillation_assign_oscillation
+\end{lstlisting}
+
+\subsubsection{Errors Analysis of the FOODiE Oscillation test results}
+
+We use the Oscillation test program (shipped with FOODiE tests suite) for the testing the above described implementation and assessing the provided solvers accuracy for this IVP.
+
+For the analysis of the accuracy of each solver, we have integrated the Oscillation equations \ref{eq:oscillation} with different, decreasing time steps in the range $[5000, 2500, 1250, 625, 320, 100]$.
 The error is estimated by the L2 norm of the difference between the exact ($U_e$) and the numerical ($U_{\Delta t}$) solutions for each time step:
@@
 \end{equation}
+where $N_s$ is the total number of time steps performed to reach the final integration time.
 
@@
   \caption{Oscillation test errors analysis}\label{tab:oscillation}
-  \resizebox{0.80\textwidth}{!}{%
+  \resizebox{0.75\textwidth}{!}{%
   \begin{tabular}{cccccc}
@@
                  &    100.0        &  0.418E-01    &  0.416E-01    & 1.44          & 1.44          \\
+    \cmidrule{2-6}
+    \multirow{6}{*}{Adams-Bashforth 4 steps}
+                 &   5000.0        &  0.384E+09    &  0.380E+09    & /             & /             \\
+                 &   2500.0        &  0.871E+01    &  0.869E+01    & 25.39         & 25.38         \\
+                 &   1250.0        &  0.314E+01    &  0.312E+01    &  1.47         &  1.48         \\
+                 &    625.0        &  0.111E+01    &  0.111E+01    &  1.49         &  1.49         \\
+                 &    320.0        &  0.409E+00    &  0.407E+00    &  1.50         &  1.50         \\
+                 &    100.0        &  0.714E-01    &  0.711E-01    &  1.50         &  1.50         \\
+
     \cmidrule{2-6}
@@
 
+{\color{red}This class of solvers does not behaves as expected.
+
+The expected observed orders of accuracy should tend to 2, 3 and 4 for the corresponding Adams-Bashforth 2, 3 and 4 steps, but the observed accuracy tend to 1.5 for all solvers.
+Is there a bug?
+Is the Adams-Bashforth not suited for the Oscillation test?
+
+Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes instability generation for some pairs scheme-$\Delta t$: odd-even phase/amplitude phenomena are present.
+A deep insight from whom is familiar with multi-step method is welcome.
+
+Note that in the summary.log file into error analysis directory I have computed also the product of $f*\Delta t$ in order to verify when the instability arises: for AB-2 and AB-4 the solution with $\Delta t=5000 \rightarrow f*\Delta t=0.5$ results to be unstable while for AB-3 it is stable, but the amplitude is dumped resulting into large errors.
+This is consistent with Duran results, where AB-3 is said to be unstable for $ f*\Delta t>0.724$.
+}
+
 \begin{figure}[!ht]
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
-  \caption{Oscillation equations solutions computed by means of Adams-Bashforth 2 steps solver}\label{fig:results-oscillation-adams-bashforth-2}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
-  \caption{Oscillation equations solutions computed by means of Adams-Bashforth 3 steps solver}\label{fig:results-oscillation-adams-bashforth-3}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
+    \caption{Adams-Bashforth 2 steps solution}\label{fig:results-oscillation-adams-bashforth-2}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
+    \caption{Adams-Bashforth 3 steps solution}\label{fig:results-oscillation-adams-bashforth-3}
+  \end{subfigure}\\
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-4.png}
+    \caption{Adams-Bashforth 4 steps solution}\label{fig:results-oscillation-adams-bashforth-4}
+  \end{subfigure}
+  \caption{Oscillation equations solutions computed by means of Adams-Bashforth solvers}\label{fig:results-oscillation-adams-bashforth}
 \end{figure}

### 2015-10-06 src/tests/oscillation/oscillation.f90

@@ -654,30 +654,30 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
 
   !---------------------------------------------------------------------------------------------------------------------------------
-  print "(A)", "Solver & Time Step & Error X & Error Y & Observed Order of Accuracy X & Observed Order of Accuracy Y"
+  print "(A)", "Solver & Time Step & f*Dt & Error X & Error Y & Observed Order of Accuracy X & Observed Order of Accuracy Y"
   do s=1, ab_steps
     if (s==1) cycle ! 1st order scheme surely not stable for this test
     print "(A)", "Adams-Bashforth, "//trim(str(.true.,s))//" steps"
     do d=1, NDt
       if (d==1) then
-        print "(A,F8.1,A,E10.3,A,E10.3,A)", "  & ", time_steps(d), " & ", &
-                                         ab_errors(1, s, d), " & " , ab_errors(2, s, d), " & / & /"
+        print "(A,F8.1,A,F10.3,A,E10.3,A,E10.3,A)", "  & ", time_steps(d), " & ", f*time_steps(d), " & ", &
+                                                 ab_errors(1, s, d), " & " , ab_errors(2, s, d), " & / & /"
       else
-        print "(A,F8.1,A,E10.3,A,E10.3,A,F9.2,A,F9.2)", "  & ", time_steps(d), " & ", &
-                                                        ab_errors(1, s, d), " & " , ab_errors(2, s, d), " & ", &
-                                                        ab_orders(1, s, d-1), " & " , ab_orders(2, s, d-1)
+        print "(A,F8.1,A,F10.3,A,E10.3,A,E10.3,A,F9.2,A,F9.2)", "  & ", time_steps(d), " & ", f*time_steps(d), " & ", &
+                                                            ab_errors(1, s, d), " & " , ab_errors(2, s, d), " & ", &
+                                                            ab_orders(1, s, d-1), " & " , ab_orders(2, s, d-1)
       endif
     enddo
   enddo
   print "(A)", "Leapfrog 2 steps"
   do d=1, NDt
     if (d==1) then
-      print "(A,F8.1,A,E10.3,A,E10.3,A)", "  & ", time_steps(d), " & ", &
-                                          lf_errors(1, d), " & " , lf_errors(2, d), " & / & /"
[…]

### 2015-10-06 src/tests/oscillation/type_oscillation.f90

@@ -48,11 +48,11 @@ type, extends(integrand) :: oscillation
   !<#### State variables organization
   !< State variables are organized as an array (rank 1) of reals of *dims* elements, in this case 2 elements.
   private
-  integer(I_P)                           :: dims=0     !< Space dimensions.
-  integer(I_P)                           :: steps=0    !< Number of time steps stored.
-  real(R_P), dimension(:),   allocatable :: U          !< Integrand (state) variables, [1:dims].
-  real(R_P), dimension(:,:), allocatable :: previous   !< Previous time steps states, [1:dims,1:steps].
-  real(R_P)                              :: f = 0._R_P !< Oscillation frequency (Hz).
+  integer(I_P)                           :: dims=0   !< Space dimensions.
+  real(R_P)                              :: f=0._R_P !< Oscillation frequency (Hz).
+  real(R_P), dimension(:),   allocatable :: U        !< Integrand (state) variables, [1:dims].
+  integer(I_P)                           :: steps=0  !< Number of time steps stored.
+  real(R_P), dimension(:,:), allocatable :: previous !< Previous time steps states, [1:dims,1:steps].
   contains
     ! auxiliary methods
     procedure, pass(self), public :: init   !< Init field.

### 2015-10-06 publish/src/euler-1D.f90
Travis CI autocommit from travis build 56

@@ -338,7 +338,7 @@ contains
   integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(euler_1D)                   :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer                          :: step                  !< Time steps counter.
   real(R_P)                        :: dt                    !< Time step.
   real(R_P)                        :: t(1:ab_steps)         !< Times.
@@ -355,14 +355,16 @@ contains
     print "(A)", ' AB-'//trim(str(.true.,s))
     title = '1D Euler equations integration, explicit Adams-Bashforth, t='//str(n=t_final)//trim(str(.true., s))//' steps'
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
     select case(s)
     case(1)
       call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, steps=s, ord=1)
+      call rk_integrator%init(stages=s)
     case(2, 3)
       call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, steps=s, ord=3)
-    case(5)
+      call rk_integrator%init(stages=s)
+    case(4, 5)
       call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, steps=s, ord=7)
+      call rk_integrator%init(stages=5)
     endselect
     t = 0._R_P
     call save_time_serie(title=title, filename=output//'-'//trim(str(.true., s))//'-time_serie.dat', t=t(s))
@@ -372,14 +374,14 @@ contains
       dt = domain%dt(Nmax=0, Tmax=t_final, t=t(s), CFL=0.1_R_P*CFL)
       if (s>=step) then
         ! the time steps from 1 to s - 1 must be computed with other scheme...
[…]

### 2015-10-06 publish/src/foodie_integrator_adams_bashforth.f90

@@ -10,14 +10,16 @@ module foodie_integrator_adams_bashforth
 !< where \(U_t = \frac{dU}{dt}\), *U* is the vector of *state* variables being a function of the time-like independent variable
 !< *t*, *R* is the (vectorial) residual function, the Adams-Bashforth class scheme implemented is:
 !<
-!< $$ U^{n+s} = U^{n+s-1} +\Delta t \left[ \sum_{i=1}^{n+s}{ b_i \cdot R(t^{n+i-1}, U^{n+i-1}) } \right] $$
+!< $$ U^{n+N_s} = U^{n+N_s-1} +\Delta t \left[ \sum_{s=1}^{n+N_s}{ b_s \cdot R(t^{n+s-1}, U^{n+s-1}) } \right] $$
+!<
+!<where \(N_s\) is the number of previous steps considered.
 !<
 !< @note The value of \(\Delta t\) must be provided, it not being computed by the integrator.
 !<
-!< The schemes are explicit. The coefficients \(\b\) define the actual scheme, that is selected accordingly to the number of
+!< The schemes are explicit. The coefficients \(b_s\) define the actual scheme, that is selected accordingly to the number of
 !< **steps** used.
 !<
-!< Currently the following schemes are available:
+!< Currently, the following schemes are available:
 !<##### 1 step, Explicit Forward Euler, 1st order
 !< This scheme is TVD and reverts to Explicit Forward Euler, it being 1st order.
 !< The *b* coefficient is:
@@ -42,12 +44,21 @@ module foodie_integrator_adams_bashforth
 !< $$ U^{n+3} = U^{n+2} +\Delta t \left[ \frac{23}{12}R(t^{n+2}, U^{n+2}) - \frac{4}{3}R(t^{n+1}, U^{n+1})
 !< +\frac{5}{12} R(t^{n}, U^{n})  \right] $$
 !<
+!<##### 4 steps
+!< This scheme is 4th order.
+!< The *b* coefficients are:
+!< $$b = \left[ {\begin{array}{*{20}{c}} b_1 & b_2 & b_3 & b_4 \end{array}} \right] =
+!<       \left[ {\begin{array}{*{20}{c}} -\frac{9}{24} & \frac{37}{24} & -\frac{59}{24} & \frac{55}{24} \end{array}} \right]$$
+!< The scheme is:
[…]

### 2015-10-06 publish/src/foodie_integrator_tvd_runge_kutta.f90

@@ -86,7 +86,7 @@ module foodie_integrator_tvd_runge_kutta
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use IR_Precision, only : R_P, I_P
+use foodie_kinds, only : R_P, I_P
 use type_integrand, only : integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
 
@@ -190,14 +190,14 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine destroy
 
-  subroutine integrate(self, field, stage, Dt, t)
+  subroutine integrate(self, U, stage, Dt, t)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with explicit TVD (or SSP) Runge-Kutta scheme.
   !<
   !< @note This method can be used **after** the integrator is created (i.e. the RK coeficients are initialized).
   !---------------------------------------------------------------------------------------------------------------------------------
   class(tvd_runge_kutta_integrator), intent(IN)    :: self      !< Actual RK integrator.
-  class(integrand),                  intent(INOUT) :: field     !< Field to be integrated.
+  class(integrand),                  intent(INOUT) :: U         !< Field to be integrated.
   class(integrand),                  intent(INOUT) :: stage(1:) !< Runge-Kutta stages [1:stages].
   real(R_P),                         intent(IN)    :: Dt        !< Time step.
   real(R_P),                         intent(IN)    :: t         !< Time.
@@ -210,7 +210,7 @@ contains
   class is(integrand)
     ! computing stages
     do s=1, self%stages
-      stage(s) = field
+      stage(s) = U
       do ss=1, s - 1
         stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
       enddo
@@ -218,7 +218,7 @@ contains
     enddo
     ! computing new time step
     do s=1, self%stages
-      field = field +  stage(s) * (Dt * self%beta(s))
+      U = U +  stage(s) * (Dt * self%beta(s))
     enddo
   endselect
   return

### 2015-10-06 publish/src/lorenz.f90

@@ -130,7 +130,7 @@ contains
   integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(lorenz)                     :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer(I_P)                     :: s                     !< AB steps counter.
   integer                          :: step                  !< Time steps counter.
   !---------------------------------------------------------------------------------------------------------------------------------
@@ -140,16 +140,21 @@ contains
   do s=1, ab_steps
     print "(A)", ' AB-'//trim(str(.true.,s))
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
+    select case(s)
+    case(1, 2, 3)
+      call rk_integrator%init(stages=s)
+    case(4)
+      call rk_integrator%init(stages=5)
+    endselect
     call attractor%init(initial_state=initial_state, sigma=sigma, rho=rho, beta=beta, steps=s)
     solution(0, 0) = 0._R_P
     solution(1:space_dimension, 0) = attractor%output()
     do step = 1, num_steps
       if (s>=step) then
         ! the time steps from 1 to s - 1 must be computed with other scheme...
-        call rk_integrator%integrate(field=attractor, stage=rk_stage(1:s), dt=dt, t=solution(0, step))
+        call rk_integrator%integrate(U=attractor, stage=rk_stage, dt=dt, t=solution(0, step))
       else
-        call ab_integrator%integrate(field=attractor, dt=dt, t=solution(0, step-s:step-1))
+        call ab_integrator%integrate(U=attractor, dt=dt, t=solution(0, step-s:step-1))
       endif
       solution(0, step) = step * dt
       solution(1:space_dimension, step) = attractor%output()
@@ -176,7 +181,7 @@ contains
   solution(0, 0) = 0._R_P
   solution(1:space_dimension, 0) = attractor%output()
   do step = 1, num_steps
-    call euler_integrator%integrate(field=attractor, dt=dt, t=solution(0, step))
+    call euler_integrator%integrate(U=attractor, dt=dt, t=solution(0, step))
[…]

### 2015-10-06 publish/src/type_integrand.f90

@@ -5,7 +5,7 @@ module type_integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use IR_Precision, only : R_P, I_P
+use foodie_kinds, only : R_P, I_P
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -29,14 +29,13 @@ type, abstract :: integrand
     procedure(symmetric_operator),   pass(lhs),  deferred, public :: add                          !< Integrand + integrand oprator.
     procedure(symmetric_operator),   pass(lhs),  deferred, public :: sub                          !< Integrand - integrand oprator.
     procedure(assignment_integrand), pass(lhs),  deferred, public :: assign_integrand             !< Integrand = integrand.
-    procedure(assignment_real),      pass(lhs),  deferred, public :: assign_real                  !< Integrand = real.
     ! operators overloading
     generic, public :: operator(+) => add                              !< Overloading + operator.
     generic, public :: operator(-) => sub                              !< Overloading - operator.
     generic, public :: operator(*) => integrand_multiply_integrand, &
                                       real_multiply_integrand, &
                                       integrand_multiply_real          !< Overloading * operator.
-    generic, public :: assignment(=) => assign_integrand, assign_real  !< Overloading = assignament.
+    generic, public :: assignment(=) => assign_integrand               !< Overloading = assignament.
 endtype integrand
 
 abstract interface
@@ -117,16 +116,6 @@ abstract interface
   class(integrand), intent(IN)    :: rhs !< Right hand side.
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine assignment_integrand
-
-  subroutine assignment_real(lhs, rhs)
-  !---------------------------------------------------------------------------------------------------------------------------------
-  !< Asymmetric assignment integrand = real.
-  !---------------------------------------------------------------------------------------------------------------------------------
-  import :: integrand, R_P
-  class(integrand), intent(INOUT) :: lhs !< Left hand side.
-  real(R_P),        intent(IN)    :: rhs !< Right hand side.
-  !---------------------------------------------------------------------------------------------------------------------------------
-  endsubroutine assignment_real
 endinterface
 !-----------------------------------------------------------------------------------------------------------------------------------
 endmodule type_integrand

### 2015-10-06 src/burgers.f90

@@ -141,7 +141,7 @@ contains
   integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(burgers)                    :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer                          :: step                  !< Time steps counter.
   real(R_P)                        :: dt                    !< Time step.
   real(R_P)                        :: t(1:ab_steps)         !< Times.
@@ -154,7 +154,12 @@ contains
   do s=1, ab_steps
     print "(A)", ' AB-'//trim(str(.true.,s))
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
+    select case(s)
+    case(1, 2, 3)
+      call rk_integrator%init(stages=s)
+    case(4)
+      call rk_integrator%init(stages=5)
+    endselect
     call domain%init(initial_state=initial_state, Ni=Ni, h=h, nu=nu, steps=s)
     dt = domain%dt(CFL=CFL)
     t = 0._R_P
@@ -162,14 +167,14 @@ contains
     do while(t(s)<t_final)
       if (s>=step) then
         ! the time steps from 1 to s - 1 must be computed with other scheme...
-        call rk_integrator%integrate(field=domain, stage=rk_stage(1:s), dt=dt, t=t(s))
+        call rk_integrator%integrate(U=domain, stage=rk_stage, dt=dt, t=t(s))
         if (step>1) then
           t(step) = t(step-1) + dt
         else
           t(step) = dt
         endif
       else
-        call ab_integrator%integrate(field=domain, dt=dt, t=t)
+        call ab_integrator%integrate(U=domain, dt=dt, t=t)
         do ss=1, s-1
           t(ss) = t(ss + 1)
         enddo
@@ -202,7 +207,7 @@ contains
   dt = domain%dt(CFL=CFL)
   t = 0._R_P
   do while(t<t_final)
-    call euler_integrator%integrate(field=domain, dt=dt, t=t)
+    call euler_integrator%integrate(U=domain, dt=dt, t=t)
     t = t + dt
   enddo
[…]

### 2015-10-06 src/foodie_integrator_euler_explicit.f90

@@ -16,7 +16,7 @@ module foodie_integrator_euler_explicit
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use IR_Precision, only : R_P
+use foodie_kinds, only : R_P
 use type_integrand, only : integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
 
@@ -36,17 +36,17 @@ type :: euler_explicit_integrator
 endtype euler_explicit_integrator
 !-----------------------------------------------------------------------------------------------------------------------------------
 contains
-  subroutine integrate(field, Dt, t)
+  subroutine integrate(U, Dt, t)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with explicit Euler scheme, 1st order.
   !---------------------------------------------------------------------------------------------------------------------------------
-  class(integrand), intent(INOUT) :: field !< Field to be integrated.
-  real(R_P),        intent(in)    :: Dt    !< Time step.
-  real(R_P),        intent(IN)    :: t     !< Time.
+  class(integrand),    intent(INOUT) :: U  !< Field to be integrated.
+  real(R_P),           intent(IN)    :: Dt !< Time step.
+  real(R_P), optional, intent(IN)    :: t  !< Time.
   !---------------------------------------------------------------------------------------------------------------------------------
 
   !---------------------------------------------------------------------------------------------------------------------------------
-  field = field + field%t(t=t) * Dt
+  U = U + U%t(t=t) * Dt
   return
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine integrate

### 2015-10-06 src/foodie_integrator_leapfrog.f90

@@ -37,7 +37,7 @@ module foodie_integrator_leapfrog
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use IR_Precision, only : R_P, I_P
+use foodie_kinds, only : R_P, I_P
 use type_integrand, only : integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
 
@@ -81,23 +81,23 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine init
 
-  subroutine integrate(self, field, filter, Dt, t)
+  subroutine integrate(self, U, filter, Dt, t)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with leapfrog class scheme.
   !---------------------------------------------------------------------------------------------------------------------------------
   class(leapfrog_integrator), intent(IN)    :: self   !< LF integrator.
-  class(integrand),           intent(INOUT) :: field  !< Field to be integrated.
+  class(integrand),           intent(INOUT) :: U      !< Field to be integrated.
   class(integrand), optional, intent(INOUT) :: filter !< Filter field displacement.
   real(R_P),                  intent(in)    :: Dt     !< Time step.
   real(R_P),                  intent(IN)    :: t      !< Time.
   !---------------------------------------------------------------------------------------------------------------------------------
 
   !---------------------------------------------------------------------------------------------------------------------------------
-  field = field%previous_step(n=1) + field%t(n=2, t=t) * (Dt * 2._R_P)
+  U = U%previous_step(n=1) + U%t(n=2, t=t) * (Dt * 2._R_P)
   if (present(filter)) then
-    filter = (field%previous_step(n=1) - field%previous_step(n=2) * 2._R_P + field) * self%nu * 0.5_R_P
+    filter = (U%previous_step(n=1) - U%previous_step(n=2) * 2._R_P + U) * self%nu * 0.5_R_P
   endif
-  call field%update_previous_steps(filter=filter, weights=[self%alpha, self%alpha - 1._R_P])
+  call U%update_previous_steps(filter=filter, weights=[self%alpha, self%alpha - 1._R_P])
   return
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine integrate

### 2015-10-06 src/foodie_integrator_low_storage_runge_kutta.f90

@@ -72,7 +72,7 @@ module foodie_integrator_low_storage_runge_kutta
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use IR_Precision, only : R_P, I_P, I8P
+use foodie_kinds, only : R_P, I_P, I8P
 use type_integrand, only : integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
 
@@ -158,14 +158,14 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine destroy
 
-  subroutine integrate(self, field, stage, Dt, t)
+  subroutine integrate(self, U, stage, Dt, t)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with explicit low storage Runge-Kutta scheme.
   !<
   !< @note This method can be used **after** the integrator is created (i.e. the RK coeficients are initialized).
   !---------------------------------------------------------------------------------------------------------------------------------
   class(ls_runge_kutta_integrator), intent(IN)    :: self       !< Actual RK integrator.
-  class(integrand),                 intent(INOUT) :: field      !< Field to be integrated.
+  class(integrand),                 intent(INOUT) :: U          !< Field to be integrated.
   class(integrand),                 intent(INOUT) :: stage(1:2) !< Runge-Kutta registers [1:2].
   real(R_P),                        intent(IN)    :: Dt         !< Time step.
   real(R_P),                        intent(IN)    :: t          !< Time.
@@ -176,13 +176,13 @@ contains
   select type(stage)
   class is(integrand)
     ! computing stages
-    stage(1) = field
-    stage(2) = field*0._R_P
+    stage(1) = U
+    stage(2) = U*0._R_P
     do s=1, self%stages
       stage(2) = stage(2) * self%A(s) + stage(1)%t(t=t + self%C(s) * Dt) * Dt
       stage(1) = stage(1) + stage(2) * self%B(s)
     enddo
-    field = stage(1)
+    U = stage(1)
   endselect
   return
   !---------------------------------------------------------------------------------------------------------------------------------

### 2015-10-06 src/foodie_kinds.f90

@@ -1,0 +1,30 @@
+!< FOODiE kinds: definition of reals and integer kind parameters of FOODiE library.
+module foodie_kinds
+!-----------------------------------------------------------------------------------------------------------------------------------
+!< FOODiE kinds: definition of reals and integer kind parameters of FOODiE library.
+!-----------------------------------------------------------------------------------------------------------------------------------
+
+!-----------------------------------------------------------------------------------------------------------------------------------
+implicit none
+private
+public:: R8P
+public:: R4P
+public:: R_P
+public:: I8P
+public:: I4P
+public:: I2P
+public:: I1P
+public:: I_P
+!-----------------------------------------------------------------------------------------------------------------------------------
+
+!-----------------------------------------------------------------------------------------------------------------------------------
+integer, parameter:: R8P  = selected_real_kind(15,307)  !< 15  digits, range \([10^{-307} , 10^{+307}  - 1]\); 64 bits.
+integer, parameter:: R4P  = selected_real_kind(6,37)    !< 6   digits, range \([10^{-37}  , 10^{+37}   - 1]\); 32 bits.
+integer, parameter:: R_P  = R8P                         !< Default real precision.
+integer, parameter:: I8P  = selected_int_kind(18)       !< Range \([-2^{63},+2^{63} - 1]\), 19 digits plus sign; 64 bits.
+integer, parameter:: I4P  = selected_int_kind(9)        !< Range \([-2^{31},+2^{31} - 1]\), 10 digits plus sign; 32 bits.
+integer, parameter:: I2P  = selected_int_kind(4)        !< Range \([-2^{15},+2^{15} - 1]\), 5  digits plus sign; 16 bits.
+integer, parameter:: I1P  = selected_int_kind(2)        !< Range \([-2^{7} ,+2^{7}  - 1]\), 3  digits plus sign; 8  bits.
+integer, parameter:: I_P  = I4P                         !< Default integer precision.
+!-----------------------------------------------------------------------------------------------------------------------------------
+endmodule foodie_kinds

### 2015-10-06 src/oscillation.f90

@@ -19,7 +19,7 @@ use pyplot_module, only :  pyplot
 !-----------------------------------------------------------------------------------------------------------------------------------
 implicit none
 type(Type_Command_Line_Interface) :: cli                                               !< Command line interface handler.
-type(oscillation)                 :: attractor                                         !< Oscillation field.
+type(oscillation)                 :: oscillator                                        !< Oscillation field.
 integer,   parameter              :: space_dimension=2                                 !< Space dimensions.
 real(R_P), parameter              :: initial_state(1:space_dimension)=[0._R_P, 1._R_P] !< Initial state.
 real(R_P)                         :: f                                                 !< Oscillation frequency.
@@ -187,10 +187,10 @@ contains
   !< Test explicit Adams-Bashforth class of ODE solvers.
   !---------------------------------------------------------------------------------------------------------------------------------
   type(tvd_runge_kutta_integrator) :: rk_integrator         !< Runge-Kutta integrator.
-  integer, parameter               :: rk_stages=3           !< Runge-Kutta stages number.
+  integer, parameter               :: rk_stages=5           !< Runge-Kutta stages number.
   type(oscillation)                :: rk_stage(1:rk_stages) !< Runge-Kutta stages.
   type(adams_bashforth_integrator) :: ab_integrator         !< Adams-Bashforth integrator.
-  integer, parameter               :: ab_steps=3            !< Adams-Bashforth steps number.
+  integer, parameter               :: ab_steps=4            !< Adams-Bashforth steps number.
   integer                          :: step                  !< Time steps counter.
   integer(I_P)                     :: s                     !< AB steps counter.
   integer(I_P)                     :: steps_range(1:2)      !< Steps used.
@@ -204,21 +204,26 @@ contains
     print "(A)", ' AB-'//trim(str(.true.,s))
     title = 'Oscillation equation integration, explicit Adams-Bashforth, t='//str(n=t_final)//' steps='//trim(str(.true., s))
     call ab_integrator%init(steps=s)
-    call rk_integrator%init(stages=s)
-    call attractor%init(initial_state=initial_state, f=f, steps=s)
[…]

### 2015-10-07 papers/announcement/manuscript.tex
minor improvements into manuscript

@@
 
+\paragraph{FOODiE numerical integration}
+
+Using the above discussed \emph{oscillation} type it is very easy to solve IVP \ref{eq:oscillation} by means of FOODiE library.
+Listing \ref{list:oscillation_leapfrog} presents the numerical integration of system \ref{eq:oscillation} by means of the Leapfrog RAW-filtered method.
+In the example, the integration is performed with $10^4$ steps with a fixed $\Delta t=10^2$ until the time $t=10^6$ is reached.
+The example shows also that for starting a multi-step scheme such as the Leapfrog one a lower-oder or equivalent order one-scheme is necessary: in the example the first 2 steps are computed by means of one-step TVD/SSP Runge-Kutta 2-stages schemes.
+Note that the memory registers for storing the Runge-Kutta stages and the RAW filter displacement must be handled by the client code.
+Listing \ref{list:oscillation_leapfrog} demonstrates how it is simple, clear and concise to solve a IVP by FOODiE solvers.
+Moreover, it proves how it is simple and effective to apply different solvers in a coupled algorithm, that greatly simplify the development of new hybrid solvers for self-adaptive time step size.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={numerical integration of the \emph{oscillation} system by means of Leapfrog RAW-filtered method},label={list:oscillation_leapfrog}]
+use foodie, only: leapfrog_integrator, tvd_runge_kutta_integrator
+type(leapfrog_integrator)        :: lf_integrator ! Leapfrog integrator.
+type(tvd_runge_kutta_integrator) :: rk_integrator ! Runge-Kutta integrator.
+type(oscillation)                :: rk_stage(1:2) ! Runge-Kutta stages.
+type(oscillation)                :: oscillator    ! Oscillation field.
+type(oscillation)                :: filter        ! Filter displacement.
+integer                          :: step          ! Time steps counter.
+real                             :: Dt            ! Time step.
+call lf_integrator%init()
+call rk_integrator%init(stages=2)
+call oscillator%init(initial_state=[0.0,1.0], f=10e4, steps=2)
+Dt = 100.0
+do step=1, 10000
+  if (2>=step) then
+    call rk_integrator%integrate(U=oscillator, stage=rk_stage, Dt=Dt, t=step*Dt)
+  else
+    call lf_integrator%integrate(U=oscillator, filter=filter, Dt=Dt, t=step*Dt)
+  endif
+enddo
+call print_results(U=oscillator)
+\end{lstlisting}
+
+\paragraph{FOODiE integration}
 \subsubsection{Errors Analysis of the FOODiE Oscillation test results}
@@
     \cmidrule{2-6}
-    \multirow{6}{*}{leapfrog 2 steps}
+    \multirow{6}{*}{Leapfrog RAW-filtered}
                  &   5000.0        &  0.151E+02    &  0.152E+02    & /             & /             \\
@@
 
-Note that in the summary.log file into error analysis directory I have computed also the product of $f*\Delta t$ in order to verify when the instability arises: for AB-2 and AB-4 the solution with $\Delta t=5000 \rightarrow f*\Delta t=0.5$ results to be unstable while for AB-3 it is stable, but the amplitude is dumped resulting into large errors.
+Note that in the summary.log file into error analysis directory I have computed also the product of $f*\Delta t$ in order to verify when the instability arises.
+
+For AB-2 and AB-4 the solution with $\Delta t=5000 \rightarrow f*\Delta t=0.5$ results to be unstable (the computational modes are not dumped) while for AB-3 it is stable, but the (physical mode) amplitude is dumped resulting into large errors.
 This is consistent with Duran results, where AB-3 is said to be unstable for $ f*\Delta t>0.724$.
+
+The AB-2 and AB-4 solutions show a phase error that decreases as the time resolution increases. The AB-3 has no phase error and the computational modes are dumped for all $\Delta t$ exercised.
 }
@@
   \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
-  \caption{Oscillation equations solutions computed by means of leapfrog 2 steps solver}\label{fig:results-oscillation-leapfrog}
+  \caption{Oscillation equations solutions computed by means of Leapfrog RAW-filtered solver}\label{fig:results-oscillation-leapfrog}
 \end{figure}

### 2015-10-07 src/lib/foodie_integrator_leapfrog.f90

@@ -17,9 +17,10 @@ module foodie_integrator_leapfrog
 !< $$ U^{n+1} = U^{n+1} + \Delta * \alpha $$
 !< $$ U^{n+2} = U^{n+2} + \Delta * (\alpha-1) $$
 !< Note that for \(\alpha=1\) the filter reverts back to the standard Robert-Asselin scheme.
-!< The filter coefficients should be taken as \(\nu \in [0,1]\) and \(\alpha \in [0.5,1]\). The default values are
-!<+ \(\nu=0.01)\)
-!<+ \(\alpha=0.5)\)
+!< The filter coefficients should be taken as \(\nu \in (0,1]\) and \(\alpha \in (0.5,1]\). The default values are
+!<
+!<  + \(\nu=0.01\)
+!<  + \(\alpha=0.53\)
 !<
 !< @note The value of \(\Delta t\) must be provided, it not being computed by the integrator.
 !<
@@ -54,8 +55,8 @@ type :: leapfrog_integrator
   !< @note The integrator could be used without initialialization (initialize the time filter coefficients) if the defulat values
   !< are suitable for the problem.
   private
-  real(R_P) :: nu=0.01_R_P   !< Robert-Asselin filter coefficient.
-  real(R_P) :: alpha=0.5_R_P !< Robert-Asselin-Williams filter coefficient.
+  real(R_P) :: nu=0.01_R_P    !< Robert-Asselin filter coefficient.
+  real(R_P) :: alpha=0.53_R_P !< Robert-Asselin-Williams filter coefficient.
   contains
     procedure, pass(self), public :: init      !< Initialize (create) the integrator.
     procedure, pass(self), public :: integrate !< Integrate integrand field.
@@ -74,7 +75,7 @@ contains
 
   !---------------------------------------------------------------------------------------------------------------------------------
   self%nu = 0.01_R_P
-  self%alpha = 0.5_R_P
+  self%alpha = 0.53_R_P
   if (present(nu)) self%nu = nu
   if (present(alpha)) self%alpha = alpha
   return

### 2015-10-07 papers/announcement/manuscript.tex

@@
     procedure(time_derivative),       pass(self), deferred, public:: t
-    procedure(update_previous_steps), pass(self), deferred, public:: update_previous_steps
-    procedure(previous_step),         pass(self), deferred, public:: previous_step
     ! operators
@@
 
-As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 3 ODE-specific procedures plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects.
+As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 1 ODE-specific procedure plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects.
 These procedures are analyzed in the following paragraphs.
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={time derivative procedure interface},label={list:time_derivative}]
-function time_derivative(self, n, t) result(dState_dt)
+function time_derivative(self, t) result(dState_dt)
 import :: integrand, R_P, I_P
 class(integrand),       intent(IN) :: self      !< Integrand field.
-integer(I_P), optional, intent(IN) :: n         !< Time level.
 real(R_P),    optional, intent(IN) :: t         !< Time.
@@
 
-This procedure-function takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optionals, and it returns an integrand object.
+This procedure-function takes two arguments, the first passed as a \emph{type bounded} argument, while the latter is optional, and it returns an integrand object.
 The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
-The optional argument \emph{n} indicates the \emph{time level} at which the residuals function must be evaluated: this argument is necessary for multi-step (often referred as multi-level) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family.
 The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.
@@
 {\color{red} Add citations to Burgers, Adams-Bashfort, leapfrog references.}
-
-\subsubsection{Update previous time steps procedure}
-
-The abstract interface of the \emph{update previous time steps} procedure is shown in listing \ref{list:update_previous_time_steps}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={update previous time steps procedure interface},label={list:update_previous_time_steps}]
-subroutine update_previous_steps(self, filter, weights)
-import :: integrand, R_P
-class(integrand),           intent(INOUT) :: self       !< Integrand field.
-class(integrand), optional, intent(IN)    :: filter     !< Filter field displacement.
-real(R_P),        optional, intent(IN)    :: weights(:) !< Weights for filtering the steps.
-endsubroutine update_previous_steps
-\end{lstlisting}
-
-This procedure is a necessary complement for using multi-step solvers.
-As its name suggests, it updates the previous time steps solutions (that must be stored on the care of the clients code).
-It takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optionals, and it returns nothing it being a subroutine directly modifying the passed argument.
-The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
-The optional argument \emph{filter} indicates a \emph{displacement} for filtering the field while the last optional argument \emph{weights} specifies the weights that must applied to the displacement for each time steps filtered.
-
-Let us consider the multi-step scheme expressed by the equation \ref{eq:AB}.
-Once the solution at time $t^{n+N_s}$ has been computed the previous $N_s-1$ steps must be updated, e.g.
-an unfiltered update can read as:
-\begin{equation}
-U\left(t^{n+s}\right) = U\left(t^{n+s+1}\right)\quad s=1,..,N_s-1
-\label{eq:previous-steps-update}
-\end{equation}
-and a filtered update can read as:
-\begin{equation}
-U\left(t^{n+s}\right) = \Delta * U\left(t^{n+s+1}\right) \omega_s\quad s=1,..,N_s-1
-\label{eq:previous-steps-update-filtered}
-\end{equation}
-where $\Delta$ is the filter displacement and $\omega_s$ the weight of the step $s-th$.
-This interface implies that the previous time steps must be embedded into the client \emph{integrand} extension.
-This procedure is automatically called by each multi-step solver after the new step $t^{n+N_s}$ has been computed, thus alleviating client codes from its explicit call.
-
-In the case the client codes will use only one-step solvers this procedure can be defined as simple \emph{empty} subroutine as shown in listing \ref{list:empty_update_previous_time_steps}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={empty implementation of the update previous time steps procedure},label={list:empty_update_previous_time_steps}]
-subroutine update_previous_steps(self, filter, weights)
-class(my_integrand),        intent(INOUT) :: self       !< Integrand field.
-class(integrand), optional, intent(IN)    :: filter     !< Filter field displacement.
-real(R_P),        optional, intent(IN)    :: weights(:) !< Weights for filtering the steps.
-! do nothing
-return
-endsubroutine update_previous_steps
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
-
-\subsubsection{Previous step procedure}
-
-The abstract interface of the \emph{previous step} procedure is shown in listing \ref{list:previous_step}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={previous step procedure interface},label={list:previous_step}]
-function previous_step(self, n) result(previous)
-import :: integrand, I_P
-class(integrand), intent(IN)  :: self     !< Integrand field.
-integer(I_P),     intent(IN)  :: n        !< Time level.
-class(integrand), allocatable :: previous !< Selected previous time integrand field.
-endfunction previous_step
-\end{lstlisting}
-
-This procedure is a necessary complement for using multi-step solvers.
-It returns the integrand solution at the specified previous time step (level).
-It takes two arguments and it returns an integrand object.
-The first argument passed as a \emph{type bounded} argument is the current instance of the integrand field object.
-The second argument, named \emph{n}, indicates the \emph{time level} at which the solution must be returned.
-
-This procedure is used in multi-step methods like the leapfrog where there is an explicit dependency on also the solution at previous time steps.
-
-For the case in which multi-step solvers are not used, this procedure can have a concrete implementation that does nothing as shown in listing \ref{list:empty_previous_step}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={empty implementation of the previous step procedure},label={list:empty_previous_step}]
-function previous_step(self, n) result(previous)
-class(my_integrand), intent(IN)  :: self     !< Integrand field.
-integer(I_P),        intent(IN)  :: n        !< Time level.
-class(integrand), allocatable    :: previous !< Selected previous time integrand field.
-! do nothing
-return
-endfunction previous_step
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT. Note that, in this case, the returned result is not allocated.
 
@@
 type(my_integrand) :: my_field
-call integrator%integrate(U=my_field, Dt=Dt, t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT and \emph{times} are the time at each 4 steps considered for the current one-step-over integration.
+type(my_integrand) :: previous(1:4)
+call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.
 
@@
 
+{\color{red} To be modified with changed API}
+
 \begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
-subroutine integrate(self, U, Dt, t)
+subroutine integrate(self, U, previous, Dt, t)
 class(adams_bashforth_integrator), intent(IN)    :: self ! Actual AB integrator.
@@
     \multirow{6}{*}{Adams-Bashforth 2 steps}
-                 &   5000.0        &  0.109E+04    &  0.112E+04    & /             & /             \\
-                 &   2500.0        &  0.246E+02    &  0.243E+02    & 5.47          & 5.53          \\
-                 &   1250.0        &  0.783E+01    &  0.789E+01    & 1.65          & 1.62          \\
-                 &    625.0        &  0.268E+01    &  0.272E+01    & 1.55          & 1.54          \\
-                 &    320.0        &  0.976E+00    &  0.990E+00    & 1.51          & 1.51          \\
-                 &    100.0        &  0.171E+00    &  0.173E+00    & 1.50          & 1.50          \\
+                 &   5000.0        &  0.596E+03    &  0.583E+03    & /             & /             \\
+                 &   2500.0        &  0.221E+02    &  0.218E+02    & 4.75          & 4.74          \\
+                 &   1250.0        &  0.764E+01    &  0.769E+01    & 1.53          & 1.50          \\
+                 &    625.0        &  0.265E+01    &  0.268E+01    & 1.53          & 1.52          \\
+                 &    320.0        &  0.968E+00    &  0.981E+00    & 1.51          & 1.50          \\
+                 &    100.0        &  0.169E+00    &  0.171E+00    & 1.50          & 1.50          \\
     \cmidrule{2-6}
     \multirow{6}{*}{Adams-Bashforth 3 steps}
-                 &   5000.0        &  0.816E+01    &  0.809E+01    & /             & /             \\
-                 &   2500.0        &  0.266E+01    &  0.264E+01    & 1.62          & 1.62          \\
-                 &   1250.0        &  0.120E+01    &  0.120E+01    & 1.15          & 1.14          \\
-                 &    625.0        &  0.545E+00    &  0.543E+00    & 1.14          & 1.14          \\
-                 &    320.0        &  0.223E+00    &  0.222E+00    & 1.34          & 1.34          \\
-                 &    100.0        &  0.418E-01    &  0.416E-01    & 1.44          & 1.44          \\
+                 &   5000.0        &  0.857E+01    &  0.854E+01    & /             & /             \\
+                 &   2500.0        &  0.391E+01    &  0.386E+01    & 1.13          & 1.14          \\
+                 &   1250.0        &  0.825E+00    &  0.814E+00    & 2.24          & 2.25          \\
+                 &    625.0        &  0.150E+00    &  0.148E+00    & 2.46          & 2.46          \\
+                 &    320.0        &  0.282E-01    &  0.278E-01    & 2.49          & 2.49          \\
+                 &    100.0        &  0.154E-02    &  0.152E-02    & 2.50          & 2.50          \\
     \cmidrule{2-6}
     \multirow{6}{*}{Adams-Bashforth 4 steps}
-                 &   5000.0        &  0.384E+09    &  0.380E+09    & /             & /             \\
-                 &   2500.0        &  0.871E+01    &  0.869E+01    & 25.39         & 25.38         \\
-                 &   1250.0        &  0.314E+01    &  0.312E+01    &  1.47         &  1.48         \\
-                 &    625.0        &  0.111E+01    &  0.111E+01    &  1.49         &  1.49         \\
-                 &    320.0        &  0.409E+00    &  0.407E+00    &  1.50         &  1.50         \\
-                 &    100.0        &  0.714E-01    &  0.711E-01    &  1.50         &  1.50         \\
+                 &   5000.0        &  0.128E+07    &  0.143E+07    & /             & /             \\
+                 &   2500.0        &  0.106E+01    &  0.107E+01    & 20.21         & 20.34         \\
+                 &   1250.0        &  0.967E-01    &  0.981E-01    &  3.45         &  3.45         \\
+                 &    625.0        &  0.859E-02    &  0.871E-02    &  3.49         &  3.49         \\
+                 &    320.0        &  0.827E-03    &  0.838E-03    &  3.50         &  3.50         \\
+                 &    100.0        &  0.141E-04    &  0.143E-04    &  3.50         &  3.50         \\
 
@@
     \multirow{6}{*}{Leapfrog RAW-filtered}
-                 &   5000.0        &  0.151E+02    &  0.152E+02    & /             & /             \\
-                 &   2500.0        &  0.500E+01    &  0.504E+01    & 1.60          & 1.59          \\
-                 &   1250.0        &  0.190E+01    &  0.193E+01    & 1.40          & 1.39          \\
-                 &    625.0        &  0.181E+01    &  0.181E+01    & 0.07          & 0.09          \\
-                 &    320.0        &  0.157E+01    &  0.157E+01    & 0.21          & 0.21          \\
-                 &    100.0        &  0.100E+01    &  0.100E+01    & 0.39          & 0.39          \\
+                 &   5000.0        &  0.156E+02    &  0.156E+02    & /             & /             \\
+                 &   2500.0        &  0.855E+01    &  0.852E+01    & 0.86          & 0.87          \\
+                 &   1250.0        &  0.303E+01    &  0.305E+01    & 1.50          & 1.48          \\
+                 &    625.0        &  0.107E+01    &  0.108E+01    & 1.51          & 1.50          \\
+                 &    320.0        &  0.390E+00    &  0.395E+00    & 1.50          & 1.50          \\
+                 &    100.0        &  0.685E-01    &  0.692E-01    & 1.50          & 1.50          \\
     \cmidrule{2-6}
@@
 
-{\color{red}This class of solvers does not behaves as expected.
-
-The expected observed orders of accuracy should tend to 2, 3 and 4 for the corresponding Adams-Bashforth 2, 3 and 4 steps, but the observed accuracy tend to 1.5 for all solvers.
-Is there a bug?
-Is the Adams-Bashforth not suited for the Oscillation test?
-
-Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes instability generation for some pairs scheme-$\Delta t$: odd-even phase/amplitude phenomena are present.
-A deep insight from whom is familiar with multi-step method is welcome.
-
-Note that in the summary.log file into error analysis directory I have computed also the product of $f*\Delta t$ in order to verify when the instability arises.
-
-For AB-2 and AB-4 the solution with $\Delta t=5000 \rightarrow f*\Delta t=0.5$ results to be unstable (the computational modes are not dumped) while for AB-3 it is stable, but the (physical mode) amplitude is dumped resulting into large errors.
-This is consistent with Duran results, where AB-3 is said to be unstable for $ f*\Delta t>0.724$.
-
-The AB-2 and AB-4 solutions show a phase error that decreases as the time resolution increases. The AB-3 has no phase error and the computational modes are dumped for all $\Delta t$ exercised.
-}
+The expected observed orders of accuracy for the Adams-Bashforth solvers using 2, 3 and 4 time steps tend to 1.5, 2.5 and 3.5 that are in agreement with the expected formal order of 2, 3 and 4, respectively.
+
+Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes the instability generation for some pairs steps number/$\Delta t$.
+The 2 and 4 steps solutions are instable for $\Delta t=5000 \rightarrow f*\Delta t=0.5$.
+On the contrary, the 3 steps solution is stable, but the amplitude is dumped and the solution vanishes as the integration proceeds.
+The 2 and 4 steps solutions show a phase error that decreases as the time resolution increases, whereas 3 steps solution has null phase error.
 

### 2015-10-08 papers/announcement/manuscript.tex
minor improvements into manuscript

@@
 
-The filter coefficients should be taken as $\nu \in [0,1]$ and $\alpha \in [0.5,1]$.
+The filter coefficients should be taken as $\nu \in (0,1]$ and $\alpha \in (0.5,1]$.
 If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude.
-The default values of the FOODiE provided scheme are $\nu=0.01$ $\alpha=0.5$, but they can be customized at runtime.
+The default values of the FOODiE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.
 
@@
 
-{\color{red} To be modified with changed API}
-
 \begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
@@
 class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
+class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
 real(R_P),                         intent(IN)    :: Dt           ! Time steps.
@@
 do s=1, self%steps
-  U = U + U%t(n=s, t=t(s)) * (Dt * self%b(s))
+  U = U + previous(s)%t() * (Dt * self%b(s))
 enddo
-call U%update_previous_steps
+do s=1, self%steps - 1
+  previous(s) = previous(s + 1)
+enddo
+previous(self%steps) = U
 return
@@
 
-This method takes four arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the time step used and the fourth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used and the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
 The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
 
-It is worth noting that there is an explicit call to the method \emph{update\_previous\_steps} for updating the values of the previously computed steps.
+It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
 
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Leapfrog integrator},label={list:init_lf_solver}]
-! default coefficients nu=0.01, alpha=0.5
+! default coefficients nu=0.01, alpha=0.53
 call integrator%init()
@@
   real(R_P) :: nu=0.01_R_P    ! Robert-Asselin filter coefficient.
-  real(R_P) :: alpha=0.5_R_P ! Robert-Asselin-Williams filter coefficient.
+  real(R_P) :: alpha=0.53_R_P ! Robert-Asselin-Williams filter coefficient.
   contains
@@
 type(my_integrand) :: my_field
-call integrator%integrate(U=my_field, filter=filter_displacement, Dt=Dt, t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.
+type(my_integrand) :: previous(1:2)
+call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{previous} are the memory registers where previous time steps solutions are saved, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.
 
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Leapfrog class},label={list:lf_integrate}]
-subroutine integrate(self, U, filter, Dt, t)
+subroutine integrate(self, U, previous, Dt, t, filter)
 class(leapfrog_integrator), intent(IN)    :: self          ! LF integrator.
 class(integrand),           intent(INOUT) :: U             ! Field to be integrated.
-class(integrand), optional, intent(INOUT) :: filter ! Filter field displacement.
+class(integrand),           intent(INOUT) :: previous(1:2) ! Previous time steps solutions.
 real(R_P),                  intent(in)    :: Dt            ! Time step.
 real(R_P),                  intent(IN)    :: t             ! Time.
-U = U%previous_step(n=1) + U%t(n=2, t=t) * (Dt * 2._R_P)
+class(integrand), optional, intent(INOUT) :: filter        ! Filter field displacement.
+U = previous(1) + previous(2)%t(t=t) * (Dt * 2._R_P)
 if (present(filter)) then
-  filter = (U%previous_step(n=1) - U%previous_step(n=2) * 2._R_P + U) * self%nu * 0.5_R_P
+  filter = (previous(1) - previous(2) * 2._R_P + U) * self%nu * 0.5_R_P
+  previous(2) = previous(2) + filter * self%alpha
+  U = U + filter * (self%alpha - 1._R_P)
 endif
-call U%update_previous_steps(filter=filter, weights=[self%alpha, self%alpha - 1._R_P])
+previous(1) = previous(2)
+previous(2) = U
 return
@@
 
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the optional filter-displacement-register, the fourth is the time step used and the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the optional filter-displacement-register, the fifth is the time step used and the sixth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
 The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
@@
 
-It is worth noting that there is an explicit call to the method \emph{update\_previous\_steps} for updating the values of the previously computed steps.
+It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
+In particular, if the filter displacement argument is passed the method performs the RAW filtering.
 
@@
   real(R_P), dimension(:),   allocatable :: U        ! Integrand (state) variables, [1:dims].
-  integer(I_P)                           :: steps=0  ! Number of time steps stored.
-  real(R_P), dimension(:,:), allocatable :: previous ! Previous time steps states, [1:dims,1:steps].
   contains
@@
     procedure, pass(self), public :: t => dOscillation_dt
-    procedure, pass(self), public :: update_previous_steps
-    procedure, pass(self), public :: previous_step
     procedure, pass(lhs),  public :: integrand_multiply_integrand => &
@@
   \item $U$ is the state vector corresponding directly to the state vector of equation \ref{eq:oscillation};
-  \item $steps$ and $previous$ are two auxiliary data members used only in case the oscillation field is integrated by means of a multi-step solver (thus they could be omitted in case of one-step solver usage):
-    \begin{itemize}
-      \item $steps$ is the number of previous time steps used for computing the next one-step-over integration with respect the current time;
-      \item $previous$ is an array where previously computed time step solutions are stored.
-      \end{itemize}
   \end{itemize}
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} residuals function},label={list:oscillation_t}]
-function dOscillation_dt(self, n, t) result(dState_dt)
+function dOscillation_dt(self, t) result(dState_dt)
 class(oscillation),     intent(IN) :: self      ! Oscillation field.
-integer(I_P), optional, intent(IN) :: n         ! Time level.
 real(R_P),    optional, intent(IN) :: t         ! Time.
@@
   dState_dt = self
-  if (self%steps>=2) then ! self\%previous should be used
-    dn = self%steps ; if (present(n)) dn = n
-    dState_dt%U(1) = -self%f * self%previous(2, dn)
-    dState_dt%U(2) =  self%f * self%previous(1, dn)
-  else ! self\%previous should not be used, use directly self\%U
   dState_dt%U(1) = -self%f * self%U(2)
   dState_dt%U(2) =  self%f * self%U(1)
-  endif
 endselect
@@
   lhs%dims = rhs%dims
-  lhs%steps = rhs%steps
+  lhs%f = rhs%f
   if (allocated(rhs%U)) lhs%U = rhs%U
-  if (allocated(rhs%previous)) lhs%previous = rhs%previous
-  lhs%f = rhs%f
 endselect
@@
 type(oscillation)                :: rk_stage(1:2) ! Runge-Kutta stages.
+type(oscillation)                :: previous(1:2) ! Previous time steps solution.
 type(oscillation)                :: oscillator    ! Oscillation field.
@@
     call rk_integrator%integrate(U=oscillator, stage=rk_stage, Dt=Dt, t=step*Dt)
+    previous(step) = oscillator
   else
-    call lf_integrator%integrate(U=oscillator, filter=filter, Dt=Dt, t=step*Dt)
+    call lf_integrator%integrate(U=oscillator, previous=previous, filter=filter, Dt=Dt, &
+                                 t=step*Dt)
   endif
@@
 
-\paragraph{FOODiE integration}
 \subsubsection{Errors Analysis of the FOODiE Oscillation test results}
@@
   \caption{Oscillation test errors analysis}\label{tab:oscillation}
-  \resizebox{0.75\textwidth}{!}{%
-  \begin{tabular}{cccccc}
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth 1 step}\label{tab:oscillation-ab-1}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
       \toprule
-    {\sc Solver} & {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{Adams-Bashforth 2 steps}
-                 &   5000.0        &  0.596E+03    &  0.583E+03    & /             & /             \\
-                 &   2500.0        &  0.221E+02    &  0.218E+02    & 4.75          & 4.74          \\
-                 &   1250.0        &  0.764E+01    &  0.769E+01    & 1.53          & 1.50          \\
-                 &    625.0        &  0.265E+01    &  0.268E+01    & 1.53          & 1.52          \\
-                 &    320.0        &  0.968E+00    &  0.981E+00    & 1.51          & 1.50          \\
-                 &    100.0        &  0.169E+00    &  0.171E+00    & 1.50          & 1.50          \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{Adams-Bashforth 3 steps}
-                 &   5000.0        &  0.857E+01    &  0.854E+01    & /             & /             \\
-                 &   2500.0        &  0.391E+01    &  0.386E+01    & 1.13          & 1.14          \\
-                 &   1250.0        &  0.825E+00    &  0.814E+00    & 2.24          & 2.25          \\
-                 &    625.0        &  0.150E+00    &  0.148E+00    & 2.46          & 2.46          \\
-                 &    320.0        &  0.282E-01    &  0.278E-01    & 2.49          & 2.49          \\
-                 &    100.0        &  0.154E-02    &  0.152E-02    & 2.50          & 2.50          \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{Adams-Bashforth 4 steps}
-                 &   5000.0        &  0.128E+07    &  0.143E+07    & /             & /             \\
-                 &   2500.0        &  0.106E+01    &  0.107E+01    & 20.21         & 20.34         \\
-                 &   1250.0        &  0.967E-01    &  0.981E-01    &  3.45         &  3.45         \\
-                 &    625.0        &  0.859E-02    &  0.871E-02    &  3.49         &  3.49         \\
-                 &    320.0        &  0.827E-03    &  0.838E-03    &  3.50         &  3.50         \\
-                 &    100.0        &  0.141E-04    &  0.143E-04    &  3.50         &  3.50         \\
-
-    \cmidrule{2-6}
-    \multirow{6}{*}{Leapfrog RAW-filtered}
-                 &   5000.0        &  0.156E+02    &  0.156E+02    & /             & /             \\
-                 &   2500.0        &  0.855E+01    &  0.852E+01    & 0.86          & 0.87          \\
-                 &   1250.0        &  0.303E+01    &  0.305E+01    & 1.50          & 1.48          \\
-                 &    625.0        &  0.107E+01    &  0.108E+01    & 1.51          & 1.50          \\
-                 &    320.0        &  0.390E+00    &  0.395E+00    & 1.50          & 1.50          \\
-                 &    100.0        &  0.685E-01    &  0.692E-01    & 1.50          & 1.50          \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{low storage Runge-Kutta 5 stages}
-                 &   5000.0        &  0.120E+00    &  0.122E+00    & /             & /             \\
-                 &   2500.0        &  0.106E-01    &  0.107E-01    & 3.51          & 3.51          \\
-                 &   1250.0        &  0.935E-03    &  0.947E-03    & 3.50          & 3.50          \\
-                 &    625.0        &  0.826E-04    &  0.836E-04    & 3.50          & 3.50          \\
-                 &    320.0        &  0.793E-05    &  0.803E-05    & 3.50          & 3.50          \\
-                 &    100.0        &  0.135E-06    &  0.137E-06    & 3.50          & 3.50          \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{TVD/SSP Runge-Kutta 2 stages}
-                 &   5000.0        &  0.316E+02    &  0.319E+02    & /             & /             \\
-                 &   2500.0        &  0.892E+01    &  0.894E+01    & 1.83          & 1.84          \\
-                 &   1250.0        &  0.301E+01    &  0.305E+01    & 1.57          & 1.55          \\
-                 &    625.0        &  0.106E+01    &  0.107E+01    & 1.51          & 1.51          \\
-                 &    320.0        &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
-                 &    100.0        &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{TVD/SSP Runge-Kutta 3 stages}
-                 &   5000.0        &  0.255E+01    &  0.252E+01    & /             & /             \\
-                 &   2500.0        &  0.523E+00    &  0.516E+00    & 2.28          & 2.29          \\
-                 &   1250.0        &  0.944E-01    &  0.931E-01    & 2.47          & 2.47          \\
-                 &    625.0        &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
-                 &    320.0        &  0.314E-02    &  0.310E-02    & 2.50          & 2.50          \\
-                 &    100.0        &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
-    \cmidrule{2-6}
-    \multirow{6}{*}{TVD/SSP Runge-Kutta 5 stages}
-                 &   5000.0        &  0.139E+00    &  0.141E+00    & /             & /             \\
-                 &   2500.0        &  0.122E-01    &  0.124E-01    & 3.50          & 3.50          \\
-                 &   1250.0        &  0.108E-02    &  0.110E-02    & 3.50          & 3.50          \\
-                 &    625.0        &  0.956E-04    &  0.969E-04    & 3.50          & 3.50          \\
-                 &    320.0        &  0.937E-05    &  0.949E-05    & 3.47          & 3.47          \\
-                 &    100.0        &  0.512E-06    &  0.519E-06    & 2.50          & 2.50          \\
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
+      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
+      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
+       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
+       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
+       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
       \bottomrule
     \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth 2 steps}\label{tab:oscillation-ab-2}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.596E+03    &  0.583E+03    & /             & /             \\
+      2500.0          &  0.221E+02    &  0.218E+02    & 4.75          & 4.74          \\
+      1250.0          &  0.764E+01    &  0.769E+01    & 1.53          & 1.50          \\
+       625.0          &  0.265E+01    &  0.268E+01    & 1.53          & 1.52          \\
+       320.0          &  0.968E+00    &  0.981E+00    & 1.51          & 1.50          \\
+       100.0          &  0.169E+00    &  0.171E+00    & 1.50          & 1.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth 3 steps}\label{tab:oscillation-ab-3}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.857E+01    &  0.854E+01    & /             & /             \\
+      2500.0          &  0.391E+01    &  0.386E+01    & 1.13          & 1.14          \\
+      1250.0          &  0.825E+00    &  0.814E+00    & 2.24          & 2.25          \\
+       625.0          &  0.150E+00    &  0.148E+00    & 2.46          & 2.46          \\
+       320.0          &  0.282E-01    &  0.278E-01    & 2.49          & 2.49          \\
+       100.0          &  0.154E-02    &  0.152E-02    & 2.50          & 2.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth 4 steps}\label{tab:oscillation-ab-4}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.128E+07    &  0.143E+07    & /             & /             \\
+      2500.0          &  0.106E+01    &  0.107E+01    & 20.21         & 20.34         \\
+      1250.0          &  0.967E-01    &  0.981E-01    &  3.45         &  3.45         \\
+       625.0          &  0.859E-02    &  0.871E-02    &  3.49         &  3.49         \\
+       320.0          &  0.827E-03    &  0.838E-03    &  3.50         &  3.50         \\
+       100.0          &  0.141E-04    &  0.143E-04    &  3.50         &  3.50         \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Leapfrog unfiltered}\label{tab:oscillation-leapfrog}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.156E+02    &  0.156E+02    & /             & /             \\
+      2500.0          &  0.849E+01    &  0.846E+01    & 0.87          & 0.88          \\
+      1250.0          &  0.300E+01    &  0.303E+01    & 1.50          & 1.48          \\
+       625.0          &  0.106E+01    &  0.107E+01    & 1.51          & 1.50          \\
+       320.0          &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
+       100.0          &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Leapfrog RAW-filtered}\label{tab:oscillation-leapfrog-raw}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.156E+02    &  0.156E+02    & /             & /             \\
+      2500.0          &  0.855E+01    &  0.852E+01    & 0.86          & 0.87          \\
+      1250.0          &  0.303E+01    &  0.305E+01    & 1.50          & 1.48          \\
+       625.0          &  0.107E+01    &  0.108E+01    & 1.51          & 1.50          \\
+       320.0          &  0.390E+00    &  0.395E+00    & 1.50          & 1.50          \\
+       100.0          &  0.685E-01    &  0.692E-01    & 1.50          & 1.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Low storage Runge-Kutta 1 stage}\label{tab:oscillation-ls-rk-1}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
+      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
+      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
+       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
+       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
+       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Low storage Runge-Kutta 5 stages}\label{tab:oscillation-ls-rk-5}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.120E+00    &  0.122E+00    & /             & /             \\
+      2500.0          &  0.106E-01    &  0.107E-01    & 3.51          & 3.51          \\
+      1250.0          &  0.935E-03    &  0.947E-03    & 3.50          & 3.50          \\
+       625.0          &  0.826E-04    &  0.836E-04    & 3.50          & 3.50          \\
+       320.0          &  0.793E-05    &  0.803E-05    & 3.50          & 3.50          \\
+       100.0          &  0.135E-06    &  0.137E-06    & 3.50          & 3.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{TVD/SSP Runge-Kutta 1 stage}\label{tab:oscillation-tvd-rk-1}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
+      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
+      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
+       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
+       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
+       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{TVD/SSP Runge-Kutta 2 stages}\label{tab:oscillation-tvd-rk-2}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.316E+02    &  0.319E+02    & /             & /             \\
+      2500.0          &  0.892E+01    &  0.894E+01    & 1.83          & 1.84          \\
+      1250.0          &  0.301E+01    &  0.305E+01    & 1.57          & 1.55          \\
+       625.0          &  0.106E+01    &  0.107E+01    & 1.51          & 1.51          \\
+       320.0          &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
+       100.0          &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{TVD/SSP Runge-Kutta 3 stages}\label{tab:oscillation-tvd-rk-3}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.255E+01    &  0.252E+01    & /             & /             \\
+      2500.0          &  0.523E+00    &  0.516E+00    & 2.28          & 2.29          \\
+      1250.0          &  0.944E-01    &  0.931E-01    & 2.47          & 2.47          \\
+       625.0          &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
+       320.0          &  0.314E-02    &  0.310E-02    & 2.50          & 2.50          \\
+       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{TVD/SSP Runge-Kutta 5 stages}\label{tab:oscillation-tvd-rk-5}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.139E+00    &  0.141E+00    & /             & /             \\
+      2500.0          &  0.122E-01    &  0.124E-01    & 3.50          & 3.50          \\
+      1250.0          &  0.108E-02    &  0.110E-02    & 3.50          & 3.50          \\
+       625.0          &  0.956E-04    &  0.969E-04    & 3.50          & 3.50          \\
+       320.0          &  0.937E-05    &  0.949E-05    & 3.47          & 3.47          \\
+       100.0          &  0.512E-06    &  0.519E-06    & 2.50          & 2.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}
 \end{table}
@@
 \paragraph{Adams-Bashforth}
+
+As expected, the Adams-Bashforth 1 step solution, that reverts back to the explicit forward Euler one, is unstable for all the $\Delta t$ exercised.
 
@@
     \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
-    \caption{Adams-Bashforth 2 steps solution}\label{fig:results-oscillation-adams-bashforth-2}
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-1.png}
+    \caption{1 step solution}\label{fig:results-oscillation-adams-bashforth-1}
   \end{subfigure}\quad%
@@
     \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
-    \caption{Adams-Bashforth 3 steps solution}\label{fig:results-oscillation-adams-bashforth-3}
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
+    \caption{2 steps solution}\label{fig:results-oscillation-adams-bashforth-2}
   \end{subfigure}\\
@@
     \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
+    \caption{3 steps solution}\label{fig:results-oscillation-adams-bashforth-3}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-4.png}
-    \caption{Adams-Bashforth 4 steps solution}\label{fig:results-oscillation-adams-bashforth-4}
+    \caption{4 steps solution}\label{fig:results-oscillation-adams-bashforth-4}
   \end{subfigure}
@@
 
+The Leapfrog solutions are in agreement with the expected results: both unfiltered and RAW-filtered solutions show an observed order of accuracy that tends to the formal $2^{nd}$ order.
+The two solutions are almost the same.
+
 \begin{figure}[!ht]
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
+    \caption{Leapfrog unfiltered solution}\label{fig:results-oscillation-leapfrog-unfiltered}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog-raw.png}
+    \caption{Leapfrog RAW-filtered solution}\label{fig:results-oscillation-leapfrog-raw}
+  \end{subfigure}
   \caption{Oscillation equations solutions computed by means of Leapfrog RAW-filtered solver}\label{fig:results-oscillation-leapfrog}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-5.png}
-  \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta 5 stages solver}\label{fig:results-oscillation-ls-runge-kutta-5}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-1.png}
+    \caption{1 stage solution}\label{fig:results-oscillation-ls-runge-kutta-1}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-5.png}
+    \caption{5 stages solution}\label{fig:results-oscillation-ls-runge-kutta-5}
+  \end{subfigure}
+  \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta solvers}\label{fig:results-oscillation-ls-runge-kutta}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-2.png}
-  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 2 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-2}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-3.png}
-  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 3 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-3}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=0.70\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-5.png}
-  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta 5 stages solver}\label{fig:results-oscillation-tvd-runge-kutta-5}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-1.png}
+    \caption{1 stage solution}\label{fig:results-oscillation-tvd-runge-kutta-1}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-2.png}
+    \caption{2 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-2}
+  \end{subfigure}\\
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-3.png}
+    \caption{3 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-3}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-5.png}
+    \caption{5 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-5}
+  \end{subfigure}\\
+  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta solvers}\label{fig:results-oscillation-tvd-runge-kutta}
 \end{figure}

### 2015-10-08 papers/announcement/errors_analysis/errors_analysis.sh

@@ -13,14 +13,9 @@ tecplot=$(which tec360)
 if [ -x "$tecplot" ] ; then
   ln -fs ../utilities-oscillation/*lay .
   ln -fs ../utilities/lay_export* .
-  ./lay_export_all_f errors_analysis-oscillation-adams-bashforth-2.lay
-  ./lay_export_all_f errors_analysis-oscillation-adams-bashforth-3.lay
-  ./lay_export_all_f errors_analysis-oscillation-adams-bashforth-4.lay
-  ./lay_export_all_f errors_analysis-oscillation-leapfrog.lay
-  ./lay_export_all_f errors_analysis-oscillation-ls-runge-kutta-5.lay
-  ./lay_export_all_f errors_analysis-oscillation-tvd-runge-kutta-2.lay
-  ./lay_export_all_f errors_analysis-oscillation-tvd-runge-kutta-3.lay
-  ./lay_export_all_f errors_analysis-oscillation-tvd-runge-kutta-5.lay
+  for file in $( ls *lay ); do
+  ./lay_export_all_f $file
+  done
 fi
 cd -
 exit 0

### 2015-10-09 src/lib/foodie_integrator_adams_bashforth_moulton.f90
Add Adams-Moulton and Adams-Bashforth-Moulton solvers
Add Adams-Moulton and Adams-Bashforth-Moulton solvers

Why:

Implicit and predictor-corrector are important for stiff problems.

This change addresses the need by:

Add the two new modules for the two new solvers.

Side effects:

Change API of AB integrate method: now the auto update of previous time
steps solution can be disabled.

@@ -1,0 +1,175 @@
+!< FOODiE integrator: provide a predictor-corrector class of Adams-Bashforth-Moutlon multi-step schemes, from 1st to 4rd order
+!< accutate.
+module foodie_integrator_adams_bashforth_moulton
+!-----------------------------------------------------------------------------------------------------------------------------------
+!< FOODiE integrator: provide a predictor-corrector class of Adams-Bashforth-Moutlon multi-step schemes, from 1st to 4rd order
+!< accutate.
+!<
+!< Considering the following ODE system:
+!<
+!< $$ U_t = R(t,U) $$
+!<
+!< where \(U_t = \frac{dU}{dt}\), *U* is the vector of *state* variables being a function of the time-like independent variable
+!< *t*, *R* is the (vectorial) residual function, the Adams-Bashforth-Moulton class scheme implemented is:
+!<
+!<##### predictor
+!< $$ U^{n+N_s^p} = U^{n+N_s^p-1} +\Delta t \left[ \sum_{s=1}^{N_s^p}{ b_s^p \cdot R(t^{n+s-1}, U^{n+s-1}) } \right] $$
+!<##### corrector
+!< $$ U^{n+N_s^c} = U^{n+N_s^c-1} +\Delta t \left[ \sum_{s=0}^{N_s^c}{ b_s^c \cdot R(t^{n+s}, U^{n+s}) } \right] $$
+!<
+!<where \(N_s^p\) is the number of previous steps considered for the predictor and \(N_s^c\) is the number of previous steps
+!<considered for the corrector.
+!<
+!< @note The value of \(\Delta t\) must be provided, it not being computed by the integrator.
+!<
+!< The coefficients \(b_s^{p,c}\) define the actual scheme, that is selected accordingly to the number of
+!< **steps** used. The predictor and corrector schemes should have the same formal order of accuracy, thus \(N_s^p=N_s^c+1\)
+!< should hold.
+!<
+!< Currently, the following schemes are available:
+!<##### P=AB(1)-C=AM(0) step, Explicit/Implicit Farward/Backward Euler, 1st order
+!< This scheme is TVD and reverts to Explicit/Implicit Farward/Backward Euler, it being 1st order.
+!< The *b* coefficient is:
+!< $$b^p = \left[b_1\right] = \left[1\right]$$
[…]

### 2015-10-09 papers/announcement/manuscript.tex

@@
 The implicit approach is of practical use for \emph{stiff} systems where the usage of explicit schemes could require an extremely small time step to evolve in a \emph{stable} way the solution.
-Currently, FOODiE provides only explicit solvers.
+Mixing together explicit and implicit schemes it is possible to build a family of \emph{predictor-corrector} methods: using an explicit scheme to predict a guess for the next step approximation it is possible to use an implicit method for correcting this guess.
+Currently, FOODiE provides explicit solvers and predictor-correct ones.
 
@@
     \begin{itemize}
-      \item Adams-Bashforth schemes:
+      \item explicit Adams-Bashforth schemes:
         \begin{itemize}
@@
           \end{itemize}
-        \item Leapfrog schemes\footnote{Leapfrog method is a 2-steps solver being, in general $2^{nd}$ order accurate, but mostly \emph{unstable}.
+      \item implicit Adams-Moulton schemes:
+        \begin{itemize}
+          \item 1-step, it being $2^{nd}$ order accurate;
+          \item 2-steps, it being $3^{rd}$ order accurate;
+          \item 3-steps, it being $4^{th}$ order accurate;
+          \end{itemize}
+      \item predictor-corrector Adams-Bashforth-Moulton schemes:
+        \begin{itemize}
+          \item 1-step, it being $2^{nd}$ order accurate;
+          \item 2-steps, it being $3^{rd}$ order accurate;
+          \item 3-steps, it being $4^{th}$ order accurate;
+          \end{itemize}
+      \item explicit Leapfrog schemes\footnote{Leapfrog method is a 2-steps solver being, in general $2^{nd}$ order accurate, but mostly \emph{unstable}.
 It is well suited for periodic/oscillatory problems, thus the error is generally categorized accordingly to the phase and amplitude errors: the unfiltered scheme provides a $2^{nd}$ order error on phase approximation while the amplitude error is null.
@@
 
-Adams-Bashforth methods belong to the more general (linear) multi-step family of schemes.
+Adams-Bashforth methods belong to the more general (linear) explicit multi-step family of schemes.
 This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
@@
 To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
+
+{\color{red} Add citations.}
+
+\subsection{The implicit Adams-Moulton class of schemes}
+
+Adams-Moulton methods belong to the more general (linear) implicit multi-step family of schemes.
+This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
+Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
+
+In general, the Adams-Moulton schemes provided by FOODiE library are written by means of the following algorithm (for only implicit schemes):
+
+\begin{equation}
+  U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=0}^{n+N_s-1}{ b_s \cdot R\left[t^{n+s}, U\left(t^{n+s}\right)\right]} + b_{N_s}\cdot R\left[t^{n+N_s}, U\left(t^{n+N_s}\right)\right]
+\label{eq:AM}
+\end{equation}
+where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.
+
+Currently FOODiE provides 1, 2, and 3 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.
+
+\begin{table}[!ht]
+  \centering
+  \caption{Implicit Adams-Moulton coefficients\label{tab:AM-coeff}}
+  \resizebox{0.30\textwidth}{!}{%
+  \begin{tabular}{c|cccc}
+    $N_s$ & $b_0$           & $b_1$           & $b_2$           & $b_3$          \\
+    \hline
+    1     & $\frac{1}{2}$   & $\frac{1}{2}$   &       /         &       /        \\
+    2     & $-\frac{1}{12}$ & $\frac{8}{12}$  & $\frac{5}{12}$  &       /        \\
+    3     & $\frac{1}{24}$  & $-\frac{5}{24}$ & $\frac{19}{24}$ & $\frac{9}{24}$ \\
+  \end{tabular}}
+\end{table}
+
+Similarly to the Runge-Kutta and Adams-Bashforth classes, the Adams-Moulton class also provides a fail-safe zero-step solver reverting back to the implicit backward Euler solver, that is useful for ROR-like frameworks.
+
+It is worth noting that for $N_s>1$ the Adams-Moulton class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided.
+To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
+
+{\color{red} Add citations.}
+
+\subsection{The predictor-corrector Adams-Bashforth-Moulton class of schemes}
+
+Adams-Bashforth-Moulton methods belong to the more general (linear) predictor-corrector multi-step family of schemes.
+This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
+Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
+
+In general, the Adams-Bashforth-Moulton schemes provided by FOODiE library are written by means of the following algorithm:
+
+\begin{equation}
+  \begin{matrix}
+    U\left(t^{n+N_s^p}\right)_p = U\left(t^{n+N_s^p-1}\right) +\Delta t \sum_{s=1}^{n+N_s^p}{ b_s^p \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}   \\
+    U\left(t^{n+N_s^c}\right)_c = U\left(t^{n+N_s^c-1}\right) +\Delta t \sum_{s=0}^{n+N_s^c-1}{ b_s^c \cdot R\left[t^{n+s}, U\left(t^{n+s}\right)\right]} + b_{N_s^c}^c\cdot R\left[t^{n+N_s^p}, U\left(t^{n+N_s^p}\right)_p\right]
+  \end{matrix}
+\label{eq:ABM}
+\end{equation}
+where $N_s^{p,c}$ is the number of time steps considered for the Adams-Bashforth predictor/Adams-Moulton corrector (respectively) and $b_s^{p,c}$ are the corresponding linear coefficients selected.
+Essentially, the Adams-Bashforth prediction $U\left(t^{n+N_s^p}\right)_p$ is corrected by means of the Adams-Moulton correction resulting in $U\left(t^{n+N_s^c}\right)_c$.
+In order to preserve the formal order of accuracy the relation $N_s^p=N_s^c+1$ always holds.
+
+Currently FOODiE provides $N_s^c=1, 2, 3 \rightarrow N_s^c=2, 3, 4$ steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively.
+The $b_s^{p,c}$ coefficients are those reported in tables \ref{tab:AB-coeff} and \ref{tab:AM-coeff}.
 
@@
                   adams_bashforth_integrator, &
+                  adams_moulton_integrator, &
+                  adams_bashforth_moulton_integrator, &
                   euler_explicit_integrator, &
@@
   \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time.
+  \item \emph{integrate}: integrate integrand field one-step-over in time;
+  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
   \end{itemize}
@@
     procedure, pass(self), public :: integrate
+    procedure, pass(self), public :: update_previous
     final                         :: finalize
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
-subroutine integrate(self, U, previous, Dt, t)
+subroutine integrate(self, U, previous, Dt, t, autoupdate)
 class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
@@
 real(R_P),                         intent(IN)    :: t(:)         ! Times.
+logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
+logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
 integer(I_P)                                     :: s            ! Steps counter.
+autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
 do s=1, self%steps
-  U = U + previous(s)%t() * (Dt * self%b(s))
+  U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
 enddo
-do s=1, self%steps - 1
-  previous(s) = previous(s + 1)
-enddo
-previous(self%steps) = U
+if (autoupdate_) call self%update_previous(U=U, previous=previous)
 return
@@
 
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used and the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions.
 The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
@@
 It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
+This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.
+
+\subsection{The implicit Adams-Moulton class of solvers}\label{subsec:solver_am}
+
+The explicit Adams-Moulton class of solvers is exposed as a single derived type named \emph{adams\_moulton\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
+  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time;
+  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
+  \end{itemize}
+
+As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_am_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an implicit Adams-Moulton integrator},label={list:define_am_solver}]
+use FOODiE, only: adams_moulton_integrator
+type(adams_moulton_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_am_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an implicit Adams-Moulton integrator},label={list:init_am_solver}]
+call integrator%init(steps=3)
+\end{lstlisting}
+
+In the listing \ref{list:init_am_solver} a 3-steps solver has been initialized.
+As a matter of facts, from the equation \ref{eq:AM} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen.
+The complete definition of the \emph{adams\_moulton\_integrator} type is reported into listing \ref{list:am_solver}.
+As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_moulton\_integrator} type},label={list:am_solver}]
+type :: adams_moulton_integrator
+  private
+  integer(I_P)           :: steps=-1 ! Number of time steps.
+  real(R_P), allocatable :: b(:)     ! \(b\) coefficients.
+  contains
+    procedure, pass(self), public :: destroy
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+    procedure, pass(self), public :: update_previous
+    final                         :: finalize
+endtype adams_moulton_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_am_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Moulton integrator},label={list:use_am_solver}]
+real               :: times(1:3)
+type(my_integrand) :: my_field
+type(my_integrand) :: previous(1:3)
+call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.
+
+The complete implementation of the \emph{integrate} method of the implicit Adams-Moulton class of solvers is reported in listing \ref{list:am_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Moulton class},label={list:am_integrate}]
+subroutine integrate(self, U, previous, Dt, t, autoupdate)
+class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
+class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
+class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
+real(R_P),                         intent(IN)    :: Dt           ! Time steps.
+real(R_P),                         intent(IN)    :: t(:)         ! Times.
+logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
+logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
+integer(I_P)                                     :: s            ! Steps counter.
+autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
+if (autoupdate_) call self%update_previous(U=U, previous=previous)
+if (self%steps>0) then
+  U = previous(self%steps) + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
+  do s=0, self%steps - 1
+    U = U + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
+  enddo
+  if (autoupdate_) call self%update_previous(U=U, previous=previous)
+else
+  U = U + U%t(t=t(s+1)) * (Dt * self%b(0))
+endif
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+
+It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
+This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.
+
+\subsection{The predictor-corrector Adams-Bashforth-Moulton class of solvers}\label{subsec:solver_abm}
+
+{\color{red} To be written.}
 
@@
 type(my_integrand) :: previous(1:2)
-call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, t=times)
+call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, &
+                          t=times)
 \end{lstlisting}
@@
 
-Table \ref{tab:oscillation} summarizes the errors analysis.
+Tables \ref{tab:oscillation_errors_explicit} and \ref{tab:oscillation_errors_pred_corr} summarize the errors analysis.
 
@@
   \centering
-  \caption{Oscillation test errors analysis}\label{tab:oscillation}
+  \caption{Oscillation test: errors analysis of explicit solvers}\label{tab:oscillation_errors_explicit}
   \begin{subtable}[b]{0.45\textwidth}
@@
 
+\begin{table}[!ht]
+  \centering
+  \caption{Oscillation test: errors analysis of predictor-corrector solvers}\label{tab:oscillation_errors_pred_corr}
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth-Moulton 1 step}\label{tab:oscillation-abm-1}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.241E+20    &  0.266E+20    & /             & /             \\
+      2500.0          &  0.664E+11    &  0.716E+11    & 28.44         & 28.47         \\
+      1250.0          &  0.952E+06    &  0.100E+07    & 16.09         & 16.12         \\
+       625.0          &  0.413E+04    &  0.407E+04    &  7.85         &  7.95         \\
+       320.0          &  0.387E+03    &  0.383E+03    &  3.54         &  3.53         \\
+       100.0          &  0.145E+03    &  0.145E+03    &  0.84         &  0.83         \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth-Moulton 2 steps}\label{tab:oscillation-abm-2}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.704E+01    &  0.701E+01    & /             & /             \\
+      2500.0          &  0.392E+01    &  0.395E+01    & 0.84          & 0.83          \\
+      1250.0          &  0.148E+01    &  0.150E+01    & 1.40          & 1.39          \\
+       625.0          &  0.526E+00    &  0.534E+00    & 1.49          & 1.49          \\
+       320.0          &  0.193E+00    &  0.196E+00    & 1.50          & 1.50          \\
+       100.0          &  0.338E-01    &  0.342E-01    & 1.50          & 1.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth-Moulton 3 steps}\label{tab:oscillation-abm-3}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.457E+01    &  0.464E+01    & /             & /             \\
+      2500.0          &  0.656E+00    &  0.654E+00    & 2.80          & 2.83          \\
+      1250.0          &  0.100E+00    &  0.987E-01    & 2.71          & 2.73          \\
+       625.0          &  0.169E-01    &  0.167E-01    & 2.56          & 2.56          \\
+       320.0          &  0.314E-02    &  0.310E-02    & 2.52          & 2.51          \\
+       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Adams-Bashforth-Moulton 4 steps}\label{tab:oscillation-abm-4}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.229E+01    &  0.225E+01    & /             & /             \\
+      2500.0          &  0.119E+00    &  0.118E+00    & 4.26          & 4.25          \\
+      1250.0          &  0.825E-02    &  0.833E-02    & 3.85          & 3.83          \\
+       625.0          &  0.671E-03    &  0.681E-03    & 3.62          & 3.61          \\
+       320.0          &  0.631E-04    &  0.640E-04    & 3.53          & 3.53          \\
+       100.0          &  0.107E-05    &  0.108E-05    & 3.51          & 3.51          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+\end{table}
+
 \paragraph{Adams-Bashforth}
@@
 
+\begin{figure}[!ht]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-1.png}
+    \caption{1 step solution}\label{fig:results-oscillation-adams-bashforth-moulton-1}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-2.png}
+    \caption{2 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-2}
+  \end{subfigure}\\
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-3.png}
+    \caption{3 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-3}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-4.png}
+    \caption{4 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-4}
+  \end{subfigure}
+  \caption{Oscillation equations solutions computed by means of Adams-Bashforth-Moulton solvers}\label{fig:results-oscillation-adams-bashforth-moulton}
+\end{figure}
 \clearpage

### 2015-10-13 papers/announcement/manuscript.tex
Add low storage RK 2N s6o4_s7o4
Add low storage RK 2N LSRK(6,4) and LSRK(7,4) from
*High-accuracy large-step explicit Runge–Kutta (HALE-RK) schemes for
computational aeroacoustics*, Vasanth Allampalli and Ray Hixon and
M. Nallasamy and Scott D. Sawyer, Journal of Computational Physics,
vol. 228, 2009, pp. 3837--3850.

Why:

High number of stages increment stability locus.

This change addresses the need by:

Add new coefficients for ls_runge_kutta_interpolator class.

Side effects:

Update oscillation test and manuscript.

@@
               \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
+              \item 6-stages 2N registers schemes, it being $4^{th}$ order accurate;
+              \item 7-stages 2N registers schemes, it being $4^{th}$ order accurate;
               \end{itemize}
@@
 
-Currently FOODiE provides only one 5 stages, $4^{th}$ order, 2N registers explicit scheme, the coefficients of which are listed in table \ref{tab:RK-ls-5}.
+Currently FOODiE provides 5/6/7 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in tables \ref{tab:RK-ls-5}, \ref{tab:RK-ls-6} and \ref{tab:RK-ls-7}.
 
@@
   \resizebox{0.40\textwidth}{!}{%
-  \begin{tabular}{ccc}
-       $A$                                    & $B$                                    & $C$                                   \\
-    \hline
-       $ 0                                  $ & $\frac{1432997174477}{9575080441755 }$ & $ 0                                 $ \\
-       $-\frac{567301805773 }{1357537059087}$ & $\frac{5161836677717}{13612068292357}$ & $\frac{1432997174477}{9575080441755}$ \\
-       $-\frac{2404267990393}{2016746695238}$ & $\frac{1720146321549}{2090206949498 }$ & $\frac{2526269341429}{6820363962896}$ \\
-       $-\frac{3550918686646}{2091501179385}$ & $\frac{3134564353537}{4481467310338 }$ & $\frac{2006345519317}{3224310063776}$ \\
-       $-\frac{1275806237668}{842570457699 }$ & $\frac{2277821191437}{14882151754819}$ & $\frac{2802321613138}{2924317926251}$ \\
+  \begin{tabular}{cccc}
+    Stage & $A$                                    & $B$                                    & $C$                                   \\
+    \hline
+    1     & $ 0                                  $ & $\frac{1432997174477}{9575080441755 }$ & $ 0                                 $ \\
+    2     & $-\frac{567301805773 }{1357537059087}$ & $\frac{5161836677717}{13612068292357}$ & $\frac{1432997174477}{9575080441755}$ \\
+    3     & $-\frac{2404267990393}{2016746695238}$ & $\frac{1720146321549}{2090206949498 }$ & $\frac{2526269341429}{6820363962896}$ \\
+    4     & $-\frac{3550918686646}{2091501179385}$ & $\frac{3134564353537}{4481467310338 }$ & $\frac{2006345519317}{3224310063776}$ \\
+    5     & $-\frac{1275806237668}{842570457699 }$ & $\frac{2277821191437}{14882151754819}$ & $\frac{2802321613138}{2924317926251}$ \\
+  \end{tabular}}
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Williamson's table of 6 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-6}}
+  \resizebox{0.40\textwidth}{!}{%
+  \begin{tabular}{cccc}
+    Stage & $A$               & $B$              & $C$              \\
+    \hline
+    1     & $ 0             $ & $0.122000000000$ & $0             $ \\
+    2     & $-0.691750960670$ & $0.477263056358$ & $0.122000000000$ \\
+    3     & $-1.727127405211$ & $0.381941220320$ & $0.269115878630$ \\
+    4     & $-0.694890150986$ & $0.447757195744$ & $0.447717183551$ \\
+    5     & $-1.039942756197$ & $0.498614246822$ & $0.749979795490$ \\
+    6     & $-1.531977447611$ & $0.186648570846$ & $0.898555413085$ \\
+  \end{tabular}}
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Williamson's table of 7 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-7}}
+  \resizebox{0.40\textwidth}{!}{%
+  \begin{tabular}{cccc}
+    Stage & $A$               & $B$              & $C$              \\
+    \hline
+    1     & $ 0.000000000000$ & $0.117322146869$ & $0.000000000000$ \\
+    2     & $-0.647900745934$ & $0.503270262127$ & $0.117322146869$ \\
+    3     & $-2.704760863204$ & $0.233663281658$ & $0.294523230758$ \\
+    4     & $-0.460080550118$ & $0.283419634625$ & $0.305658622131$ \\
+    5     & $-0.500581787785$ & $0.540367414023$ & $0.582864148403$ \\
+    6     & $-1.906532255913$ & $0.371499414620$ & $0.858664273599$ \\
+    7     & $-1.450000000000$ & $0.136670099385$ & $0.868664273599$ \\
   \end{tabular}}
@@
 
-Tables \ref{tab:oscillation_errors_explicit} and \ref{tab:oscillation_errors_pred_corr} summarize the errors analysis.
+Tables \ref{tab:oscillation_errors_ab}, \ref{tab:oscillation_errors_lf}, \ref{tab:oscillation_errors_ls_rk}, \ref{tab:oscillation_errors_tvd_rk} and \ref{tab:oscillation_errors_abm} summarize the errors analysis.
 
@@
   \centering
-  \caption{Oscillation test: errors analysis of explicit solvers}\label{tab:oscillation_errors_explicit}
+  \caption{Oscillation test: errors analysis of explicit Adams-Bashforth solvers}\label{tab:oscillation_errors_ab}
   \begin{subtable}[b]{0.45\textwidth}
@@
   \end{subtable}\\
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Oscillation test: errors analysis of explicit Leapfrog solvers}\label{tab:oscillation_errors_lf}
   \begin{subtable}[b]{0.45\textwidth}
@@
   \end{subtable}\\
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Oscillation test: errors analysis of explicit Low Storage Runge-Kutta solvers}\label{tab:oscillation_errors_ls_rk}
   \begin{subtable}[b]{0.45\textwidth}
@@
     \centering
+    \caption{Low storage Runge-Kutta 6 stages}\label{tab:oscillation-ls-rk-s}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.979E-01    &  0.994E-01    & /             & /             \\
+      2500.0          &  0.876E-02    &  0.888E-02    & 3.48          & 3.48          \\
+      1250.0          &  0.776E-03    &  0.786E-03    & 3.50          & 3.50          \\
+       625.0          &  0.686E-04    &  0.695E-04    & 3.50          & 3.50          \\
+       320.0          &  0.659E-05    &  0.667E-05    & 3.50          & 3.50          \\
+       100.0          &  0.112E-06    &  0.114E-06    & 3.50          & 3.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{Low storage Runge-Kutta 7 stages}\label{tab:oscillation-ls-rk-7}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.238E-01    &  0.240E-01    & /             & /             \\
+      2500.0          &  0.203E-02    &  0.205E-02    & 3.55          & 3.55          \\
+      1250.0          &  0.177E-03    &  0.180E-03    & 3.51          & 3.51          \\
+       625.0          &  0.156E-04    &  0.158E-04    & 3.50          & 3.50          \\
+       320.0          &  0.150E-05    &  0.152E-05    & 3.50          & 3.50          \\
+       100.0          &  0.269E-07    &  0.273E-07    & 3.46          & 3.46          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \caption{Oscillation test: errors analysis of explicit TVD/SSP Runge-Kutta}\label{tab:oscillation_errors_tvd_rk}
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
     \caption{TVD/SSP Runge-Kutta 1 stage}\label{tab:oscillation-tvd-rk-1}
@@
   \centering
-  \caption{Oscillation test: errors analysis of predictor-corrector solvers}\label{tab:oscillation_errors_pred_corr}
+  \caption{Oscillation test: errors analysis of predictor-corrector Adams-Bashforth-Moulton solvers}\label{tab:oscillation_errors_abm}
   \begin{subtable}[b]{0.45\textwidth}
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-1.png}
-    \caption{1 step solution}\label{fig:results-oscillation-adams-bashforth-1}
+    \caption{1 step}\label{fig:results-oscillation-adams-bashforth-1}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
-    \caption{2 steps solution}\label{fig:results-oscillation-adams-bashforth-2}
+    \caption{2 steps}\label{fig:results-oscillation-adams-bashforth-2}
   \end{subfigure}\\
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
-    \caption{3 steps solution}\label{fig:results-oscillation-adams-bashforth-3}
+    \caption{3 steps}\label{fig:results-oscillation-adams-bashforth-3}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-4.png}
-    \caption{4 steps solution}\label{fig:results-oscillation-adams-bashforth-4}
+    \caption{4 steps}\label{fig:results-oscillation-adams-bashforth-4}
   \end{subfigure}
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
-    \caption{Leapfrog unfiltered solution}\label{fig:results-oscillation-leapfrog-unfiltered}
+    \caption{Leapfrog unfiltered}\label{fig:results-oscillation-leapfrog-unfiltered}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog-raw.png}
-    \caption{Leapfrog RAW-filtered solution}\label{fig:results-oscillation-leapfrog-raw}
+    \caption{Leapfrog RAW-filtered}\label{fig:results-oscillation-leapfrog-raw}
   \end{subfigure}
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-1.png}
-    \caption{1 stage solution}\label{fig:results-oscillation-ls-runge-kutta-1}
+    \caption{1 stage}\label{fig:results-oscillation-ls-runge-kutta-1}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-5.png}
-    \caption{5 stages solution}\label{fig:results-oscillation-ls-runge-kutta-5}
+    \caption{5 stages}\label{fig:results-oscillation-ls-runge-kutta-5}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-6.png}
+    \caption{6 stages}\label{fig:results-oscillation-ls-runge-kutta-6}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-7.png}
+    \caption{7 stages}\label{fig:results-oscillation-ls-runge-kutta-7}
   \end{subfigure}
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-1.png}
-    \caption{1 stage solution}\label{fig:results-oscillation-tvd-runge-kutta-1}
+    \caption{1 stage}\label{fig:results-oscillation-tvd-runge-kutta-1}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-2.png}
-    \caption{2 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-2}
+    \caption{2 stages}\label{fig:results-oscillation-tvd-runge-kutta-2}
   \end{subfigure}\\
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-3.png}
-    \caption{3 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-3}
+    \caption{3 stages}\label{fig:results-oscillation-tvd-runge-kutta-3}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-5.png}
-    \caption{5 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-5}
+    \caption{5 stages}\label{fig:results-oscillation-tvd-runge-kutta-5}
   \end{subfigure}\\
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-1.png}
-    \caption{1 step solution}\label{fig:results-oscillation-adams-bashforth-moulton-1}
+    \caption{1 step}\label{fig:results-oscillation-adams-bashforth-moulton-1}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-2.png}
-    \caption{2 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-2}
+    \caption{2 steps}\label{fig:results-oscillation-adams-bashforth-moulton-2}
   \end{subfigure}\\
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-3.png}
-    \caption{3 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-3}
+    \caption{3 steps}\label{fig:results-oscillation-adams-bashforth-moulton-3}
   \end{subfigure}\quad%
@@
     \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-4.png}
-    \caption{4 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-4}
+    \caption{4 steps}\label{fig:results-oscillation-adams-bashforth-moulton-4}
   \end{subfigure}

### 2015-10-13 src/lib/foodie_integrator_low_storage_runge_kutta.f90

@@ -32,36 +32,50 @@ module foodie_integrator_low_storage_runge_kutta
 !<
 !< The different schemes are selected accordingly to the number of stages used. Currently the following schemes are available:
 !<
-!<##### 1 stage, Explicit Forward Euler, 1st order
+!<#### 1 stage, Explicit Forward Euler, 1st order
 !< This scheme is TVD and reverts to Explicit Forward Euler, it being 1st order. It is not a real low storage scheme, this being
 !< meaningless for a first order scheme. However it is added for safety reason.
-!< $$A = \left[ 0 \right]$$
-!< $$B = \left[ 1 \right]$$
-!< $$C = \left[ 0 \right]$$
 !<
-!<##### 5 stages, SSP, 4th order
+!< | ` Stage ` | ` A ` | ` B ` | ` C ` |
+!< |-----------|-------|-------|-------|
+!< | ` 1 `     | ` 0 ` | ` 1 ` | ` 0 ` |
+!<
+!<#### 5 stages, SSP, 4th order
 !< This scheme is a low storage RK(5, 4), based on the *solution 3* proposed in [2].
-!<$$A = \left[ {\begin{matrix}
-!< 0                                   \\
-!<-\frac{567301805773 }{1357537059087} \\
-!<-\frac{2404267990393}{2016746695238} \\
-!<-\frac{3550918686646}{2091501179385} \\
-!<-\frac{1275806237668}{842570457699 }
-!<\end{matrix}} \right]$$
-!<$$B = \left[ {\begin{matrix}
-!<\frac{1432997174477}{9575080441755 } \\
-!<\frac{5161836677717}{13612068292357} \\
-!<\frac{1720146321549}{2090206949498 } \\
-!<\frac{3134564353537}{4481467310338 } \\
-!<\frac{2277821191437}{14882151754819}
-!<\end{matrix}} \right]$$
-!<$$C = \left[ {\begin{matrix}
-!< 0                                  \\
-!<\frac{1432997174477}{9575080441755} \\
-!<\frac{2526269341429}{6820363962896} \\
-!<\frac{2006345519317}{3224310063776} \\
-!<\frac{2802321613138}{2924317926251}
-!<\end{matrix}} \right]$$
+!<
+!< | ` Stage ` | ` A `                            | ` B `                            | ` C `                           |
[…]

### 2015-10-13 src/tests/oscillation/oscillation.f90

@@ -296,7 +296,7 @@ contains
   !< Test explicit low storage Runge-Kutta class of ODE solvers.
   !---------------------------------------------------------------------------------------------------------------------------------
   integer(I_P), intent(IN)      :: stages            !< Number of stages used: if negative all RK solvers are used.
-  integer, parameter            :: rk_stages=5       !< Runge-Kutta stages number.
+  integer, parameter            :: rk_stages=7       !< Runge-Kutta stages number.
   real(R_P), allocatable        :: solution(:,:)     !< Solution at each time step.
   integer(I_P)                  :: s                 !< RK stages counter.
   integer(I_P)                  :: stages_range(1:2) !< Stages used.
@@ -590,53 +590,40 @@ contains
   !<
   !<where \(\frac{\Delta t_1}{\Delta t_2}>1\).
   !---------------------------------------------------------------------------------------------------------------------------------
-  integer(I_P), parameter          :: NDt=6                        !< Number of time steps exercised.
-  real(R_P),    parameter          :: time_steps(1:NDt)=[5000._R_P, &
-                                                         2500._R_P, &
-                                                         1250._R_P, &
-                                                         625._R_P,  &
-                                                         320._R_P,  &
-                                                         100._R_P] !< Time steps exercised.
-  integer, parameter               :: ab_steps=4                   !< Adams-Bashforth steps number.
-  integer, parameter               :: rk_stages=5                  !< Runge-Kutta stages number.
-  real(R_P)                        :: ab_errors(1:space_dimension,&
-                                                1:ab_steps,       &
-                                                1:NDt)             !< Adams-Bashforth errors.
-  real(R_P)                        :: abm_errors(1:space_dimension,&
-                                                 1:ab_steps,       &
-                                                 1:NDt)            !< Adams-Bashforth-Moulton errors.
-  real(R_P)                        :: lf_errors(1:space_dimension,&
-                                                1:2,              &! unfiltered and RAW-filtered
[…]

### 2015-10-13 papers/announcement/manuscript.tex
Add low storage RK 2N s12/13/14o4
Add low storage RK 2N LSRK(12,4), LSRK(13,4) and LSRK(14,4) from
*High-accuracy large-step explicit Runge–Kutta (HALE-RK) schemes for
computational aeroacoustics*, Vasanth Allampalli and Ray Hixon and
M. Nallasamy and Scott D. Sawyer, Journal of Computational Physics,
vol. 228, 2009, pp. 3837--3850.

Why:

High number of stages increment stability locus.

This change addresses the need by:

Add new coefficients for ls_runge_kutta_interpolator class.

Side effects:

Update oscillation test and manuscript.

@@
               \item 7-stages 2N registers schemes, it being $4^{th}$ order accurate;
+              \item 12-stages 2N registers schemes, it being $4^{th}$ order accurate;
+              \item 13-stages 2N registers schemes, it being $4^{th}$ order accurate;
+              \item 14-stages 2N registers schemes, it being $4^{th}$ order accurate;
               \end{itemize}
@@
 
-Currently FOODiE provides 5/6/7 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in tables \ref{tab:RK-ls-5}, \ref{tab:RK-ls-6} and \ref{tab:RK-ls-7}.
+Currently FOODiE provides 5/6/7/12/13/14 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in table \ref{tab:RK-ls}.
 
@@
   \centering
-  \caption{Williamson's table of 5 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-5}}
-  \resizebox{0.40\textwidth}{!}{%
+  \caption{Williamson's table of low storage Runge-Kutta schemes\label{tab:RK-ls}}
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{5 stages, $4^{th}$ order\label{tab:RK-ls-5}}
+    \resizebox{1.00\textwidth}{!}{%
     \begin{tabular}{cccc}
@@
     \end{tabular}}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \caption{Williamson's table of 6 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-6}}
-  \resizebox{0.40\textwidth}{!}{%
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{6 stages, $4^{th}$ order\label{tab:RK-ls-6}}
+    \resizebox{1.00\textwidth}{!}{%
     \begin{tabular}{cccc}
@@
     \end{tabular}}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \caption{Williamson's table of 7 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-7}}
-  \resizebox{0.40\textwidth}{!}{%
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{7 stages, $4^{th}$ order\label{tab:RK-ls-7}}
+    \resizebox{1.00\textwidth}{!}{%
     \begin{tabular}{cccc}
@@
     \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{12 stages, $4^{th}$ order\label{tab:RK-ls-12}}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{cccc}
+      Stage & $A$                   & $B$                  & $C$                  \\
+      \hline
+      1     & $ 0.0000000000000000$ & $0.0650008435125904$ & $0.0000000000000000$ \\
+      2     & $-0.0923311242368072$ & $0.0161459902249842$ & $0.0650008435125904$ \\
+      3     & $-0.9441056581158819$ & $0.5758627178358159$ & $0.0796560563081853$ \\
+      4     & $-4.3271273247576394$ & $0.1649758848361671$ & $0.1620416710085376$ \\
+      5     & $-2.1557771329026072$ & $0.3934619494248182$ & $0.2248877362907778$ \\
+      6     & $-0.9770727190189062$ & $0.0443509641602719$ & $0.2952293985641261$ \\
+      7     & $-0.7581835342571139$ & $0.2074504268408778$ & $0.3318332506149405$ \\
+      8     & $-1.7977525470825499$ & $0.6914247433015102$ & $0.4094724050198658$ \\
+      9     & $-2.6915667972700770$ & $0.3766646883450449$ & $0.6356954475753369$ \\
+      10    & $-4.6466798960268143$ & $0.0757190350155483$ & $0.6806551557645497$ \\
+      11    & $-0.1539613783825189$ & $0.2027862031054088$ & $0.7143773712418350$ \\
+      12    & $-0.5943293901830616$ & $0.2167029365631842$ & $0.9032588871651854$ \\
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{13 stages, $4^{th}$ order\label{tab:RK-ls-13}}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{cccc}
+      Stage & $A$                   & $B$                  & $C$                  \\
+      \hline
+      1     & $ 0.0000000000000000$ & $0.0271990297818803$ & $0.0000000000000000$ \\
+      2     & $-0.6160178650170565$ & $0.1772488819905108$ & $0.0271990297818803$ \\
+      3     & $-0.4449487060774118$ & $0.0378528418949694$ & $0.0952594339119365$ \\
+      4     & $-1.0952033345276178$ & $0.6086431830142991$ & $0.1266450286591127$ \\
+      5     & $-1.2256030785959187$ & $0.2154313974316100$ & $0.1825883045699772$ \\
+      6     & $-0.2740182222332805$ & $0.2066152563885843$ & $0.3737511439063931$ \\
+      7     & $-0.0411952089052647$ & $0.0415864076069797$ & $0.5301279418422206$ \\
+      8     & $-0.1797084899153560$ & $0.0219891884310925$ & $0.5704177433952291$ \\
+      9     & $-1.1771530652064288$ & $0.9893081222650993$ & $0.5885784947099155$ \\
+      10    & $-0.4078831463120878$ & $0.0063199019859826$ & $0.6160769826246714$ \\
+      11    & $-0.8295636426191777$ & $0.3749640721105318$ & $0.6223252334314046$ \\
+      12    & $-4.7895970584252288$ & $1.6080235151003195$ & $0.6897593128753419$ \\
+      13    & $-0.6606671432964504$ & $0.0961209123818189$ & $0.9126827615920843$ \\
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{14 stages, $4^{th}$ order\label{tab:RK-ls-14}}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{cccc}
+      Stage & $A$                   & $B$                  & $C$                  \\
+      \hline
+      1     & $ 0.0000000000000000$ & $0.0367762454319673$ & $0.0000000000000000$ \\
+      2     & $-0.7188012108672410$ & $0.3136296607553959$ & $0.0367762454319673$ \\
+      3     & $-0.7785331173421570$ & $0.1531848691869027$ & $0.1249685262725025$ \\
+      4     & $-0.0053282796654044$ & $0.0030097086818182$ & $0.2446177702277698$ \\
+      5     & $-0.8552979934029281$ & $0.3326293790646110$ & $0.2476149531070420$ \\
+      6     & $-3.9564138245774565$ & $0.2440251405350864$ & $0.2969311120382472$ \\
+      7     & $-1.5780575380587385$ & $0.3718879239592277$ & $0.3978149645802642$ \\
+      8     & $-2.0837094552574054$ & $0.6204126221582444$ & $0.5270854589440328$ \\
+      9     & $-0.7483334182761610$ & $0.1524043173028741$ & $0.6981269994175695$ \\
+      10    & $-0.7032861106563359$ & $0.0760894927419266$ & $0.8190890835352128$ \\
+      11    & $ 0.0013917096117681$ & $0.0077604214040978$ & $0.8527059887098624$ \\
+      12    & $-0.0932075369637460$ & $0.0024647284755382$ & $0.8604711817462826$ \\
+      13    & $-0.9514200470875948$ & $0.0780348340049386$ & $0.8627060376969976$ \\
+      14    & $-7.1151571693922548$ & $5.5059777270269628$ & $0.8734213127600976$ \\
+    \end{tabular}}
+  \end{subtable}\\
 \end{table}
@@
     \centering
-    \caption{Low storage Runge-Kutta 1 stage}\label{tab:oscillation-ls-rk-1}
+    \caption{1 stage}\label{tab:oscillation-ls-rk-1}
     \resizebox{1.00\textwidth}{!}{%
@@
     \centering
-    \caption{Low storage Runge-Kutta 5 stages}\label{tab:oscillation-ls-rk-5}
+    \caption{5 stages}\label{tab:oscillation-ls-rk-5}
     \resizebox{1.00\textwidth}{!}{%
@@
     \centering
-    \caption{Low storage Runge-Kutta 6 stages}\label{tab:oscillation-ls-rk-s}
+    \caption{6 stages}\label{tab:oscillation-ls-rk-6}
     \resizebox{1.00\textwidth}{!}{%
@@
     \centering
-    \caption{Low storage Runge-Kutta 7 stages}\label{tab:oscillation-ls-rk-7}
+    \caption{7 stages}\label{tab:oscillation-ls-rk-7}
     \resizebox{1.00\textwidth}{!}{%
@@
   \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{12 stages}\label{tab:oscillation-ls-rk-12}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.195E-01    &  0.198E-01    & /             & /             \\
+      2500.0          &  0.175E-02    &  0.177E-02    & 3.48          & 3.48          \\
+      1250.0          &  0.155E-03    &  0.157E-03    & 3.50          & 3.50          \\
+       625.0          &  0.137E-04    &  0.139E-04    & 3.50          & 3.50          \\
+       320.0          &  0.132E-05    &  0.133E-05    & 3.50          & 3.50          \\
+       100.0          &  0.225E-07    &  0.228E-07    & 3.50          & 3.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{13 stages}\label{tab:oscillation-ls-rk-13}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.795E-02    &  0.805E-02    & /             & /             \\
+      2500.0          &  0.703E-03    &  0.712E-03    & 3.50          & 3.50          \\
+      1250.0          &  0.621E-04    &  0.629E-04    & 3.50          & 3.50          \\
+       625.0          &  0.549E-05    &  0.556E-05    & 3.50          & 3.50          \\
+       320.0          &  0.527E-06    &  0.534E-06    & 3.50          & 3.50          \\
+       100.0          &  0.899E-08    &  0.911E-08    & 3.50          & 3.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.45\textwidth}
+    \centering
+    \caption{14 stages}\label{tab:oscillation-ls-rk-14}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      \toprule
+      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
+      \hline
+      5000.0          &  0.849E-02    &  0.860E-02    & /             & /             \\
+      2500.0          &  0.750E-03    &  0.759E-03    & 3.50          & 3.50          \\
+      1250.0          &  0.662E-04    &  0.671E-04    & 3.50          & 3.50          \\
+       625.0          &  0.585E-05    &  0.593E-05    & 3.50          & 3.50          \\
+       320.0          &  0.562E-06    &  0.569E-06    & 3.50          & 3.50          \\
+       100.0          &  0.959E-08    &  0.972E-08    & 3.50          & 3.50          \\
+      \bottomrule
+    \end{tabular}}
+  \end{subtable}\quad%
 \end{table}
@@
   \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-12.png}
+    \caption{12 stages}\label{fig:results-oscillation-ls-runge-kutta-12}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-13.png}
+    \caption{13 stages}\label{fig:results-oscillation-ls-runge-kutta-13}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-14.png}
+    \caption{14 stages}\label{fig:results-oscillation-ls-runge-kutta-14}
+  \end{subfigure}\quad%
   \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta solvers}\label{fig:results-oscillation-ls-runge-kutta}

### 2015-10-13 src/lib/foodie_integrator_low_storage_runge_kutta.f90

@@ -69,13 +69,70 @@ module foodie_integrator_low_storage_runge_kutta
 !<
 !< | ` Stage ` | ` A `               | ` B `              | ` C `              |
 !< |-----------|---------------------|--------------------|--------------------|
-!< | ` 1 `     | `  0.000000000000 ` | ` 0.117322146869 ` | ` 0.000000000000 ` |
+!< | ` 1 `     | `  0              ` | ` 0.117322146869 ` | ` 0              ` |
 !< | ` 2 `     | ` -0.647900745934 ` | ` 0.503270262127 ` | ` 0.117322146869 ` |
 !< | ` 3 `     | ` -2.704760863204 ` | ` 0.233663281658 ` | ` 0.294523230758 ` |
 !< | ` 4 `     | ` -0.460080550118 ` | ` 0.283419634625 ` | ` 0.305658622131 ` |
 !< | ` 5 `     | ` -0.500581787785 ` | ` 0.540367414023 ` | ` 0.582864148403 ` |
 !< | ` 6 `     | ` -1.906532255913 ` | ` 0.371499414620 ` | ` 0.858664273599 ` |
 !< | ` 7 `     | ` -1.450000000000 ` | ` 0.136670099385 ` | ` 0.868664273599 ` |
+!<
+!<#### 12 stages, 4th order
+!< This scheme is a low storage RK(12, 4), by [4].
+!<
+!< | ` Stage ` | ` A `                   | ` B `                  | ` C `                  |
+!< |-----------|-------------------------|------------------------|------------------------|
+!< | ` 1  `    | `  0                  ` | ` 0.0650008435125904 ` | ` 0                  ` |
+!< | ` 2  `    | ` -0.0923311242368072 ` | ` 0.0161459902249842 ` | ` 0.0650008435125904 ` |
+!< | ` 3  `    | ` -0.9441056581158819 ` | ` 0.5758627178358159 ` | ` 0.0796560563081853 ` |
+!< | ` 4  `    | ` -4.3271273247576394 ` | ` 0.1649758848361671 ` | ` 0.1620416710085376 ` |
+!< | ` 5  `    | ` -2.1557771329026072 ` | ` 0.3934619494248182 ` | ` 0.2248877362907778 ` |
[…]

### 2015-10-13 src/tests/oscillation/oscillation.f90

@@ -296,7 +296,7 @@ contains
   !< Test explicit low storage Runge-Kutta class of ODE solvers.
   !---------------------------------------------------------------------------------------------------------------------------------
   integer(I_P), intent(IN)      :: stages            !< Number of stages used: if negative all RK solvers are used.
-  integer, parameter            :: rk_stages=7       !< Runge-Kutta stages number.
+  integer, parameter            :: rk_stages=14      !< Runge-Kutta stages number.
   real(R_P), allocatable        :: solution(:,:)     !< Solution at each time step.
   integer(I_P)                  :: s                 !< RK stages counter.
   integer(I_P)                  :: stages_range(1:2) !< Stages used.
@@ -313,6 +313,10 @@ contains
     if (s==2) cycle ! 2 stages not yet implemented
     if (s==3) cycle ! 3 stages not yet implemented
     if (s==4) cycle ! 4 stages not yet implemented
+    if (s==8) cycle ! 8 stages not yet implemented
+    if (s==9) cycle ! 9 stages not yet implemented
+    if (s==10) cycle ! 10 stages not yet implemented
+    if (s==11) cycle ! 11 stages not yet implemented
     print "(A)", ' RK-'//trim(str(.true.,s))
     title = 'Oscillation equation integration, explicit low storage Runge-Kutta, t='//str(n=t_final)//&
       ' steps='//trim(str(.true., s))
@@ -599,7 +603,7 @@ contains
                                                       100._R_P] !< Time steps exercised.
   integer, parameter            :: ab_steps=4                   !< Adams-Bashforth steps number.
   integer, parameter            :: tvd_rk_stages=5              !< TVD/SSP Runge-Kutta stages number.
-  integer, parameter            :: ls_rk_stages=7               !< Low storage Runge-Kutta stages number.
+  integer, parameter            :: ls_rk_stages=14              !< Low storage Runge-Kutta stages number.
   integer(I_P)                  :: s                            !< Steps/stages counter.
   integer(I_P)                  :: d                            !< Time steps-exercised counter.
   character(len=:), allocatable :: output                       !< Output files basename.
@@ -691,6 +695,10 @@ contains
       if (s==2) cycle ! 2 stages not yet implemented
       if (s==3) cycle ! 3 stages not yet implemented
[…]

### 2015-10-14 papers/announcement/manuscript.tex
Minor improvments on manuscript

--- 2015-10-09 papers/announcement/manuscript.tex
@@
 \begin{abstract}
-To be written.
+The (numerical) solution of Ordinary Differential Equations (ODEs) problems is of paramount relevance, ODEs system being an ubiquitous mathematical formulation of many physical phenomena (such as those involved in fluid dynamics, chemistry, biology, evolutionary-anthropology, \dots).
+The present paper is the first \emph{manifesto} of FOODiE, a library aimed to numerically solve ODEs problems by means of a clear, concise and efficient \emph{abstract} interface.
+FOODiE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has manifolds aims: to provide a set to built-in numerical schemes that are accurate, robust, validated and efficient and to allow easy application of these schemes to (almost) all ODEs problems by means of an effective Abstract Calculus Pattern.
+The key idea is to allow the same solver-implementation to be applied to all ODEs problems thus avoiding the re-implementation of the ODEs solver for each different ODEs problem: code re-usability is consequently maximized, FOODiE being a general robust framework.
+Besides, the same framework also allows rapid development of new ODEs solvers due to the high abstraction level of the library itself.
+
+FOODiE is a modern Fortran library which main features are:
+\begin{description}
+  \item[Free] FOODiE is a free software;
+  \item[OOP] FOODiE is based on Object Oriented Programming paradigm;
+  \item[TDD] the FOODiE development follows the Test Driven Development software process;
+  \item[Accurately documented] the FOODiE documentation is based on high quality, first class solutions embedding detailed (mathematical) descriptions directly inside code sources; moreover, comprehensive hyper-linked documentations is also provided;
+  \item[Comphrensive] FOODiE provides a comprehensive set of built-in solvers, namely one-step, multi-steps, explicit, implicit, predictor-corrector solvers; moreover, its Abstract Calculus Pattern allows the solution of virtually all kinds of ODEs problems;
+  \item[Collaborative] the development of FOODiE takes advantage of web communications, the main project being hosted on GitHub.
+\end{description}
+
+The present paper is the first announcement of FOODiE project: the current implementation is extensively discussed and its capabilities are proved by means of tests and examples.
 \end{abstract}
@@
   Abstract Calculus Pattern (ACP) \sep
-  Fortran \sep
+  Fortran
 \end{keyword}
@@
 
-\section{Introduction}\label{sec:introduction}
-\subsection{Background}
-
-Initial Value Problem (IVP, or Cauchy problem) constitutes a class of mathematical models of paramount relevance, it being applied to the modelling of a wide range of problems.
-Briefly, an IVP is an Ordinary Differential Equations system (ODE) coupled with specified initial values of the unknown state variables, the solution of which are searched at a given time after the initial time considered.
-
-The prototype of IVP can be expressed as:
-\begin{equation}
-  \begin{matrix}
-  U_t = R(t,U) \\
-  U_0 = U(t_0)
-  \end{matrix}
-\label{eq:IVP}
-\end{equation}
-where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residuals function and $U(t_0)$ is the (vectorial) initial conditions, namely the state variables function evaluated at the initial time $t_0$.
-In general, the residuals function $R$ is a function of the state variable $U$ through which it is a function of time, but it can be also a direct function of time, thus in general $R=R(t,U(t))$ holds.
-
-The problem prototype \ref{eq:IVP} is ubiquitous in the mathematical modelling of physical problems: essentially whenever an \emph{evolutionary} phenomenon is considered the prevision (simulation) of the future solutions involves the solution of an IVP.
-As a matter of facts, many physical problems (fluid dynamics, chemistry, biology, evolutionary-anthropology, etc...) are described by means of an IVP.
-
-It is worth noting that the state vector variables $U$ and its corresponding residuals function $U_t = \frac{dU}{dt} = R(t,U)$ are \emph{problem dependent}: the number and the meaning of the state variables as well as the equations governing their evolution (that are embedded into the residuals function) are different for the Navier-Stokes conservation laws with respect the Burgers one, as an example.
-Nevertheless, the \emph{solvers} used for the prediction of the Navier-Stokes equations evolution (hereafter the \emph{solution}) are the same that are used for Burgers equations time-integration.
-As a consequence, the solution of the IVP model prototype can be generalized, allowing the application of the same solver to many different problems, thus eliminating the necessity to re-implement the same solver for each different problem.
-
-FOODiE library is designed for solving the generalized IVP \ref{eq:IVP}, it being completely unaware of the actual problem's definition.
-FOODiE library provides a high-level, well-documented, simple Application Program Interface (API) for many well-known ODE integration solvers, its aims being twofold:
-
-\begin{itemize}
-  \item provide a robust set of ODE solvers ready to be applied to a wide range of different problems;
-  \item provide a simple framework for the rapid development of new ODE solvers.
-\end{itemize}
-
-{\color{red} Add citations to IVP, Cauchy, ODE references.}
-
-\subsection{Related Works}
-
-{\color{red} To be written.}
-
-\subsection{Motivations and Aims}
-
-FOODiE library must:
-
-\begin{itemize}
-  \item be written in modern Fortran (standard 2008 or newer);
-  \item be written by means of Object Oriented Programming (OOP) paradigm;
-  \item be well documented;
-  \item be Tests Driven Developed (TDD);
-  \item be collaboratively developed;
-  \item be free.
-\end{itemize}
-
-FOODiE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has been developed with the aim to satisfy the above specifications.
-The present paper is its first comprehensive presentation.
-
-The Fortran (Formula Translator) programming language is the \emph{de facto} standard into computer science field: it strongly facilitates the effective and efficient translation of (even complex) mathematical and numerical models into an operative software without compromise on computations speed and accuracy.
-Moreover, its simple syntax is suitable for scientific researchers that are interested (and skilled) in the physical aspects of the numerical computations rather than computer technicians.
-Consequently, we develop FOODiE using Fortran language: FOODiE is written by research scientists for research scientists.
-
-One key-point of the FOODiE development is the \emph{problem generalization}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its actual definitions.
-Consequently, we must rely on a \emph{generic} implementation of the solvers.
-To this aim, OOP is very useful: it allows to express IVP \ref{eq:IVP} in a very concise and clear formulation that is really generic.
-In particular, our implementation is based on \emph{Abstract Calculus Pattern} (ACP) concept.
-
-\paragraph{The Abstract Calculus Pattern}
-The abstract calculus pattern provides a simple solution for the connection between the very high-level expression of IVP \ref{eq:IVP} and the eventual concrete (low-level) implementation of the ODE problem being solved.
-ACP essentially constitutes a \emph{contract} based on an Abstract Data Type (ADT): we specify an ADT supporting a certain set of mathematical operators (differential and integral ones) and implement FOODiE solvers only on the basis of this ADT.
-FOODiE clients must formulate the ODE problem under integration defining their own concrete extensions of our ADT (implementing all the deferred operators).
-Such an approach defines the abstract calculus pattern: FOODiE solvers are aware of only the ADT, while FOODiE clients extend the ADT for defining the concrete ODE problem.
-
-Is is worth noting that this ACP emancipates the solvers implementations from any low-level problem-dependent details: the ODE solvers developed with this pattern are extremely concise, clear, maintainable and less errors-prone with respect a low-level (non abstract) pattern.
-Moreover, the FOODiE clients can use solvers being extremely robust: as a matter of facts, FOODiE solvers are expressed in a formulation very close to the mathematical one and are tested on an extremely varying family of problems.
-As shown in the following, such a great flexibility does not compromise the computational efficiency.
-
-The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODiE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODiE applications is presented; section \ref{sec:parallel} provides an analysis of FOODiE performances under parallel frameworks scenario like the OpenMP and MPI paradigms; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.
-
-{\color{red} Add citations to Fortran standards, OOP, TDD, free software, ACP, ADT, Rouson's book references.}
-
-\section{Mathematical and Numerical Models}\label{sec:MNmodels}
-
-In many (most) circumstances, the solution of equation \ref{eq:IVP} cannot be computed in a closed, exact form (even if it exists and is unique) due to the complexity and nature of the residuals functions, that is often non linear.
-Consequently, the problem is often solved relying on a numerical approach: the solution of system \ref{eq:IVP} at a time $t^n$, namely $U(t^n)$, is approximated by a subsequent time-marching approximations $U_0=u_0 \rightarrow u_1 \rightarrow u_2 \rightarrow ...
-\rightarrow u_N\approx U(t^n)$ where the relation $u_i \rightarrow u_{i+1}$ implies a \emph{stepping, numerical integration} from the time $t^i$ to time $t^{i+1}$ and $N$ is the total number of numerical time steps necessary to evolve the initial conditions toward the searched solution $U(t^n)$.
-To this aim, many numerical schemes have been devised.
-Notably, the numerical schemes of practical usefulness must posses some necessary proprieties such as \emph{consistency} and \emph{stability} to ensure that the numerical approximation \emph{converges} to the exact solution as the numerical time step tends to zero.
-A detailed discussion of these details is out the scope of the present work and is omitted.
-Here, we briefly recall some classifications necessary to introduce the schemes implemented into the FOODiE library.
-
-A non comprehensive classification of the most widely used schemes could distinguish between \emph{multi-step} versus \emph{one-step} schemes and between \emph{explicit} versus \emph{implicit} schemes.
-
-Essentially, the multi-step schemes have been developed to obtain an accurate approximation of the subsequent numerical steps using the informations contained into the previously computed steps, thus this approach relates the next step approximation to a set of the previously computed steps.
-On the contrary, a one-step scheme evolves the solution toward the next step using only the information coming from the current time approximation.
-In the framework of one-step schemes family an equivalent accurate approximation can be obtained by means of a multi-stage approach as the one due to Runge-Kutta.
-The current version of FOODiE provides schemes belonging to both these families.
-
-The other ODE solvers classification concerns with explicit or implicit nature of the schemes employed.
-Briefly, an explicit scheme computes the next step approximation using the previously computed steps at most to the current time, whereas an implicit scheme uses also the next step approximation (that is the unknown), thus it requires extra computations.
-The implicit approach is of practical use for \emph{stiff} systems where the usage of explicit schemes could require an extremely small time step to evolve in a \emph{stable} way the solution.
-Mixing together explicit and implicit schemes it is possible to build a family of \emph{predictor-corrector} methods: using an explicit scheme to predict a guess for the next step approximation it is possible to use an implicit method for correcting this guess.
-Currently, FOODiE provides explicit solvers and predictor-correct ones.
-
-FOODiE currently implements the following ODE schemes:
-
-\begin{itemize}
-  \item one-step schemes:
-    \begin{itemize}
-      \item explicit forward Euler scheme, it being $1^{st}$ order accurate;
-      \item explicit Runge-Kutta schemes:
-        \begin{itemize}
-          \item TVD/SSP Runge-Kutta schemes:
-            \begin{itemize}
-              \item 2-stages, it being $2^{nd}$ order accurate;
-              \item 3-stages, it being $3^{rd}$ order accurate;
-              \item 5-stages, it being $4^{th}$ order accurate;
-              \end{itemize}
-          \item low storage Runge-Kutta schemes:
-            \begin{itemize}
-              \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \end{itemize}
-          \end{itemize}
-      \end{itemize}
-  \item multi-step schemes:
-    \begin{itemize}
-      \item explicit Adams-Bashforth schemes:
-        \begin{itemize}
-          \item 2-steps, it being $2^{nd}$ order accurate;
-          \item 3-steps, it being $3^{rd}$ order accurate;
-          \item 4-steps, it being $4^{th}$ order accurate;
-          \end{itemize}
-      \item implicit Adams-Moulton schemes:
-        \begin{itemize}
-          \item 1-step, it being $2^{nd}$ order accurate;
-          \item 2-steps, it being $3^{rd}$ order accurate;
-          \item 3-steps, it being $4^{th}$ order accurate;
-          \end{itemize}
-      \item predictor-corrector Adams-Bashforth-Moulton schemes:
-        \begin{itemize}
-          \item 1-step, it being $2^{nd}$ order accurate;
-          \item 2-steps, it being $3^{rd}$ order accurate;
-          \item 3-steps, it being $4^{th}$ order accurate;
-          \end{itemize}
-      \item explicit Leapfrog schemes\footnote{Leapfrog method is a 2-steps solver being, in general $2^{nd}$ order accurate, but mostly \emph{unstable}.
-It is well suited for periodic/oscillatory problems, thus the error is generally categorized accordingly to the phase and amplitude errors: the unfiltered scheme provides a $2^{nd}$ order error on phase approximation while the amplitude error is null.
-However, the unfiltered scheme is mostly unstable, thus many filters have been devised.
-In general, the filters retain the accuracy on the phase error, while affect the amplitude one.}:
-        \begin{itemize}
-          \item 2-steps unfiltered, it being $2^{nd}$ order accurate, but mostly \emph{unstable};
-          \item 2-steps Robert-Asselin filtered, it being $1^{st}$ order accurate (on \emph{amplitude error});
-          \item 2-steps Robert-Asselin-Williams filtered, it being $3^{rd}$ order accurate (on \emph{amplitude error});
-          \end{itemize}
-      \end{itemize}
-  \end{itemize}
-
-{\color{red} Add citations to all solvers reference papers.}
-
-\subsection{The explicit forward Euler scheme}
-
-The explicit forward Euler scheme for ODE integration is probably the simplest solver ever devised.
-Considering the system \ref{eq:IVP}, the solution (approximation) of the state vector $U$ at the time $t^{n+1}=t^n+\Delta t$ assuming to known the solution at time $t^n$ is:
-
-\begin{equation}
-  U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]
-\label{eq:solver-euler}
-\end{equation}
-where the solution at the new time step is computed by means of only the current time solution, thus this is an explicit scheme.
-The solution is an approximation of $1^{st}$ order, the local truncation error being $O(Dt^2)$.
-As well known, this scheme has an absolute (linear) stability locus equals to $|1+\Delta t \lambda|\le 1$ where $\lambda$ contains the eigenvalues of the linear (or linearized) Jacobian matrix of the system.
-
-This scheme is Total Variation Diminishing (TVD), thus satisfies the maximum principle (or the equivalent positivity preserving property).
-
-{\color{red} Add citations.}
-
-\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
-
-Runge-Kutta methods belong to the more general multi-stage family of schemes.
-This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme, but without increasing the number of time steps used, as it is done with the multi-step schemes.
-Essentially, the high order of accuracy is obtained by means of \emph{intermediate values} (the stages) of the solution and its derivative are generated and used within a single time step.
-This commonly implies the allocation of some auxiliary memory registers for storing the intermediate stages.
-
-Notably, the multi-stage schemes class has the attractive property to be \emph{self-starting}: the high order accurate solution can be obtained directly from the previous one, without the necessity to compute \emph{before} a certain number of previous steps, as it happens for the multi-step schemes.
-Moreover, one-step multi-stage methods are suitable for adaptively-varying time-step size (that is also possible for multi-step schemes, but at a cost of more complexity) and for discontinuous solutions, namely discontinued solutions happening at a certain time $t*$ (that in a multi-step framework can involve an overall accuracy degradation).
-
-In general, the TVD/SSP Runge-Kutta schemes provided by FOODiE library are written by means of the following algorithm:
-\begin{equation}
-  U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
-\label{eq:RK}
-\end{equation}
-where $N_s$ is the number of Runge-Kutta stages used and $K_s$ is the $s^{th}$ stage defined as:
-\begin{equation}
-  K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right)
-\label{eq:RK-stage}
-\end{equation}
-It is worth noting that the equations \ref{eq:RK} and \ref{eq:RK-stage} contain also implicit schemes.
-A scheme belonging to this family is operative once the coefficients $\alpha,\, \beta,\, \gamma$ are provided.
-We represent these coefficients using the Butcher's table, that for an explicit scheme where $\gamma_1=\alpha_{1,*}=\alpha_{i,i}=0$ has the form reported in table \ref{tab:butcher-table}.
-
-\begin{table}[!ht]
-\caption{Butcher's table for explicit Runge-Kutta schemes}\label{tab:butcher-table}
-\centering
-\resizebox{0.40\textwidth}{!}{%
-\begin{tabular}{c|ccccc}
- $\gamma_2$     & $\alpha_{2,1}$   &                  &          &                      &              \\
- $\gamma_3$     & $\alpha_{3,1}$   & $\alpha_{3,2}$   &          &                      &              \\
- $\vdots$       & $\vdots$         &                  & $\ddots$ &                      &              \\
- $\gamma_{N_s}$ & $\alpha_{N_s,1}$ & $\alpha_{N_s,2}$ & $\cdots$ & $\alpha_{N_s,N_s-1}$ &              \\
- \hline
-                & $\beta_1$        & $\beta_2$        & $\cdots$ & $\beta_{N_s-1}$      & $\beta_{N_s}$\\
-\end{tabular}}
-\end{table}
-
-The equations \ref{eq:RK} and \ref{eq:RK-stage} show that Runge-Kutta methods do not require any additional differentiations of the ODE system for achieving high order accuracy, rather they require additional evaluations of the residuals function $R$.
-
-The nature of the scheme and the properties of the solutions obtained depend on the number of stages and on the value of the coefficients selected.
-Currently, FOODiE provides 3 Runge-Kutta schemes having TVD or Strong Stability Preserving (SSP) propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Butcher's table of 2 stages, $2^{nd}$ order, Runge-Kutta TVD scheme\label{tab:RK-TVD-butcher-2}}
-  \resizebox{0.15\textwidth}{!}{%
-  \begin{tabular}{c|cc}
-   1 & 1 & 0       \\
-   \hline
-       & 1/2 & 1/2   \\
-  \end{tabular}}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \caption{Butcher's table of 3 stages, $3^{rd}$ order, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-3}}
-  \resizebox{0.20\textwidth}{!}{%
-  \begin{tabular}{c|ccc}
-   1   & 1   &     &      \\
-   1/2 & 1/4 & 1/4 &      \\
-   \hline
-       & 1/6 & 1/6 & 2/3  \\
-  \end{tabular}}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \caption{Butcher's table of 5 stages, $4^{th}$ order accurate, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-5}}
-  \resizebox{0.99\textwidth}{!}{%
-  \begin{tabular}{c|ccccc}
-   0.39175222700392 & 0.39175222700392 &                  &                  &                  &                  \\
-   0.58607968896779 & 0.21766909633821 & 0.36841059262959 &                  &                  &                  \\
-   0.47454236302687 & 0.08269208670950 & 0.13995850206999 & 0.25189177424738 &                  &                  \\
-   0.93501063100924 & 0.06796628370320 & 0.11503469844438 & 0.20703489864929 & 0.54497475021237 &                  \\
-   \hline
-                    & 0.14681187618661 & 0.24848290924556 & 0.10425883036650 & 0.27443890091960 & 0.22600748319395 \\
-  \end{tabular}}
-\end{table}
-The absolute stability locus depends on the coefficients selected, however, as a general principle, we can assume that greater is the stages number and wider is the stability locus on equal accuracy orders.
-
-It is worth noting that FODDiE also provides a one-stage TVD Runge-Kutta solver that reverts back to the explicit forward Euler scheme: it can be used, for example, into a Recursive Order Reduction (ROR) framework that automatically checks some properties of the solution and, in case, reduces the order of the Runge-Kutta solver until those properties are obtained.
-
-{\color{red} Add citations.}
-
-\subsection{The explicit low storage Runge-Kutta class of schemes}
-
-As aforementioned, standard Runge-Kutta schemes have the drawback to require $N_S$ auxiliary memory registers to store the necessary stages data.
-In order to make an efficient use of the available limited computer memory, the class of low storage Runge-Kutta scheme was devised.
-Essentially, the standard Runge-Kutta class (under some specific conditions) can be reformulated allowing a more efficient memory management.
-Currently FOODiE provides a class of \emph{2N registers} storage Runge-Kutta schemes, meaning that the storage of all stages requires only 2 registers of memory with a \emph{word} length $N$ (namely the length of the state vector) in contrast to the standard formulation where $N_s$ registers of the same length $N$ are required.
-This is a dramatic improvement of memory efficiency especially for schemes using a high number of stages ($N_s \ge 4$) where the memory necessary is an half with respect the original formulation.
-Unfortunately, not all standard Runge-Kutta schemes can be reformulated as a low storage one.
-
-Following the Williamson's approach the standard coefficients are reformulated to the coefficients vectors $A$, $B$ and $C$ and the Runge-Kutta algorithm becomes:
-
-\begin{equation}
-\begin{matrix}
-  K_1 = U\left(t^n\right) \\
-  K_2 = 0 \\
-  \left.\begin{matrix}
-    K_2 = A_s K_2 + \Delta t \cdot R\left(t^n + C_s \Delta t, K_1\right) \\
-    K_1 = K_1 + B_s K_2
-  \end{matrix}\right\} s=1,2,...N_s\\
-  U\left(t^{n+1}\right) = K_1
-  \end{matrix}
-\label{eq:RK-ls}
-\end{equation}
-
-Currently FOODiE provides only one 5 stages, $4^{th}$ order, 2N registers explicit scheme, the coefficients of which are listed in table \ref{tab:RK-ls-5}.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Williamson's table of 5 stages, $4^{th}$ order, Runge-Kutta low storage scheme\label{tab:RK-ls-5}}
-  \resizebox{0.40\textwidth}{!}{%
-  \begin{tabular}{ccc}
-       $A$                                    & $B$                                    & $C$                                   \\
-    \hline
-       $ 0                                  $ & $\frac{1432997174477}{9575080441755 }$ & $ 0                                 $ \\
-       $-\frac{567301805773 }{1357537059087}$ & $\frac{5161836677717}{13612068292357}$ & $\frac{1432997174477}{9575080441755}$ \\
-       $-\frac{2404267990393}{2016746695238}$ & $\frac{1720146321549}{2090206949498 }$ & $\frac{2526269341429}{6820363962896}$ \\
-       $-\frac{3550918686646}{2091501179385}$ & $\frac{3134564353537}{4481467310338 }$ & $\frac{2006345519317}{3224310063776}$ \\
-       $-\frac{1275806237668}{842570457699 }$ & $\frac{2277821191437}{14882151754819}$ & $\frac{2802321613138}{2924317926251}$ \\
-  \end{tabular}}
-\end{table}
-
-Similarly to the TVD/SSP Runge-Kutta class, the low storage class also provides a fail-safe one-stage solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
-
-{\color{red} Add citations.}
-
-\subsection{The explicit Adams-Bashforth class of schemes}
-
-Adams-Bashforth methods belong to the more general (linear) explicit multi-step family of schemes.
-This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
-Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
-
-In general, the Adams-Bashforth schemes provided by FOODiE library are written by means of the following algorithm (for only explicit schemes):
-
-\begin{equation}
-U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}
-\label{eq:AB}
-\end{equation}
-where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.
-
-Currently FOODiE provides 2, 3, and 4 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Explicit Adams-Bashforth coefficients\label{tab:AB-coeff}}
-  \resizebox{0.30\textwidth}{!}{%
-  \begin{tabular}{c|cccc}
-    $N_s$ & $b_1$           & $b_2$            & $b_3$            & $b_4$           \\
-    \hline
-    2     & $-\frac{1}{2}$  & $\frac{3}{2}$    &       /          &       /         \\
-    3     & $\frac{5}{12}$  & $-\frac{16}{12}$ & $\frac{23}{12}$  &       /         \\
-    4     & $-\frac{9}{24}$ & $\frac{37}{24}$  & $-\frac{59}{24}$ & $\frac{55}{24}$ \\
-  \end{tabular}}
-\end{table}
-
-Similarly to the Runge-Kutta classes, the Adams-Bashforth class also provides a fail-safe one-step solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
-
-It is worth noting that for $N_s>1$ the Adams-Bashforth class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided.
-To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
-
-{\color{red} Add citations.}
-
-\subsection{The implicit Adams-Moulton class of schemes}
-
-Adams-Moulton methods belong to the more general (linear) implicit multi-step family of schemes.
-This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
-Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
-
-In general, the Adams-Moulton schemes provided by FOODiE library are written by means of the following algorithm (for only implicit schemes):
-
-\begin{equation}
-  U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=0}^{n+N_s-1}{ b_s \cdot R\left[t^{n+s}, U\left(t^{n+s}\right)\right]} + b_{N_s}\cdot R\left[t^{n+N_s}, U\left(t^{n+N_s}\right)\right]
-\label{eq:AM}
-\end{equation}
-where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.
-
-Currently FOODiE provides 1, 2, and 3 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Implicit Adams-Moulton coefficients\label{tab:AM-coeff}}
-  \resizebox{0.30\textwidth}{!}{%
-  \begin{tabular}{c|cccc}
-    $N_s$ & $b_0$           & $b_1$           & $b_2$           & $b_3$          \\
-    \hline
-    1     & $\frac{1}{2}$   & $\frac{1}{2}$   &       /         &       /        \\
-    2     & $-\frac{1}{12}$ & $\frac{8}{12}$  & $\frac{5}{12}$  &       /        \\
-    3     & $\frac{1}{24}$  & $-\frac{5}{24}$ & $\frac{19}{24}$ & $\frac{9}{24}$ \\
-  \end{tabular}}
-\end{table}
-
-Similarly to the Runge-Kutta and Adams-Bashforth classes, the Adams-Moulton class also provides a fail-safe zero-step solver reverting back to the implicit backward Euler solver, that is useful for ROR-like frameworks.
-
-It is worth noting that for $N_s>1$ the Adams-Moulton class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided.
-To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
-
-{\color{red} Add citations.}
-
-\subsection{The predictor-corrector Adams-Bashforth-Moulton class of schemes}
-
-Adams-Bashforth-Moulton methods belong to the more general (linear) predictor-corrector multi-step family of schemes.
-This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
-Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
-
-In general, the Adams-Bashforth-Moulton schemes provided by FOODiE library are written by means of the following algorithm:
-
-\begin{equation}
-  \begin{matrix}
-    U\left(t^{n+N_s^p}\right)_p = U\left(t^{n+N_s^p-1}\right) +\Delta t \sum_{s=1}^{n+N_s^p}{ b_s^p \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}   \\
-    U\left(t^{n+N_s^c}\right)_c = U\left(t^{n+N_s^c-1}\right) +\Delta t \sum_{s=0}^{n+N_s^c-1}{ b_s^c \cdot R\left[t^{n+s}, U\left(t^{n+s}\right)\right]} + b_{N_s^c}^c\cdot R\left[t^{n+N_s^p}, U\left(t^{n+N_s^p}\right)_p\right]
-  \end{matrix}
-\label{eq:ABM}
-\end{equation}
-where $N_s^{p,c}$ is the number of time steps considered for the Adams-Bashforth predictor/Adams-Moulton corrector (respectively) and $b_s^{p,c}$ are the corresponding linear coefficients selected.
-Essentially, the Adams-Bashforth prediction $U\left(t^{n+N_s^p}\right)_p$ is corrected by means of the Adams-Moulton correction resulting in $U\left(t^{n+N_s^c}\right)_c$.
-In order to preserve the formal order of accuracy the relation $N_s^p=N_s^c+1$ always holds.
-
-Currently FOODiE provides $N_s^c=1, 2, 3 \rightarrow N_s^c=2, 3, 4$ steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively.
-The $b_s^{p,c}$ coefficients are those reported in tables \ref{tab:AB-coeff} and \ref{tab:AM-coeff}.
-
-{\color{red} Add citations.}
-
-\subsection{The leapfrog solver}
-
-The \emph{leapfrog} scheme belongs to the multi-step family, it being formally a centered second order approximation in time.
-The leapfrog method (in its original formulation) is mostly unstable, however it is well suited for periodic-oscillatory problems providing a null error on the amplitude value and a formal second order error on the phase one, under the satisfaction of the time-step size stable limit.
-Commonly, the leapfrog methods are said to provide a $2 \Delta t$ computational mode that can generate unphysical, unstable solutions.
-As consequence, the original leapfrog scheme is generally \emph{filtered} in order to suppress these computational modes.
-
-The unfiltered leapfrog scheme provided by FOODiE is:
-\begin{equation}
-  U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]
-\label{eq:leapfrog}
-\end{equation}
-
-FOODiE provides, in a \emph{seamless} API, also filtered leapfrog schemes.
-A widely used filter is due to Robert and Asselin, that suppress the computational modes at the cost of accuracy reduction resulting into a $1^{st}$ order error in amplitude value.
-A more accurate filter, able to provide a $3^{rd}$ order error on amplitude, is a modification of the Robert-Asselin filter due to Williams known as Robert-Asselin-Williams (RAW) filter, that filters the approximation of $U\left(t^{n+1}\right)$ and $U\left(t^{n+2}\right)$ by the following scalar coefficient:
-
-\begin{equation}
-  \begin{matrix}
-    U\left(t^{n+1}\right) = U\left(t^{n+1}\right) + \Delta * \alpha     \\
-    U\left(t^{n+2}\right) = U\left(t^{n+2}\right) + \Delta * (\alpha-1) \\
-    where \\
-    \Delta = \frac{\nu}{2}(U^{n} - 2 U^{n+1} + U^{n+2})
-  \end{matrix}
-\label{eq:leapfrog-RAW}
-\end{equation}
-
-The filter coefficients should be taken as $\nu \in (0,1]$ and $\alpha \in (0.5,1]$.
-If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude.
-The default values of the FOODiE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.
-
-{\color{red} Add citations.}
+\input{src/introduction/introduction.tex}
 
@@
 
-\section{Application Program Interface}\label{sec:API}
-
-In this section we review the FOODiE API providing a detailed discussion of the implementation choices.
-
-As aforementioned, the programming language used is the Fortran 2008 standard, that is a minor revision of the previous Fortran 2003 standard.
-Such a new Fortran idioms provide (among other useful features) an almost complete support for OOP, in particular for ADT concept.
-Fortran 2003 has introduced the \emph{abstract derived type}: it is a derived type suitable to serve as \emph{contract} for concrete type-extensions that has not any actual implementations, rather it provides a well-defined set of type bound procedures interfaces, that in Fortran nomenclature are called \emph{deferred} procedures.
-Using such an abstract definition, we can implement algorithms operating on only this abstract type and on \emph{all its concrete extensions}.
-This is the key feature of FOODiE library: all the above described ODE solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax.
-In the meanwhile, client codes must implement their own IVPs extending only one simple abstract type.
-
-In the subsection \ref{subsec:type_integrand} a review of the FOODiE main ADT, the \emph{integrand} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.
-
-It is worth noting that all FOODiE public \emph{entities} (ADT and solvers) must be accessed by the FOODiE module, see listing \ref{list:main_module} for an example on how access to all public FOODiE entities.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={usage example importing all public entities of FOODiE main module},label={list:main_module}]
-use foodie, only: integrand, &
-                  adams_bashforth_integrator, &
-                  adams_moulton_integrator, &
-                  adams_bashforth_moulton_integrator, &
-                  euler_explicit_integrator, &
-                  leapfrog_integrator, &
-                  ls_runge_kutta_integrator, &
-                  tvd_runge_kutta_integrator
-! or simply
-use foodie
-\end{lstlisting}
-
-\subsection{The main FOODiE Abstract Data Type: the integrand type}\label{subsec:type_integrand}
-
-The implemented ACP is based on one main ADT, the \emph{integrand} type, the definition of which is shown in listing~\ref{list:type_integrand}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={integrand type definition},label={list:type_integrand}]
-type, abstract :: integrand
-  !< Abstract type for building FOODiE ODE integrators.
-  contains
-    ! public deferred procedures that concrete integrand-field must implement
-    procedure(time_derivative),       pass(self), deferred, public:: t
-    ! operators
-    procedure(symmetric_operator),   pass(lhs), deferred, public:: integrand_multiply_integrand
-    procedure(integrand_op_real),    pass(lhs), deferred, public:: integrand_multiply_real
-    procedure(real_op_integrand),    pass(rhs), deferred, public:: real_multiply_integrand
-    procedure(symmetric_operator),   pass(lhs), deferred, public:: add
-    procedure(symmetric_operator),   pass(lhs), deferred, public:: sub
-    procedure(assignment_integrand), pass(lhs), deferred, public:: assign_integrand
-    ! operators overloading
-    generic, public:: operator(+) => add
-    generic, public:: operator(-) => sub
-    generic, public:: operator(*) => integrand_multiply_integrand, &
-                                     real_multiply_integrand, &
-                                     integrand_multiply_real
-    generic, public:: assignment(=) => assign_integrand
-endtype integrand
-\end{lstlisting}
-
-The \emph{integrand} type does not implement any actual integrand field, it being and abstract type.
-It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODiE solver.
-
-As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 1 ODE-specific procedure plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects.
-These procedures are analyzed in the following paragraphs.
-
-\subsubsection{Time derivative procedure, the residuals function}
-
-The abstract interface of the time derivative procedure \emph{t} is shown in listing \ref{list:time_derivative}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={time derivative procedure interface},label={list:time_derivative}]
-function time_derivative(self, t) result(dState_dt)
-import :: integrand, R_P, I_P
-class(integrand),       intent(IN) :: self      !< Integrand field.
-real(R_P),    optional, intent(IN) :: t         !< Time.
-class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.
-endfunction time_derivative
-\end{lstlisting}
-
-This procedure-function takes two arguments, the first passed as a \emph{type bounded} argument, while the latter is optional, and it returns an integrand object.
-The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
-The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.
-
-Commonly, into the concrete implementation of this deferred abstract procedure clients embed the actual ODE equations being solved.
-As an example, for the Burgers equation, that is a Partial Differential Equations (PDE) system involving also a boundary value problem, this procedure embeds the spatial operator that convert the PDE to a system of algebraic ODE.
-As a consequence, the eventual concrete implementation of this procedure can be very complex and errors-prone.
-Nevertheless, the FOODiE solvers are implemented only on the above abstract interface, thus emancipating the solvers implementation from any concrete complexity.
-
-{\color{red} Add citations to Burgers, Adams-Bashfort, leapfrog references.}
-
-\subsubsection{Symmetric operators procedures}
-
-The abstract interface of \emph{symmetric} procedures is shown in listing \ref{list:symmetric_operator}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={symmetric operator procedure interface},label={list:symmetric_operator}]
-function symmetric_operator(lhs, rhs) result(operator_result)
-import :: integrand
-class(integrand), intent(IN)  :: lhs             !< Left hand side.
-class(integrand), intent(IN)  :: rhs             !< Right hand side.
-class(integrand), allocatable :: operator_result !< Operator result.
-endfunction symmetric_operator
-\end{lstlisting}
-
-This interface defines a class of procedures operating on 2 \emph{integrand} objects, namely it is used for the definition of the operators \emph{multiplication}, \emph{summation} and \emph{subtraction} of integrand objects.
-These operators are used into the above described ODE solvers, for example see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
-The implementation details of such a procedures class are strictly dependent on the concrete extension of the integrand type.
-From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are two integrand object and the returned object is still an integrand one.
-
-\subsubsection{Integrand/real and real/integrand operators procedures}
-
-The abstract interfaces of \emph{Integrand/real and real/integrand operators} procedures are shown in listing \ref{list:integrand_real_ops}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={Integrand/real and real/integrand operators procedure interfaces},label={list:integrand_real_ops}]
-function integrand_op_real(lhs, rhs) result(operator_result)
-import :: integrand, R_P
-class(integrand), intent(IN)  :: lhs              !< Left hand side.
-real(R_P),        intent(IN)  :: rhs              !< Right hand side.
-class(integrand), allocatable :: operator_result  !< Operator result.
-endfunction integrand_op_real
-
-function real_op_integrand(lhs, rhs) result(operator_result)
-import :: integrand, R_P
-real(R_P),        intent(IN)  :: lhs             !< Left hand side.
-class(integrand), intent(IN)  :: rhs             !< Right hand side.
-class(integrand), allocatable :: operator_result !< Operator result.
-endfunction real_op_integrand
-\end{lstlisting}
-
-These two interfaces are necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the multiplication of an integrand object for a real number, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
-The implementation details of these procedures are strictly dependent on the concrete extension of the integrand type.
-From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are an integrand object and real number of viceversa and the returned object is still an integrand one.
-
-\subsubsection{Integrand assignment procedure}
-
-The abstract interface of \emph{integrand assignment} procedure is shown in listing \ref{list:integrand_assignment}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={integrand assignment procedure interface},label={list:integrand_assignment}]
-subroutine assignment_integrand(lhs, rhs)
-import :: integrand
-class(integrand), intent(INOUT) :: lhs !< Left hand side.
-class(integrand), intent(IN)    :: rhs !< Right hand side.
-endsubroutine assignment_integrand
-\end{lstlisting}
-
-The assignment statement is necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the assignment of an integrand object by another one, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
-The implementation details of this assignment is strictly dependent on the concrete extension of the integrand type.
-From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the assignment, and the second argument, the right-hand-side of the assignment, are two integrand objects.
+\input{src/mathematical-numerical-models/mathematical-numerical-models.tex}
 
@@
 
-\subsection{The explicit forward Euler solver}\label{subsec:solver_euler}
-
-The explicit forward Euler solver is exposed (by the FOODiE main module that must imported, see listing \ref{list:main_module}) as a single derived type (that is a standard convention for all FOODiE solvers) named \emph{euler\_explicit\_integrator}.
-It provides the type bound procedure (also referred as \emph{method}) \emph{integrate} for integrating in time an \emph{integrand} object, or any of its polymorphic concrete extensions.
-Consequently, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_euler_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit forward Euler integrator},label={list:define_euler_solver}]
-use FOODiE, only: euler_explicit_integrator
-type(euler_explicit_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it can be directly used without any initialization, for example see listing \ref{list:use_euler_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an explicit forward Euler integrator},label={list:use_euler_solver}]
-type(my_integrand) :: my_field
-call integrator%integrate(U=my_field, Dt=0.1)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
-
-The complete implementation of the \emph{integrate} method of the explicit forward Euler solver is reported in listing \ref{list:euler_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of Euler solver},label={list:euler_integrate}]
-subroutine integrate(U, Dt, t)
-class(integrand),    intent(INOUT) :: U  !< Field to be integrated.
-real(R_P),           intent(IN)    :: Dt !< Time step.
-real(R_P), optional, intent(IN)    :: t  !< Time.
-U = U + U%t(t=t) * Dt
-return
-endsubroutine integrate
-\end{lstlisting}
-This method takes three arguments, the first argument is an integrand class, it being the integrand field that must integrated one-step-over in time, the second is the time step used and the third, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-
-\subsection{The explicit TVD/SSP Runge-Kutta class of solvers}\label{subsec:solver_tvd_rk}
-
-The TVD/SSP Runge-Kutta class of solvers is exposed as a single derived type named \emph{tvd\_runge\_kutta\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time.
-  \end{itemize}
-
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_tvd_rk_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit TVD/SSP Runge-Kutta integrator},label={list:define_tvd_rk_solver}]
-use FOODiE, only: tvd_runge_kutta_integrator
-type(tvd_runge_kutta_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_tvd_rk_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit TVD/SSP Runge-Kutta integrator},label={list:init_tvd_rk_solver}]
-call integrator%init(stages=3)
-\end{lstlisting}
-
-In the listing \ref{list:init_tvd_rk_solver} a 3-stages solver has been initialized.
-As a matter of facts, from the equations \ref{eq:RK} and \ref{eq:RK-stage} a solver belonging to this class is completely defined once the number of stages adopted has been chosen.
-The complete definition of the \emph{tvd\_runge\_kutta\_integrator} type is reported into listing \ref{list:tvd_rk_solver}.
-As shown, the Butcher's coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{tvd\_runge\_kutta\_integrator} type},label={list:tvd_rk_solver}]
-type :: tvd_runge_kutta_integrator
-  integer(I_P)           :: stages=0  ! Number of stages.
-  real(R_P), allocatable :: alph(:,:) ! alpha Butcher's coefficients.
-  real(R_P), allocatable :: beta(:)   ! beta Butcher's coefficients.
-  real(R_P), allocatable :: gamm(:)   ! gamma Butcher's coefficients.
-  contains
-    procedure, pass(self), public :: destroy
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-    final                         :: finalize
-endtype tvd_runge_kutta_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_tvd_rk_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a TVD/SSP Runge-Kutta integrator},label={list:use_tvd_rk_solver}]
-type(my_integrand) :: my_field
-type(my_integrand) :: my_stages(1:3)
-call integrator%integrate(U=my_field, stage=my_stage, Dt=0.1)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
-Listing \ref{list:use_tvd_rk_solver} shows that the memory registers necessary for storing the Runge-Kutta stages must be supplied by the client code.
-
-The complete implementation of the \emph{integrate} method of the explicit TVD/SSP Runge-Kutta class of solvers is reported in listing \ref{list:tvd_rk_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit TVD/SSP Runge-Kutta class},label={list:tvd_rk_integrate}]
-subroutine integrate(self, U, stage, Dt, t)
-class(tvd_runge_kutta_integrator), intent(IN)    :: self      ! Actual RK integrator.
-class(integrand),                  intent(INOUT) :: U         ! Field to be integrated.
-class(integrand),                  intent(INOUT) :: stage(1:) ! Runge-Kutta stages [1:stages].
-real(R_P),                         intent(IN)    :: Dt        ! Time step.
-real(R_P),                         intent(IN)    :: t         ! Time.
-integer(I_P)                                     :: s         ! First stages counter.
-integer(I_P)                                     :: ss        ! Second stages counter.
-select type(stage)
-class is(integrand)
-  do s=1, self%stages
-    stage(s) = U
-    do ss=1, s - 1
-      stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
-    enddo
-    stage(s) = stage(s)%t(t=t + self%gamm(s) * Dt)
-  enddo
-  do s=1, self%stages
-    U = U +  stage(s) * (Dt * self%beta(s))
-  enddo
-endselect
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-
-It is worth noting that the stages memory registers, namely the array \emph{stage}, must be passed as argument because it is defined as a \emph{not-passed} polymorphic argument, thus we are not allowed to define it as an automatic array of the \emph{integrate} method.
-
-\subsection{The explicit low storage Runge-Kutta class of solvers}\label{subsec:solver_ls_rk}
-
-The low storage variant of Runge-Kutta class of solvers is exposed as a single derived type named \emph{ls\_runge\_kutta\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time.
-  \end{itemize}
-
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ls_rk_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit low storage Runge-Kutta integrator},label={list:define_ls_rk_solver}]
-use FOODiE, only: ls_runge_kutta_integrator
-type(ls_runge_kutta_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ls_rk_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit low storage Runge-Kutta integrator},label={list:init_ls_rk_solver}]
-call integrator%init(stages=5)
-\end{lstlisting}
-
-In the listing \ref{list:init_ls_rk_solver} a 5-stages solver has been initialized.
-As a matter of facts, from the equation \ref{eq:RK-ls} a solver belonging to this class is completely defined once the number of stages adopted has been chosen.
-The complete definition of the \emph{ls\_runge\_kutta\_integrator} type is reported into listing \ref{list:ls_rk_solver}.
-As shown, the Williamson's coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{ls\_runge\_kutta\_integrator} type},label={list:ls_rk_solver}]
-type :: ls_runge_kutta_integrator
-  integer(I_P)           :: stages=0 ! Number of stages.
-  real(R_P), allocatable :: A(:)     ! Low storage *A* coefficients.
-  real(R_P), allocatable :: B(:)     ! Low storage *B* coefficients.
-  real(R_P), allocatable :: C(:)     ! Low storage *C* coefficients.
-  contains
-    procedure, pass(self), public :: destroy
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-    final                         :: finalize
-endtype ls_runge_kutta_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ls_rk_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a low storage Runge-Kutta integrator},label={list:use_ls_rk_solver}]
-type(my_integrand) :: my_field
-type(my_integrand) :: my_stages(1:2)
-call integrator%integrate(U=my_field, stage=my_stage, Dt=0.1)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.
-Listing \ref{list:use_ls_rk_solver} shows that the memory registers necessary for storing the Runge-Kutta stages must be supplied by the client code, as it happens of the TVD/SSP Runge-Kutta class.
-However, now the registers necessary is always 2, independently on the number of stages used, that in the example considered are 5.
-
-The complete implementation of the \emph{integrate} method of the explicit low storage Runge-Kutta class of solvers is reported in listing \ref{list:ls_rk_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit low storage Runge-Kutta class},label={list:ls_rk_integrate}]
-subroutine integrate(self, U, stage, Dt, t)
-class(ls_runge_kutta_integrator), intent(IN)    :: self       ! Actual RK integrator.
-class(integrand),                 intent(INOUT) :: U          ! Field to be integrated.
-class(integrand),                 intent(INOUT) :: stage(1:2) ! Runge-Kutta registers [1:2].
-real(R_P),                        intent(IN)    :: Dt         ! Time step.
-real(R_P),                        intent(IN)    :: t          ! Time.
-integer(I_P)                                    :: s          ! First stages counter.
-select type(stage)
-class is(integrand)
-  stage(1) = U
-  stage(2) = U*0._R_P
-  do s=1, self%stages
-    stage(2) = stage(2) * self%A(s) + stage(1)%t(t=t + self%C(s) * Dt) * Dt
-    stage(1) = stage(1) + stage(2) * self%B(s)
-  enddo
-  U = stage(1)
-endselect
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-
-It is worth noting that the stages memory registers, namely the array \emph{stage}, must be passed as argument because it is defined as a \emph{not-passed} polymorphic argument, thus we are not allowed to define it as an automatic array of the \emph{integrate} method.
-
-\subsection{The explicit Adams-Bashforth class of solvers}\label{subsec:solver_ab}
-
-The explicit Adams-Bashforth class of solvers is exposed as a single derived type named \emph{adams\_bashforth\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time;
-  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
-  \end{itemize}
-
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ab_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Adams-Bashforth integrator},label={list:define_ab_solver}]
-use FOODiE, only: adams_bashforth_integrator
-type(adams_bashforth_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ab_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Adams-Bashforth integrator},label={list:init_ab_solver}]
-call integrator%init(steps=4)
-\end{lstlisting}
-
-In the listing \ref{list:init_ab_solver} a 4-steps solver has been initialized.
-As a matter of facts, from the equation \ref{eq:AB} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen.
-The complete definition of the \emph{adams\_bashforth\_integrator} type is reported into listing \ref{list:ab_solver}.
-As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_bashforth\_integrator} type},label={list:ab_solver}]
-type :: adams_bashforth_integrator
-  private
-  integer(I_P)           :: steps=0 ! Number of time steps.
-  real(R_P), allocatable :: b(:)    ! \(b\) coefficients.
-  contains
-    procedure, pass(self), public :: destroy
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-    procedure, pass(self), public :: update_previous
-    final                         :: finalize
-endtype adams_bashforth_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ab_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
-real               :: times(1:4)
-type(my_integrand) :: my_field
-type(my_integrand) :: previous(1:4)
-call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.
-
-The complete implementation of the \emph{integrate} method of the explicit Adams-Bashforth class of solvers is reported in listing \ref{list:ab_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
-subroutine integrate(self, U, previous, Dt, t, autoupdate)
-class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
-class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
-class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
-real(R_P),                         intent(IN)    :: Dt           ! Time steps.
-real(R_P),                         intent(IN)    :: t(:)         ! Times.
-logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
-logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
-integer(I_P)                                     :: s            ! Steps counter.
-autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
-do s=1, self%steps
-  U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
-enddo
-if (autoupdate_) call self%update_previous(U=U, previous=previous)
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-
-It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
-This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.
-
-\subsection{The implicit Adams-Moulton class of solvers}\label{subsec:solver_am}
-
-The explicit Adams-Moulton class of solvers is exposed as a single derived type named \emph{adams\_moulton\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time;
-  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
-  \end{itemize}
-
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_am_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an implicit Adams-Moulton integrator},label={list:define_am_solver}]
-use FOODiE, only: adams_moulton_integrator
-type(adams_moulton_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_am_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an implicit Adams-Moulton integrator},label={list:init_am_solver}]
-call integrator%init(steps=3)
-\end{lstlisting}
-
-In the listing \ref{list:init_am_solver} a 3-steps solver has been initialized.
-As a matter of facts, from the equation \ref{eq:AM} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen.
-The complete definition of the \emph{adams\_moulton\_integrator} type is reported into listing \ref{list:am_solver}.
-As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_moulton\_integrator} type},label={list:am_solver}]
-type :: adams_moulton_integrator
-  private
-  integer(I_P)           :: steps=-1 ! Number of time steps.
-  real(R_P), allocatable :: b(:)     ! \(b\) coefficients.
-  contains
-    procedure, pass(self), public :: destroy
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-    procedure, pass(self), public :: update_previous
-    final                         :: finalize
-endtype adams_moulton_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_am_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Moulton integrator},label={list:use_am_solver}]
-real               :: times(1:3)
-type(my_integrand) :: my_field
-type(my_integrand) :: previous(1:3)
-call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.
-
-The complete implementation of the \emph{integrate} method of the implicit Adams-Moulton class of solvers is reported in listing \ref{list:am_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Moulton class},label={list:am_integrate}]
-subroutine integrate(self, U, previous, Dt, t, autoupdate)
-class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
-class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
-class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
-real(R_P),                         intent(IN)    :: Dt           ! Time steps.
-real(R_P),                         intent(IN)    :: t(:)         ! Times.
-logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
-logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
-integer(I_P)                                     :: s            ! Steps counter.
-autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
-if (autoupdate_) call self%update_previous(U=U, previous=previous)
-if (self%steps>0) then
-  U = previous(self%steps) + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
-  do s=0, self%steps - 1
-    U = U + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
-  enddo
-  if (autoupdate_) call self%update_previous(U=U, previous=previous)
-else
-  U = U + U%t(t=t(s+1)) * (Dt * self%b(0))
-endif
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-
-It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
-This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.
-
-\subsection{The predictor-corrector Adams-Bashforth-Moulton class of solvers}\label{subsec:solver_abm}
-
-{\color{red} To be written.}
-
-\subsection{The leapfrog solver}\label{subsec:solver_leapfrog}
-
-The explicit Leapfrog class of solvers is exposed as a single derived type named \emph{leapfrog\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time.
-  \end{itemize}
-
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_lf_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Leapfrog integrator},label={list:define_lf_solver}]
-use FOODiE, only: leapfrog_integrator
-type(leapfrog_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_lf_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Leapfrog integrator},label={list:init_lf_solver}]
-! default coefficients nu=0.01, alpha=0.53
-call integrator%init()
-! custom coefficients
-call integrator%init(nu=0.015, alpha=0.6)
-\end{lstlisting}
-
-The complete definition of the \emph{leapfrog\_integrator} type is reported into listing \ref{list:lf_solver}.
-As shown, the filter coefficients are initialized to zero, suitable values are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{leapfrog\_integrator} type},label={list:lf_solver}]
-type :: leapfrog_integrator
-  private
-  real(R_P) :: nu=0.01_R_P    ! Robert-Asselin filter coefficient.
-  real(R_P) :: alpha=0.53_R_P ! Robert-Asselin-Williams filter coefficient.
-  contains
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-endtype leapfrog_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_lf_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a Leapfrog integrator},label={list:use_lf_solver}]
-real               :: times(1:2)
-type(my_integrand) :: filter_displacement
-type(my_integrand) :: my_field
-type(my_integrand) :: previous(1:2)
-call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, &
-                          t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{previous} are the memory registers where previous time steps solutions are saved, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.
-
-The complete implementation of the \emph{integrate} method of the explicit Leapfrog class of solvers is reported in listing \ref{list:lf_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Leapfrog class},label={list:lf_integrate}]
-subroutine integrate(self, U, previous, Dt, t, filter)
-class(leapfrog_integrator), intent(IN)    :: self          ! LF integrator.
-class(integrand),           intent(INOUT) :: U             ! Field to be integrated.
-class(integrand),           intent(INOUT) :: previous(1:2) ! Previous time steps solutions.
-real(R_P),                  intent(in)    :: Dt            ! Time step.
-real(R_P),                  intent(IN)    :: t             ! Time.
-class(integrand), optional, intent(INOUT) :: filter        ! Filter field displacement.
-U = previous(1) + previous(2)%t(t=t) * (Dt * 2._R_P)
-if (present(filter)) then
-  filter = (previous(1) - previous(2) * 2._R_P + U) * self%nu * 0.5_R_P
-  previous(2) = previous(2) + filter * self%alpha
-  U = U + filter * (self%alpha - 1._R_P)
-endif
-previous(1) = previous(2)
-previous(2) = U
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the optional filter-displacement-register, the fifth is the time step used and the sixth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-It is worth noting that if the filter displacement argument is not passed, the solver reverts back to the standard unfiltered Leapfrog method.
-
-It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
-In particular, if the filter displacement argument is passed the method performs the RAW filtering.
-
-\subsection{General Remarks}\label{subsec:API-remarks}
-
-Table \ref{tab:foodie-vs-math} presents a comparison of the relevant parts of equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:RK-stage}, \ref{eq:RK-ls}, \ref{eq:AB} and \ref{eq:leapfrog} with the corresponding FOODiE implementations reported in listings \ref{list:euler_integrate}, \ref{list:tvd_rk_integrate}, \ref{list:ls_rk_integrate}, \ref{list:ab_integrate} and \ref{list:lf_integrate}, respectively.
-This comparison proves that the \emph{integrand} ADT has allowed a very high-level implementation syntax.
-The Fortran implementation is almost equivalent to the rigorous mathematical formulation.
-This aspect directly implies that the implementation of a ODE solver into the FOODiE library is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the solvers must be implemented for each specific definition of the integrand type, it being not abstract.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Comparison between rigorous mathematical formulation and FOODiE high-level implementation; the syntax "$(s)$" and "$(ss)$" imply the summation operation}\label{tab:foodie-vs-math}
-  \resizebox{0.99\textwidth}{!}{%
-  \begin{tabular}{c|cc}
-    \toprule
-    {\sc Solver} & {\sc Mathematical formulation} & {\sc FOODiE implementation} \\
-    \hline
-    explicit forward Euler & $U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]$ & $U = U + U\%t(t=t) * Dt$\\
-    \hline
-    TVD/SSP Runge-Kutta & $\begin{matrix}
-  K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right) \\
-  U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
-\end{matrix}$ & $\begin{matrix}
-  stage(s) = stage(s) + stage(ss) * (Dt * self\%alph(s, ss)) \\
-  U = U +  stage(s) * (Dt * self\%beta(s))
-\end{matrix}$\\
-    \hline
-    low storage Runge-Kutta & $\begin{matrix}
-  K_2 = A_s K_2 + \Delta t \cdot R\left(t^n + C_s \Delta t, K_1\right) \\
-  K_1 = K_1 + B_s K_2
-\end{matrix}$ & $\begin{matrix}
-  stage(2) = stage(2) * self\%A(s) + \\ + stage(1)\%t(t=t + self\%C(s) * Dt) * Dt \\
-  stage(1) = stage(1) + stage(2) * self\%B(s)
-\end{matrix}$\\
-    \hline
-  explicit Adams-Bashforth & $\begin{matrix}U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) + \\ +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}\end{matrix}$ & $U = U + U\%t(n=s, t=t(s)) * (Dt * self\%b(s))$\\
-    \hline
-    explicit Leapfrog & $ U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]$ & $U = U\%previous\_step(n=1) + U\%t(n=2, t=t) * (Dt * 2.)$\\
-
-  \end{tabular}}
-\end{table}
+\input{src/API/API.tex}
 
@@
 
-\section{Tests and Examples}\label{sec:Tests}
+\input{src/tests/tests.tex}
 
-{\color{red} To be written.}
-
-\subsection{Oscillation equations test}
-
-Let us consider the \emph{oscillation} problem, it being a simple, yet interesting IVP.
-Briefly, the oscillation problem is a prototype problem of non dissipative, oscillatory phenomena.
-For example, let us consider a pendulum subjected to the Coriolis accelerations without dissipation, the motion equations of which can be described by the ODE system \ref{eq:oscillation}.
-
-\begin{equation}
-\begin{matrix}
-U_t = R(U)  \\
-U = \begin{bmatrix}
-x \\
-y
-\end{bmatrix}\;\;\;
-R(U) = \begin{bmatrix}
--f y \\
-f x
-\end{bmatrix}
-\end{matrix}
-\label{eq:oscillation}
-\end{equation}
-where the frequency is chosen as $f=10^4$. The ODE system \ref{eq:oscillation} is completed by the following initial conditions:
-
-\begin{equation}
-\begin{matrix}
-  x(t_0) = 0 \\
-  y(t_0) = 1
-\end{matrix}
-\label{eq:oscillation-ic}
-\end{equation}
-where $t_0=0$ is the initial time considered.
-
-The IVP constituted by equations \ref{eq:oscillation} and \ref{eq:oscillation-ic} is (apparently) simple and its exact solution is known:
-
-\begin{equation}
-\begin{matrix}
-  x(t_0 + \Delta t) = X_0 cos(f \Delta t) - y_0 sin(f \Delta t) \\
-  y(t_0 + \Delta t) = X_0 sin(f \Delta t) + y_0 cos(f \Delta t)
-\end{matrix}
-\label{eq:oscillation-exact}
-\end{equation}
-where $\Delta t$ is an arbitrary time step.
-
-This problem is only apparently simple.
-As a matter of facts, in a non dissipative oscillatory problem the eventual errors in the amplitude approximation can rapidly drive the subsequent series of approximations to an unphysical solution.
-This is of particular relevance if the solution (that is numerically approximated) constitutes a \emph{prediction} far from the initial conditions, that is the common case in weather forecasting.
-
-\subsubsection{FOODiE aware implementation of an oscillation numerical solver}
-The IVP \ref{eq:oscillation} can be easily solved by means of FOODiE library.
-The first block of a FOODiE aware solution consists to define an \emph{oscillation integrand field} defining a concrete extension of the FOODiE \emph{integrand} type.
-Listing \ref{list:oscillation_type} reports the implementation of such an integrand field that is contained into the tests suite shipped within the FOODiE library.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} type},label={list:oscillation_type}]
-type, extends(integrand) :: oscillation
-  private
-  integer(I_P)                           :: dims=0   ! Space dimensions.
-  real(R_P)                              :: f=0._R_P ! Oscillation frequency (Hz).
-  real(R_P), dimension(:),   allocatable :: U        ! Integrand (state) variables, [1:dims].
-  contains
-    ! auxiliary methods
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: output
-    ! type\_integrand deferred methods
-    procedure, pass(self), public :: t => dOscillation_dt
-    procedure, pass(lhs),  public :: integrand_multiply_integrand => &
-                                     oscillation_multiply_oscillation
-    procedure, pass(lhs),  public :: integrand_multiply_real => oscillation_multiply_real
-    procedure, pass(rhs),  public :: real_multiply_integrand => real_multiply_oscillation
-    procedure, pass(lhs),  public :: add => add_oscillation
-    procedure, pass(lhs),  public :: sub => sub_oscillation
-    procedure, pass(lhs),  public :: assign_integrand => oscillation_assign_oscillation
-    procedure, pass(lhs),  public :: assign_real => oscillation_assign_real
-endtype oscillation
-\end{lstlisting}
-
-The \emph{oscillation} field extends the \emph{integrand} ADT making it a concrete type.
-This derived type is very simple: it has 5 data members for storing the state vector and some auxiliary variables, and it implements all the deferred methods necessary for defining a valid concrete extension of the \emph{integrand} ADT (plus 2 auxiliary methods that are not relevant for our discussion).
-The key point is here constituted by the implementation of the deferred methods: the \emph{integrand} ADT does not impose any structure for the data members, that are consequently free to be customized by the client code.
-In this example the data members have a very simple, clean and concise structure:
-\begin{itemize}
-  \item $dims$ is the number of space dimensions; in the case of equation \ref{eq:oscillation} we have $dims=2$, however the test implementation has been kept more general parametrizing this dimension in order to easily allow future modification of the test-program itself;
-  \item $f$ stores the frequency of the oscillatory problem solved, that is here set to $10^{4}$, but it can be changed at runtime in the test-program;
-  \item $U$ is the state vector corresponding directly to the state vector of equation \ref{eq:oscillation};
-  \end{itemize}
-
-As the listing \ref{list:oscillation_type} shows, the FOODiE implementation strictly corresponds to the mathematical formulation embracing all the relevant mathematical aspects into one derived type, a single \emph{object}.
-Here we not review the implementation of all deferred methods, this being out of the scope of the present work: the interested reader can see the tests suite sources shipped within the FOODiE library.
-However, some of these methods are relevant for our discussion, thus they are reviewed.
-
-\paragraph{dOscillation\_dt, the oscillation residuals function}
-
-Probably, the most important methods for an IVP solver is the residuals function.
-As a matter of facts, the ODE equations are implemented into the residuals function.
-However, the FOODiE ADT strongly alleviates the subtle problems that could arise when the ODE solver is hard-implemented within the specific ODE equations.
-As a matter of facts, the \emph{integrand} ADT specifies the precise interface the residuals function must have: if the client code implements a compliant interface, the FOODiE solvers will work as expected, reducing the possible errors location into the ODE equations, having designed the solvers on the ADT and not on the concrete type.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} residuals function},label={list:oscillation_t}]
-function dOscillation_dt(self, t) result(dState_dt)
-class(oscillation),     intent(IN) :: self      ! Oscillation field.
-real(R_P),    optional, intent(IN) :: t         ! Time.
-class(integrand),  allocatable     :: dState_dt ! Oscillation field time derivative.
-integer(I_P)                       :: dn        ! Time level, dummy variable.
-allocate(oscillation :: dState_dt)
-select type(dState_dt)
-class is(oscillation)
-  dState_dt = self
-  dState_dt%U(1) = -self%f * self%U(2)
-  dState_dt%U(2) =  self%f * self%U(1)
-endselect
-return
-endfunction dOscillation_dt
-\end{lstlisting}
-
-Listing \ref{list:oscillation_t} reports the implementation of the oscillation residuals function: it is very clear and concise.
-Moreover, comparing this listing with the equation \ref{eq:oscillation} the close correspondence between the mathematical formulation and Fortran implementation is evident.
-
-\paragraph{add method, an example of oscillation symmetric operator}
-
-As a prototype of the operators overloading let us consider the \emph{add} operator, it being a prototype of symmetric operators, the implementation of which is presented in listing \ref{list:oscillation_add}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} add operator},label={list:oscillation_add}]
-function add_oscillation(lhs, rhs) result(opr)
-class(oscillation), intent(IN) :: lhs ! Left hand side.
-class(integrand),   intent(IN) :: rhs ! Right hand side.
-class(integrand), allocatable  :: opr ! Operator result.
-allocate(oscillation :: opr)
-select type(opr)
-class is(oscillation)
-  opr = lhs
-  select type(rhs)
-  class is (oscillation)
-    opr%U = lhs%U + rhs%U
-  endselect
-endselect
-return
-endfunction add_Oscillation
-\end{lstlisting}
-It is very simple and clear: firstly all the auxiliary data are copied into the operator result, then the state vector of the result is populated with the addiction between the state vectors of the left-hand-side and right-hand-side.
-This is very intuitive from the mathematical point of view and it helps to reduce implementation errors.
-Similar implementations are possible for all the other operators necessary to define a valid \emph{intregrand} ADT concrete extension.
-
-\paragraph{assignment of an oscillation object}
-
-The assignment overloading of the \emph{oscillation} type is the last key-method that enforces the conciseness of the FOODiE aware implementation.
-Listing \ref{list:oscillation_assign} reports the implementation of the assignment overloading.
-Essentially, to all the data members of the left-hand-side are assigned the values of the corresponding right-hand-side.
-Notably, for the assignment of the state vector and of the previous time steps solution array we take advantage of the automatic re-allocation of the left-hand-side variables when they are not allocated or allocated differently from the right-hand-side, that is a Fortran 2003 feature.
-In spite its simplicity, the assignment overloading is a key-method enabling the usage of FOODiE solver: effectively, the assignment between two \emph{integrand} ADT variables is ubiquitous into the solvers implementations, see equation \ref{eq:RK} for example.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} assignment},label={list:oscillation_assign}]
-subroutine oscillation_assign_oscillation(lhs, rhs)
-class(oscillation), intent(INOUT) :: lhs ! Left hand side.
-class(integrand),   intent(IN)    :: rhs ! Right hand side.
-select type(rhs)
-class is (oscillation)
-  lhs%dims = rhs%dims
-  lhs%f = rhs%f
-  if (allocated(rhs%U)) lhs%U = rhs%U
-endselect
-return
-endsubroutine oscillation_assign_oscillation
-\end{lstlisting}
-
-\paragraph{FOODiE numerical integration}
-
-Using the above discussed \emph{oscillation} type it is very easy to solve IVP \ref{eq:oscillation} by means of FOODiE library.
-Listing \ref{list:oscillation_leapfrog} presents the numerical integration of system \ref{eq:oscillation} by means of the Leapfrog RAW-filtered method.
-In the example, the integration is performed with $10^4$ steps with a fixed $\Delta t=10^2$ until the time $t=10^6$ is reached.
-The example shows also that for starting a multi-step scheme such as the Leapfrog one a lower-oder or equivalent order one-scheme is necessary: in the example the first 2 steps are computed by means of one-step TVD/SSP Runge-Kutta 2-stages schemes.
-Note that the memory registers for storing the Runge-Kutta stages and the RAW filter displacement must be handled by the client code.
-Listing \ref{list:oscillation_leapfrog} demonstrates how it is simple, clear and concise to solve a IVP by FOODiE solvers.
-Moreover, it proves how it is simple and effective to apply different solvers in a coupled algorithm, that greatly simplify the development of new hybrid solvers for self-adaptive time step size.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={numerical integration of the \emph{oscillation} system by means of Leapfrog RAW-filtered method},label={list:oscillation_leapfrog}]
-use foodie, only: leapfrog_integrator, tvd_runge_kutta_integrator
-type(leapfrog_integrator)        :: lf_integrator ! Leapfrog integrator.
-type(tvd_runge_kutta_integrator) :: rk_integrator ! Runge-Kutta integrator.
-type(oscillation)                :: rk_stage(1:2) ! Runge-Kutta stages.
-type(oscillation)                :: previous(1:2) ! Previous time steps solution.
-type(oscillation)                :: oscillator    ! Oscillation field.
-type(oscillation)                :: filter        ! Filter displacement.
-integer                          :: step          ! Time steps counter.
-real                             :: Dt            ! Time step.
-call lf_integrator%init()
-call rk_integrator%init(stages=2)
-call oscillator%init(initial_state=[0.0,1.0], f=10e4, steps=2)
-Dt = 100.0
-do step=1, 10000
-  if (2>=step) then
-    call rk_integrator%integrate(U=oscillator, stage=rk_stage, Dt=Dt, t=step*Dt)
-    previous(step) = oscillator
-  else
-    call lf_integrator%integrate(U=oscillator, previous=previous, filter=filter, Dt=Dt, &
-                                 t=step*Dt)
-  endif
-enddo
-call print_results(U=oscillator)
-\end{lstlisting}
-
-\subsubsection{Errors Analysis of the FOODiE Oscillation test results}
-
-We use the Oscillation test program (shipped with FOODiE tests suite) for the testing the above described implementation and assessing the provided solvers accuracy for this IVP.
-
-For the analysis of the accuracy of each solver, we have integrated the Oscillation equations \ref{eq:oscillation} with different, decreasing time steps in the range $[5000, 2500, 1250, 625, 320, 100]$.
-The error is estimated by the L2 norm of the difference between the exact ($U_e$) and the numerical ($U_{\Delta t}$) solutions for each time step:
-
-\begin{equation}
-  \varepsilon (\Delta t) = || U_e - U_{\Delta t} ||_2 = \sqrt{ \sum_{s=1}^{N_s} { \left(U_e(t_0 + s * \Delta t) - U_{\Delta t}(t_0 + s * \Delta t) \right)^2 }}
-\label{eq:oscillation-error}
-\end{equation}
-where $N_s$ is the total number of time steps performed to reach the final integration time.
-
-Using two pairs of subsequent-decreasing time steps solution is possible to estimate the order of accuracy of the solver employed computing the \emph{observed order} of accuracy:
-
-\begin{equation}
-  p = \frac{log10 \left( \frac{\varepsilon (\Delta t_1)}{\varepsilon (\Delta t_2)} \right)}{log10 \left( \frac{\Delta t_1}{\Delta t_2} \right)}
-\label{eq:oscillation-observed-order}
-\end{equation}
-
-where $\frac{\Delta t_1}{\Delta t_2}>1$.
-
-Tables \ref{tab:oscillation_errors_explicit} and \ref{tab:oscillation_errors_pred_corr} summarize the errors analysis.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Oscillation test: errors analysis of explicit solvers}\label{tab:oscillation_errors_explicit}
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth 1 step}\label{tab:oscillation-ab-1}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
-      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
-      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
-       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
-       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
-       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth 2 steps}\label{tab:oscillation-ab-2}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.596E+03    &  0.583E+03    & /             & /             \\
-      2500.0          &  0.221E+02    &  0.218E+02    & 4.75          & 4.74          \\
-      1250.0          &  0.764E+01    &  0.769E+01    & 1.53          & 1.50          \\
-       625.0          &  0.265E+01    &  0.268E+01    & 1.53          & 1.52          \\
-       320.0          &  0.968E+00    &  0.981E+00    & 1.51          & 1.50          \\
-       100.0          &  0.169E+00    &  0.171E+00    & 1.50          & 1.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth 3 steps}\label{tab:oscillation-ab-3}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.857E+01    &  0.854E+01    & /             & /             \\
-      2500.0          &  0.391E+01    &  0.386E+01    & 1.13          & 1.14          \\
-      1250.0          &  0.825E+00    &  0.814E+00    & 2.24          & 2.25          \\
-       625.0          &  0.150E+00    &  0.148E+00    & 2.46          & 2.46          \\
-       320.0          &  0.282E-01    &  0.278E-01    & 2.49          & 2.49          \\
-       100.0          &  0.154E-02    &  0.152E-02    & 2.50          & 2.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth 4 steps}\label{tab:oscillation-ab-4}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.128E+07    &  0.143E+07    & /             & /             \\
-      2500.0          &  0.106E+01    &  0.107E+01    & 20.21         & 20.34         \\
-      1250.0          &  0.967E-01    &  0.981E-01    &  3.45         &  3.45         \\
-       625.0          &  0.859E-02    &  0.871E-02    &  3.49         &  3.49         \\
-       320.0          &  0.827E-03    &  0.838E-03    &  3.50         &  3.50         \\
-       100.0          &  0.141E-04    &  0.143E-04    &  3.50         &  3.50         \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Leapfrog unfiltered}\label{tab:oscillation-leapfrog}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.156E+02    &  0.156E+02    & /             & /             \\
-      2500.0          &  0.849E+01    &  0.846E+01    & 0.87          & 0.88          \\
-      1250.0          &  0.300E+01    &  0.303E+01    & 1.50          & 1.48          \\
-       625.0          &  0.106E+01    &  0.107E+01    & 1.51          & 1.50          \\
-       320.0          &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
-       100.0          &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Leapfrog RAW-filtered}\label{tab:oscillation-leapfrog-raw}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.156E+02    &  0.156E+02    & /             & /             \\
-      2500.0          &  0.855E+01    &  0.852E+01    & 0.86          & 0.87          \\
-      1250.0          &  0.303E+01    &  0.305E+01    & 1.50          & 1.48          \\
-       625.0          &  0.107E+01    &  0.108E+01    & 1.51          & 1.50          \\
-       320.0          &  0.390E+00    &  0.395E+00    & 1.50          & 1.50          \\
-       100.0          &  0.685E-01    &  0.692E-01    & 1.50          & 1.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Low storage Runge-Kutta 1 stage}\label{tab:oscillation-ls-rk-1}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
-      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
-      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
-       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
-       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
-       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Low storage Runge-Kutta 5 stages}\label{tab:oscillation-ls-rk-5}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.120E+00    &  0.122E+00    & /             & /             \\
-      2500.0          &  0.106E-01    &  0.107E-01    & 3.51          & 3.51          \\
-      1250.0          &  0.935E-03    &  0.947E-03    & 3.50          & 3.50          \\
-       625.0          &  0.826E-04    &  0.836E-04    & 3.50          & 3.50          \\
-       320.0          &  0.793E-05    &  0.803E-05    & 3.50          & 3.50          \\
-       100.0          &  0.135E-06    &  0.137E-06    & 3.50          & 3.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{TVD/SSP Runge-Kutta 1 stage}\label{tab:oscillation-tvd-rk-1}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
-      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
-      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
-       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
-       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
-       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{TVD/SSP Runge-Kutta 2 stages}\label{tab:oscillation-tvd-rk-2}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.316E+02    &  0.319E+02    & /             & /             \\
-      2500.0          &  0.892E+01    &  0.894E+01    & 1.83          & 1.84          \\
-      1250.0          &  0.301E+01    &  0.305E+01    & 1.57          & 1.55          \\
-       625.0          &  0.106E+01    &  0.107E+01    & 1.51          & 1.51          \\
-       320.0          &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
-       100.0          &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{TVD/SSP Runge-Kutta 3 stages}\label{tab:oscillation-tvd-rk-3}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.255E+01    &  0.252E+01    & /             & /             \\
-      2500.0          &  0.523E+00    &  0.516E+00    & 2.28          & 2.29          \\
-      1250.0          &  0.944E-01    &  0.931E-01    & 2.47          & 2.47          \\
-       625.0          &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
-       320.0          &  0.314E-02    &  0.310E-02    & 2.50          & 2.50          \\
-       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{TVD/SSP Runge-Kutta 5 stages}\label{tab:oscillation-tvd-rk-5}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.139E+00    &  0.141E+00    & /             & /             \\
-      2500.0          &  0.122E-01    &  0.124E-01    & 3.50          & 3.50          \\
-      1250.0          &  0.108E-02    &  0.110E-02    & 3.50          & 3.50          \\
-       625.0          &  0.956E-04    &  0.969E-04    & 3.50          & 3.50          \\
-       320.0          &  0.937E-05    &  0.949E-05    & 3.47          & 3.47          \\
-       100.0          &  0.512E-06    &  0.519E-06    & 2.50          & 2.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \caption{Oscillation test: errors analysis of predictor-corrector solvers}\label{tab:oscillation_errors_pred_corr}
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth-Moulton 1 step}\label{tab:oscillation-abm-1}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.241E+20    &  0.266E+20    & /             & /             \\
-      2500.0          &  0.664E+11    &  0.716E+11    & 28.44         & 28.47         \\
-      1250.0          &  0.952E+06    &  0.100E+07    & 16.09         & 16.12         \\
-       625.0          &  0.413E+04    &  0.407E+04    &  7.85         &  7.95         \\
-       320.0          &  0.387E+03    &  0.383E+03    &  3.54         &  3.53         \\
-       100.0          &  0.145E+03    &  0.145E+03    &  0.84         &  0.83         \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth-Moulton 2 steps}\label{tab:oscillation-abm-2}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.704E+01    &  0.701E+01    & /             & /             \\
-      2500.0          &  0.392E+01    &  0.395E+01    & 0.84          & 0.83          \\
-      1250.0          &  0.148E+01    &  0.150E+01    & 1.40          & 1.39          \\
-       625.0          &  0.526E+00    &  0.534E+00    & 1.49          & 1.49          \\
-       320.0          &  0.193E+00    &  0.196E+00    & 1.50          & 1.50          \\
-       100.0          &  0.338E-01    &  0.342E-01    & 1.50          & 1.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth-Moulton 3 steps}\label{tab:oscillation-abm-3}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.457E+01    &  0.464E+01    & /             & /             \\
-      2500.0          &  0.656E+00    &  0.654E+00    & 2.80          & 2.83          \\
-      1250.0          &  0.100E+00    &  0.987E-01    & 2.71          & 2.73          \\
-       625.0          &  0.169E-01    &  0.167E-01    & 2.56          & 2.56          \\
-       320.0          &  0.314E-02    &  0.310E-02    & 2.52          & 2.51          \\
-       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{Adams-Bashforth-Moulton 4 steps}\label{tab:oscillation-abm-4}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{ccccc}
-      \toprule
-      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
-      \hline
-      5000.0          &  0.229E+01    &  0.225E+01    & /             & /             \\
-      2500.0          &  0.119E+00    &  0.118E+00    & 4.26          & 4.25          \\
-      1250.0          &  0.825E-02    &  0.833E-02    & 3.85          & 3.83          \\
-       625.0          &  0.671E-03    &  0.681E-03    & 3.62          & 3.61          \\
-       320.0          &  0.631E-04    &  0.640E-04    & 3.53          & 3.53          \\
-       100.0          &  0.107E-05    &  0.108E-05    & 3.51          & 3.51          \\
-      \bottomrule
-    \end{tabular}}
-  \end{subtable}\\
-\end{table}
-
-\paragraph{Adams-Bashforth}
-
-As expected, the Adams-Bashforth 1 step solution, that reverts back to the explicit forward Euler one, is unstable for all the $\Delta t$ exercised.
-
-The expected observed orders of accuracy for the Adams-Bashforth solvers using 2, 3 and 4 time steps tend to 1.5, 2.5 and 3.5 that are in agreement with the expected formal order of 2, 3 and 4, respectively.
-
-Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes the instability generation for some pairs steps number/$\Delta t$.
-The 2 and 4 steps solutions are instable for $\Delta t=5000 \rightarrow f*\Delta t=0.5$.
-On the contrary, the 3 steps solution is stable, but the amplitude is dumped and the solution vanishes as the integration proceeds.
-The 2 and 4 steps solutions show a phase error that decreases as the time resolution increases, whereas 3 steps solution has null phase error.
-
-\begin{figure}[!ht]
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-1.png}
-    \caption{1 step solution}\label{fig:results-oscillation-adams-bashforth-1}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
-    \caption{2 steps solution}\label{fig:results-oscillation-adams-bashforth-2}
-  \end{subfigure}\\
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
-    \caption{3 steps solution}\label{fig:results-oscillation-adams-bashforth-3}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-4.png}
-    \caption{4 steps solution}\label{fig:results-oscillation-adams-bashforth-4}
-  \end{subfigure}
-  \caption{Oscillation equations solutions computed by means of Adams-Bashforth solvers}\label{fig:results-oscillation-adams-bashforth}
-\end{figure}
-
-\paragraph{Leapfrog}
-
-The Leapfrog solutions are in agreement with the expected results: both unfiltered and RAW-filtered solutions show an observed order of accuracy that tends to the formal $2^{nd}$ order.
-The two solutions are almost the same.
-
-\begin{figure}[!ht]
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
-    \caption{Leapfrog unfiltered solution}\label{fig:results-oscillation-leapfrog-unfiltered}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog-raw.png}
-    \caption{Leapfrog RAW-filtered solution}\label{fig:results-oscillation-leapfrog-raw}
-  \end{subfigure}
-  \caption{Oscillation equations solutions computed by means of Leapfrog RAW-filtered solver}\label{fig:results-oscillation-leapfrog}
-\end{figure}
-
-\paragraph{Low Storage Runge-Kutta}
-
-\begin{figure}[!ht]
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-1.png}
-    \caption{1 stage solution}\label{fig:results-oscillation-ls-runge-kutta-1}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-5.png}
-    \caption{5 stages solution}\label{fig:results-oscillation-ls-runge-kutta-5}
-  \end{subfigure}
-  \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta solvers}\label{fig:results-oscillation-ls-runge-kutta}
-\end{figure}
-
-\paragraph{TVD/SSP Runge-Kutta}
-
-\begin{figure}[!ht]
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-1.png}
-    \caption{1 stage solution}\label{fig:results-oscillation-tvd-runge-kutta-1}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-2.png}
-    \caption{2 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-2}
-  \end{subfigure}\\
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-3.png}
-    \caption{3 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-3}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-5.png}
-    \caption{5 stages solution}\label{fig:results-oscillation-tvd-runge-kutta-5}
-  \end{subfigure}\\
-  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta solvers}\label{fig:results-oscillation-tvd-runge-kutta}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-1.png}
-    \caption{1 step solution}\label{fig:results-oscillation-adams-bashforth-moulton-1}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-2.png}
-    \caption{2 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-2}
-  \end{subfigure}\\
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-3.png}
-    \caption{3 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-3}
-  \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-4.png}
-    \caption{4 steps solution}\label{fig:results-oscillation-adams-bashforth-moulton-4}
-  \end{subfigure}
-  \caption{Oscillation equations solutions computed by means of Adams-Bashforth-Moulton solvers}\label{fig:results-oscillation-adams-bashforth-moulton}
-\end{figure}
 \clearpage
 
-\section{Benchmarks on parallel frameworks}\label{sec:parallel}
+\input{src/parallel/parallel.tex}
 
-{\color{red} To be written.}
+\clearpage
 
-\subsection{OpenMP benchmark}\label{subsec:openmp}
-
-{\color{red} To be written.}
-
-\subsection{MPI benchmark}\label{subsec:mpi}
-
-{\color{red} To be written.}
-
-\section{Concluding Remarks and Perspectives}\label{sec:conclusions}
-
-{\color{red} To be written.}
+\input{src/conclusions/conclusions.tex}
 

### 2015-10-14 papers/announcement/src/API/API.tex
Minor improvments on manuscript

\section{Application Program Interface}\label{sec:API}

In this section we review the FOODiE API providing a detailed discussion of the implementation choices.

As aforementioned, the programming language used is the Fortran 2008 standard, that is a minor revision of the previous Fortran 2003 standard. Such a new Fortran idioms provide (among other useful features) an almost complete support for OOP, in particular for ADT concept. Fortran 2003 has introduced the \emph{abstract derived type}: it is a derived type suitable to serve as \emph{contract} for concrete type-extensions that has not any actual implementations, rather it provides a well-defined set of type bound procedures interfaces, that in Fortran nomenclature are called \emph{deferred} procedures. Using such an abstract definition, we can implement algorithms operating on only this abstract type and on \emph{all its concrete extensions}. This is the key feature of FOODiE library: all the above described ODE solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax. In the meanwhile, client codes must implement their own IVPs extending only one simple abstract type.

In the subsection \ref{subsec:type_integrand} a review of the FOODiE main ADT, the \emph{integrand} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.

It is worth noting that all FOODiE public \emph{entities} (ADT and solvers) must be accessed by the FOODiE module, see listing \ref{list:main_module} for an example on how access to all public FOODiE entities.

\begin{lstlisting}[firstnumber=1,style=code,caption={usage example importing all public entities of FOODiE main module},label={list:main_module}]
use foodie, only: integrand, &
                  adams_bashforth_integrator, &
                  adams_moulton_integrator, &
                  adams_bashforth_moulton_integrator, &
                  euler_explicit_integrator, &
                  leapfrog_integrator, &
                  ls_runge_kutta_integrator, &
                  tvd_runge_kutta_integrator
! or simply
use foodie
\end{lstlisting}

\subsection{The main FOODiE Abstract Data Type: the integrand type}\label{subsec:type_integrand}

The implemented ACP is based on one main ADT, the \emph{integrand} type, the definition of which is shown in listing~\ref{list:type_integrand}.

\begin{lstlisting}[firstnumber=1,style=code,caption={integrand type definition},label={list:type_integrand}]
type, abstract :: integrand
  !< Abstract type for building FOODiE ODE integrators.
  contains
    ! public deferred procedures that concrete integrand-field must implement
    procedure(time_derivative),       pass(self), deferred, public:: t
    ! operators
    procedure(symmetric_operator),   pass(lhs), deferred, public:: integrand_multiply_integrand
    procedure(integrand_op_real),    pass(lhs), deferred, public:: integrand_multiply_real
    procedure(real_op_integrand),    pass(rhs), deferred, public:: real_multiply_integrand
    procedure(symmetric_operator),   pass(lhs), deferred, public:: add
    procedure(symmetric_operator),   pass(lhs), deferred, public:: sub
    procedure(assignment_integrand), pass(lhs), deferred, public:: assign_integrand
    ! operators overloading
    generic, public:: operator(+) => add
    generic, public:: operator(-) => sub
    generic, public:: operator(*) => integrand_multiply_integrand, &
                                     real_multiply_integrand, &
                                     integrand_multiply_real
    generic, public:: assignment(=) => assign_integrand
endtype integrand
\end{lstlisting}

The \emph{integrand} type does not implement any actual integrand field, it being and abstract type. It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODiE solver.

As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 1 ODE-specific procedure plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects. These procedures are analyzed in the following paragraphs.

\subsubsection{Time derivative procedure, the residuals function}

The abstract interface of the time derivative procedure \emph{t} is shown in listing \ref{list:time_derivative}.

\begin{lstlisting}[firstnumber=1,style=code,caption={time derivative procedure interface},label={list:time_derivative}]
function time_derivative(self, t) result(dState_dt)
import :: integrand, R_P, I_P
class(integrand),       intent(IN) :: self      !< Integrand field.
real(R_P),    optional, intent(IN) :: t         !< Time.
class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.
endfunction time_derivative
\end{lstlisting}

This procedure-function takes two arguments, the first passed as a \emph{type bounded} argument, while the latter is optional, and it returns an integrand object. The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT. The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.

Commonly, into the concrete implementation of this deferred abstract procedure clients embed the actual ODE equations being solved. As an example, for the Burgers equation, that is a Partial Differential Equations (PDE) system involving also a boundary value problem, this procedure embeds the spatial operator that convert the PDE to a system of algebraic ODE. As a consequence, the eventual concrete implementation of this procedure can be very complex and errors-prone. Nevertheless, the FOODiE solvers are implemented only on the above abstract interface, thus emancipating the solvers implementation from any concrete complexity.

{\color{red} Add citations to Burgers, Adams-Bashfort, leapfrog references.}

\subsubsection{Symmetric operators procedures}

The abstract interface of \emph{symmetric} procedures is shown in listing \ref{list:symmetric_operator}.

\begin{lstlisting}[firstnumber=1,style=code,caption={symmetric operator procedure interface},label={list:symmetric_operator}]
function symmetric_operator(lhs, rhs) result(operator_result)
import :: integrand
class(integrand), intent(IN)  :: lhs             !< Left hand side.
class(integrand), intent(IN)  :: rhs             !< Right hand side.
class(integrand), allocatable :: operator_result !< Operator result.
endfunction symmetric_operator
\end{lstlisting}

This interface defines a class of procedures operating on 2 \emph{integrand} objects, namely it is used for the definition of the operators \emph{multiplication}, \emph{summation} and \emph{subtraction} of integrand objects. These operators are used into the above described ODE solvers, for example see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}. The implementation details of such a procedures class are strictly dependent on the concrete extension of the integrand type. From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are two integrand object and the returned object is still an integrand one.

\subsubsection{Integrand/real and real/integrand operators procedures}

The abstract interfaces of \emph{Integrand/real and real/integrand operators} procedures are shown in listing \ref{list:integrand_real_ops}.

\begin{lstlisting}[firstnumber=1,style=code,caption={Integrand/real and real/integrand operators procedure interfaces},label={list:integrand_real_ops}]
function integrand_op_real(lhs, rhs) result(operator_result)
import :: integrand, R_P
class(integrand), intent(IN)  :: lhs              !< Left hand side.
real(R_P),        intent(IN)  :: rhs              !< Right hand side.
class(integrand), allocatable :: operator_result  !< Operator result.
endfunction integrand_op_real

function real_op_integrand(lhs, rhs) result(operator_result)
import :: integrand, R_P
real(R_P),        intent(IN)  :: lhs             !< Left hand side.
class(integrand), intent(IN)  :: rhs             !< Right hand side.
class(integrand), allocatable :: operator_result !< Operator result.
endfunction real_op_integrand
\end{lstlisting}

These two interfaces are necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the multiplication of an integrand object for a real number, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}. The implementation details of these procedures are strictly dependent on the concrete extension of the integrand type. From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are an integrand object and real number of viceversa and the returned object is still an integrand one.

\subsubsection{Integrand assignment procedure}

The abstract interface of \emph{integrand assignment} procedure is shown in listing \ref{list:integrand_assignment}.

\begin{lstlisting}[firstnumber=1,style=code,caption={integrand assignment procedure interface},label={list:integrand_assignment}]
subroutine assignment_integrand(lhs, rhs)
import :: integrand
class(integrand), intent(INOUT) :: lhs !< Left hand side.
class(integrand), intent(IN)    :: rhs !< Right hand side.
endsubroutine assignment_integrand
\end{lstlisting}

The assignment statement is necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the assignment of an integrand object by another one, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}. The implementation details of this assignment is strictly dependent on the concrete extension of the integrand type. From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the assignment, and the second argument, the right-hand-side of the assignment, are two integrand objects.

\clearpage

\subsection{The explicit forward Euler solver}\label{subsec:solver_euler}

The explicit forward Euler solver is exposed (by the FOODiE main module that must imported, see listing \ref{list:main_module}) as a single derived type (that is a standard convention for all FOODiE solvers) named \emph{euler\_explicit\_integrator}. It provides the type bound procedure (also referred as \emph{method}) \emph{integrate} for integrating in time an \emph{integrand} object, or any of its polymorphic concrete extensions. Consequently, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_euler_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit forward Euler integrator},label={list:define_euler_solver}]
use FOODiE, only: euler_explicit_integrator
type(euler_explicit_integrator) :: integrator
\end{lstlisting}

Once an integrator of this type has been instantiated, it can be directly used without any initialization, for example see listing \ref{list:use_euler_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an explicit forward Euler integrator},label={list:use_euler_solver}]
type(my_integrand) :: my_field
call integrator%integrate(U=my_field, Dt=0.1)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.

The complete implementation of the \emph{integrate} method of the explicit forward Euler solver is reported in listing \ref{list:euler_integrate}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of Euler solver},label={list:euler_integrate}]
subroutine integrate(U, Dt, t)
class(integrand),    intent(INOUT) :: U  !< Field to be integrated.
real(R_P),           intent(IN)    :: Dt !< Time step.
real(R_P), optional, intent(IN)    :: t  !< Time.
U = U + U%t(t=t) * Dt
return
endsubroutine integrate
\end{lstlisting}
This method takes three arguments, the first argument is an integrand class, it being the integrand field that must integrated one-step-over in time, the second is the time step used and the third, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time. The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.

\subsection{The explicit TVD/SSP Runge-Kutta class of solvers}\label{subsec:solver_tvd_rk}

The TVD/SSP Runge-Kutta class of solvers is exposed as a single derived type named \emph{tvd\_runge\_kutta\_integrator}. This type provides three methods:

\begin{itemize}
  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
  \item \emph{integrate}: integrate integrand field one-step-over in time.
  \end{itemize}

As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_tvd_rk_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit TVD/SSP Runge-Kutta integrator},label={list:define_tvd_rk_solver}]
use FOODiE, only: tvd_runge_kutta_integrator
type(tvd_runge_kutta_integrator) :: integrator
\end{lstlisting}

Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_tvd_rk_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit TVD/SSP Runge-Kutta integrator},label={list:init_tvd_rk_solver}]
call integrator%init(stages=3)
\end{lstlisting}

In the listing \ref{list:init_tvd_rk_solver} a 3-stages solver has been initialized. As a matter of facts, from the equations \ref{eq:RK} and \ref{eq:RK-stage} a solver belonging to this class is completely defined once the number of stages adopted has been chosen. The complete definition of the \emph{tvd\_runge\_kutta\_integrator} type is reported into listing \ref{list:tvd_rk_solver}. As shown, the Butcher's coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{tvd\_runge\_kutta\_integrator} type},label={list:tvd_rk_solver}]
type :: tvd_runge_kutta_integrator
  integer(I_P)           :: stages=0  ! Number of stages.
  real(R_P), allocatable :: alph(:,:) ! alpha Butcher's coefficients.
  real(R_P), allocatable :: beta(:)   ! beta Butcher's coefficients.
  real(R_P), allocatable :: gamm(:)   ! gamma Butcher's coefficients.
  contains
    procedure, pass(self), public :: destroy
    procedure, pass(self), public :: init
    procedure, pass(self), public :: integrate
    final                         :: finalize
endtype tvd_runge_kutta_integrator
\end{lstlisting}

After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_tvd_rk_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a TVD/SSP Runge-Kutta integrator},label={list:use_tvd_rk_solver}]
type(my_integrand) :: my_field
type(my_integrand) :: my_stages(1:3)
call integrator%integrate(U=my_field, stage=my_stage, Dt=0.1)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT. Listing \ref{list:use_tvd_rk_solver} shows that the memory registers necessary for storing the Runge-Kutta stages must be supplied by the client code.

The complete implementation of the \emph{integrate} method of the explicit TVD/SSP Runge-Kutta class of solvers is reported in listing \ref{list:tvd_rk_integrate}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit TVD/SSP Runge-Kutta class},label={list:tvd_rk_integrate}]
subroutine integrate(self, U, stage, Dt, t)
class(tvd_runge_kutta_integrator), intent(IN)    :: self      ! Actual RK integrator.
class(integrand),                  intent(INOUT) :: U         ! Field to be integrated.
class(integrand),                  intent(INOUT) :: stage(1:) ! Runge-Kutta stages [1:stages].
real(R_P),                         intent(IN)    :: Dt        ! Time step.
real(R_P),                         intent(IN)    :: t         ! Time.
integer(I_P)                                     :: s         ! First stages counter.
integer(I_P)                                     :: ss        ! Second stages counter.
select type(stage)
class is(integrand)
  do s=1, self%stages
    stage(s) = U
    do ss=1, s - 1
      stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
    enddo
    stage(s) = stage(s)%t(t=t + self%gamm(s) * Dt)
  enddo
  do s=1, self%stages
    U = U +  stage(s) * (Dt * self%beta(s))
  enddo
endselect
return
endsubroutine integrate
\end{lstlisting}

This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time. The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.

It is worth noting that the stages memory registers, namely the array \emph{stage}, must be passed as argument because it is defined as a \emph{not-passed} polymorphic argument, thus we are not allowed to define it as an automatic array of the \emph{integrate} method.

\subsection{The explicit low storage Runge-Kutta class of solvers}\label{subsec:solver_ls_rk}

The low storage variant of Runge-Kutta class of solvers is exposed as a single derived type named \emph{ls\_runge\_kutta\_integrator}. This type provides three methods:

\begin{itemize}
  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
  \item \emph{integrate}: integrate integrand field one-step-over in time.
  \end{itemize}

As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ls_rk_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit low storage Runge-Kutta integrator},label={list:define_ls_rk_solver}]
use FOODiE, only: ls_runge_kutta_integrator
type(ls_runge_kutta_integrator) :: integrator
\end{lstlisting}

Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ls_rk_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit low storage Runge-Kutta integrator},label={list:init_ls_rk_solver}]
call integrator%init(stages=5)
\end{lstlisting}

In the listing \ref{list:init_ls_rk_solver} a 5-stages solver has been initialized. As a matter of facts, from the equation \ref{eq:RK-ls} a solver belonging to this class is completely defined once the number of stages adopted has been chosen. The complete definition of the \emph{ls\_runge\_kutta\_integrator} type is reported into listing \ref{list:ls_rk_solver}. As shown, the Williamson's coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{ls\_runge\_kutta\_integrator} type},label={list:ls_rk_solver}]
type :: ls_runge_kutta_integrator
  integer(I_P)           :: stages=0 ! Number of stages.
  real(R_P), allocatable :: A(:)     ! Low storage *A* coefficients.
  real(R_P), allocatable :: B(:)     ! Low storage *B* coefficients.
  real(R_P), allocatable :: C(:)     ! Low storage *C* coefficients.
  contains
    procedure, pass(self), public :: destroy
    procedure, pass(self), public :: init
    procedure, pass(self), public :: integrate
    final                         :: finalize
endtype ls_runge_kutta_integrator
\end{lstlisting}

After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ls_rk_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a low storage Runge-Kutta integrator},label={list:use_ls_rk_solver}]
type(my_integrand) :: my_field
type(my_integrand) :: my_stages(1:2)
call integrator%integrate(U=my_field, stage=my_stage, Dt=0.1)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT. Listing \ref{list:use_ls_rk_solver} shows that the memory registers necessary for storing the Runge-Kutta stages must be supplied by the client code, as it happens of the TVD/SSP Runge-Kutta class. However, now the registers necessary is always 2, independently on the number of stages used, that in the example considered are 5.

The complete implementation of the \emph{integrate} method of the explicit low storage Runge-Kutta class of solvers is reported in listing \ref{list:ls_rk_integrate}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit low storage Runge-Kutta class},label={list:ls_rk_integrate}]
subroutine integrate(self, U, stage, Dt, t)
class(ls_runge_kutta_integrator), intent(IN)    :: self       ! Actual RK integrator.
class(integrand),                 intent(INOUT) :: U          ! Field to be integrated.
class(integrand),                 intent(INOUT) :: stage(1:2) ! Runge-Kutta registers [1:2].
real(R_P),                        intent(IN)    :: Dt         ! Time step.
real(R_P),                        intent(IN)    :: t          ! Time.
integer(I_P)                                    :: s          ! First stages counter.
select type(stage)
class is(integrand)
  stage(1) = U
  stage(2) = U*0._R_P
  do s=1, self%stages
    stage(2) = stage(2) * self%A(s) + stage(1)%t(t=t + self%C(s) * Dt) * Dt
    stage(1) = stage(1) + stage(2) * self%B(s)
  enddo
  U = stage(1)
endselect
return
endsubroutine integrate
\end{lstlisting}

This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third is the stages array for storing the stages computations, the fourth is the time step used and the fifth, that is optional, is the current time value that is passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time. The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.

It is worth noting that the stages memory registers, namely the array \emph{stage}, must be passed as argument because it is defined as a \emph{not-passed} polymorphic argument, thus we are not allowed to define it as an automatic array of the \emph{integrate} method.

\subsection{The explicit Adams-Bashforth class of solvers}\label{subsec:solver_ab}

The explicit Adams-Bashforth class of solvers is exposed as a single derived type named \emph{adams\_bashforth\_integrator}. This type provides three methods:

\begin{itemize}
  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
  \item \emph{integrate}: integrate integrand field one-step-over in time;
  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
  \end{itemize}

As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ab_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Adams-Bashforth integrator},label={list:define_ab_solver}]
use FOODiE, only: adams_bashforth_integrator
type(adams_bashforth_integrator) :: integrator
\end{lstlisting}

Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ab_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Adams-Bashforth integrator},label={list:init_ab_solver}]
call integrator%init(steps=4)
\end{lstlisting}

In the listing \ref{list:init_ab_solver} a 4-steps solver has been initialized. As a matter of facts, from the equation \ref{eq:AB} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen. The complete definition of the \emph{adams\_bashforth\_integrator} type is reported into listing \ref{list:ab_solver}. As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_bashforth\_integrator} type},label={list:ab_solver}]
type :: adams_bashforth_integrator
  private
  integer(I_P)           :: steps=0 ! Number of time steps.
  real(R_P), allocatable :: b(:)    ! \(b\) coefficients.
  contains
    procedure, pass(self), public :: destroy
    procedure, pass(self), public :: init
    procedure, pass(self), public :: integrate
    procedure, pass(self), public :: update_previous
    final                         :: finalize
endtype adams_bashforth_integrator
\end{lstlisting}

After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ab_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
real               :: times(1:4)
type(my_integrand) :: my_field
type(my_integrand) :: previous(1:4)
call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.

The complete implementation of the \emph{integrate} method of the explicit Adams-Bashforth class of solvers is reported in listing \ref{list:ab_integrate}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
subroutine integrate(self, U, previous, Dt, t, autoupdate)
class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
real(R_P),                         intent(IN)    :: Dt           ! Time steps.
real(R_P),                         intent(IN)    :: t(:)         ! Times.
logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
integer(I_P)                                     :: s            ! Steps counter.
autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
do s=1, self%steps
  U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
enddo
if (autoupdate_) call self%update_previous(U=U, previous=previous)
return
endsubroutine integrate
\end{lstlisting}

This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions. The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.

It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers. This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.

\subsection{The implicit Adams-Moulton class of solvers}\label{subsec:solver_am}

The explicit Adams-Moulton class of solvers is exposed as a single derived type named \emph{adams\_moulton\_integrator}. This type provides three methods:

\begin{itemize}
  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
  \item \emph{integrate}: integrate integrand field one-step-over in time;
  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
  \end{itemize}

As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_am_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an implicit Adams-Moulton integrator},label={list:define_am_solver}]
use FOODiE, only: adams_moulton_integrator
type(adams_moulton_integrator) :: integrator
\end{lstlisting}

Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_am_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an implicit Adams-Moulton integrator},label={list:init_am_solver}]
call integrator%init(steps=3)
\end{lstlisting}

In the listing \ref{list:init_am_solver} a 3-steps solver has been initialized. As a matter of facts, from the equation \ref{eq:AM} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen. The complete definition of the \emph{adams\_moulton\_integrator} type is reported into listing \ref{list:am_solver}. As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_moulton\_integrator} type},label={list:am_solver}]
type :: adams_moulton_integrator
  private
  integer(I_P)           :: steps=-1 ! Number of time steps.
  real(R_P), allocatable :: b(:)     ! \(b\) coefficients.
  contains
    procedure, pass(self), public :: destroy
    procedure, pass(self), public :: init
    procedure, pass(self), public :: integrate
    procedure, pass(self), public :: update_previous
    final                         :: finalize
endtype adams_moulton_integrator
\end{lstlisting}

After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_am_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Moulton integrator},label={list:use_am_solver}]
real               :: times(1:3)
type(my_integrand) :: my_field
type(my_integrand) :: previous(1:3)
call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.

The complete implementation of the \emph{integrate} method of the implicit Adams-Moulton class of solvers is reported in listing \ref{list:am_integrate}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Moulton class},label={list:am_integrate}]
subroutine integrate(self, U, previous, Dt, t, autoupdate)
class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
real(R_P),                         intent(IN)    :: Dt           ! Time steps.
real(R_P),                         intent(IN)    :: t(:)         ! Times.
logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
integer(I_P)                                     :: s            ! Steps counter.
autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
if (autoupdate_) call self%update_previous(U=U, previous=previous)
if (self%steps>0) then
  U = previous(self%steps) + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
  do s=0, self%steps - 1
    U = U + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
  enddo
  if (autoupdate_) call self%update_previous(U=U, previous=previous)
else
  U = U + U%t(t=t(s+1)) * (Dt * self%b(0))
endif
return
endsubroutine integrate
\end{lstlisting}

This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions. The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.

It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers. This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.

\subsection{The predictor-corrector Adams-Bashforth-Moulton class of solvers}\label{subsec:solver_abm}

{\color{red} To be written.}

\subsection{The leapfrog solver}\label{subsec:solver_leapfrog}

The explicit Leapfrog class of solvers is exposed as a single derived type named \emph{leapfrog\_integrator}. This type provides three methods:

\begin{itemize}
  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
  \item \emph{integrate}: integrate integrand field one-step-over in time.
  \end{itemize}

As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_lf_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Leapfrog integrator},label={list:define_lf_solver}]
use FOODiE, only: leapfrog_integrator
type(leapfrog_integrator) :: integrator
\end{lstlisting}

Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_lf_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Leapfrog integrator},label={list:init_lf_solver}]
! default coefficients nu=0.01, alpha=0.53
call integrator%init()
! custom coefficients
call integrator%init(nu=0.015, alpha=0.6)
\end{lstlisting}

The complete definition of the \emph{leapfrog\_integrator} type is reported into listing \ref{list:lf_solver}. As shown, the filter coefficients are initialized to zero, suitable values are initialized by the \emph{init} method.

\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{leapfrog\_integrator} type},label={list:lf_solver}]
type :: leapfrog_integrator
  private
  real(R_P) :: nu=0.01_R_P    ! Robert-Asselin filter coefficient.
  real(R_P) :: alpha=0.53_R_P ! Robert-Asselin-Williams filter coefficient.
  contains
    procedure, pass(self), public :: init
    procedure, pass(self), public :: integrate
endtype leapfrog_integrator
\end{lstlisting}

After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_lf_solver}.

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a Leapfrog integrator},label={list:use_lf_solver}]
real               :: times(1:2)
type(my_integrand) :: filter_displacement
type(my_integrand) :: my_field
type(my_integrand) :: previous(1:2)
call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, &
                          t=times)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{previous} are the memory registers where previous time steps solutions are saved, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.

The complete implementation of the \emph{integrate} method of the explicit Leapfrog class of solvers is reported in listing \ref{list:lf_integrate}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Leapfrog class},label={list:lf_integrate}]
subroutine integrate(self, U, previous, Dt, t, filter)
class(leapfrog_integrator), intent(IN)    :: self          ! LF integrator.
class(integrand),           intent(INOUT) :: U             ! Field to be integrated.
class(integrand),           intent(INOUT) :: previous(1:2) ! Previous time steps solutions.
real(R_P),                  intent(in)    :: Dt            ! Time step.
real(R_P),                  intent(IN)    :: t             ! Time.
class(integrand), optional, intent(INOUT) :: filter        ! Filter field displacement.
U = previous(1) + previous(2)%t(t=t) * (Dt * 2._R_P)
if (present(filter)) then
  filter = (previous(1) - previous(2) * 2._R_P + U) * self%nu * 0.5_R_P
  previous(2) = previous(2) + filter * self%alpha
  U = U + filter * (self%alpha - 1._R_P)
endif
previous(1) = previous(2)
previous(2) = U
return
endsubroutine integrate
\end{lstlisting}

This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the optional filter-displacement-register, the fifth is the time step used and the sixth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time. The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method. It is worth noting that if the filter displacement argument is not passed, the solver reverts back to the standard unfiltered Leapfrog method.

It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers. In particular, if the filter displacement argument is passed the method performs the RAW filtering.

\subsection{General Remarks}\label{subsec:API-remarks}

Table \ref{tab:foodie-vs-math} presents a comparison of the relevant parts of equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:RK-stage}, \ref{eq:RK-ls}, \ref{eq:AB} and \ref{eq:leapfrog} with the corresponding FOODiE implementations reported in listings \ref{list:euler_integrate}, \ref{list:tvd_rk_integrate}, \ref{list:ls_rk_integrate}, \ref{list:ab_integrate} and \ref{list:lf_integrate}, respectively. This comparison proves that the \emph{integrand} ADT has allowed a very high-level implementation syntax. The Fortran implementation is almost equivalent to the rigorous mathematical formulation. This aspect directly implies that the implementation of a ODE solver into the FOODiE library is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the solvers must be implemented for each specific definition of the integrand type, it being not abstract.

\begin{table}[!ht]
  \centering
  \caption{Comparison between rigorous mathematical formulation and FOODiE high-level implementation; the syntax "$(s)$" and "$(ss)$" imply the summation operation}\label{tab:foodie-vs-math}
  \resizebox{0.99\textwidth}{!}{%
  \begin{tabular}{c|cc}
    \toprule
    {\sc Solver} & {\sc Mathematical formulation} & {\sc FOODiE implementation} \\
    \hline
    explicit forward Euler & $U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]$ & $U = U + U\%t(t=t) * Dt$\\
    \hline
    TVD/SSP Runge-Kutta & $\begin{matrix}
  K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right) \\
  U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
\end{matrix}$ & $\begin{matrix}
  stage(s) = stage(s) + stage(ss) * (Dt * self\%alph(s, ss)) \\
  U = U +  stage(s) * (Dt * self\%beta(s))
\end{matrix}$\\
    \hline
    low storage Runge-Kutta & $\begin{matrix}
  K_2 = A_s K_2 + \Delta t \cdot R\left(t^n + C_s \Delta t, K_1\right) \\
  K_1 = K_1 + B_s K_2
\end{matrix}$ & $\begin{matrix}
  stage(2) = stage(2) * self\%A(s) + \\ + stage(1)\%t(t=t + self\%C(s) * Dt) * Dt \\
  stage(1) = stage(1) + stage(2) * self\%B(s)
\end{matrix}$\\
    \hline
  explicit Adams-Bashforth & $\begin{matrix}U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) + \\ +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}\end{matrix}$ & $U = U + U\%t(n=s, t=t(s)) * (Dt * self\%b(s))$\\
    \hline
    explicit Leapfrog & $ U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]$ & $U = U\%previous\_step(n=1) + U\%t(n=2, t=t) * (Dt * 2.)$\\

  \end{tabular}}
\end{table}

### 2015-10-14 papers/announcement/src/introduction/introduction.tex
Minor improvments on manuscript

\section{Introduction}\label{sec:introduction}
\subsection{Background}

Initial Value Problem (IVP, or Cauchy problem) constitutes a class of mathematical models of paramount relevance, it being applied to the modelling of a wide range of problems. Briefly, an IVP is an Ordinary Differential Equations system (ODE) coupled with specified initial values of the unknown state variables, the solution of which are searched at a given time after the initial time considered.

The prototype of IVP can be expressed as:
\begin{equation}
  \begin{matrix}
  U_t = R(t,U) \\
  U_0 = U(t_0)
  \end{matrix}
\label{eq:IVP}
\end{equation}
where $U(t)$ is the vector of state variables being a function of the time-like independent variable $t$, $U_t = \frac{dU}{dt} = R(t,U)$ is the (vectorial) residuals function and $U(t_0)$ is the (vectorial) initial conditions, namely the state variables function evaluated at the initial time $t_0$. In general, the residuals function $R$ is a function of the state variable $U$ through which it is a function of time, but it can be also a direct function of time, thus in general $R=R(t,U(t))$ holds.

The problem prototype \ref{eq:IVP} is ubiquitous in the mathematical modelling of physical problems: essentially whenever an \emph{evolutionary} phenomenon is considered the prevision (simulation) of the future solutions involves the solution of an IVP. As a matter of facts, many physical problems (fluid dynamics, chemistry, biology, evolutionary-anthropology, \dots) are described by means of an IVP.

It is worth noting that the state vector variables $U$ and its corresponding residuals function $U_t = \frac{dU}{dt} = R(t,U)$ are \emph{problem dependent}: the number and the meaning of the state variables as well as the equations governing their evolution (that are embedded into the residuals function) are different for the Navier-Stokes conservation laws with respect the Burgers one, as an example. Nevertheless, the \emph{solvers} used for the prediction of the Navier-Stokes equations evolution (hereafter the \emph{solution}) are the same that are used for Burgers equations time-integration. As a consequence, the solution of the IVP model prototype can be generalized, allowing the application of the same solver to many different problems, thus eliminating the necessity to re-implement the same solver for each different problem.

FOODiE library is designed for solving the generalized IVP \ref{eq:IVP}, it being completely unaware of the actual problem's definition. FOODiE library provides a high-level, well-documented, simple Application Program Interface (API) for many well-known ODE integration solvers, its aims being twofold:

\begin{itemize}
  \item provide a robust set of ODE solvers ready to be applied to a wide range of different problems;
  \item provide a simple framework for the rapid development of new ODE solvers.
\end{itemize}

{\color{red} Add citations to IVP, Cauchy, ODE references.}

\subsection{Related Works}

{\color{red} To be written.}

\subsection{Motivations and Aims}

FOODiE library must:

\begin{itemize}
  \item be written in modern Fortran (standard 2008 or newer);
  \item be written by means of Object Oriented Programming (OOP) paradigm;
  \item be well documented;
  \item be Tests Driven Developed (TDD);
  \item be collaboratively developed;
  \item be free.
\end{itemize}

FOODiE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has been developed with the aim to satisfy the above specifications. The present paper is its first comprehensive presentation.

The Fortran (Formula Translator) programming language is the \emph{de facto} standard into computer science field: it strongly facilitates the effective and efficient translation of (even complex) mathematical and numerical models into an operative software without compromise on computations speed and accuracy. Moreover, its simple syntax is suitable for scientific researchers that are interested (and skilled) in the physical aspects of the numerical computations rather than computer technicians. Consequently, we develop FOODiE using Fortran language: FOODiE is written by research scientists for research scientists.

One key-point of the FOODiE development is the \emph{problem generalization}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its actual definitions. Consequently, we must rely on a \emph{generic} implementation of the solvers. To this aim, OOP is very useful: it allows to express IVP \ref{eq:IVP} in a very concise and clear formulation that is really generic. In particular, our implementation is based on \emph{Abstract Calculus Pattern} (ACP) concept.

\paragraph{The Abstract Calculus Pattern}
The abstract calculus pattern provides a simple solution for the connection between the very high-level expression of IVP \ref{eq:IVP} and the eventual concrete (low-level) implementation of the ODE problem being solved. ACP essentially constitutes a \emph{contract} based on an Abstract Data Type (ADT): we specify an ADT supporting a certain set of mathematical operators (differential and integral ones) and implement FOODiE solvers only on the basis of this ADT. FOODiE clients must formulate the ODE problem under integration defining their own concrete extensions of our ADT (implementing all the deferred operators). Such an approach defines the abstract calculus pattern: FOODiE solvers are aware of only the ADT, while FOODiE clients extend the ADT for defining the concrete ODE problem.

Is is worth noting that this ACP emancipates the solvers implementations from any low-level problem-dependent details: the ODE solvers developed with this pattern are extremely concise, clear, maintainable and less errors-prone with respect a low-level (non abstract) pattern. Moreover, the FOODiE clients can use solvers being extremely robust: as a matter of facts, FOODiE solvers are expressed in a formulation very close to the mathematical one and are tested on an extremely varying family of problems. As shown in the following, such a great flexibility does not compromise the computational efficiency.

The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODiE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODiE applications is presented; section \ref{sec:parallel} provides an analysis of FOODiE performances under parallel frameworks scenario like the OpenMP and MPI paradigms; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.

{\color{red} Add citations to Fortran standards, OOP, TDD, free software, ACP, ADT, Rouson's book references.}

### 2015-10-14 papers/announcement/src/mathematical-numerical-models/mathematical-numerical-models.tex
Minor improvments on manuscript

\section{Mathematical and Numerical Models}\label{sec:MNmodels}

In many (most) circumstances, the solution of equation \ref{eq:IVP} cannot be computed in a closed, exact form (even if it exists and is unique) due to the complexity and nature of the residuals functions, that is often non linear. Consequently, the problem is often solved relying on a numerical approach: the solution of system \ref{eq:IVP} at a time $t^n$, namely $U(t^n)$, is approximated by a subsequent time-marching approximations $U_0=u_0 \rightarrow u_1 \rightarrow u_2 \rightarrow ... \rightarrow u_N\approx U(t^n)$ where the relation $u_i \rightarrow u_{i+1}$ implies a \emph{stepping, numerical integration} from the time $t^i$ to time $t^{i+1}$ and $N$ is the total number of numerical time steps necessary to evolve the initial conditions toward the searched solution $U(t^n)$. To this aim, many numerical schemes have been devised. Notably, the numerical schemes of practical usefulness must posses some necessary proprieties such as \emph{consistency} and \emph{stability} to ensure that the numerical approximation \emph{converges} to the exact solution as the numerical time step tends to zero. A detailed discussion of these details is out the scope of the present work and is omitted. Here, we briefly recall some classifications necessary to introduce the schemes implemented into the FOODiE library.

A non comprehensive classification of the most widely used schemes could distinguish between \emph{multi-step} versus \emph{one-step} schemes and between \emph{explicit} versus \emph{implicit} schemes.

Essentially, the multi-step schemes have been developed to obtain an accurate approximation of the subsequent numerical steps using the informations contained into the previously computed steps, thus this approach relates the next step approximation to a set of the previously computed steps. On the contrary, a one-step scheme evolves the solution toward the next step using only the information coming from the current time approximation. In the framework of one-step schemes family an equivalent accurate approximation can be obtained by means of a multi-stage approach as the one due to Runge-Kutta. The current version of FOODiE provides schemes belonging to both these families.

The other ODE solvers classification concerns with explicit or implicit nature of the schemes employed. Briefly, an explicit scheme computes the next step approximation using the previously computed steps at most to the current time, whereas an implicit scheme uses also the next step approximation (that is the unknown), thus it requires extra computations. The implicit approach is of practical use for \emph{stiff} systems where the usage of explicit schemes could require an extremely small time step to evolve in a \emph{stable} way the solution. Mixing together explicit and implicit schemes it is possible to build a family of \emph{predictor-corrector} methods: using an explicit scheme to predict a guess for the next step approximation it is possible to use an implicit method for correcting this guess. Currently, FOODiE provides explicit solvers and predictor-correct ones.

FOODiE currently implements the following ODE schemes:

\begin{itemize}
  \item one-step schemes:
    \begin{itemize}
      \item explicit forward Euler scheme, it being $1^{st}$ order accurate;
      \item explicit Runge-Kutta schemes:
        \begin{itemize}
          \item TVD/SSP Runge-Kutta schemes:
            \begin{itemize}
              \item 2-stages, it being $2^{nd}$ order accurate;
              \item 3-stages, it being $3^{rd}$ order accurate;
              \item 5-stages, it being $4^{th}$ order accurate;
              \end{itemize}
          \item low storage Runge-Kutta schemes:
            \begin{itemize}
              \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
              \item 6-stages 2N registers schemes, it being $4^{th}$ order accurate;
              \item 7-stages 2N registers schemes, it being $4^{th}$ order accurate;
              \item 12-stages 2N registers schemes, it being $4^{th}$ order accurate;
              \item 13-stages 2N registers schemes, it being $4^{th}$ order accurate;
              \item 14-stages 2N registers schemes, it being $4^{th}$ order accurate;
              \end{itemize}
          \end{itemize}
      \end{itemize}
  \item multi-step schemes:
    \begin{itemize}
      \item explicit Adams-Bashforth schemes:
        \begin{itemize}
          \item 2-steps, it being $2^{nd}$ order accurate;
          \item 3-steps, it being $3^{rd}$ order accurate;
          \item 4-steps, it being $4^{th}$ order accurate;
          \end{itemize}
      \item implicit Adams-Moulton schemes:
        \begin{itemize}
          \item 1-step, it being $2^{nd}$ order accurate;
          \item 2-steps, it being $3^{rd}$ order accurate;
          \item 3-steps, it being $4^{th}$ order accurate;
          \end{itemize}
      \item predictor-corrector Adams-Bashforth-Moulton schemes:
        \begin{itemize}
          \item 1-step, it being $2^{nd}$ order accurate;
          \item 2-steps, it being $3^{rd}$ order accurate;
          \item 3-steps, it being $4^{th}$ order accurate;
          \end{itemize}
      \item explicit Leapfrog schemes\footnote{Leapfrog method is a 2-steps solver being, in general $2^{nd}$ order accurate, but mostly \emph{unstable}. It is well suited for periodic/oscillatory problems, thus the error is generally categorized accordingly to the phase and amplitude errors: the unfiltered scheme provides a $2^{nd}$ order error on phase approximation while the amplitude error is null. However, the unfiltered scheme is mostly unstable, thus many filters have been devised. In general, the filters retain the accuracy on the phase error, while affect the amplitude one.}:
        \begin{itemize}
          \item 2-steps unfiltered, it being $2^{nd}$ order accurate, but mostly \emph{unstable};
          \item 2-steps Robert-Asselin filtered, it being $1^{st}$ order accurate (on \emph{amplitude error});
          \item 2-steps Robert-Asselin-Williams filtered, it being $3^{rd}$ order accurate (on \emph{amplitude error});
          \end{itemize}
      \end{itemize}
  \end{itemize}

{\color{red} Add citations to all solvers reference papers.}

\subsection{The explicit forward Euler scheme}

The explicit forward Euler scheme for ODE integration is probably the simplest solver ever devised. Considering the system \ref{eq:IVP}, the solution (approximation) of the state vector $U$ at the time $t^{n+1}=t^n+\Delta t$ assuming to known the solution at time $t^n$ is:

\begin{equation}
  U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]
\label{eq:solver-euler}
\end{equation}
where the solution at the new time step is computed by means of only the current time solution, thus this is an explicit scheme. The solution is an approximation of $1^{st}$ order, the local truncation error being $O(Dt^2)$. As well known, this scheme has an absolute (linear) stability locus equals to $|1+\Delta t \lambda|\le 1$ where $\lambda$ contains the eigenvalues of the linear (or linearized) Jacobian matrix of the system.

This scheme is Total Variation Diminishing (TVD), thus satisfies the maximum principle (or the equivalent positivity preserving property).

{\color{red} Add citations.}

\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}

Runge-Kutta methods belong to the more general multi-stage family of schemes. This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme, but without increasing the number of time steps used, as it is done with the multi-step schemes. Essentially, the high order of accuracy is obtained by means of \emph{intermediate values} (the stages) of the solution and its derivative are generated and used within a single time step. This commonly implies the allocation of some auxiliary memory registers for storing the intermediate stages.

Notably, the multi-stage schemes class has the attractive property to be \emph{self-starting}: the high order accurate solution can be obtained directly from the previous one, without the necessity to compute \emph{before} a certain number of previous steps, as it happens for the multi-step schemes. Moreover, one-step multi-stage methods are suitable for adaptively-varying time-step size (that is also possible for multi-step schemes, but at a cost of more complexity) and for discontinuous solutions, namely discontinued solutions happening at a certain time $t*$ (that in a multi-step framework can involve an overall accuracy degradation).

In general, the TVD/SSP Runge-Kutta schemes provided by FOODiE library are written by means of the following algorithm:
\begin{equation}
  U^{n+1} = U^{n} + \Delta t \cdot\sum\limits_{s=1}^{N_{s}}{\beta_s K_s}
\label{eq:RK}
\end{equation}
where $N_s$ is the number of Runge-Kutta stages used and $K_s$ is the $s^{th}$ stage defined as:
\begin{equation}
  K_s = R\left(t^n + \gamma_s \Delta t, U^n+\Delta t\sum\limits_{l=1}^{s-1}{\alpha_{s,l} K_l} \right)
\label{eq:RK-stage}
\end{equation}
It is worth noting that the equations \ref{eq:RK} and \ref{eq:RK-stage} contain also implicit schemes. A scheme belonging to this family is operative once the coefficients $\alpha,\, \beta,\, \gamma$ are provided. We represent these coefficients using the Butcher's table, that for an explicit scheme where $\gamma_1=\alpha_{1,*}=\alpha_{i,i}=0$ has the form reported in table \ref{tab:butcher-table}.

\begin{table}[!ht]
\caption{Butcher's table for explicit Runge-Kutta schemes}\label{tab:butcher-table}
\centering
\resizebox{0.40\textwidth}{!}{%
\begin{tabular}{c|ccccc}
 $\gamma_2$     & $\alpha_{2,1}$   &                  &          &                      &              \\
 $\gamma_3$     & $\alpha_{3,1}$   & $\alpha_{3,2}$   &          &                      &              \\
 $\vdots$       & $\vdots$         &                  & $\ddots$ &                      &              \\
 $\gamma_{N_s}$ & $\alpha_{N_s,1}$ & $\alpha_{N_s,2}$ & $\cdots$ & $\alpha_{N_s,N_s-1}$ &              \\
 \hline
                & $\beta_1$        & $\beta_2$        & $\cdots$ & $\beta_{N_s-1}$      & $\beta_{N_s}$\\
\end{tabular}}
\end{table}

The equations \ref{eq:RK} and \ref{eq:RK-stage} show that Runge-Kutta methods do not require any additional differentiations of the ODE system for achieving high order accuracy, rather they require additional evaluations of the residuals function $R$.

The nature of the scheme and the properties of the solutions obtained depend on the number of stages and on the value of the coefficients selected. Currently, FOODiE provides 3 Runge-Kutta schemes having TVD or Strong Stability Preserving (SSP) propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.

\begin{table}[!ht]
  \centering
  \caption{Butcher's table of 2 stages, $2^{nd}$ order, Runge-Kutta TVD scheme\label{tab:RK-TVD-butcher-2}}
  \resizebox{0.15\textwidth}{!}{%
  \begin{tabular}{c|cc}
   1 & 1 & 0       \\
   \hline
       & 1/2 & 1/2   \\
  \end{tabular}}
\end{table}

\begin{table}[!ht]
  \centering
  \caption{Butcher's table of 3 stages, $3^{rd}$ order, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-3}}
  \resizebox{0.20\textwidth}{!}{%
  \begin{tabular}{c|ccc}
   1   & 1   &     &      \\
   1/2 & 1/4 & 1/4 &      \\
   \hline
       & 1/6 & 1/6 & 2/3  \\
  \end{tabular}}
\end{table}

\begin{table}[!ht]
  \centering
  \caption{Butcher's table of 5 stages, $4^{th}$ order accurate, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-5}}
  \resizebox{0.99\textwidth}{!}{%
  \begin{tabular}{c|ccccc}
   0.39175222700392 & 0.39175222700392 &                  &                  &                  &                  \\
   0.58607968896779 & 0.21766909633821 & 0.36841059262959 &                  &                  &                  \\
   0.47454236302687 & 0.08269208670950 & 0.13995850206999 & 0.25189177424738 &                  &                  \\
   0.93501063100924 & 0.06796628370320 & 0.11503469844438 & 0.20703489864929 & 0.54497475021237 &                  \\
   \hline
                    & 0.14681187618661 & 0.24848290924556 & 0.10425883036650 & 0.27443890091960 & 0.22600748319395 \\
  \end{tabular}}
\end{table}
The absolute stability locus depends on the coefficients selected, however, as a general principle, we can assume that greater is the stages number and wider is the stability locus on equal accuracy orders.

It is worth noting that FODDiE also provides a one-stage TVD Runge-Kutta solver that reverts back to the explicit forward Euler scheme: it can be used, for example, into a Recursive Order Reduction (ROR) framework that automatically checks some properties of the solution and, in case, reduces the order of the Runge-Kutta solver until those properties are obtained.

{\color{red} Add citations.}

\subsection{The explicit low storage Runge-Kutta class of schemes}

As aforementioned, standard Runge-Kutta schemes have the drawback to require $N_S$ auxiliary memory registers to store the necessary stages data. In order to make an efficient use of the available limited computer memory, the class of low storage Runge-Kutta scheme was devised. Essentially, the standard Runge-Kutta class (under some specific conditions) can be reformulated allowing a more efficient memory management. Currently FOODiE provides a class of \emph{2N registers} storage Runge-Kutta schemes, meaning that the storage of all stages requires only 2 registers of memory with a \emph{word} length $N$ (namely the length of the state vector) in contrast to the standard formulation where $N_s$ registers of the same length $N$ are required. This is a dramatic improvement of memory efficiency especially for schemes using a high number of stages ($N_s \ge 4$) where the memory necessary is an half with respect the original formulation. Unfortunately, not all standard Runge-Kutta schemes can be reformulated as a low storage one.

Following the Williamson's approach the standard coefficients are reformulated to the coefficients vectors $A$, $B$ and $C$ and the Runge-Kutta algorithm becomes:

\begin{equation}
\begin{matrix}
  K_1 = U\left(t^n\right) \\
  K_2 = 0 \\
  \left.\begin{matrix}
    K_2 = A_s K_2 + \Delta t \cdot R\left(t^n + C_s \Delta t, K_1\right) \\
    K_1 = K_1 + B_s K_2
  \end{matrix}\right\} s=1,2,...N_s\\
  U\left(t^{n+1}\right) = K_1
  \end{matrix}
\label{eq:RK-ls}
\end{equation}

Currently FOODiE provides 5/6/7/12/13/14 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in table \ref{tab:RK-ls}.

\begin{table}[!ht]
  \centering
  \caption{Williamson's table of low storage Runge-Kutta schemes\label{tab:RK-ls}}
  \begin{subtable}[b]{0.45\textwidth}
    \centering
    \caption{5 stages, $4^{th}$ order\label{tab:RK-ls-5}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccc}
      Stage & $A$                                    & $B$                                    & $C$                                   \\
      \hline
      1     & $ 0                                  $ & $\frac{1432997174477}{9575080441755 }$ & $ 0                                 $ \\
      2     & $-\frac{567301805773 }{1357537059087}$ & $\frac{5161836677717}{13612068292357}$ & $\frac{1432997174477}{9575080441755}$ \\
      3     & $-\frac{2404267990393}{2016746695238}$ & $\frac{1720146321549}{2090206949498 }$ & $\frac{2526269341429}{6820363962896}$ \\
      4     & $-\frac{3550918686646}{2091501179385}$ & $\frac{3134564353537}{4481467310338 }$ & $\frac{2006345519317}{3224310063776}$ \\
      5     & $-\frac{1275806237668}{842570457699 }$ & $\frac{2277821191437}{14882151754819}$ & $\frac{2802321613138}{2924317926251}$ \\
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.45\textwidth}
    \centering
    \caption{6 stages, $4^{th}$ order\label{tab:RK-ls-6}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccc}
      Stage & $A$               & $B$              & $C$              \\
      \hline
      1     & $ 0             $ & $0.122000000000$ & $0             $ \\
      2     & $-0.691750960670$ & $0.477263056358$ & $0.122000000000$ \\
      3     & $-1.727127405211$ & $0.381941220320$ & $0.269115878630$ \\
      4     & $-0.694890150986$ & $0.447757195744$ & $0.447717183551$ \\
      5     & $-1.039942756197$ & $0.498614246822$ & $0.749979795490$ \\
      6     & $-1.531977447611$ & $0.186648570846$ & $0.898555413085$ \\
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.45\textwidth}
    \centering
    \caption{7 stages, $4^{th}$ order\label{tab:RK-ls-7}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccc}
      Stage & $A$               & $B$              & $C$              \\
      \hline
      1     & $ 0.000000000000$ & $0.117322146869$ & $0.000000000000$ \\
      2     & $-0.647900745934$ & $0.503270262127$ & $0.117322146869$ \\
      3     & $-2.704760863204$ & $0.233663281658$ & $0.294523230758$ \\
      4     & $-0.460080550118$ & $0.283419634625$ & $0.305658622131$ \\
      5     & $-0.500581787785$ & $0.540367414023$ & $0.582864148403$ \\
      6     & $-1.906532255913$ & $0.371499414620$ & $0.858664273599$ \\
      7     & $-1.450000000000$ & $0.136670099385$ & $0.868664273599$ \\
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.45\textwidth}
    \centering
    \caption{12 stages, $4^{th}$ order\label{tab:RK-ls-12}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccc}
      Stage & $A$                   & $B$                  & $C$                  \\
      \hline
      1     & $ 0.0000000000000000$ & $0.0650008435125904$ & $0.0000000000000000$ \\
      2     & $-0.0923311242368072$ & $0.0161459902249842$ & $0.0650008435125904$ \\
      3     & $-0.9441056581158819$ & $0.5758627178358159$ & $0.0796560563081853$ \\
      4     & $-4.3271273247576394$ & $0.1649758848361671$ & $0.1620416710085376$ \\
      5     & $-2.1557771329026072$ & $0.3934619494248182$ & $0.2248877362907778$ \\
      6     & $-0.9770727190189062$ & $0.0443509641602719$ & $0.2952293985641261$ \\
      7     & $-0.7581835342571139$ & $0.2074504268408778$ & $0.3318332506149405$ \\
      8     & $-1.7977525470825499$ & $0.6914247433015102$ & $0.4094724050198658$ \\
      9     & $-2.6915667972700770$ & $0.3766646883450449$ & $0.6356954475753369$ \\
      10    & $-4.6466798960268143$ & $0.0757190350155483$ & $0.6806551557645497$ \\
      11    & $-0.1539613783825189$ & $0.2027862031054088$ & $0.7143773712418350$ \\
      12    & $-0.5943293901830616$ & $0.2167029365631842$ & $0.9032588871651854$ \\
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.45\textwidth}
    \centering
    \caption{13 stages, $4^{th}$ order\label{tab:RK-ls-13}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccc}
      Stage & $A$                   & $B$                  & $C$                  \\
      \hline
      1     & $ 0.0000000000000000$ & $0.0271990297818803$ & $0.0000000000000000$ \\
      2     & $-0.6160178650170565$ & $0.1772488819905108$ & $0.0271990297818803$ \\
      3     & $-0.4449487060774118$ & $0.0378528418949694$ & $0.0952594339119365$ \\
      4     & $-1.0952033345276178$ & $0.6086431830142991$ & $0.1266450286591127$ \\
      5     & $-1.2256030785959187$ & $0.2154313974316100$ & $0.1825883045699772$ \\
      6     & $-0.2740182222332805$ & $0.2066152563885843$ & $0.3737511439063931$ \\
      7     & $-0.0411952089052647$ & $0.0415864076069797$ & $0.5301279418422206$ \\
      8     & $-0.1797084899153560$ & $0.0219891884310925$ & $0.5704177433952291$ \\
      9     & $-1.1771530652064288$ & $0.9893081222650993$ & $0.5885784947099155$ \\
      10    & $-0.4078831463120878$ & $0.0063199019859826$ & $0.6160769826246714$ \\
      11    & $-0.8295636426191777$ & $0.3749640721105318$ & $0.6223252334314046$ \\
      12    & $-4.7895970584252288$ & $1.6080235151003195$ & $0.6897593128753419$ \\
      13    & $-0.6606671432964504$ & $0.0961209123818189$ & $0.9126827615920843$ \\
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.45\textwidth}
    \centering
    \caption{14 stages, $4^{th}$ order\label{tab:RK-ls-14}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccc}
      Stage & $A$                   & $B$                  & $C$                  \\
      \hline
      1     & $ 0.0000000000000000$ & $0.0367762454319673$ & $0.0000000000000000$ \\
      2     & $-0.7188012108672410$ & $0.3136296607553959$ & $0.0367762454319673$ \\
      3     & $-0.7785331173421570$ & $0.1531848691869027$ & $0.1249685262725025$ \\
      4     & $-0.0053282796654044$ & $0.0030097086818182$ & $0.2446177702277698$ \\
      5     & $-0.8552979934029281$ & $0.3326293790646110$ & $0.2476149531070420$ \\
      6     & $-3.9564138245774565$ & $0.2440251405350864$ & $0.2969311120382472$ \\
      7     & $-1.5780575380587385$ & $0.3718879239592277$ & $0.3978149645802642$ \\
      8     & $-2.0837094552574054$ & $0.6204126221582444$ & $0.5270854589440328$ \\
      9     & $-0.7483334182761610$ & $0.1524043173028741$ & $0.6981269994175695$ \\
      10    & $-0.7032861106563359$ & $0.0760894927419266$ & $0.8190890835352128$ \\
      11    & $ 0.0013917096117681$ & $0.0077604214040978$ & $0.8527059887098624$ \\
      12    & $-0.0932075369637460$ & $0.0024647284755382$ & $0.8604711817462826$ \\
      13    & $-0.9514200470875948$ & $0.0780348340049386$ & $0.8627060376969976$ \\
      14    & $-7.1151571693922548$ & $5.5059777270269628$ & $0.8734213127600976$ \\
    \end{tabular}}
  \end{subtable}\\
\end{table}

Similarly to the TVD/SSP Runge-Kutta class, the low storage class also provides a fail-safe one-stage solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.

{\color{red} Add citations.}

\subsection{The explicit Adams-Bashforth class of schemes}

Adams-Bashforth methods belong to the more general (linear) explicit multi-step family of schemes. This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps. Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.

In general, the Adams-Bashforth schemes provided by FOODiE library are written by means of the following algorithm (for only explicit schemes):

\begin{equation}
U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}
\label{eq:AB}
\end{equation}
where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.

Currently FOODiE provides 2, 3, and 4 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.

\begin{table}[!ht]
  \centering
  \caption{Explicit Adams-Bashforth coefficients\label{tab:AB-coeff}}
  \resizebox{0.30\textwidth}{!}{%
  \begin{tabular}{c|cccc}
    $N_s$ & $b_1$           & $b_2$            & $b_3$            & $b_4$           \\
    \hline
    2     & $-\frac{1}{2}$  & $\frac{3}{2}$    &       /          &       /         \\
    3     & $\frac{5}{12}$  & $-\frac{16}{12}$ & $\frac{23}{12}$  &       /         \\
    4     & $-\frac{9}{24}$ & $\frac{37}{24}$  & $-\frac{59}{24}$ & $\frac{55}{24}$ \\
  \end{tabular}}
\end{table}

Similarly to the Runge-Kutta classes, the Adams-Bashforth class also provides a fail-safe one-step solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.

It is worth noting that for $N_s>1$ the Adams-Bashforth class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided. To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.

{\color{red} Add citations.}

\subsection{The implicit Adams-Moulton class of schemes}

Adams-Moulton methods belong to the more general (linear) implicit multi-step family of schemes. This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps. Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.

In general, the Adams-Moulton schemes provided by FOODiE library are written by means of the following algorithm (for only implicit schemes):

\begin{equation}
  U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=0}^{n+N_s-1}{ b_s \cdot R\left[t^{n+s}, U\left(t^{n+s}\right)\right]} + b_{N_s}\cdot R\left[t^{n+N_s}, U\left(t^{n+N_s}\right)\right]
\label{eq:AM}
\end{equation}
where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.

Currently FOODiE provides 1, 2, and 3 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.

\begin{table}[!ht]
  \centering
  \caption{Implicit Adams-Moulton coefficients\label{tab:AM-coeff}}
  \resizebox{0.30\textwidth}{!}{%
  \begin{tabular}{c|cccc}
    $N_s$ & $b_0$           & $b_1$           & $b_2$           & $b_3$          \\
    \hline
    1     & $\frac{1}{2}$   & $\frac{1}{2}$   &       /         &       /        \\
    2     & $-\frac{1}{12}$ & $\frac{8}{12}$  & $\frac{5}{12}$  &       /        \\
    3     & $\frac{1}{24}$  & $-\frac{5}{24}$ & $\frac{19}{24}$ & $\frac{9}{24}$ \\
  \end{tabular}}
\end{table}

Similarly to the Runge-Kutta and Adams-Bashforth classes, the Adams-Moulton class also provides a fail-safe zero-step solver reverting back to the implicit backward Euler solver, that is useful for ROR-like frameworks.

It is worth noting that for $N_s>1$ the Adams-Moulton class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided. To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.

{\color{red} Add citations.}

\subsection{The predictor-corrector Adams-Bashforth-Moulton class of schemes}

Adams-Bashforth-Moulton methods belong to the more general (linear) predictor-corrector multi-step family of schemes. This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps. Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.

In general, the Adams-Bashforth-Moulton schemes provided by FOODiE library are written by means of the following algorithm:

\begin{equation}
  \begin{matrix}
    U\left(t^{n+N_s^p}\right)_p = U\left(t^{n+N_s^p-1}\right) +\Delta t \sum_{s=1}^{n+N_s^p}{ b_s^p \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}   \\
    U\left(t^{n+N_s^c}\right)_c = U\left(t^{n+N_s^c-1}\right) +\Delta t \sum_{s=0}^{n+N_s^c-1}{ b_s^c \cdot R\left[t^{n+s}, U\left(t^{n+s}\right)\right]} + b_{N_s^c}^c\cdot R\left[t^{n+N_s^p}, U\left(t^{n+N_s^p}\right)_p\right]
  \end{matrix}
\label{eq:ABM}
\end{equation}
where $N_s^{p,c}$ is the number of time steps considered for the Adams-Bashforth predictor/Adams-Moulton corrector (respectively) and $b_s^{p,c}$ are the corresponding linear coefficients selected. Essentially, the Adams-Bashforth prediction $U\left(t^{n+N_s^p}\right)_p$ is corrected by means of the Adams-Moulton correction resulting in $U\left(t^{n+N_s^c}\right)_c$. In order to preserve the formal order of accuracy the relation $N_s^p=N_s^c+1$ always holds.

Currently FOODiE provides $N_s^c=1, 2, 3 \rightarrow N_s^c=2, 3, 4$ steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s^{p,c}$ coefficients are those reported in tables \ref{tab:AB-coeff} and \ref{tab:AM-coeff}.

{\color{red} Add citations.}

\subsection{The leapfrog solver}

The \emph{leapfrog} scheme belongs to the multi-step family, it being formally a centered second order approximation in time. The leapfrog method (in its original formulation) is mostly unstable, however it is well suited for periodic-oscillatory problems providing a null error on the amplitude value and a formal second order error on the phase one, under the satisfaction of the time-step size stable limit. Commonly, the leapfrog methods are said to provide a $2 \Delta t$ computational mode that can generate unphysical, unstable solutions. As consequence, the original leapfrog scheme is generally \emph{filtered} in order to suppress these computational modes.

The unfiltered leapfrog scheme provided by FOODiE is:
\begin{equation}
  U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]
\label{eq:leapfrog}
\end{equation}

FOODiE provides, in a \emph{seamless} API, also filtered leapfrog schemes. A widely used filter is due to Robert and Asselin, that suppress the computational modes at the cost of accuracy reduction resulting into a $1^{st}$ order error in amplitude value. A more accurate filter, able to provide a $3^{rd}$ order error on amplitude, is a modification of the Robert-Asselin filter due to Williams known as Robert-Asselin-Williams (RAW) filter, that filters the approximation of $U\left(t^{n+1}\right)$ and $U\left(t^{n+2}\right)$ by the following scalar coefficient:

\begin{equation}
  \begin{matrix}
    U\left(t^{n+1}\right) = U\left(t^{n+1}\right) + \Delta * \alpha     \\
    U\left(t^{n+2}\right) = U\left(t^{n+2}\right) + \Delta * (\alpha-1) \\
    where \\
    \Delta = \frac{\nu}{2}(U^{n} - 2 U^{n+1} + U^{n+2})
  \end{matrix}
\label{eq:leapfrog-RAW}
\end{equation}

The filter coefficients should be taken as $\nu \in (0,1]$ and $\alpha \in (0.5,1]$. If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude. The default values of the FOODiE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.

{\color{red} Add citations.}

### 2015-10-14 papers/announcement/src/parallel/parallel.tex
Minor improvments on manuscript

\section{Benchmarks on parallel frameworks}\label{sec:parallel}

{\color{red} To be written.}

\subsection{OpenMP benchmark}\label{subsec:openmp}

{\color{red} To be written.}

\subsection{MPI benchmark}\label{subsec:mpi}

{\color{red} To be written.}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-API.tex
Minor improvments on manuscript

\subsubsection{FOODiE aware implementation of an oscillation numerical solver}

The IVP \ref{eq:oscillation} can be easily solved by means of FOODiE library. The first block of a FOODiE aware solution consists to define an \emph{oscillation integrand field} defining a concrete extension of the FOODiE \emph{integrand} type. Listing \ref{list:oscillation_type} reports the implementation of such an integrand field that is contained into the tests suite shipped within the FOODiE library.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} type},label={list:oscillation_type}]
type, extends(integrand) :: oscillation
  private
  integer(I_P)                           :: dims=0   ! Space dimensions.
  real(R_P)                              :: f=0._R_P ! Oscillation frequency (Hz).
  real(R_P), dimension(:),   allocatable :: U        ! Integrand (state) variables, [1:dims].
  contains
    ! auxiliary methods
    procedure, pass(self), public :: init
    procedure, pass(self), public :: output
    ! type\_integrand deferred methods
    procedure, pass(self), public :: t => dOscillation_dt
    procedure, pass(lhs),  public :: integrand_multiply_integrand => &
                                     oscillation_multiply_oscillation
    procedure, pass(lhs),  public :: integrand_multiply_real => oscillation_multiply_real
    procedure, pass(rhs),  public :: real_multiply_integrand => real_multiply_oscillation
    procedure, pass(lhs),  public :: add => add_oscillation
    procedure, pass(lhs),  public :: sub => sub_oscillation
    procedure, pass(lhs),  public :: assign_integrand => oscillation_assign_oscillation
    procedure, pass(lhs),  public :: assign_real => oscillation_assign_real
endtype oscillation
\end{lstlisting}

The \emph{oscillation} field extends the \emph{integrand} ADT making it a concrete type. This derived type is very simple: it has 5 data members for storing the state vector and some auxiliary variables, and it implements all the deferred methods necessary for defining a valid concrete extension of the \emph{integrand} ADT (plus 2 auxiliary methods that are not relevant for our discussion). The key point is here constituted by the implementation of the deferred methods: the \emph{integrand} ADT does not impose any structure for the data members, that are consequently free to be customized by the client code. In this example the data members have a very simple, clean and concise structure:
\begin{itemize}
  \item $dims$ is the number of space dimensions; in the case of equation \ref{eq:oscillation} we have $dims=2$, however the test implementation has been kept more general parametrizing this dimension in order to easily allow future modification of the test-program itself;
  \item $f$ stores the frequency of the oscillatory problem solved, that is here set to $10^{4}$, but it can be changed at runtime in the test-program;
  \item $U$ is the state vector corresponding directly to the state vector of equation \ref{eq:oscillation};
  \end{itemize}

As the listing \ref{list:oscillation_type} shows, the FOODiE implementation strictly corresponds to the mathematical formulation embracing all the relevant mathematical aspects into one derived type, a single \emph{object}. Here we not review the implementation of all deferred methods, this being out of the scope of the present work: the interested reader can see the tests suite sources shipped within the FOODiE library. However, some of these methods are relevant for our discussion, thus they are reviewed.

\paragraph{dOscillation\_dt, the oscillation residuals function}

Probably, the most important methods for an IVP solver is the residuals function. As a matter of facts, the ODE equations are implemented into the residuals function. However, the FOODiE ADT strongly alleviates the subtle problems that could arise when the ODE solver is hard-implemented within the specific ODE equations. As a matter of facts, the \emph{integrand} ADT specifies the precise interface the residuals function must have: if the client code implements a compliant interface, the FOODiE solvers will work as expected, reducing the possible errors location into the ODE equations, having designed the solvers on the ADT and not on the concrete type.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} residuals function},label={list:oscillation_t}]
function dOscillation_dt(self, t) result(dState_dt)
class(oscillation),     intent(IN) :: self      ! Oscillation field.
real(R_P),    optional, intent(IN) :: t         ! Time.
class(integrand),  allocatable     :: dState_dt ! Oscillation field time derivative.
integer(I_P)                       :: dn        ! Time level, dummy variable.
allocate(oscillation :: dState_dt)
select type(dState_dt)
class is(oscillation)
  dState_dt = self
  dState_dt%U(1) = -self%f * self%U(2)
  dState_dt%U(2) =  self%f * self%U(1)
endselect
return
endfunction dOscillation_dt
\end{lstlisting}

Listing \ref{list:oscillation_t} reports the implementation of the oscillation residuals function: it is very clear and concise. Moreover, comparing this listing with the equation \ref{eq:oscillation} the close correspondence between the mathematical formulation and Fortran implementation is evident.

\paragraph{add method, an example of oscillation symmetric operator}

As a prototype of the operators overloading let us consider the \emph{add} operator, it being a prototype of symmetric operators, the implementation of which is presented in listing \ref{list:oscillation_add}.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} add operator},label={list:oscillation_add}]
function add_oscillation(lhs, rhs) result(opr)
class(oscillation), intent(IN) :: lhs ! Left hand side.
class(integrand),   intent(IN) :: rhs ! Right hand side.
class(integrand), allocatable  :: opr ! Operator result.
allocate(oscillation :: opr)
select type(opr)
class is(oscillation)
  opr = lhs
  select type(rhs)
  class is (oscillation)
    opr%U = lhs%U + rhs%U
  endselect
endselect
return
endfunction add_Oscillation
\end{lstlisting}
It is very simple and clear: firstly all the auxiliary data are copied into the operator result, then the state vector of the result is populated with the addiction between the state vectors of the left-hand-side and right-hand-side. This is very intuitive from the mathematical point of view and it helps to reduce implementation errors. Similar implementations are possible for all the other operators necessary to define a valid \emph{intregrand} ADT concrete extension.

\paragraph{assignment of an oscillation object}

The assignment overloading of the \emph{oscillation} type is the last key-method that enforces the conciseness of the FOODiE aware implementation. Listing \ref{list:oscillation_assign} reports the implementation of the assignment overloading. Essentially, to all the data members of the left-hand-side are assigned the values of the corresponding right-hand-side. Notably, for the assignment of the state vector and of the previous time steps solution array we take advantage of the automatic re-allocation of the left-hand-side variables when they are not allocated or allocated differently from the right-hand-side, that is a Fortran 2003 feature. In spite its simplicity, the assignment overloading is a key-method enabling the usage of FOODiE solver: effectively, the assignment between two \emph{integrand} ADT variables is ubiquitous into the solvers implementations, see equation \ref{eq:RK} for example.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{oscillation integrand} assignment},label={list:oscillation_assign}]
subroutine oscillation_assign_oscillation(lhs, rhs)
class(oscillation), intent(INOUT) :: lhs ! Left hand side.
class(integrand),   intent(IN)    :: rhs ! Right hand side.
select type(rhs)
class is (oscillation)
  lhs%dims = rhs%dims
  lhs%f = rhs%f
  if (allocated(rhs%U)) lhs%U = rhs%U
endselect
return
endsubroutine oscillation_assign_oscillation
\end{lstlisting}

\paragraph{FOODiE numerical integration}

Using the above discussed \emph{oscillation} type it is very easy to solve IVP \ref{eq:oscillation} by means of FOODiE library. Listing \ref{list:oscillation_leapfrog} presents the numerical integration of system \ref{eq:oscillation} by means of the Leapfrog RAW-filtered method. In the example, the integration is performed with $10^4$ steps with a fixed $\Delta t=10^2$ until the time $t=10^6$ is reached. The example shows also that for starting a multi-step scheme such as the Leapfrog one a lower-oder or equivalent order one-scheme is necessary: in the example the first 2 steps are computed by means of one-step TVD/SSP Runge-Kutta 2-stages schemes. Note that the memory registers for storing the Runge-Kutta stages and the RAW filter displacement must be handled by the client code. Listing \ref{list:oscillation_leapfrog} demonstrates how it is simple, clear and concise to solve a IVP by FOODiE solvers. Moreover, it proves how it is simple and effective to apply different solvers in a coupled algorithm, that greatly simplify the development of new hybrid solvers for self-adaptive time step size.

\begin{lstlisting}[firstnumber=1,style=code,caption={numerical integration of the \emph{oscillation} system by means of Leapfrog RAW-filtered method},label={list:oscillation_leapfrog}]
use foodie, only: leapfrog_integrator, tvd_runge_kutta_integrator
type(leapfrog_integrator)        :: lf_integrator ! Leapfrog integrator.
type(tvd_runge_kutta_integrator) :: rk_integrator ! Runge-Kutta integrator.
type(oscillation)                :: rk_stage(1:2) ! Runge-Kutta stages.
type(oscillation)                :: previous(1:2) ! Previous time steps solution.
type(oscillation)                :: oscillator    ! Oscillation field.
type(oscillation)                :: filter        ! Filter displacement.
integer                          :: step          ! Time steps counter.
real                             :: Dt            ! Time step.
call lf_integrator%init()
call rk_integrator%init(stages=2)
call oscillator%init(initial_state=[0.0,1.0], f=10e4, steps=2)
Dt = 100.0
do step=1, 10000
  if (2>=step) then
    call rk_integrator%integrate(U=oscillator, stage=rk_stage, Dt=Dt, t=step*Dt)
    previous(step) = oscillator
  else
    call lf_integrator%integrate(U=oscillator, previous=previous, filter=filter, Dt=Dt, &
                                 t=step*Dt)
  endif
enddo
call print_results(U=oscillator)
\end{lstlisting}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-solver-ab.tex
Minor improvments on manuscript

\subsubsection{Adams-Bashforth}

Table \ref{tab:oscillation_errors_ab} summarizes the Adams-Bashforth error analysis. As expected, the Adams-Bashforth 1 step solution, that reverts back to the explicit forward Euler one, is unstable for all the $\Delta t$ exercised.

The expected observed orders of accuracy for the Adams-Bashforth solvers using 2, 3 and 4 time steps tend to 1.5, 2.5 and 3.5 that are in agreement with the expected formal order of 2, 3 and 4, respectively. Comparing the errors of the finest time resolution, i.e. $\Delta t=100$, we find that the L2 norm decreases of the 2 orders of magnitude as the solver's accuracy increases by 1 order. This also means that fixing a tolerance on the errors, the higher is the solver's accuracy the larger is the time resolution available. As an example, assuming that admissible errors are of $O(10^{-2})$ with the 4-steps solver we can use $\Delta t=625$ performing $N_s=t_final/625$ numerical integration steps, whereas using a 3-steps solvers we must adopt $\Delta t=100$ performing $6.25 \times N_s$ numerical integration steps instead of $N_s$. Considering that the computational costs is only slightly affected by the number of previous time steps considered\footnote{Recalling equation \ref{eq:AB} one can observe that there is only one new evaluation of the residuals function $R$ independently of the previous time steps considered. Thus, the computational costs is affected only by the increasing number of residuals summations, the costs of which are typically negligible with respect the cost of $R$ evaluation.}, the accuracy order has strong impact on the overall numerical efficiency: to improve the numerical efficiency reducing the computational costs, the usage of high order Adams-Bashforth solvers with larger time steps should be preferred instead of low order solvers with smaller time steps.

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of explicit Adams-Bashforth solvers}\label{tab:oscillation_errors_ab}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{1 step}\label{tab:oscillation-ab-1}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{2 steps}\label{tab:oscillation-ab-2}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.596E+03    &  0.583E+03    & /             & /             \\
      2500.0          &  0.221E+02    &  0.218E+02    & 4.75          & 4.74          \\
      1250.0          &  0.764E+01    &  0.769E+01    & 1.53          & 1.50          \\
       625.0          &  0.265E+01    &  0.268E+01    & 1.53          & 1.52          \\
       320.0          &  0.968E+00    &  0.981E+00    & 1.51          & 1.50          \\
       100.0          &  0.169E+00    &  0.171E+00    & 1.50          & 1.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{3 steps}\label{tab:oscillation-ab-3}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.857E+01    &  0.854E+01    & /             & /             \\
      2500.0          &  0.391E+01    &  0.386E+01    & 1.13          & 1.14          \\
      1250.0          &  0.825E+00    &  0.814E+00    & 2.24          & 2.25          \\
       625.0          &  0.150E+00    &  0.148E+00    & 2.46          & 2.46          \\
       320.0          &  0.282E-01    &  0.278E-01    & 2.49          & 2.49          \\
       100.0          &  0.154E-02    &  0.152E-02    & 2.50          & 2.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{4 steps}\label{tab:oscillation-ab-4}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.128E+07    &  0.143E+07    & /             & /             \\
      2500.0          &  0.106E+01    &  0.107E+01    & 20.21         & 20.34         \\
      1250.0          &  0.967E-01    &  0.981E-01    &  3.45         &  3.45         \\
       625.0          &  0.859E-02    &  0.871E-02    &  3.49         &  3.49         \\
       320.0          &  0.827E-03    &  0.838E-03    &  3.50         &  3.50         \\
       100.0          &  0.141E-04    &  0.143E-04    &  3.50         &  3.50         \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
\end{table}

Figure \ref{fig:results-oscillation-adams-bashforth} shows, for each solver exercised, the $X(t)$ and $Y(t)$ solution for $t \in [0, 10^6]$: the plots into the figure report a global overview of the solution for all the instants considered (left subplots) and a detailed zoom over the last instants of the integration (right subplots) for evaluating the numerical errors accumulation. For the sake of clarity, the strongly unstable solutions are omitted into the subplots concerning the final integration instants, namely the solutions for large $\Delta t$. Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes the instability generation for some pairs steps number/$\Delta t$. The 2 and 4 steps solutions are instable for $\Delta t=5000 \rightarrow f*\Delta t=0.5$. On the contrary, the 3 steps solution is stable, but the amplitude is dumped and the solution vanishes as the integration proceeds. The 2 and 4 steps solutions show a phase error that decreases as the time resolution increases, whereas 3 steps solution has null phase error.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-1.png}
    \caption{1 step}\label{fig:results-oscillation-adams-bashforth-1}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-2.png}
    \caption{2 steps}\label{fig:results-oscillation-adams-bashforth-2}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-3.png}
    \caption{3 steps}\label{fig:results-oscillation-adams-bashforth-3}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-4.png}
    \caption{4 steps}\label{fig:results-oscillation-adams-bashforth-4}
  \end{subfigure}
  \caption{Oscillation equations solutions computed by means of Adams-Bashforth solvers}\label{fig:results-oscillation-adams-bashforth}
\end{figure}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-solver-abm.tex
Minor improvments on manuscript

\subsubsection{Adams-Bashforth-Moulton}

Table \ref{tab:oscillation_errors_abm} summarizes the Adams-Bashforth-Moulton error analysis. The same considerations done for the Adams-Bashforth solutions can repeated for the Adams-Bashforth-Moulton ones, thus they are omitted for the sake of conciseness. An interesting result concerns the observed errors: the $O(10^{-2})$ error is now obtained with $\Delta t=1250$ for the 4-steps solver, thus it is 2 times faster than the corresponding Adams-Bashforth 4-step solver. Considering that the computational costs of a single Adams-Bashforth-Moulton step is only slightly greater than the corresponding Adamas-Bashforth step, the efficiency increasing is not negligible.

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of predictor-corrector Adams-Bashforth-Moulton solvers}\label{tab:oscillation_errors_abm}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{1 step}\label{tab:oscillation-abm-1}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.241E+20    &  0.266E+20    & /             & /             \\
      2500.0          &  0.664E+11    &  0.716E+11    & 28.44         & 28.47         \\
      1250.0          &  0.952E+06    &  0.100E+07    & 16.09         & 16.12         \\
       625.0          &  0.413E+04    &  0.407E+04    &  7.85         &  7.95         \\
       320.0          &  0.387E+03    &  0.383E+03    &  3.54         &  3.53         \\
       100.0          &  0.145E+03    &  0.145E+03    &  0.84         &  0.83         \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{2 steps}\label{tab:oscillation-abm-2}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.704E+01    &  0.701E+01    & /             & /             \\
      2500.0          &  0.392E+01    &  0.395E+01    & 0.84          & 0.83          \\
      1250.0          &  0.148E+01    &  0.150E+01    & 1.40          & 1.39          \\
       625.0          &  0.526E+00    &  0.534E+00    & 1.49          & 1.49          \\
       320.0          &  0.193E+00    &  0.196E+00    & 1.50          & 1.50          \\
       100.0          &  0.338E-01    &  0.342E-01    & 1.50          & 1.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{3 steps}\label{tab:oscillation-abm-3}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.457E+01    &  0.464E+01    & /             & /             \\
      2500.0          &  0.656E+00    &  0.654E+00    & 2.80          & 2.83          \\
      1250.0          &  0.100E+00    &  0.987E-01    & 2.71          & 2.73          \\
       625.0          &  0.169E-01    &  0.167E-01    & 2.56          & 2.56          \\
       320.0          &  0.314E-02    &  0.310E-02    & 2.52          & 2.51          \\
       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{4 steps}\label{tab:oscillation-abm-4}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.229E+01    &  0.225E+01    & /             & /             \\
      2500.0          &  0.119E+00    &  0.118E+00    & 4.26          & 4.25          \\
      1250.0          &  0.825E-02    &  0.833E-02    & 3.85          & 3.83          \\
       625.0          &  0.671E-03    &  0.681E-03    & 3.62          & 3.61          \\
       320.0          &  0.631E-04    &  0.640E-04    & 3.53          & 3.53          \\
       100.0          &  0.107E-05    &  0.108E-05    & 3.51          & 3.51          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
\end{table}

Figure \ref{fig:results-oscillation-adams-bashforth-moulton} shows similar plots of figure \ref{fig:results-oscillation-adams-bashforth} above discussed. Differently from the Adams-Bashforth class, the amplitude damping feature is now possessed by the 2-steps solver, see plot \ref{fig:results-oscillation-adams-bashforth-moulton-2}, while all solutions show phase errors that decrease as the time resolution increases.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-1.png}
    \caption{1 step}\label{fig:results-oscillation-adams-bashforth-moulton-1}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-2.png}
    \caption{2 steps}\label{fig:results-oscillation-adams-bashforth-moulton-2}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-3.png}
    \caption{3 steps}\label{fig:results-oscillation-adams-bashforth-moulton-3}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-bashforth-moulton-4.png}
    \caption{4 steps}\label{fig:results-oscillation-adams-bashforth-moulton-4}
  \end{subfigure}
  \caption{Oscillation equations solutions computed by means of Adams-Bashforth-Moulton solvers}\label{fig:results-oscillation-adams-bashforth-moulton}
\end{figure}
\clearpage

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-solver-am.tex
Minor improvments on manuscript

\subsubsection{Adams-Moulton}

Table \ref{tab:oscillation_errors_am} summarizes the Adams-Moulton error analysis. The implicit Adams-Moulton solvers behave much like the Adams-Bashforth-Moulton ones: they have similar errors and observed orders for the same formal order considered. However, the implicit Adams-Moulton class uses one less step with respect the corresponding Adams-Bashforth-Moulton class: this could lead to the promise of higher computational efficiency. Notwithstanding, for solving the implicit non-linearity embedded into the Adams-Moulton solvers an iterative algorithm must be employed: for the results presented, a 5 iterations of \emph{fixed point algorithm} have been computed. This strongly reduces the eventual gain of computational efficiency with respect the Adams-Bashforth-Moulton class.

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of explicit Adams-Moulton solvers; the implicit non-linearity is solved by 5 iterations of \emph{fixed point algorithm}}\label{tab:oscillation_errors_am}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{1 step}\label{tab:oscillation-am-1}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{2 steps}\label{tab:oscillation-am-2}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.108E+02    &  0.109E+02    & /             & /             \\
      2500.0          &  0.412E+01    &  0.419E+01    & 1.39          & 1.38          \\
      1250.0          &  0.148E+01    &  0.150E+01    & 1.48          & 1.48          \\
       625.0          &  0.527E+00    &  0.533E+00    & 1.49          & 1.49          \\
       320.0          &  0.193E+00    &  0.196E+00    & 1.50          & 1.50          \\
       100.0          &  0.338E-01    &  0.342E-01    & 1.50          & 1.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{3 steps}\label{tab:oscillation-am-3}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.390E+01    &  0.384E+01    & /             & /             \\
      2500.0          &  0.551E+00    &  0.544E+00    & 2.82          & 2.82          \\
      1250.0          &  0.947E-01    &  0.934E-01    & 2.54          & 2.54          \\
       625.0          &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
       320.0          &  0.313E-02    &  0.309E-02    & 2.50          & 2.50          \\
       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{4 steps}\label{tab:oscillation-am-4}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.983E+00    &  0.999E+00    & /             & /             \\
      2500.0          &  0.832E-01    &  0.845E-01    & 3.56          & 3.56          \\
      1250.0          &  0.736E-02    &  0.746E-02    & 3.50          & 3.50          \\
       625.0          &  0.652E-03    &  0.660E-03    & 3.50          & 3.50          \\
       320.0          &  0.626E-04    &  0.635E-04    & 3.50          & 3.50          \\
       100.0          &  0.107E-05    &  0.108E-05    & 3.50          & 3.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
\end{table}

Figure \ref{fig:results-oscillation-adams-moulton} shows similar plots of figure \ref{fig:results-oscillation-adams-bashforth-moulton} above discussed: there are not relevant differences between the 2 classes of solvers.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-moulton-0.png}
    \caption{0 step}\label{fig:results-oscillation-adams-moulton-0}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-moulton-1.png}
    \caption{1 step}\label{fig:results-oscillation-adams-moulton-1}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-moulton-2.png}
    \caption{2 steps}\label{fig:results-oscillation-adams-moulton-2}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-adams-moulton-3.png}
    \caption{3 steps}\label{fig:results-oscillation-adams-moulton-3}
  \end{subfigure}
  \caption{Oscillation equations solutions computed by means of Adams-Moulton solvers}\label{fig:results-oscillation-adams-moulton}
\end{figure}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-solver-lf.tex
Minor improvments on manuscript

\subsubsection{Leapfrog}

The Leapfrog solutions are in agreement with the expected results: both unfiltered and RAW-filtered solutions show an observed order of accuracy that tends to the formal $2^{nd}$ order. The two solutions are almost the same.

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of explicit Leapfrog solvers}\label{tab:oscillation_errors_lf}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{Unfiltered}\label{tab:oscillation-leapfrog}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.156E+02    &  0.156E+02    & /             & /             \\
      2500.0          &  0.849E+01    &  0.846E+01    & 0.87          & 0.88          \\
      1250.0          &  0.300E+01    &  0.303E+01    & 1.50          & 1.48          \\
       625.0          &  0.106E+01    &  0.107E+01    & 1.51          & 1.50          \\
       320.0          &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
       100.0          &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{RAW-filtered}\label{tab:oscillation-leapfrog-raw}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.156E+02    &  0.156E+02    & /             & /             \\
      2500.0          &  0.855E+01    &  0.852E+01    & 0.86          & 0.87          \\
      1250.0          &  0.303E+01    &  0.305E+01    & 1.50          & 1.48          \\
       625.0          &  0.107E+01    &  0.108E+01    & 1.51          & 1.50          \\
       320.0          &  0.390E+00    &  0.395E+00    & 1.50          & 1.50          \\
       100.0          &  0.685E-01    &  0.692E-01    & 1.50          & 1.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
\end{table}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog.png}
    \caption{Unfiltered}\label{fig:results-oscillation-leapfrog-unfiltered}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-leapfrog-raw.png}
    \caption{RAW-filtered}\label{fig:results-oscillation-leapfrog-raw}
  \end{subfigure}
  \caption{Oscillation equations solutions computed by means of Leapfrog solvers}\label{fig:results-oscillation-leapfrog}
\end{figure}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-solver-ls-rk.tex
Minor improvments on manuscript

\subsubsection{Low Storage Runge-Kutta}

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of explicit Low Storage Runge-Kutta solvers}\label{tab:oscillation_errors_ls_rk}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{1 stage}\label{tab:oscillation-ls-rk-1}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{5 stages}\label{tab:oscillation-ls-rk-5}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.120E+00    &  0.122E+00    & /             & /             \\
      2500.0          &  0.106E-01    &  0.107E-01    & 3.51          & 3.51          \\
      1250.0          &  0.935E-03    &  0.947E-03    & 3.50          & 3.50          \\
       625.0          &  0.826E-04    &  0.836E-04    & 3.50          & 3.50          \\
       320.0          &  0.793E-05    &  0.803E-05    & 3.50          & 3.50          \\
       100.0          &  0.135E-06    &  0.137E-06    & 3.50          & 3.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{6 stages}\label{tab:oscillation-ls-rk-6}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.979E-01    &  0.994E-01    & /             & /             \\
      2500.0          &  0.876E-02    &  0.888E-02    & 3.48          & 3.48          \\
      1250.0          &  0.776E-03    &  0.786E-03    & 3.50          & 3.50          \\
       625.0          &  0.686E-04    &  0.695E-04    & 3.50          & 3.50          \\
       320.0          &  0.659E-05    &  0.667E-05    & 3.50          & 3.50          \\
       100.0          &  0.112E-06    &  0.114E-06    & 3.50          & 3.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{7 stages}\label{tab:oscillation-ls-rk-7}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.238E-01    &  0.240E-01    & /             & /             \\
      2500.0          &  0.203E-02    &  0.205E-02    & 3.55          & 3.55          \\
      1250.0          &  0.177E-03    &  0.180E-03    & 3.51          & 3.51          \\
       625.0          &  0.156E-04    &  0.158E-04    & 3.50          & 3.50          \\
       320.0          &  0.150E-05    &  0.152E-05    & 3.50          & 3.50          \\
       100.0          &  0.269E-07    &  0.273E-07    & 3.46          & 3.46          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{12 stages}\label{tab:oscillation-ls-rk-12}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.195E-01    &  0.198E-01    & /             & /             \\
      2500.0          &  0.175E-02    &  0.177E-02    & 3.48          & 3.48          \\
      1250.0          &  0.155E-03    &  0.157E-03    & 3.50          & 3.50          \\
       625.0          &  0.137E-04    &  0.139E-04    & 3.50          & 3.50          \\
       320.0          &  0.132E-05    &  0.133E-05    & 3.50          & 3.50          \\
       100.0          &  0.225E-07    &  0.228E-07    & 3.50          & 3.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{13 stages}\label{tab:oscillation-ls-rk-13}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.795E-02    &  0.805E-02    & /             & /             \\
      2500.0          &  0.703E-03    &  0.712E-03    & 3.50          & 3.50          \\
      1250.0          &  0.621E-04    &  0.629E-04    & 3.50          & 3.50          \\
       625.0          &  0.549E-05    &  0.556E-05    & 3.50          & 3.50          \\
       320.0          &  0.527E-06    &  0.534E-06    & 3.50          & 3.50          \\
       100.0          &  0.899E-08    &  0.911E-08    & 3.50          & 3.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{14 stages}\label{tab:oscillation-ls-rk-14}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.849E-02    &  0.860E-02    & /             & /             \\
      2500.0          &  0.750E-03    &  0.759E-03    & 3.50          & 3.50          \\
      1250.0          &  0.662E-04    &  0.671E-04    & 3.50          & 3.50          \\
       625.0          &  0.585E-05    &  0.593E-05    & 3.50          & 3.50          \\
       320.0          &  0.562E-06    &  0.569E-06    & 3.50          & 3.50          \\
       100.0          &  0.959E-08    &  0.972E-08    & 3.50          & 3.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
\end{table}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-1.png}
    \caption{1 stage}\label{fig:results-oscillation-ls-runge-kutta-1}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-5.png}
    \caption{5 stages}\label{fig:results-oscillation-ls-runge-kutta-5}
  \end{subfigure}
  \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta solvers}\label{fig:results-oscillation-ls-runge-kutta-1-5}
\end{figure}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-6.png}
    \caption{6 stages}\label{fig:results-oscillation-ls-runge-kutta-6}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-7.png}
    \caption{7 stages}\label{fig:results-oscillation-ls-runge-kutta-7}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-12.png}
    \caption{12 stages}\label{fig:results-oscillation-ls-runge-kutta-12}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-13.png}
    \caption{13 stages}\label{fig:results-oscillation-ls-runge-kutta-13}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-ls-runge-kutta-14.png}
    \caption{14 stages}\label{fig:results-oscillation-ls-runge-kutta-14}
  \end{subfigure}\quad%
  \caption{Oscillation equations solutions computed by means of low storage Runge-Kutta solvers}\label{fig:results-oscillation-ls-runge-kutta-6-14}
\end{figure}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation-solver-tvd-rk.tex
Minor improvments on manuscript

\subsubsection{TVD/SSP Runge-Kutta}

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of explicit TVD/SSP Runge-Kutta}\label{tab:oscillation_errors_tvd_rk}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{1 stage}\label{tab:oscillation-tvd-rk-1}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.840E+10    &  0.706E+10    & /             & /             \\
      2500.0          &  0.503E+06    &  0.570E+06    & 14.03         & 13.60         \\
      1250.0          &  0.289E+04    &  0.272E+04    &  7.45         &  7.71         \\
       625.0          &  0.239E+03    &  0.232E+03    &  3.59         &  3.55         \\
       320.0          &  0.737E+02    &  0.722E+02    &  1.76         &  1.74         \\
       100.0          &  0.250E+02    &  0.247E+02    &  0.93         &  0.92         \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{2 stages}\label{tab:oscillation-tvd-rk-2}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.316E+02    &  0.319E+02    & /             & /             \\
      2500.0          &  0.892E+01    &  0.894E+01    & 1.83          & 1.84          \\
      1250.0          &  0.301E+01    &  0.305E+01    & 1.57          & 1.55          \\
       625.0          &  0.106E+01    &  0.107E+01    & 1.51          & 1.51          \\
       320.0          &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
       100.0          &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{3 stages}\label{tab:oscillation-tvd-rk-3}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.255E+01    &  0.252E+01    & /             & /             \\
      2500.0          &  0.523E+00    &  0.516E+00    & 2.28          & 2.29          \\
      1250.0          &  0.944E-01    &  0.931E-01    & 2.47          & 2.47          \\
       625.0          &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
       320.0          &  0.314E-02    &  0.310E-02    & 2.50          & 2.50          \\
       100.0          &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\quad%
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{5 stages}\label{tab:oscillation-tvd-rk-5}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      5000.0          &  0.139E+00    &  0.141E+00    & /             & /             \\
      2500.0          &  0.122E-01    &  0.124E-01    & 3.50          & 3.50          \\
      1250.0          &  0.108E-02    &  0.110E-02    & 3.50          & 3.50          \\
       625.0          &  0.956E-04    &  0.969E-04    & 3.50          & 3.50          \\
       320.0          &  0.937E-05    &  0.949E-05    & 3.47          & 3.47          \\
       100.0          &  0.512E-06    &  0.519E-06    & 2.50          & 2.50          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}
\end{table}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-1.png}
    \caption{1 stage}\label{fig:results-oscillation-tvd-runge-kutta-1}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-2.png}
    \caption{2 stages}\label{fig:results-oscillation-tvd-runge-kutta-2}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-3.png}
    \caption{3 stages}\label{fig:results-oscillation-tvd-runge-kutta-3}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-tvd-runge-kutta-5.png}
    \caption{5 stages}\label{fig:results-oscillation-tvd-runge-kutta-5}
  \end{subfigure}\\
  \caption{Oscillation equations solutions computed by means of TVD/SSP Runge-Kutta solvers}\label{fig:results-oscillation-tvd-runge-kutta}
\end{figure}

### 2015-10-14 papers/announcement/src/tests/oscillation/oscillation.tex
Minor improvments on manuscript

\subsection{Oscillation equations test}

Let us consider the \emph{oscillation} problem, it being a simple, yet interesting IVP. Briefly, the oscillation problem is a prototype problem of non dissipative, oscillatory phenomena. For example, let us consider a pendulum subjected to the Coriolis accelerations without dissipation, the motion equations of which can be described by the ODE system \ref{eq:oscillation}.

\begin{equation}
\begin{matrix}
U_t = R(U)  \\
U = \begin{bmatrix}
x \\
y
\end{bmatrix}\;\;\;
R(U) = \begin{bmatrix}
-f y \\
f x
\end{bmatrix}
\end{matrix}
\label{eq:oscillation}
\end{equation}
where the frequency is chosen as $f=10^4$. The ODE system \ref{eq:oscillation} is completed by the following initial conditions:

\begin{equation}
\begin{matrix}
  x(t_0) = 0 \\
  y(t_0) = 1
\end{matrix}
\label{eq:oscillation-ic}
\end{equation}
where $t_0=0$ is the initial time considered.

The IVP constituted by equations \ref{eq:oscillation} and \ref{eq:oscillation-ic} is (apparently) simple and its exact solution is known:

\begin{equation}
\begin{matrix}
  x(t_0 + \Delta t) = X_0 cos(f \Delta t) - y_0 sin(f \Delta t) \\
  y(t_0 + \Delta t) = X_0 sin(f \Delta t) + y_0 cos(f \Delta t)
\end{matrix}
\label{eq:oscillation-exact}
\end{equation}
where $\Delta t$ is an arbitrary time step. This problem is non-stiff meaning that the solution is constituted by only one time-scale, namely the single frequency $f$.

This problem is only apparently simple. As a matter of facts, in a non dissipative oscillatory problem the eventual errors in the amplitude approximation can rapidly drive the subsequent series of approximations to an unphysical solution. This is of particular relevance if the solution (that is numerically approximated) constitutes a \emph{prediction} far from the initial conditions, that is the common case in weather forecasting.

Because the Oscillation system \ref{eq:oscillation} posses a closed exact solution, the discussion on this test has twofolds aims: to assess the accuracy of the FOODiE's built-in solvers comparing the numerical solutions with the exact one and to demonstrate how it is simple to solve this prototypical problem by means of FOODiE.

\subsubsection{Errors Analysis}

For the analysis of the accuracy of each solver, we have integrated the Oscillation equations \ref{eq:oscillation} with different, decreasing time steps in the range $[5000, 2500, 1250, 625, 320, 100]$. The error is estimated by the L2 norm of the difference between the exact ($U_e$) and the numerical ($U_{\Delta t}$) solutions for each time step:

\begin{equation}
  \varepsilon (\Delta t) = || U_e - U_{\Delta t} ||_2 = \sqrt{ \sum_{s=1}^{N_s} { \left(U_e(t_0 + s * \Delta t) - U_{\Delta t}(t_0 + s * \Delta t) \right)^2 }}
\label{eq:oscillation-error}
\end{equation}
where $N_s$ is the total number of time steps performed to reach the final integration time.

Using two pairs of subsequent-decreasing time steps solution is possible to estimate the order of accuracy of the solver employed computing the \emph{observed order} of accuracy:

\begin{equation}
  p = \frac{log10 \left( \frac{\varepsilon (\Delta t_1)}{\varepsilon (\Delta t_2)} \right)}{log10 \left( \frac{\Delta t_1}{\Delta t_2} \right)}
\label{eq:oscillation-observed-order}
\end{equation}

where $\frac{\Delta t_1}{\Delta t_2}>1$.

\input{src/tests/oscillation/oscillation-API.tex}

\input{src/tests/oscillation/oscillation-solver-ab.tex}

\input{src/tests/oscillation/oscillation-solver-abm.tex}

\input{src/tests/oscillation/oscillation-solver-am.tex}

\input{src/tests/oscillation/oscillation-solver-lf.tex}

\input{src/tests/oscillation/oscillation-solver-ls-rk.tex}

\input{src/tests/oscillation/oscillation-solver-tvd-rk.tex}

### 2015-10-14 papers/announcement/src/tests/tests.tex
Minor improvments on manuscript

\section{Tests and Examples}\label{sec:Tests}

For the assessment of FOODiE capabilities three different tests are discussed:

\begin{itemize}
  \item Oscillation equations IVP: this test is a pure ODEs, non-stiff problem;
  \item Lorenz equations IVP: this test is a pure ODEs, stiff problem;
  \item 1D Euler equations IVP: this test is a PDEs problem;
  \end{itemize}

With the first two pure ODEs tests we assess the accuracy of the FOODiE built-in solvers and demonstrate how to use the library in real applications. The last test discussion is mainly devoted to the demonstration of FOODiE application into a real, complex PDEs problem.

\input{src/tests/oscillation/oscillation.tex}

### 2015-10-14 src/lib/foodie_integrator_adams_moulton.f90

@@ -136,7 +136,7 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine destroy
 
-  subroutine integrate(self, U, previous, Dt, t, autoupdate)
+  subroutine integrate(self, U, previous, Dt, t, iterations, autoupdate)
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Integrate field with Adams-Moulton class scheme.
   !---------------------------------------------------------------------------------------------------------------------------------
@@ -145,18 +145,30 @@ contains
   class(integrand),                intent(INOUT) :: previous(1:) !< Previous time steps solutions of integrand field.
   real(R_P),                       intent(IN)    :: Dt           !< Time steps.
   real(R_P),                       intent(IN)    :: t(:)         !< Times.
-  logical, optional,               intent(IN)    :: autoupdate   !< Perform cyclic autoupdate of previous time steps.
+  integer(I_P), optional,          intent(IN)    :: iterations   !< Fixed point iterations.
+  logical,      optional,          intent(IN)    :: autoupdate   !< Perform cyclic autoupdate of previous time steps.
   logical                                        :: autoupdate_  !< Perform cyclic autoupdate of previous time steps, dummy var.
+  class(integrand), allocatable                  :: delta        !< Delta RHS for fixex point iterations.
   integer(I_P)                                   :: s            !< Steps counter.
   !---------------------------------------------------------------------------------------------------------------------------------
 
   !---------------------------------------------------------------------------------------------------------------------------------
   autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
   if (self%steps>0) then
-    U = previous(self%steps) + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
-    do s=0, self%steps - 1
-      U = U + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
-    enddo
+    if (present(iterations)) then ! perform fixed point iterations
+      allocate(delta, source=previous(self%steps))
+      do s=0, self%steps - 1
+        delta = delta + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
+      enddo
+      do s=1, iterations
+        U = delta + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
+      enddo
+    else
[…]

### 2015-10-14 src/tests/oscillation/oscillation.f90

@@ -10,6 +10,7 @@ use type_oscillation, only : oscillation
 use Data_Type_Command_Line_Interface, only : Type_Command_Line_Interface
 use foodie, only : adams_bashforth_integrator, &
                    adams_bashforth_moulton_integrator, &
+                   adams_moulton_integrator, &
                    euler_explicit_integrator, &
                    leapfrog_integrator, &
                    ls_runge_kutta_integrator, &
@@ -27,6 +28,7 @@ real(R_P)                         :: t_final                                    
 real(R_P)                         :: Dt                                                !< Time step.
 integer(I_P)                      :: error                                             !< Error handler.
 integer(I_P)                      :: stages_steps                                      !< Number of stages/steps used.
+integer(I_P)                      :: implicit_iterations                               !< Number of iterations for implicit solvers.
 character(99)                     :: solver                                            !< Solver used.
 logical                           :: plots                                             !< Flag for activating plots saving.
 logical                           :: results                                           !< Flag for activating results saving.
@@ -46,6 +48,7 @@ call cli%init(progname    = 'oscillation',                                      
                              "oscillation --solver all --plots -r   "])
 call cli%add(switch='--solver', switch_ab='-s', help='ODE solver used', required=.false., def='all', act='store')
 call cli%add(switch='--ss', help='Stages/steps used', required=.false., act='store', def='-1')
+call cli%add(switch='--iterations', help='Number of iterations for implicit solvers', required=.false., act='store', def='5')
 call cli%add(switch='--frequency', switch_ab='-f', help='Oscillation frequency', required=.false., def='1e-4', act='store')
 call cli%add(switch='--time_step', switch_ab='-Dt', help='Integration time step', required=.false., def='100.d0', act='store')
 call cli%add(switch='--t_final', switch_ab='-tf', help='Final integration time', required=.false., def='1e6', act='store')
@@ -57,6 +60,7 @@ call cli%add(switch='--errors_analysis', help='Peform errors analysis', required
 call cli%parse(error=error)
 call cli%get(switch='-s', val=solver, error=error) ; if (error/=0) stop
 call cli%get(switch='--ss', val=stages_steps, error=error) ; if (error/=0) stop
[…]

### 2015-10-14 src/foodie_integrator_adams_bashforth_moulton.f90
Travis CI autocommit from travis build 64

@@ -112,7 +112,7 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Create the actual Adams-Bashforth-Moulton integrator: initialize the *b* coefficients.
   !---------------------------------------------------------------------------------------------------------------------------------
-  class(adams_bashforth_moulton_integrator), intent(INOUT) :: self  !< AB integrator.
+  class(adams_bashforth_moulton_integrator), intent(INOUT) :: self  !< ABM integrator.
   integer(I_P), intent(IN)                                 :: steps !< Number of time steps used.
   !---------------------------------------------------------------------------------------------------------------------------------
 
@@ -128,7 +128,7 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Destoy the integrator.
   !---------------------------------------------------------------------------------------------------------------------------------
-  class(adams_bashforth_moulton_integrator), intent(INOUT) :: self !< AB integrator.
+  class(adams_bashforth_moulton_integrator), intent(INOUT) :: self !< ABM integrator.
   !---------------------------------------------------------------------------------------------------------------------------------
 
   !---------------------------------------------------------------------------------------------------------------------------------
@@ -139,21 +139,22 @@ contains
   !---------------------------------------------------------------------------------------------------------------------------------
   endsubroutine destroy
 
-  subroutine integrate(self, U, previous, Dt, t)
+  subroutine integrate(self, U, previous, Dt, t, iterations)
   !---------------------------------------------------------------------------------------------------------------------------------
-  !< Integrate field with Adams-Bashforth_Moulton class scheme.
+  !< Integrate field with Adams-Bashforth-Moulton class scheme.
   !---------------------------------------------------------------------------------------------------------------------------------
-  class(adams_bashforth_moulton_integrator), intent(IN)    :: self         !< Actual AB integrator.
+  class(adams_bashforth_moulton_integrator), intent(IN)    :: self         !< Actual ABM integrator.
   class(integrand),                          intent(INOUT) :: U            !< Field to be integrated.
   class(integrand),                          intent(INOUT) :: previous(1:) !< Previous time steps solutions of integrand field.
   real(R_P),                                 intent(IN)    :: Dt           !< Time steps.
   real(R_P),                                 intent(IN)    :: t(:)         !< Times.
+  integer(I_P), optional,                    intent(IN)    :: iterations   !< Fixed point iterations of AM scheme.
   integer(I_P)                                             :: s            !< Steps counter.
[…]

### 2015-10-16 papers/announcement/src/parallel/openmp.tex
improve openmp benchmark

\subsection{OpenMP benchmark}\label{subsec:openmp}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.95\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-scaling.png}
    \caption{Strong scaling, number of cells 5000}\label{fig:strong-scaling-openmp}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.95\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-scaling.png}
    \caption{Weak scaling, minimum number of cells 500}\label{fig:weak-scaling-openmp}
  \end{subfigure}\\
  \caption{Scaling efficiency with OpenMP}\label{fig:scaling-openmp}
\end{figure}

### 2015-10-16 papers/announcement/errors_analysis/errors_analysis.sh

@@ -1,9 +1,10 @@
 #!/bin/bash
 # script automotion for FOODiE's errors analysis of the provided tests
-cd ../../../
 echo "Building tests executables"
-FoBiS.py build -mode errors-analysis --build_dir papers/announcement/errors_analysis/tests > buils.log
+cd ../../../
+FoBiS.py build -mode errors-analysis --build_dir papers/announcement/errors_analysis/tests > builds.log
 cd -
+
 echo "Oscillation test errors analysis"
 rm -rf results-oscillation
 mkdir results-oscillation
@@ -12,10 +13,11 @@ cd results-oscillation
 tecplot=$(which tec360)
 if [ -x "$tecplot" ] ; then
   ln -fs ../utilities-oscillation/*lay .
-  ln -fs ../utilities/lay_export* .
+  ln -fs ../../utilities/lay_export* .
   for file in $( ls *lay ); do
   ./lay_export_all_f $file
   done
 fi
 cd -
+
 exit 0

### 2015-10-16 papers/announcement/openmp_benchmark/openmp_benchmark.sh

@@ -1,15 +1,138 @@
 #!/bin/bash
 # script automotion for FOODiE's OpenMP benchmark using provided tests
-cd ../../../
+
+# defaults
+bench='all'
+ripetitions=1
+Ni=100
+
+# functions
+function print_usage {
+  # print help message for correct CLI usage
+  echo
+  echo "`basename $0`"
+  echo "script automotion for FOODiE OpenMP benchmark"
+  echo "Usage: `basename $0` [opts [args]]"
+  echo "  [ -b #benchmark_type -r #ripetitions -Ni #cells ]"
+  echo
+  echo "Defaults of optional arguments:"
+  echo "  -b all # type of benchmark: valid value are 'all', 'strong', 'weak'"
+  echo "  -r 1 # number of ripetitions for averaging the benchmarks results"
+  echo "  -Ni 1000 # test size for strong scaling or minimum test size for weak scaling benchamrk"
+  echo
+  echo "Examples:"
+  echo "  `basename $0` # perform $ripetitions ripetitions with default test size of $Ni cells"
+  echo "  `basename $0` -r 5 # perform 5 ripetitions and compute the average"
+  echo "  `basename $0` -Ni 2000 -r 10 # use a test size of 2000 cells"
+  echo
+}
+function run_exe {
+  # perform the code execution
+  exe=$1
+  thd=$2
+  size=$3
+  average_cpu_time=0.
+  for r in $( seq $ripetitions ); do
+    cpu_time=$($exe --Ni $size --omp_threads $thd | awk '{print $2}')
+    average_cpu_time=`echo "scale=10; $average_cpu_time+$cpu_time" | bc -l`
+  done
+  average_cpu_time=`echo "scale=10; $average_cpu_time/$ripetitions" | bc -l`
+  echo $average_cpu_time
+}
+
+# parsing CLI
+while [ $# -gt 0 ]; do
+  case "$1" in
+  "-b")
+    shift; bench=$1
+    if [ "$bench" != "all" ] && [ "$bench" != "strong" ] && [ "$bench" != "weak" ] ; then
+      echo; echo "invalid value $1"; print_usage; exit 1
+    fi
+    ;;
+  "-r")
+    shift; ripetitions=$1
+    ;;
+  "-Ni")
+    shift; Ni=$1
+    ;;
+  *)
[…]

### 2015-10-16 src/tests/euler-1D-openmp/euler-1D-openmp.f90

@@ -90,10 +90,8 @@ call omp_set_num_threads(omp_threads)
 !$OMP SHARED(omp_threads)
 omp_threads = omp_get_num_threads()
 !$OMP END PARALLEL
-print "(A)", 'Testing FOODiE with '//trim(str(.true., omp_threads))//' OpenMP threads for Ni: '//trim(str(.true., Ni))//' cells'
 #endif
 call init()
-print "(A)", 'Integrating 1D Euler equations by means of TVD/SSP Runge-Kutta class of solvers'
 call rk_integrator%init(stages=rk_stages)
 call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, ord=7)
 t = 0._R_P
@@ -120,7 +118,7 @@ call save_time_serie(t=t, finish=.true.)
 call save_results(title='FOODiE test: 1D Euler equations integration, explicit TVD Runge-Kutta, t='//str(n=t_final)// &
                         trim(str(.true., rk_stages))//' stages', &
                   filename='euler_1D_openmp_integration-tvdrk-'//trim(str(.true., rk_stages)))
-print "(A)", 'Finish! Average system seconds spent for 1 step integration: '//trim(str(.true., system_clocks))
+print "(I5,A,F23.15)", omp_threads, ' ', system_clocks
 stop
 !-----------------------------------------------------------------------------------------------------------------------------------
 contains

### 2015-10-16 src/tests/euler-1D-openmp/type_euler-1D-openmp.f90

@@ -340,12 +340,13 @@ contains
   enddo
   ! compute residuals
   allocate(euler_1D_openmp :: dState_dt)
+  select type(dState_dt)
+  class is(euler_1D_openmp)
+    dState_dt = self
+  endselect
   !$OMP PARALLEL PRIVATE(i) SHARED(self, dState_dt, F)
   select type(dState_dt)
   class is(euler_1D_openmp)
-    !$OMP SINGLE
-    dState_dt = self
-    !$OMP END SINGLE
     !$OMP DO
     do i=1, self%Ni
       dState_dt%U(:, i) = (F(:, i - 1) - F(:, i)) / self%Dx
@@ -368,12 +369,13 @@ contains
 
   !---------------------------------------------------------------------------------------------------------------------------------
   allocate(euler_1D_openmp :: opr)
+  select type(opr)
+  class is(euler_1D_openmp)
+    opr = lhs
+  endselect
   !$OMP PARALLEL DEFAULT(NONE) PRIVATE(i) SHARED(lhs, rhs, opr)
   select type(opr)
   class is(euler_1D_openmp)
-    !$OMP SINGLE
-    opr = lhs
-    !$OMP END SINGLE
     select type(rhs)
     class is (euler_1D_openmp)
       !$OMP DO
@@ -399,12 +401,13 @@ contains
 
   !---------------------------------------------------------------------------------------------------------------------------------
   allocate(euler_1D_openmp :: opr)
+  select type(opr)
+  class is(euler_1D_openmp)
+    opr = lhs
+  endselect
   !$OMP PARALLEL DEFAULT(NONE) PRIVATE(i) SHARED(lhs, rhs, opr)
   select type(opr)
   class is(euler_1D_openmp)
-    !$OMP SINGLE
-    opr = lhs
-    !$OMP END SINGLE
     !$OMP DO
     do i=1, lhs%Ni
       opr%U(:, i) = lhs%U(:, i) * rhs
@@ -427,12 +430,13 @@ contains
 
   !---------------------------------------------------------------------------------------------------------------------------------
   allocate(euler_1D_openmp :: opr)
+  select type(opr)
+  class is(euler_1D_openmp)
+    opr = rhs
+  endselect
   !$OMP PARALLEL DEFAULT(NONE) PRIVATE(i) SHARED(lhs, rhs, opr)
   select type(opr)
   class is(euler_1D_openmp)
[…]

### 2015-10-16 publish/src/euler-1D-openmp.f90
Travis CI autocommit from travis build 65

@@ -90,11 +90,8 @@ call omp_set_num_threads(omp_threads)
 !$OMP SHARED(omp_threads)
 omp_threads = omp_get_num_threads()
 !$OMP END PARALLEL
-print "(A)", 'Testing FOODiE with '//trim(str(.true., omp_threads))//' OpenMP threads for Ni: '//trim(str(.true., Ni))//' cells'
 #endif
-! create Euler field initial state
 call init()
-print "(A)", 'Integrating 1D Euler equations by means of TVD/SSP Runge-Kutta class of solvers'
 call rk_integrator%init(stages=rk_stages)
 call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, ord=7)
 t = 0._R_P
@@ -121,7 +118,7 @@ call save_time_serie(t=t, finish=.true.)
 call save_results(title='FOODiE test: 1D Euler equations integration, explicit TVD Runge-Kutta, t='//str(n=t_final)// &
                         trim(str(.true., rk_stages))//' stages', &
                   filename='euler_1D_openmp_integration-tvdrk-'//trim(str(.true., rk_stages)))
-print "(A)", 'Finish! Average system seconds spent for 1 step integration: '//trim(str(.true., system_clocks))
+print "(I5,A,F23.15)", omp_threads, ' ', system_clocks
 stop
 !-----------------------------------------------------------------------------------------------------------------------------------
 contains

### 2015-10-16 src/foodie_adt_integrand.f90

@@ -5,7 +5,7 @@ module foodie_adt_integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
-use foodie_kinds, only : R_P, I_P
+use foodie_kinds, only : R_P
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -42,7 +42,7 @@ abstract interface
   !---------------------------------------------------------------------------------------------------------------------------------
   !< Time derivative function of integrand class, i.e. the residuals function.
   !---------------------------------------------------------------------------------------------------------------------------------
-  import :: integrand, R_P, I_P
+  import :: integrand, R_P
   class(integrand),       intent(IN) :: self      !< Integrand field.
   real(R_P),    optional, intent(IN) :: t         !< Time.
   class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.

### 2015-10-16 src/type_euler-1D-openmp.f90

@@ -126,7 +126,6 @@ type, extends(integrand) :: euler_1D_openmp
   !< + Ns+1 : momentum conservation             (r*u)
   !< + Ns+2 : energy conservation               (r*E)
   private
-  integer(I_P)                   :: steps=0         !< Number of time steps stored.
   integer(I_P)                   :: ord=0           !< Space accuracy formal order.
   integer(I_P)                   :: Ni=0            !< Space dimension.
   integer(I_P)                   :: Ng=0            !< Number of ghost cells for boundary conditions handling.
@@ -136,7 +135,6 @@ type, extends(integrand) :: euler_1D_openmp
   real(R_P)                      :: Dx=0._R_P       !< Space step.
   type(weno_interpolator_upwind) :: weno            !< WENO interpolator.
   real(R_P),    allocatable      :: U(:,:)          !< Integrand (state) variables, whole physical domain [1:Nc,1-Ng:Ni+Ng].
-  real(R_P),    allocatable      :: previous(:,:,:) !< Previous time steps states [1:Nc,1-Ng:Ni+Ng,1:steps].
   real(R_P),    allocatable      :: cp0(:)          !< Specific heat cp of initial species [1:Ns].
   real(R_P),    allocatable      :: cv0(:)          !< Specific heat cv of initial species [1:Ns].
   character(:), allocatable      :: BC_L            !< Left boundary condition type.
@@ -149,8 +147,6 @@ type, extends(integrand) :: euler_1D_openmp
     procedure, pass(self), public :: dt => compute_dt !< Compute the current time step, by means of CFL condition.
     ! ADT integrand deferred methods
     procedure, pass(self), public :: t => dEuler_dt                                       !< Time derivative, residuals function.
-    procedure, pass(self), public :: update_previous_steps                                !< Update previous time steps.
-    procedure, pass(self), public :: previous_step                                        !< Get a previous time step.
     procedure, pass(lhs),  public :: integrand_multiply_integrand => euler_multiply_euler !< Euler * Euler operator.
     procedure, pass(lhs),  public :: integrand_multiply_real => euler_multiply_real       !< Euler * real operator.
     procedure, pass(rhs),  public :: real_multiply_integrand => real_multiply_euler       !< Real * Euler operator.
@@ -169,7 +165,7 @@ endtype euler_1D_openmp
[…]

### 2015-10-19 papers/announcement/manuscript.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
 
-\title{FOODiE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern}
+\title{FOODIE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern}
 
@@
 The (numerical) solution of Ordinary Differential Equations (ODEs) problems is of paramount relevance, ODEs system being an ubiquitous mathematical formulation of many physical phenomena (such as those involved in fluid dynamics, chemistry, biology, evolutionary-anthropology, \dots).
-The present paper is the first \emph{manifesto} of FOODiE, a library aimed to numerically solve ODEs problems by means of a clear, concise and efficient \emph{abstract} interface.
-FOODiE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has manifolds aims: to provide a set to built-in numerical schemes that are accurate, robust, validated and efficient and to allow easy application of these schemes to (almost) all ODEs problems by means of an effective Abstract Calculus Pattern.
-The key idea is to allow the same solver-implementation to be applied to all ODEs problems thus avoiding the re-implementation of the ODEs solver for each different ODEs problem: code re-usability is consequently maximized, FOODiE being a general robust framework.
+The present paper is the first \emph{manifesto} of FOODIE, a library aimed to numerically solve ODEs problems by means of a clear, concise and efficient \emph{abstract} interface.
+FOODIE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has manifolds aims: to provide a set to built-in numerical schemes that are accurate, robust, validated and efficient and to allow easy application of these schemes to (almost) all ODEs problems by means of an effective Abstract Calculus Pattern.
+The key idea is to allow the same solver-implementation to be applied to all ODEs problems thus avoiding the re-implementation of the ODEs solver for each different ODEs problem: code re-usability is consequently maximized, FOODIE being a general robust framework.
 Besides, the same framework also allows rapid development of new ODEs solvers due to the high abstraction level of the library itself.
 
-FOODiE is a modern Fortran library which main features are:
+FOODIE is a modern Fortran library which main features are:
 \begin{description}
-  \item[Free] FOODiE is a free software;
-  \item[OOP] FOODiE is based on Object Oriented Programming paradigm;
-  \item[TDD] the FOODiE development follows the Test Driven Development software process;
-  \item[Accurately documented] the FOODiE documentation is based on high quality, first class solutions embedding detailed (mathematical) descriptions directly inside code sources; moreover, comprehensive hyper-linked documentations is also provided;
-  \item[Comphrensive] FOODiE provides a comprehensive set of built-in solvers, namely one-step, multi-steps, explicit, implicit, predictor-corrector solvers; moreover, its Abstract Calculus Pattern allows the solution of virtually all kinds of ODEs problems;
-  \item[Collaborative] the development of FOODiE takes advantage of web communications, the main project being hosted on GitHub.
+  \item[Free] FOODIE is a free software;
+  \item[OOP] FOODIE is based on Object Oriented Programming paradigm;
+  \item[TDD] the FOODIE development follows the Test Driven Development software process;
+  \item[Accurately documented] the FOODIE documentation is based on high quality, first class solutions embedding detailed (mathematical) descriptions directly inside code sources; moreover, comprehensive hyper-linked documentations is also provided;
+  \item[Comphrensive] FOODIE provides a comprehensive set of built-in solvers, namely one-step, multi-steps, explicit, implicit, predictor-corrector solvers; moreover, its Abstract Calculus Pattern allows the solution of virtually all kinds of ODEs problems;
+  \item[Collaborative] the development of FOODIE takes advantage of web communications, the main project being hosted on GitHub.
 \end{description}
 
-The present paper is the first announcement of FOODiE project: the current implementation is extensively discussed and its capabilities are proved by means of tests and examples.
+The present paper is the first announcement of FOODIE project: the current implementation is extensively discussed and its capabilities are proved by means of tests and examples.
 \end{abstract}
@@
 \noindent
-\emph{Manuscript Title:} FOODiE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern \\
+\emph{Manuscript Title:} FOODIE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern \\
 \emph{Authors:} Zaghi, S., Curcic, M., Rouson, D., Beekman, I. \\
-\emph{Program title:} FOODiE \\
+\emph{Program title:} FOODIE \\
 \emph{Journal Reference:} \\

### 2015-10-19 papers/announcement/src/API/API.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
 
-In this section we review the FOODiE API providing a detailed discussion of the implementation choices.
+In this section we review the FOODIE API providing a detailed discussion of the implementation choices.
 
@@
 Using such an abstract definition, we can implement algorithms operating on only this abstract type and on \emph{all its concrete extensions}.
-This is the key feature of FOODiE library: all the above described ODE solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax.
+This is the key feature of FOODIE library: all the above described ODE solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax.
 In the meanwhile, client codes must implement their own IVPs extending only one simple abstract type.
 
-In the subsection \ref{subsec:type_integrand} a review of the FOODiE main ADT, the \emph{integrand} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.
+In the subsection \ref{subsec:type_integrand} a review of the FOODIE main ADT, the \emph{integrand} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.
 
-It is worth noting that all FOODiE public \emph{entities} (ADT and solvers) must be accessed by the FOODiE module, see listing \ref{list:main_module} for an example on how access to all public FOODiE entities.
+It is worth noting that all FOODIE public \emph{entities} (ADT and solvers) must be accessed by the FOODIE module, see listing \ref{list:main_module} for an example on how access to all public FOODIE entities.
 
-\begin{lstlisting}[firstnumber=1,style=code,caption={usage example importing all public entities of FOODiE main module},label={list:main_module}]
+\begin{lstlisting}[firstnumber=1,style=code,caption={usage example importing all public entities of FOODIE main module},label={list:main_module}]
 use foodie, only: integrand, &
@@
 
-\subsection{The main FOODiE Abstract Data Type: the integrand type}\label{subsec:type_integrand}
+\subsection{The main FOODIE Abstract Data Type: the integrand type}\label{subsec:type_integrand}
 
@@
 type, abstract :: integrand
-  !< Abstract type for building FOODiE ODE integrators.
+  !< Abstract type for building FOODIE ODE integrators.
   contains
@@
 The \emph{integrand} type does not implement any actual integrand field, it being and abstract type.
-It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODiE solver.
+It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODIE solver.
 
@@
 As a consequence, the eventual concrete implementation of this procedure can be very complex and errors-prone.
-Nevertheless, the FOODiE solvers are implemented only on the above abstract interface, thus emancipating the solvers implementation from any concrete complexity.
+Nevertheless, the FOODIE solvers are implemented only on the above abstract interface, thus emancipating the solvers implementation from any concrete complexity.
 
@@
 The implementation details of such a procedures class are strictly dependent on the concrete extension of the integrand type.
-From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are two integrand object and the returned object is still an integrand one.
+From the FOODIE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are two integrand object and the returned object is still an integrand one.
 
@@
 The implementation details of these procedures are strictly dependent on the concrete extension of the integrand type.
-From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are an integrand object and real number of viceversa and the returned object is still an integrand one.
+From the FOODIE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are an integrand object and real number of viceversa and the returned object is still an integrand one.
 
@@
 The implementation details of this assignment is strictly dependent on the concrete extension of the integrand type.
-From the FOODiE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the assignment, and the second argument, the right-hand-side of the assignment, are two integrand objects.
+From the FOODIE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the assignment, and the second argument, the right-hand-side of the assignment, are two integrand objects.
 
@@
 
-The explicit forward Euler solver is exposed (by the FOODiE main module that must imported, see listing \ref{list:main_module}) as a single derived type (that is a standard convention for all FOODiE solvers) named \emph{euler\_explicit\_integrator}.
+The explicit forward Euler solver is exposed (by the FOODIE main module that must imported, see listing \ref{list:main_module}) as a single derived type (that is a standard convention for all FOODIE solvers) named \emph{euler\_explicit\_integrator}.
 It provides the type bound procedure (also referred as \emph{method}) \emph{integrate} for integrating in time an \emph{integrand} object, or any of its polymorphic concrete extensions.
-Consequently, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_euler_solver}.
+Consequently, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_euler_solver}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit forward Euler integrator},label={list:define_euler_solver}]
-use FOODiE, only: euler_explicit_integrator
+use FOODIE, only: euler_explicit_integrator
 type(euler_explicit_integrator) :: integrator
@@
 
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_tvd_rk_solver}.
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_tvd_rk_solver}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit TVD/SSP Runge-Kutta integrator},label={list:define_tvd_rk_solver}]
-use FOODiE, only: tvd_runge_kutta_integrator
+use FOODIE, only: tvd_runge_kutta_integrator
 type(tvd_runge_kutta_integrator) :: integrator
@@
 
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ls_rk_solver}.
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_ls_rk_solver}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit low storage Runge-Kutta integrator},label={list:define_ls_rk_solver}]
-use FOODiE, only: ls_runge_kutta_integrator
+use FOODIE, only: ls_runge_kutta_integrator
 type(ls_runge_kutta_integrator) :: integrator
@@
 
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_ab_solver}.
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_ab_solver}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Adams-Bashforth integrator},label={list:define_ab_solver}]
-use FOODiE, only: adams_bashforth_integrator
+use FOODIE, only: adams_bashforth_integrator
 type(adams_bashforth_integrator) :: integrator
@@
 
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_am_solver}.
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_am_solver}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={definition of an implicit Adams-Moulton integrator},label={list:define_am_solver}]
-use FOODiE, only: adams_moulton_integrator
+use FOODIE, only: adams_moulton_integrator
 type(adams_moulton_integrator) :: integrator
@@
 
-As common for FOODiE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODiE integrator type, see listing \ref{list:define_lf_solver}.
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_lf_solver}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Leapfrog integrator},label={list:define_lf_solver}]
-use FOODiE, only: leapfrog_integrator
+use FOODIE, only: leapfrog_integrator
 type(leapfrog_integrator) :: integrator
@@
 
-Table \ref{tab:foodie-vs-math} presents a comparison of the relevant parts of equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:RK-stage}, \ref{eq:RK-ls}, \ref{eq:AB} and \ref{eq:leapfrog} with the corresponding FOODiE implementations reported in listings \ref{list:euler_integrate}, \ref{list:tvd_rk_integrate}, \ref{list:ls_rk_integrate}, \ref{list:ab_integrate} and \ref{list:lf_integrate}, respectively.
+Table \ref{tab:foodie-vs-math} presents a comparison of the relevant parts of equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:RK-stage}, \ref{eq:RK-ls}, \ref{eq:AB} and \ref{eq:leapfrog} with the corresponding FOODIE implementations reported in listings \ref{list:euler_integrate}, \ref{list:tvd_rk_integrate}, \ref{list:ls_rk_integrate}, \ref{list:ab_integrate} and \ref{list:lf_integrate}, respectively.
 This comparison proves that the \emph{integrand} ADT has allowed a very high-level implementation syntax.
 The Fortran implementation is almost equivalent to the rigorous mathematical formulation.
-This aspect directly implies that the implementation of a ODE solver into the FOODiE library is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the solvers must be implemented for each specific definition of the integrand type, it being not abstract.
+This aspect directly implies that the implementation of a ODE solver into the FOODIE library is very clear, concise and less-errors-prone than an \emph{hard-coded} implementation where the solvers must be implemented for each specific definition of the integrand type, it being not abstract.
 
@@
   \centering
-  \caption{Comparison between rigorous mathematical formulation and FOODiE high-level implementation; the syntax "$(s)$" and "$(ss)$" imply the summation operation}\label{tab:foodie-vs-math}
+  \caption{Comparison between rigorous mathematical formulation and FOODIE high-level implementation; the syntax "$(s)$" and "$(ss)$" imply the summation operation}\label{tab:foodie-vs-math}
   \resizebox{0.99\textwidth}{!}{%
@@
     \toprule
-    {\sc Solver} & {\sc Mathematical formulation} & {\sc FOODiE implementation} \\
+    {\sc Solver} & {\sc Mathematical formulation} & {\sc FOODIE implementation} \\
     \hline

### 2015-10-19 papers/announcement/src/introduction/introduction.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
 
-FOODiE library is designed for solving the generalized IVP \ref{eq:IVP}, it being completely unaware of the actual problem's definition.
-FOODiE library provides a high-level, well-documented, simple Application Program Interface (API) for many well-known ODE integration solvers, its aims being twofold:
+FOODIE library is designed for solving the generalized IVP \ref{eq:IVP}, it being completely unaware of the actual problem's definition.
+FOODIE library provides a high-level, well-documented, simple Application Program Interface (API) for many well-known ODE integration solvers, its aims being twofold:
 
@@
 
-FOODiE library must:
+FOODIE library must:
 
@@
 
-FOODiE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has been developed with the aim to satisfy the above specifications.
+FOODIE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has been developed with the aim to satisfy the above specifications.
 The present paper is its first comprehensive presentation.
@@
 Moreover, its simple syntax is suitable for scientific researchers that are interested (and skilled) in the physical aspects of the numerical computations rather than computer technicians.
-Consequently, we develop FOODiE using Fortran language: FOODiE is written by research scientists for research scientists.
+Consequently, we develop FOODIE using Fortran language: FOODIE is written by research scientists for research scientists.
 
-One key-point of the FOODiE development is the \emph{problem generalization}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its actual definitions.
+One key-point of the FOODIE development is the \emph{problem generalization}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its actual definitions.
 Consequently, we must rely on a \emph{generic} implementation of the solvers.
@@
 The abstract calculus pattern provides a simple solution for the connection between the very high-level expression of IVP \ref{eq:IVP} and the eventual concrete (low-level) implementation of the ODE problem being solved.
-ACP essentially constitutes a \emph{contract} based on an Abstract Data Type (ADT): we specify an ADT supporting a certain set of mathematical operators (differential and integral ones) and implement FOODiE solvers only on the basis of this ADT.
-FOODiE clients must formulate the ODE problem under integration defining their own concrete extensions of our ADT (implementing all the deferred operators).
-Such an approach defines the abstract calculus pattern: FOODiE solvers are aware of only the ADT, while FOODiE clients extend the ADT for defining the concrete ODE problem.
+ACP essentially constitutes a \emph{contract} based on an Abstract Data Type (ADT): we specify an ADT supporting a certain set of mathematical operators (differential and integral ones) and implement FOODIE solvers only on the basis of this ADT.
+FOODIE clients must formulate the ODE problem under integration defining their own concrete extensions of our ADT (implementing all the deferred operators).
+Such an approach defines the abstract calculus pattern: FOODIE solvers are aware of only the ADT, while FOODIE clients extend the ADT for defining the concrete ODE problem.
 
 Is is worth noting that this ACP emancipates the solvers implementations from any low-level problem-dependent details: the ODE solvers developed with this pattern are extremely concise, clear, maintainable and less errors-prone with respect a low-level (non abstract) pattern.
-Moreover, the FOODiE clients can use solvers being extremely robust: as a matter of facts, FOODiE solvers are expressed in a formulation very close to the mathematical one and are tested on an extremely varying family of problems.
+Moreover, the FOODIE clients can use solvers being extremely robust: as a matter of facts, FOODIE solvers are expressed in a formulation very close to the mathematical one and are tested on an extremely varying family of problems.
 As shown in the following, such a great flexibility does not compromise the computational efficiency.
 
-The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODiE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODiE applications is presented; section \ref{sec:parallel} provides an analysis of FOODiE performances under parallel frameworks scenario like the OpenMP and MPI paradigms; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.
+The present paper is organized as following: in section \ref{sec:MNmodels} a brief description of the mathematical and numerical methods currently implemented into FOODIE is presented; in section \ref{sec:API} a detailed discussion on the implementation specifications is provided by means of an analytical code-listings review; in section \ref{sec:Tests} a verification analysis on the results of FOODIE applications is presented; section \ref{sec:parallel} provides an analysis of FOODIE performances under parallel frameworks scenario like the OpenMP and MPI paradigms; finally, in section \ref{sec:conclusions} concluding remarks and perspectives are depicted.
 

### 2015-10-19 papers/announcement/src/mathematical-numerical-models/mathematical-numerical-models.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
 A detailed discussion of these details is out the scope of the present work and is omitted.
-Here, we briefly recall some classifications necessary to introduce the schemes implemented into the FOODiE library.
+Here, we briefly recall some classifications necessary to introduce the schemes implemented into the FOODIE library.
 
@@
 In the framework of one-step schemes family an equivalent accurate approximation can be obtained by means of a multi-stage approach as the one due to Runge-Kutta.
-The current version of FOODiE provides schemes belonging to both these families.
+The current version of FOODIE provides schemes belonging to both these families.
 
@@
 Mixing together explicit and implicit schemes it is possible to build a family of \emph{predictor-corrector} methods: using an explicit scheme to predict a guess for the next step approximation it is possible to use an implicit method for correcting this guess.
-Currently, FOODiE provides explicit solvers and predictor-correct ones.
+Currently, FOODIE provides explicit solvers and predictor-correct ones.
 
-FOODiE currently implements the following ODE schemes:
+FOODIE currently implements the following ODE schemes:
 
@@
 
-In general, the TVD/SSP Runge-Kutta schemes provided by FOODiE library are written by means of the following algorithm:
+In general, the TVD/SSP Runge-Kutta schemes provided by FOODIE library are written by means of the following algorithm:
 \begin{equation}
@@
 The nature of the scheme and the properties of the solutions obtained depend on the number of stages and on the value of the coefficients selected.
-Currently, FOODiE provides 3 Runge-Kutta schemes having TVD or Strong Stability Preserving (SSP) propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.
+Currently, FOODIE provides 3 Runge-Kutta schemes having TVD or Strong Stability Preserving (SSP) propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.
 
@@
 Essentially, the standard Runge-Kutta class (under some specific conditions) can be reformulated allowing a more efficient memory management.
-Currently FOODiE provides a class of \emph{2N registers} storage Runge-Kutta schemes, meaning that the storage of all stages requires only 2 registers of memory with a \emph{word} length $N$ (namely the length of the state vector) in contrast to the standard formulation where $N_s$ registers of the same length $N$ are required.
+Currently FOODIE provides a class of \emph{2N registers} storage Runge-Kutta schemes, meaning that the storage of all stages requires only 2 registers of memory with a \emph{word} length $N$ (namely the length of the state vector) in contrast to the standard formulation where $N_s$ registers of the same length $N$ are required.
 This is a dramatic improvement of memory efficiency especially for schemes using a high number of stages ($N_s \ge 4$) where the memory necessary is an half with respect the original formulation.
@@
 
-Currently FOODiE provides 5/6/7/12/13/14 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in table \ref{tab:RK-ls}.
+Currently FOODIE provides 5/6/7/12/13/14 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in table \ref{tab:RK-ls}.
 
@@
 
-In general, the Adams-Bashforth schemes provided by FOODiE library are written by means of the following algorithm (for only explicit schemes):
+In general, the Adams-Bashforth schemes provided by FOODIE library are written by means of the following algorithm (for only explicit schemes):
 
@@
 
-Currently FOODiE provides 2, 3, and 4 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.
+Currently FOODIE provides 2, 3, and 4 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.
 
@@
 
-In general, the Adams-Moulton schemes provided by FOODiE library are written by means of the following algorithm (for only implicit schemes):
+In general, the Adams-Moulton schemes provided by FOODIE library are written by means of the following algorithm (for only implicit schemes):
 
@@
 
-Currently FOODiE provides 1, 2, and 3 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.
+Currently FOODIE provides 1, 2, and 3 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.
 
@@
 
-In general, the Adams-Bashforth-Moulton schemes provided by FOODiE library are written by means of the following algorithm:
+In general, the Adams-Bashforth-Moulton schemes provided by FOODIE library are written by means of the following algorithm:
 
@@
 
-Currently FOODiE provides $N_s^c=1, 2, 3 \rightarrow N_s^c=2, 3, 4$ steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively.
+Currently FOODIE provides $N_s^c=1, 2, 3 \rightarrow N_s^c=2, 3, 4$ steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively.
 The $b_s^{p,c}$ coefficients are those reported in tables \ref{tab:AB-coeff} and \ref{tab:AM-coeff}.
@@
 
-The unfiltered leapfrog scheme provided by FOODiE is:
+The unfiltered leapfrog scheme provided by FOODIE is:
 \begin{equation}
@@
 
-FOODiE provides, in a \emph{seamless} API, also filtered leapfrog schemes.
+FOODIE provides, in a \emph{seamless} API, also filtered leapfrog schemes.
 A widely used filter is due to Robert and Asselin, that suppress the computational modes at the cost of accuracy reduction resulting into a $1^{st}$ order error in amplitude value.
@@
 If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude.
-The default values of the FOODiE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.
+The default values of the FOODIE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.
 

### 2015-10-19 papers/announcement/src/tests/oscillation/oscillation-API.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
-\subsubsection{FOODiE aware implementation of an oscillation numerical solver}
+\subsubsection{FOODIE aware implementation of an oscillation numerical solver}
 
-The IVP \ref{eq:oscillation} can be easily solved by means of FOODiE library.
-The first block of a FOODiE aware solution consists to define an \emph{oscillation integrand field} defining a concrete extension of the FOODiE \emph{integrand} type.
-Listing \ref{list:oscillation_type} reports the implementation of such an integrand field that is contained into the tests suite shipped within the FOODiE library.
+The IVP \ref{eq:oscillation} can be easily solved by means of FOODIE library.
+The first block of a FOODIE aware solution consists to define an \emph{oscillation integrand field} defining a concrete extension of the FOODIE \emph{integrand} type.
+Listing \ref{list:oscillation_type} reports the implementation of such an integrand field that is contained into the tests suite shipped within the FOODIE library.
 
@@
 
-As the listing \ref{list:oscillation_type} shows, the FOODiE implementation strictly corresponds to the mathematical formulation embracing all the relevant mathematical aspects into one derived type, a single \emph{object}.
-Here we not review the implementation of all deferred methods, this being out of the scope of the present work: the interested reader can see the tests suite sources shipped within the FOODiE library.
+As the listing \ref{list:oscillation_type} shows, the FOODIE implementation strictly corresponds to the mathematical formulation embracing all the relevant mathematical aspects into one derived type, a single \emph{object}.
+Here we not review the implementation of all deferred methods, this being out of the scope of the present work: the interested reader can see the tests suite sources shipped within the FOODIE library.
 However, some of these methods are relevant for our discussion, thus they are reviewed.
@@
 As a matter of facts, the ODE equations are implemented into the residuals function.
-However, the FOODiE ADT strongly alleviates the subtle problems that could arise when the ODE solver is hard-implemented within the specific ODE equations.
-As a matter of facts, the \emph{integrand} ADT specifies the precise interface the residuals function must have: if the client code implements a compliant interface, the FOODiE solvers will work as expected, reducing the possible errors location into the ODE equations, having designed the solvers on the ADT and not on the concrete type.
+However, the FOODIE ADT strongly alleviates the subtle problems that could arise when the ODE solver is hard-implemented within the specific ODE equations.
+As a matter of facts, the \emph{integrand} ADT specifies the precise interface the residuals function must have: if the client code implements a compliant interface, the FOODIE solvers will work as expected, reducing the possible errors location into the ODE equations, having designed the solvers on the ADT and not on the concrete type.
 
@@
 
-The assignment overloading of the \emph{oscillation} type is the last key-method that enforces the conciseness of the FOODiE aware implementation.
+The assignment overloading of the \emph{oscillation} type is the last key-method that enforces the conciseness of the FOODIE aware implementation.
 Listing \ref{list:oscillation_assign} reports the implementation of the assignment overloading.
@@
 Notably, for the assignment of the state vector and of the previous time steps solution array we take advantage of the automatic re-allocation of the left-hand-side variables when they are not allocated or allocated differently from the right-hand-side, that is a Fortran 2003 feature.
-In spite its simplicity, the assignment overloading is a key-method enabling the usage of FOODiE solver: effectively, the assignment between two \emph{integrand} ADT variables is ubiquitous into the solvers implementations, see equation \ref{eq:RK} for example.
+In spite its simplicity, the assignment overloading is a key-method enabling the usage of FOODIE solver: effectively, the assignment between two \emph{integrand} ADT variables is ubiquitous into the solvers implementations, see equation \ref{eq:RK} for example.
 
@@
 
-\paragraph{FOODiE numerical integration}
+\paragraph{FOODIE numerical integration}
 
-Using the above discussed \emph{oscillation} type it is very easy to solve IVP \ref{eq:oscillation} by means of FOODiE library.
+Using the above discussed \emph{oscillation} type it is very easy to solve IVP \ref{eq:oscillation} by means of FOODIE library.
 Listing \ref{list:oscillation_leapfrog} presents the numerical integration of system \ref{eq:oscillation} by means of the Leapfrog RAW-filtered method.
@@
 Note that the memory registers for storing the Runge-Kutta stages and the RAW filter displacement must be handled by the client code.
-Listing \ref{list:oscillation_leapfrog} demonstrates how it is simple, clear and concise to solve a IVP by FOODiE solvers.
+Listing \ref{list:oscillation_leapfrog} demonstrates how it is simple, clear and concise to solve a IVP by FOODIE solvers.
 Moreover, it proves how it is simple and effective to apply different solvers in a coupled algorithm, that greatly simplify the development of new hybrid solvers for self-adaptive time step size.

### 2015-10-19 papers/announcement/src/tests/oscillation/oscillation.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
 
-Because the Oscillation system \ref{eq:oscillation} posses a closed exact solution, the discussion on this test has twofolds aims: to assess the accuracy of the FOODiE's built-in solvers comparing the numerical solutions with the exact one and to demonstrate how it is simple to solve this prototypical problem by means of FOODiE.
+Because the Oscillation system \ref{eq:oscillation} posses a closed exact solution, the discussion on this test has twofolds aims: to assess the accuracy of the FOODIE's built-in solvers comparing the numerical solutions with the exact one and to demonstrate how it is simple to solve this prototypical problem by means of FOODIE.
 

### 2015-10-19 papers/announcement/src/tests/tests.tex
Change project name to FOODIE
FOODIE => Fortran Object-Oriented Differential-equations Integration Environment

Why:

Due to a great idea of Damian.

This change addresses the need by:

Change all references to old FOODiE acronym.

Side effects:

A lot...

@@
 
-For the assessment of FOODiE capabilities three different tests are discussed:
+For the assessment of FOODIE capabilities three different tests are discussed:
 
@@
 
-With the first two pure ODEs tests we assess the accuracy of the FOODiE built-in solvers and demonstrate how to use the library in real applications.
-The last test discussion is mainly devoted to the demonstration of FOODiE application into a real, complex PDEs problem.
+With the first two pure ODEs tests we assess the accuracy of the FOODIE built-in solvers and demonstrate how to use the library in real applications.
+The last test discussion is mainly devoted to the demonstration of FOODIE application into a real, complex PDEs problem.
 

### 2015-10-19 papers/announcement/errors_analysis/errors_analysis.sh

@@ -1,5 +1,5 @@
 #!/bin/bash
-# script automotion for FOODiE's errors analysis of the provided tests
+# script automotion for FOODIE's errors analysis of the provided tests
 echo "Building tests executables"
 cd ../../../
 FoBiS.py build -mode errors-analysis --build_dir papers/announcement/errors_analysis/tests > builds.log

### 2015-10-19 papers/announcement/openmp_benchmark/openmp_benchmark.sh

@@ -1,5 +1,5 @@
 #!/bin/bash
-# script automotion for FOODiE's OpenMP benchmark using provided tests
+# script automotion for FOODIE's OpenMP benchmark using provided tests
 
 # defaults
 bench='all'
@@ -11,7 +11,7 @@ function print_usage {
   # print help message for correct CLI usage
   echo
   echo "`basename $0`"
-  echo "script automotion for FOODiE OpenMP benchmark"
+  echo "script automotion for FOODIE OpenMP benchmark"
   echo "Usage: `basename $0` [opts [args]]"
   echo "  [ -b #benchmark_type -r #ripetitions -Ni #cells ]"
   echo

### 2015-10-19 src/lib/foodie_adt_integrand.f90

@@ -1,7 +1,7 @@
-!< Define the abstract type *integrand* for building FOODiE ODE integrators.
+!< Define the abstract type *integrand* for building FOODIE ODE integrators.
 module foodie_adt_integrand
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< Define the abstract type *integrand* for building FOODiE ODE integrators.
+!< Define the abstract type *integrand* for building FOODIE ODE integrators.
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -16,7 +16,7 @@ public :: integrand
 
 !-----------------------------------------------------------------------------------------------------------------------------------
 type, abstract :: integrand
-  !< Abstract type for building FOODiE ODE integrators.
+  !< Abstract type for building FOODIE ODE integrators.
   contains
     ! public deferred procedures that concrete integrand-field must implement
     procedure(time_derivative),      pass(self), deferred, public :: t !< Time derivative, residuals.

### 2015-10-19 src/lib/foodie_integrator_adams_bashforth_moulton.f90

@@ -1,8 +1,8 @@
-!< FOODiE integrator: provide a predictor-corrector class of Adams-Bashforth-Moutlon multi-step schemes, from 1st to 4rd order
+!< FOODIE integrator: provide a predictor-corrector class of Adams-Bashforth-Moutlon multi-step schemes, from 1st to 4rd order
 !< accutate.
 module foodie_integrator_adams_bashforth_moulton
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< FOODiE integrator: provide a predictor-corrector class of Adams-Bashforth-Moutlon multi-step schemes, from 1st to 4rd order
+!< FOODIE integrator: provide a predictor-corrector class of Adams-Bashforth-Moutlon multi-step schemes, from 1st to 4rd order
 !< accutate.
 !<
 !< Considering the following ODE system:
@@ -92,7 +92,7 @@ public :: adams_bashforth_moulton_integrator
 
 !-----------------------------------------------------------------------------------------------------------------------------------
 type :: adams_bashforth_moulton_integrator
-  !< FOODiE integrator: provide an explicit class of Adams-Bashforth-Moulton multi-step schemes, from 1st to 4rd order accurate.
+  !< FOODIE integrator: provide an explicit class of Adams-Bashforth-Moulton multi-step schemes, from 1st to 4rd order accurate.
   !<
   !< @note The integrator must be created or initialized (predictor and corrector schemes selection) before used.
   private

### 2015-10-19 src/tests/euler-1D-openmp/euler-1D-openmp.f90

@@ -1,7 +1,7 @@
-!< Test FOODiE with the integration of Euler 1D PDEs system.
+!< Test FOODIE with the integration of Euler 1D PDEs system.
 program integrate_euler_1D_openmp
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< Test FOODiE with the integration of Euler 1D PDEs system.
+!< Test FOODIE with the integration of Euler 1D PDEs system.
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -59,7 +59,7 @@ integer(I_P)                      :: steps                 !< Time steps counter
 call cli%init(progname    = 'euler-1D-openmp',                                                       &
               authors     = 'Fortran-FOSS-Programmers',                                              &
               license     = 'GNU GPLv3',                                                             &
-              description = 'Test FOODiE library on 1D Euler equations integration, OpenMP enabled', &
+              description = 'Test FOODIE library on 1D Euler equations integration, OpenMP enabled', &
               examples    = ["euler-1D --results  ",                                                 &
                              "euler-1D -r -t -v -p",                                                 &
                              "euler-1D            ",                                                 &
@@ -96,7 +96,7 @@ call rk_integrator%init(stages=rk_stages)
 call domain%init(Ni=Ni, Ns=Ns, Dx=Dx, BC_L=BC_L, BC_R=BC_R, initial_state=initial_state, cp0=cp0, cv0=cv0, ord=7)
 t = 0._R_P
 steps = 0
-call save_time_serie(title='FOODiE test: 1D Euler equations integration, explicit TVD Runge-Kutta, t='//str(n=t_final)// &
+call save_time_serie(title='FOODIE test: 1D Euler equations integration, explicit TVD Runge-Kutta, t='//str(n=t_final)// &
                            trim(str(.true., rk_stages))//' stages', &
                      filename='euler_1D_openmp_integration-tvdrk-'//trim(str(.true., rk_stages))//'-time_serie.dat', &
                      t=t)
@@ -115,7 +115,7 @@ enddo
 system_clocks = system_clocks / steps
 if (verbose) print "(A)", '    Time step: '//str(n=dt)//', Time: '//str(n=t)
 call save_time_serie(t=t, finish=.true.)
[…]

### 2015-10-19 src/tests/euler-1D-openmp/type_euler-1D-openmp.f90

@@ -20,7 +20,7 @@ public :: euler_1D_openmp
 type, extends(integrand) :: euler_1D_openmp
   !< Euler 1D (OpenMP enabled) PDEs system field.
   !<
-  !< It is a FOODiE integrand class concrete extension.
+  !< It is a FOODIE integrand class concrete extension.
   !<
   !<### 1D Euler PDEs system
   !< The 1D Euler PDEs system considered is a non linear, hyperbolic (inviscid) system of conservation laws for compressible gas

### 2015-10-19 src/tests/euler-1D/error_analysis.py

@@ -49,7 +49,7 @@ def compute_error(s_coarse, s_fine, Dx):
 
 
 if __name__ == '__main__':
-  cliparser = argparse.ArgumentParser(prog='error_analysis.py', description='Perform error analysis of 1D Euler test of FOODiE library')
+  cliparser = argparse.ArgumentParser(prog='error_analysis.py', description='Perform error analysis of 1D Euler test of FOODIE library')
   cliparser.add_argument('-ex', '--execute', required=False, action='store_true', default=False, help='Execute the euler-1D program (avoided if results files already generated)')
   cliparser.add_argument('-p', '--problem', required=False, action='store', default='sod', choices=['sod', 'smooth'], help='Problem to be solved')
   cliparser.add_argument('--Ni_grids', required=False, action='store', nargs='+', default=[100, 200, 400, 800, 1600, 3200], help='Grids number of cells')

### 2015-10-19 src/tests/euler-1D/euler-1D.f90

@@ -1,7 +1,7 @@
-!< Test FOODiE with the integration of Euler 1D PDEs system.
+!< Test FOODIE with the integration of Euler 1D PDEs system.
 program integrate_euler_1D
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< Test FOODiE with the integration of Euler 1D PDEs system.
+!< Test FOODIE with the integration of Euler 1D PDEs system.
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -56,7 +56,7 @@ integer(I_P)                      :: av_Ni              !< Average the solution 
 call cli%init(progname    = 'euler-1D',                                              &
               authors     = 'Fortran-FOSS-Programmers',                              &
               license     = 'GNU GPLv3',                                             &
-              description = 'Test FOODiE library on 1D Euler equations integration', &
+              description = 'Test FOODIE library on 1D Euler equations integration', &
               examples    = ["euler-1D --solver euler --results  ",                  &
                              "euler-1D --solver ls-runge-kutta -r",                  &
                              "euler-1D --solver adams-bashforth  ",                  &
@@ -261,7 +261,7 @@ contains
     open(newunit=rawfile, file=basename//'-'//trim(str(.true.,Ni))//'_cells.dat')
     write(rawfile, '(A)')'TITLE="'//title//'"'
     write(rawfile, '(A)')variables
-    write(rawfile, '(A)')'ZONE T="FOODiE: '//trim(str(.true.,Ni))//' cells", I='//trim(str(.true.,Ni+1))//&
+    write(rawfile, '(A)')'ZONE T="FOODIE: '//trim(str(.true.,Ni))//' cells", I='//trim(str(.true.,Ni+1))//&
                          ', J=1, K=1, DATAPACKING=BLOCK, VARLOCATION=([1]=NODAL,[2-'//trim(str(.true.,Np+1)) //']=CELLCENTERED)'
     write(rawfile, '('//trim(str(.true.,Ni+1))//'('//FR_P//',1X))')xnode
     do v=1, Np
@@ -270,7 +270,7 @@ contains
     if ((av_Ni>0).and.(av_Ni/=Ni)) then
[…]

### 2015-10-19 src/tests/lorenz/lorenz.f90

@@ -1,7 +1,7 @@
-!< Test FOODiE with the integration of Lorenz equations.
+!< Test FOODIE with the integration of Lorenz equations.
 program integrate_lorenz
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< Test FOODiE with the integration of Lorenz equations.
+!< Test FOODIE with the integration of Lorenz equations.
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -39,7 +39,7 @@ logical                           :: results                                    
 call cli%init(progname    = 'lorenz',                                              &
               authors     = 'Fortran-FOSS-Programmers',                            &
               license     = 'GNU GPLv3',                                           &
-              description = 'Test FOODiE library on Lorenz equations integration', &
+              description = 'Test FOODIE library on Lorenz equations integration', &
               examples    = ["lorenz --solver euler --results  ",                  &
                              "lorenz --solver ls-runge-kutta -r",                  &
                              "lorenz --solver adams-bashforth  ",                  &
@@ -161,7 +161,7 @@ contains
       solution(0, step) = step * dt
       solution(1:space_dimension, step) = attractor%output()
     enddo
-    call save_results(title='FOODiE test: Lorenz equation integration, explicit Adams-Bashforth '//trim(str(.true., s))//' steps', &
+    call save_results(title='FOODIE test: Lorenz equation integration, explicit Adams-Bashforth '//trim(str(.true., s))//' steps', &
                       filename='lorenz_integration-ab-'//trim(str(.true., s)))
   enddo
   print "(A)", 'Finish!'
@@ -187,7 +187,7 @@ contains
     solution(0, step) = step * dt
     solution(1:space_dimension, step) = attractor%output()
   enddo
-  call save_results(title='FOODiE test: Lorenz equation integration, explicit Euler', &
+  call save_results(title='FOODIE test: Lorenz equation integration, explicit Euler', &
                     filename='lorenz_integration-euler')
   print "(A)", 'Finish!'
   return
@@ -225,7 +225,7 @@ contains
     solution(0, step) = step * dt
     solution(1:space_dimension, step) = attractor%output()
   enddo
[…]

### 2015-10-19 src/tests/lorenz/type_lorenz.f90

@@ -19,11 +19,11 @@ public :: lorenz
 type, extends(integrand) :: lorenz
   !< Lorenz equations field.
   !<
-  !< It is a FOODiE integrand class concrete extension.
+  !< It is a FOODIE integrand class concrete extension.
   !<
   !<### Lorenz ODEs system
   !<The Lorenz' equations system [1] is a non linear system of pure ODEs that retains a reasonable-complex behaviour: such a
-  !<system, for a certain parameters-region exhibits a chaotic dynamics useful for testing FOODiE solvers.
+  !<system, for a certain parameters-region exhibits a chaotic dynamics useful for testing FOODIE solvers.
   !<
   !<The Lorenz' ODEs system can be written as:
   !<
@@ -49,7 +49,7 @@ type, extends(integrand) :: lorenz
   !< \beta = \frac{8}{3}
   !<\end{matrix}$$
   !<
-  !<These values are chaos-inducing thus they magnify the eventual numerical inaccuracies of FOODiE solvers, see [2].
+  !<These values are chaos-inducing thus they magnify the eventual numerical inaccuracies of FOODIE solvers, see [2].
   !<
   !<#### Bibliography
   !<

### 2015-10-19 src/tests/oscillation/oscillation.f90

@@ -1,7 +1,7 @@
-!< Test FOODiE with the integration of Oscillation equations.
+!< Test FOODIE with the integration of Oscillation equations.
 program integrate_oscillation
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< Test FOODiE with the integration of Oscillation equations.
+!< Test FOODIE with the integration of Oscillation equations.
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------
@@ -41,7 +41,7 @@ logical                           :: errors_analysis                            
 call cli%init(progname    = 'oscillation',                                              &
               authors     = 'Fortran-FOSS-Programmers',                                 &
               license     = 'GNU GPLv3',                                                &
-              description = 'Test FOODiE library on Oscillation equations integration', &
+              description = 'Test FOODIE library on Oscillation equations integration', &
               examples    = ["oscillation --solver euler --results  ",                  &
                              "oscillation --solver ls-runge-kutta -r",                  &
                              "oscillation --solver adams-bashforth  ",                  &
@@ -172,7 +172,7 @@ contains
     open(newunit=rawfile, file=basename//'.dat')
     write(rawfile, '(A)')'TITLE="'//title//'"'
     write(rawfile, '(A)')'VARIABLES="t" "x" "y"'
-    write(rawfile, '(A)')'ZONE T="FOODiE time serie"'
+    write(rawfile, '(A)')'ZONE T="FOODIE time serie"'
     do s=0, ubound(solution, dim=2)
       write(rawfile, '(3('//FR_P//',1X))')(solution(i, s), i=0, space_dimension)
     enddo

### 2015-10-19 src/tests/oscillation/type_oscillation.f90

@@ -19,7 +19,7 @@ public :: oscillation
 type, extends(integrand) :: oscillation
   !< Oscillation equations field.
   !<
-  !< It is a FOODiE integrand class concrete extension.
+  !< It is a FOODIE integrand class concrete extension.
   !<
   !<### Oscillation ODEs system
   !<The (inertial) Oscillation equations system is a non linear system of pure ODEs and it can be written as:

### 2015-10-19 papers/announcement/src/parallel/openmp.tex
Add OpenMP Euler 1D without FOODIE
Add the OpenMP Euler 1D test solved without FOODIE.

Why:

To provide a reference for parallel scaling analysis.

@@
 
+Figure \ref{fig:scaling-openmp} reports the parallel scaling analysis performed with OpenMP directives-based paradigm.
+This benchmark has been done on a \emph{dual Intel(R) Xeon(R) CPU X5650} exacores workstation for a total of 12 physical cores, coupled with 24GB of RAM.
+The codes (with and without FOODIE) have been compiled by means of the GNU gfortran compiler v5.2.0 with \emph{-O2 -fopenmp} compilation flags.
+
+The Euler conservation laws are integrated for 10 time steps by means of the TVD RK(5,4) solver: the measured CPU time used for computing the scaling efficiencies is the average of the 10 integrations, thus representing the mean CPU time for computing one time step integration.
+
+For the strong scaling, the benchmark has been conducted with 100000 finite volumes.
+Figure \ref{fig:strong-scaling-openmp} summarizes the strong scaling analysis: it shows that FOODIE-based code scales similarly to the baseline code without FOODIE.
+
+For the weak scaling the minimum size if 10000 finite volumes and the size is scaled linearly with the OpenMP threads, thus $N_{12} = 120000$ cells.
+Figure \ref{fig:weak-scaling-openmp} summarizes the weak scaling analysis and it essentially confirms that FOODIE-based code scales similarly to the baseline code without FOODIE.
+
+Both strong and weak scaling analysis point out that for the computing architecture considered the parallel scaling is reasonable up to 8 cores: using 12 cores the measured efficiencies become unsatisfactory, reducing below the 50\%.
 \begin{figure}[!ht]
@@
     \centering
-    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-scaling.png}
-    \caption{Strong scaling, number of cells 5000}\label{fig:strong-scaling-openmp}
+    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-scaling-comparison.png}
+    \caption{Strong scaling, number of cells 100000}\label{fig:strong-scaling-openmp}
   \end{subfigure}\\
@@
     \centering
-    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-scaling.png}
-    \caption{Weak scaling, minimum number of cells 500}\label{fig:weak-scaling-openmp}
+    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-scaling-comparison.png}
+    \caption{Weak scaling, minimum number of cells 10000}\label{fig:weak-scaling-openmp}
   \end{subfigure}\\
-  \caption{Scaling efficiency with OpenMP}\label{fig:scaling-openmp}
+  \caption{Scaling efficiency with OpenMP programming model}\label{fig:scaling-openmp}
 \end{figure}

### 2015-10-19 papers/announcement/src/parallel/parallel.tex
Add OpenMP Euler 1D without FOODIE
Add the OpenMP Euler 1D test solved without FOODIE.

Why:

To provide a reference for parallel scaling analysis.

@@
 \section{Benchmarks on parallel frameworks}\label{sec:parallel}
+
+As aforementioned, FOODIE is unaware of any parallel paradigms or programming models the client codes adopt.
+As a consequence, the parallel performances measurements presented into this section are aimed only to prove that FOODIE environment does not destroy the parallel scaling of the baseline code implemented without FOODIE.
+
+To obtain such a prove, the 1D Euler PDE system described previously is numerically solved with FOODIE-aware test codes that in turn exploit parallel resources by means:
+
+\begin{itemize}
+  \item CoArray Fortran (CAF) model, for shared and distributed memory architectures;
+  \item OpenMP directive-based model, for only shared memory architectures;
+  \item Message Passing Interface (MPI) library-based model, for only distributed memory architectures;
+  \end{itemize}
+
+In order to measure the performances of the parallel-enabled FOODIE tests, the \emph{strong} and \emph{weak} scaling have been considered. For the strong scaling the \emph{speedup} has been computed:
+
+\begin{equation}
+  speedup(N, k) = \frac{T_{serial}(N)}{T_{parallel}(N, k)}
+  \label{eq:strong-scaling-speedup}
+\end{equation}
+where $N$ is the problem size, $K$ the number of parallel resources used (namely the physical cores), $T_{serial}$ is the CPU time of the serial code and $T_{parallel}$ the one of the parallel code.
+The ideal speedup is linear with slop equals to 1.
+The efficiency correlated to the strong scaling measurement is defined as:
+
+\begin{equation}
+  efficiency(N, k) = \frac{speedup(N, k)}{k}
+  \label{eq:strong-scaling-efficiency}
+\end{equation}
+The maximum ideal efficiency is obviously the unity.
+
+For the of weak scaling measurement the \emph{sizeup} has been computed:
+
+\begin{equation}
+  sizeup(N, k) = \frac{N_k}{N_1} \cdot \frac{T_{serial}(N_1)}{T_{parallel}(N_k, k)}
+  \label{eq:weak-scaling-sizeup}
+\end{equation}
+where $N_1$ is the minimum size considered and $N_K$ is the size used for the test computed with $k$ parallel resources.
+If $N_K$ is scaled proportional to $N_1$, the ideal sizeup is again linear and if $N_k = k \cdot N_1$ the slope is again linear.
+The efficiency correlated to the weak scaling is defined as:
+
+\begin{equation}
+  efficiency(N, k) = \frac{sizeup(N, k)}{k}
+  \label{eq:weak-scaling-efficiency}
+\end{equation}
+The maximum ideal efficiency is obviously the unity.
+
+The same 1D Euler PDE problem is also solved by parallel-enabled codes that are not based on FOODIE: their solutions provide a reference for measuring the effect of FOODIE abstraction on the parallel scaling.
+
+\subsection{CAF benchmark}\label{subsec:caf}
 
@@
 
-\subsection{OpenMP benchmark}\label{subsec:openmp}
-
-{\color{red} To be written.}
+\input{src/parallel/openmp.tex}
 

### 2015-10-19 papers/announcement/openmp_benchmark/openmp_benchmark.sh

@@ -2,6 +2,7 @@
 # script automotion for FOODIE's OpenMP benchmark using provided tests
 
 # defaults
+code='all'
 bench='all'
 ripetitions=1
 Ni=100
@@ -13,9 +14,10 @@ function print_usage {
   echo "`basename $0`"
   echo "script automotion for FOODIE OpenMP benchmark"
   echo "Usage: `basename $0` [opts [args]]"
-  echo "  [ -b #benchmark_type -r #ripetitions -Ni #cells ]"
+  echo "  [ -c #code_type -b #benchmark_type -r #ripetitions -Ni #cells ]"
   echo
   echo "Defaults of optional arguments:"
+  echo "  -c all # type of code: valid value are 'all', 'foodie', 'no-foodie'"
   echo "  -b all # type of benchmark: valid value are 'all', 'strong', 'weak'"
   echo "  -r 1 # number of ripetitions for averaging the benchmarks results"
   echo "  -Ni 1000 # test size for strong scaling or minimum test size for weak scaling benchamrk"
@@ -39,10 +41,77 @@ function run_exe {
   average_cpu_time=`echo "scale=10; $average_cpu_time/$ripetitions" | bc -l`
   echo $average_cpu_time
 }
+function strong_scaling {
+  # perform strong scaling analysis
+  foodie=$1
+  if [ "$foodie" == "yes" ]; then
+    exe='../tests-openmp/euler-1D-openmp'
+    output='strong-scaling.dat'
+  else
+    exe='../tests-openmp/euler-1D-openmp-no-foodie'
+    output='strong-scaling-no-foodie.dat'
+  fi
+  echo 'TITLE="Strong scaling analysis, cells '$Ni'"' > $output
+  echo 'VARIABLES="OpenMP threads number" "CPU time"' >> $output
+  echo "Benchmarking without OpenMP threads"
+  average_cpu_time=$(run_exe $exe'-off' 1 $Ni)
+  echo 0 $average_cpu_time >> $output
+  c=1
+  while [ "$c" -lt $ncores ]; do
+    echo "Benchmarking with "$c" OpenMP threads"
+    average_cpu_time=$(run_exe $exe'-on' $c $Ni)
+    echo $c $average_cpu_time >> $output
+    c=$((c*2))
+    if [ "$c" -gt $ncores ]; then
+      # doing the last benchmark
[…]

### 2015-10-19 src/tests/euler-1D-openmp-no-foodie/type_euler-1D-openmp.f90

@@ -1,0 +1,942 @@
+!< Define Euler 1D (OpenMP enabled) field.
+module type_euler_1D_openmp_no_foodie
+!-----------------------------------------------------------------------------------------------------------------------------------
+!< Define Euler 1D (OpenMP enabled) field.
+!-----------------------------------------------------------------------------------------------------------------------------------
+
+!-----------------------------------------------------------------------------------------------------------------------------------
+use IR_Precision, only : R_P, I_P
+use wenoof, only : weno_factory, weno_constructor_upwind, weno_interpolator, weno_interpolator_upwind
+!-----------------------------------------------------------------------------------------------------------------------------------
+
+!-----------------------------------------------------------------------------------------------------------------------------------
+implicit none
+private
+public :: euler_1D_openmp
+public :: tvd_runge_kutta_integrator
+!-----------------------------------------------------------------------------------------------------------------------------------
+
+!-----------------------------------------------------------------------------------------------------------------------------------
+type :: euler_1D_openmp
+  !< Euler 1D (OpenMP enabled) PDEs system field.
+  !<
+  !<### 1D Euler PDEs system
+  !< The 1D Euler PDEs system considered is a non linear, hyperbolic (inviscid) system of conservation laws for compressible gas
+  !< dynamics, that reads as
+  !<$$
+  !<\begin{matrix}
+  !<U_t = R(U)  \Leftrightarrow U_t = F(U)_x \\
+  !<U = \begin{bmatrix}
+  !<\rho \\
+  !<\rho u \\
+  !<\rho E
+  !<\end{bmatrix}\;\;\;
+  !<F(U) = \begin{bmatrix}
+  !<\rho u \\
+  !<\rho u^2 + p \\
+  !<\rho u H
+  !<\end{bmatrix}
+  !<\end{matrix}
+  !<$$
+  !< where \(\rho\) is the density, \(u\) is the velocity, \(p\) the pressure, \(E\) the total internal specific energy and \(H\)
+  !< the total specific enthalpy. The PDEs system must completed with the proper initial and boundary conditions. Moreover, an ideal
+  !< (thermally and calorically perfect) gas is considered
+  !<$$
+  !<\begin{matrix}
[…]

### 2015-10-19 src/tests/euler-1D-openmp/euler-1D-openmp.f90

@@ -29,7 +29,6 @@ real(R_P)                         :: dt                    !< Time step.
 real(R_P)                         :: t                     !< Time.
 type(euler_1D_openmp)             :: domain                !< Domain of Euler equations.
 real(R_P),    parameter           :: CFL=0.7_R_P           !< CFL value.
-real(R_P),    parameter           :: t_final=0.15_R_P      !< Final time.
 integer(I_P), parameter           :: Ns=1                  !< Number of differnt initial gas species.
 integer(I_P), parameter           :: Nc=Ns+2               !< Number of conservative variables.
 integer(I_P), parameter           :: Np=Ns+4               !< Number of primitive variables.
@@ -43,6 +42,7 @@ real(R_P), allocatable            :: initial_state(:,:)    !< Initial state of p
 real(R_P), allocatable            :: x(:)                  !< Cell center x-abscissa values.
 real(R_P), allocatable            :: final_state(:,:)      !< Final state.
 integer(I_P)                      :: error                 !< Error handler.
+integer(I_P)                      :: steps_max             !< Maximum number of time steps.
 integer(I_P)                      :: omp_threads           !< Number of OpenMP threads.
 logical                           :: plots                 !< Flag for activating plots saving.
 logical                           :: results               !< Flag for activating results saving.
@@ -65,6 +65,7 @@ call cli%init(progname    = 'euler-1D-openmp',                                  
                              "euler-1D            ",                                                 &
                              "euler-1D --plots -r "])
 call cli%add(switch='--Ni', help='Number finite volumes used', required=.false., act='store', def='100', error=error)
+call cli%add(switch='--steps', help='Number time steps performed', required=.false., act='store', def='10', error=error)
 call cli%add(switch='--results', switch_ab='-r', help='Save results', required=.false., act='store_true', def='.false.', &
              error=error)
 call cli%add(switch='--plots', switch_ab='-p', help='Save plots of results', required=.false., act='store_true', def='.false.', &
[…]

### 2015-10-19 publish/src/foodie_integrator_tvd_runge_kutta.f90
Travis CI autocommit from travis build 66

@@ -1,7 +1,7 @@
-!< FOODiE integrator: provide an explicit class of TVD or SSP Runge-Kutta schemes, from 1st to 4th order accutate.
+!< FOODIE integrator: provide an explicit class of TVD or SSP Runge-Kutta schemes, from 1st to 4th order accutate.
 module foodie_integrator_tvd_runge_kutta
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< FOODiE integrator: provide an explicit class of TVD or SSP Runge-Kutta schemes, from 1st to 4th order accutate.
+!< FOODIE integrator: provide an explicit class of TVD or SSP Runge-Kutta schemes, from 1st to 4th order accutate.
 !<
 !< The class of integrators provided have the Total Variation Diminishing (TVD) property or the Strong Stability Preserving (SSP)
 !< one. The schemes are explicit and defined through the Butcher's table syntax, see[1] .
@@ -98,7 +98,7 @@ public :: tvd_runge_kutta_integrator
 
 !-----------------------------------------------------------------------------------------------------------------------------------
 type :: tvd_runge_kutta_integrator
-  !< FOODiE integrator: provide an explicit class of TVD or SSP Runge-Kutta schemes, from 1st to 4th order accutate.
+  !< FOODIE integrator: provide an explicit class of TVD or SSP Runge-Kutta schemes, from 1st to 4th order accutate.
   !<
   !< @note The integrator must be created or initialized (initialize the RK coeficients) before used.
   integer(I_P)           :: stages=0  !< Number of stages.

### 2015-10-19 publish/src/foodie_kinds.f90

@@ -1,7 +1,7 @@
-!< FOODiE kinds: definition of reals and integer kind parameters of FOODiE library.
+!< FOODIE kinds: definition of reals and integer kind parameters of FOODIE library.
 module foodie_kinds
 !-----------------------------------------------------------------------------------------------------------------------------------
-!< FOODiE kinds: definition of reals and integer kind parameters of FOODiE library.
+!< FOODIE kinds: definition of reals and integer kind parameters of FOODIE library.
 !-----------------------------------------------------------------------------------------------------------------------------------
 
 !-----------------------------------------------------------------------------------------------------------------------------------

### 2015-10-19 publish/src/type_burgers.f90

@@ -19,7 +19,7 @@ public :: burgers
 type, extends(integrand) :: burgers
   !< Burgers equations field.
   !<
-  !< It is a FOODiE integrand class concrete extension.
+  !< It is a FOODIE integrand class concrete extension.
   !<
   !<### Burgers PDE equation
   !<The Burgers PDE equation is a non linear PDE widely used as numerical benchmark that can be applied to describe

### 2015-10-22 papers/announcement/src/parallel/openmp.tex
check openmp tests

@@
 
-Figure \ref{fig:scaling-openmp} reports the parallel scaling analysis performed with OpenMP directives-based paradigm.
-This benchmark has been done on a \emph{dual Intel(R) Xeon(R) CPU X5650} exacores workstation for a total of 12 physical cores, coupled with 24GB of RAM.
-The codes (with and without FOODIE) have been compiled by means of the GNU gfortran compiler v5.2.0 with \emph{-O2 -fopenmp} compilation flags.
+This subsection reports the parallel scaling analysis performed with OpenMP directives-based paradigm.
+These benchmarks have been done on a \emph{dual Intel(R) Xeon(R) CPU X5650} exacores workstation for a total of 12 physical cores, coupled with 24GB of RAM.
+In order to perform an accurate analysis 4 different codes have considered:
+
+\begin{itemize}
+  \item FOODIE-aware codes:
+    \begin{itemize}
+      \item serial code;
+      \item OpenMP-enabled code;
+      \end{itemize}
+  \item procedural codes without using FOODIE library:
+    \begin{itemize}
+      \item serial code;
+      \item OpenMP-enabled code;
+      \end{itemize}
+  \end{itemize}
+
+These codes (see appendix \ref{sec:euler-1D-OpenMP-API} for the implementation details) have been compiled by means of the GNU gfortran compiler v5.2.0 with \emph{-O2 -fopenmp} compilation flags.
 
@@
 
-For the strong scaling, the benchmark has been conducted with 100000 finite volumes.
-Figure \ref{fig:strong-scaling-openmp} summarizes the strong scaling analysis: it shows that FOODIE-based code scales similarly to the baseline code without FOODIE.
-
-For the weak scaling the minimum size if 10000 finite volumes and the size is scaled linearly with the OpenMP threads, thus $N_{12} = 120000$ cells.
-Figure \ref{fig:weak-scaling-openmp} summarizes the weak scaling analysis and it essentially confirms that FOODIE-based code scales similarly to the baseline code without FOODIE.
-
-Both strong and weak scaling analysis point out that for the computing architecture considered the parallel scaling is reasonable up to 8 cores: using 12 cores the measured efficiencies become unsatisfactory, reducing below the 50\%.
 \begin{figure}[!ht]
   \centering
-  \begin{subfigure}[b]{0.95\textwidth}
+  \begin{subfigure}[b]{0.85\textwidth}
     \centering
@@
   \end{subfigure}\\
-  \begin{subfigure}[b]{0.95\textwidth}
+  \begin{subfigure}[b]{0.85\textwidth}
     \centering
@@
 \end{figure}
+
+For the strong scaling, the benchmark has been conducted with 100000 finite volumes.
+Figure \ref{fig:strong-scaling-openmp} summarizes the strong scaling analysis: it shows that FOODIE-based code scales similarly to the baseline code without FOODIE.
+
+For the weak scaling the minimum size is 10000 finite volumes and the size is scaled linearly with the OpenMP threads, thus $N_{12} = 120000$ cells.
+Figure \ref{fig:weak-scaling-openmp} summarizes the weak scaling analysis and it essentially confirms that FOODIE-based code scales similarly to the baseline code without FOODIE.
+
+Both strong and weak scaling analysis point out that for the computing architecture considered the parallel scaling is reasonable up to 8 cores: using 12 cores the measured efficiencies become unsatisfactory, reducing below the 50\%.
+
+To complete the comparison, the absolute CPU-time consumed by the two families of codes (with and without FOODIE) must be considered.
+Table \ref{tab:openmp-results} summarizes the benchmarks results.
+As shown, procedural and FOODIE-aware codes consume a very similar CPU-time for both the strong and the weak benchmarks.
+The same results are shown in figure \ref{fig:cpu-time-openmp}.
+These results prove that the abstraction of FOODIE environment does not degrade the computational efficiency.
+
+\begin{table}[!ht]
+  \centering
+  \caption{OpenMP benchmarks results\label{tab:openmp-results}}
+  \begin{subtable}[b]{0.80\textwidth}
+    \centering
+    \caption{Strong benchmarks, number of cells 100000\label{tab:openmp-results-strong}}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{ccccc}
+      {\sc Number of OpenMP threads} & \multicolumn{4}{c}{\sc CPU time for 1 time step integration}              \\
+      \hline
+                                     & FOODIE serial & FOODIE parallel & procedural serial & procedural parallel \\
+      \cmidrule{2-5}
+      1                              & 1.2941        & 1.2942          & 1.2820            & 1.2769              \\
+      2                              & /             & 0.7472          & /                 & 0.7784              \\
+      4                              & /             & 0.3988          & /                 & 0.4218              \\
+      8                              & /             & 0.2185          & /                 & 0.2559              \\
+      12                             & /             & 0.2352          & /                 & 0.2272              \\
+      \hline
+    \end{tabular}}
+  \end{subtable}\\
+  \begin{subtable}[b]{0.80\textwidth}
+    \centering
+    \caption{Weak benchmarks, minimum number of cells 10000\label{tab:openmp-results-weak}}
+    \resizebox{1.00\textwidth}{!}{%
+    \begin{tabular}{cccccc}
+      {\sc Number of OpenMP threads} & {\sc Number of Cells} & \multicolumn{4}{c}{\sc CPU time for 1 time step integration} \\
+      \hline
+                                     &                       & FOODIE serial & FOODIE parallel & procedural serial & procedural parallel \\
+      \cmidrule{3-6}
+      1                              & 10000                 & 0.1496        & 0.1419          & 1.2820            & 1.2769              \\
+      2                              & 20000                 & /             & 0.1643          & /                 & 0.7784              \\
+      4                              & 40000                 & /             & 0.1671          & /                 & 0.4218              \\
+      8                              & 80000                 & /             & 0.1752          & /                 & 0.2559              \\
+      12                             & 120000                & /             & 0.2883          & /                 & 0.2272              \\
+      \hline
+    \end{tabular}}
+  \end{subtable}
+\end{table}
+
+\begin{figure}[!ht]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-cpu-time-comparison.png}
+    \caption{Strong benchmark, number of cells 100000}\label{fig:strong-cpu-time-openmp}
+  \end{subfigure}\quad%
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-cpu-time-comparison.png}
+    \caption{Weak benchmark, minimum number of cells 10000}\label{fig:weak-cpu-time-openmp}
+  \end{subfigure}\\
+  \caption{CPU time consumed with OpenMP programming model}\label{fig:cpu-time-openmp}
+\end{figure}

### 2015-10-22 papers/announcement/errors_analysis/errors_analysis.sh

@@ -1,8 +1,9 @@
 #!/bin/bash
 # script automotion for FOODIE's errors analysis of the provided tests
 echo "Building tests executables"
+date > builds.log
 cd ../../../
-FoBiS.py build -f src/tests/accuracy/oscillation/fobos -mode errors-analysis --build_dir papers/announcement/errors_analysis/tests > builds.log
+FoBiS.py build -f src/tests/accuracy/oscillation/fobos -mode errors-analysis --build_dir papers/announcement/errors_analysis/tests >> papers/announcement/errors_analysis/builds.log
 cd -
 
 echo "Oscillation test errors analysis"

### 2015-10-22 papers/announcement/openmp_benchmark/openmp_benchmark.sh

@@ -133,22 +133,23 @@ done
 
 # building codes
 echo "Building 1D Euler OpenMP test executable"
+date > builds.log
 cd ../../../
 if [ "$code" == "all" ] || [ "$code" == "foodie" ]; then
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/obj
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/mod
-  FoBiS.py build -mode benchmark-openmp-on --build_dir papers/announcement/openmp_benchmark/tests-openmp > papers/announcement/openmp_benchmark/builds.log
+  FoBiS.py build -f src/tests/parallel/euler-1D-openmp/fobos -mode benchmark-openmp-on --build_dir papers/announcement/openmp_benchmark/tests-openmp >> papers/announcement/openmp_benchmark/builds.log
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/obj
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/mod
-  FoBiS.py build -mode benchmark-openmp-off --build_dir papers/announcement/openmp_benchmark/tests-openmp >> papers/announcement/openmp_benchmark/builds.log
+  FoBiS.py build -f src/tests/parallel/euler-1D-openmp/fobos -mode benchmark-openmp-off --build_dir papers/announcement/openmp_benchmark/tests-openmp >> papers/announcement/openmp_benchmark/builds.log
 fi
 if [ "$code" == "all" ] || [ "$code" == "no-foodie" ]; then
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/obj
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/mod
-  FoBiS.py build -mode benchmark-openmp-no-foodie-on --build_dir papers/announcement/openmp_benchmark/tests-openmp >> papers/announcement/openmp_benchmark/builds.log
+  FoBiS.py build -f src/tests/parallel/euler-1D-openmp-no-foodie/fobos -mode benchmark-openmp-on --build_dir papers/announcement/openmp_benchmark/tests-openmp >> papers/announcement/openmp_benchmark/builds.log
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/obj
   rm -rf papers/announcement/openmp_benchmark/tests-openmp/mod
[…]

### 2015-10-22 papers/announcement/src/tests/oscillation/oscillation-solver-emd-rk.tex
Add some errors analysis for embedded RK

\subsubsection{Embedded Runge-Kutta}

\begin{table}[!ht]
  \centering
  \caption{Oscillation test: errors analysis of explicit embedded Runge-Kutta solvers}\label{tab:oscillation_errors_emd_rk}
  \begin{subtable}[b]{0.40\textwidth}
    \centering
    \caption{7 stages}\label{tab:oscillation-emd-rk-7}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      \toprule
      {\sc Mean Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
      \hline
      7352.9               &  0.626E-01    &  0.622E-01    & /             & /             \\
      3759.4               &  0.482E-02    &  0.480E-02    & 3.82          & 3.82          \\
      2272.7               &  0.793E-03    &  0.786E-03    & 3.59          & 3.59          \\
      1379.3               &  0.393E-04    &  0.390E-04    & 6.02          & 6.02          \\
       618.8               &  0.216E-05    &  0.213E-05    & 3.62          & 3.62          \\
       356.0               &  0.252E-06    &  0.249E-06    & 3.89          & 3.88          \\
      \bottomrule
    \end{tabular}}
  \end{subtable}\\
\end{table}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{errors-analysis/oscillation/errors_analysis-oscillation-emd-runge-kutta-7.png}
    \caption{7 stages}\label{fig:results-oscillation-emd-runge-kutta-7}
  \end{subfigure}\quad%
  \caption{Oscillation equations solutions computed by means of embedded Runge-Kutta solvers}\label{fig:results-oscillation-emd-runge-kutta-1-7}
\end{figure}

### 2015-11-05 papers/announcement/manuscript.tex
improve manuscript

@@
 \usepackage{listings}
-\lstdefinelanguage{MyFortran}[95]{Fortran}{morecomment=[l]{\#},morestring=[d]',morekeywords={procedure,pass,deferred,non_overridable,generic,class,is}}
+\lstdefinelanguage{MyFortran}[08]{Fortran}{morecomment=[l]{\#},morestring=[d]',morekeywords={procedure,pass,deferred,non_overridable,generic,class,is}}
 \lstdefinestyle{code}{%
@@
   columns=fixed,%
+  keepspaces=true,%
   xleftmargin=10pt,%
@@
   keywordstyle=\color{RoyalBlue},%
-  stringstyle={\sffamily},%
+  % stringstyle={\sffamily},%
   texcl=true,%
@@
   commentstyle=\color{Maroon}%
+}
+
+\lstdefinestyle{codesimple}{%
+  basicstyle=\footnotesize,%
+  backgroundcolor=\color{gray},%
+  captionpos=b,%
+  columns=fixed,%
+  keepspaces=true,%
+  xleftmargin=10pt,%
+  numbers=none,%
 }
@@
 
+\appendix
+
+\input{src/appendix/euler-1D-parallel-tests-API.tex}
+
 \bibliographystyle{mycpc2}

### 2015-11-05 papers/announcement/src/appendix/euler-1D-parallel-tests-API.tex
improve manuscript

\section{Euler 1D Parallel Tests API}

In subsections \ref{subsec:caf} and \ref{subsec:openmp} it has been proved that FOODIE usage does not penalize the parallel scaling of an equivalent procedural code implemented without FOODIE. To this aim, we have solved the Euler's conservation laws (in one dimension) by means of FOODIE: as a matter of fact, Euler 1D PDEs constitutes a complex test retaining many difficulties of real applications, but it is still simple enough to serve as benchmark test. In this section we report the implementation details of the codes developed to solve (with serial and parallel models) the Euler 1D PDEs system.

\subsection{Euler 1D baseline API}\label{subsec:euler-1D-API}

The 1D Euler PDEs system is a non linear, hyperbolic (inviscid) system of conservation laws for compressible gas dynamics, that reads
\begin{equation}
\begin{matrix}
U_t = R(U)  \Leftrightarrow U_t = F(U)_x \\
U = \begin{bmatrix}
\rho \\
\rho u \\
\rho E
\end{bmatrix}\;\;\;
F(U) = \begin{bmatrix}
\rho u \\
\rho u^2 + p \\
\rho u H
\end{bmatrix}
\end{matrix}
\label{eq:euler-1D}
\end{equation}
where $\rho$ is the density, $u$ is the velocity, $p$ the pressure, $E$ the total internal specific energy and $H$ the total specific enthalpy. The PDEs system must completed with the proper initial and boundary conditions. Moreover, an ideal (thermally and calorically perfect) gas is considered
\begin{equation}
\begin{matrix}
R = c_p - c_v \\
\gamma = \frac{c_p}{c_v}\\
e = c_v T \\
h = c_p T
\end{matrix}
\end{equation}
where $R$ is the gas constant, $c_p$ and $c_v$ are the specific heats at constant pressure and volume (respectively), $e$ is the internal energy, $h$ is the internal enthalpy and *T* is the temperature. The following addition equations of state hold:
\begin{equation}
\begin{matrix}
T = \frac{p}{\rho R} \\
E = \rho e + \frac{1}{2} \rho u^2 \\
H = \rho h + \frac{1}{2} \rho u^2 \\
a = \sqrt{\frac{\gamma p}{\rho}}
\end{matrix}
\end{equation}

An extension of the above Euler system is considered allowing the modelling of a multi-fluid mixture of different gas (with different physical characteristics). The well known Standard Thermodynamic Model is used to model the gas mixture replacing the density with the density fraction of each specie composing the mixture. This led to the following system:
\begin{equation}
\begin{matrix}
U_t = R(U)  \Leftrightarrow U_t = F(U)_x \\
U = \begin{bmatrix}
\rho_s \\
\rho u \\
\rho E
\end{bmatrix}\;\;\;
F(U) = \begin{bmatrix}
\rho_s u \\
\rho u^2 + p \\
\rho u H
\end{bmatrix}\;\;\; for\; s=1,2,...N_s \\
\rho = \sum_{s=1}^{N_s}\rho_s \\
c_p = \sum_{s=1}^{N_S} \frac{\rho_s}{\rho} c_{p,s} \quad  c_v = \sum_{s=1}^{N_S} \frac{\rho_s}{\rho} c_{v,s}
\end{matrix}
\end{equation}
where $N_s$ is the number of initial species composing the gas mixture.

\subsubsection{Memory organization}
The finite volume, Godunov's like approach is employed. Essentially, the method of Lines is used to decouple the space operator from the time one. Firstly, the space operator (the residual function of equation \ref{eq:euler-1D}) is numerically solved in order to reduce the original PDEs system to a system of ODEs that is then integrated over time by means of FOODIE solvers. Here we omit the details of the numerical models, interested readers can see \cite{}. On the contrary, some details on the memory organization is necessary to explaining the implemented API.

The conservative variables are co-located at the cell center. The cell and (inter)faces numeration is as shown in listing \ref{list:grid}.
\begin{lstlisting}[firstnumber=1,style=codesimple,caption={Numerical grid organization},label={list:grid}]
        cell   (inter)faces
         |            |
         v            v
 |----|----|-...-|----|---|---|---|-...-|-----|-----|-----|-...-|-------|-------|
 |1-Ng|2-Ng| ... |-1  | 0 | 1 | 2 | ... |  Ni | Ni+1| Ni+1| ... |Ni+Ng-1| Ni+Ng |
 |----|----|-...-|----|---|---|---|-...-|-----|-----|-----|-...-|-------|-------|
0-Ng                 -1   0   1   2    Ni-1   Ni                      Ni+Ng
\end{lstlisting}
In listing \ref{list:grid} $Ni$ is the number of finite volumes (cells) used for discretizing the domain and $Ng$ is the number of ghost cells used for imposing the left and right boundary conditions (for a total of $2Ng$ cells).
For each cell the conservative variables must be stored: this is done by means of of rank 2 array where the first index refers to the conservative variables (densities, momentum or energy) while the second index refers to the space location, namely the cell index.

The most CPU time consuming part of a finite volume scheme is the fluxes computation across the cells interfaces. Such a computation corresponds to a loop over all the cells interfaces. Listing \ref{list:pseudo-fluxes} shows a pseudo-code example of such a computation.

\begin{lstlisting}[firstnumber=1,style=code,caption={Pseudo-code example of fluxes computation},label={list:pseudo-fluxes}]
do i=0, Ni
  F(:, i) = compute_fluxes(U(:, i), U(:, i+1))
enddo
\end{lstlisting}

In the pseudo-code of listing \ref{list:pseudo-fluxes} it has been emphasized that the fluxes across an interface depends on the cells at left and right of the interface itself. The key point for the parallelization of such an algorithm is to compute the fluxes concurrently using as much as possible the parallel resources provided by the running architecture. As a consequence, the above showed loop over the whole domain is split into sub-domains (explicitly or implicitly accordingly to the parallel model adopted) and the fluxed of each sub-domain are computed concurrently.

\subsubsection{The serial API}

The conservative variables of 1D Euler's system can be easily implemented as a \emph{FOODIE integrand field} defining a concrete extension of the FOODIE \emph{integrand} type. Listing \ref{list:euler_1D_type} reports the implementation of such an integrand field that is contained into the tests suite shipped within the FOODIE library.

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{Euler 1D integrand} type},label={list:euler_1D_type}]
type, extends(integrand) :: euler_1D
  private
  integer(I_P)                   :: ord=0     ! Space accuracy formal order.
  integer(I_P)                   :: Ni=0      ! Space dimension.
  integer(I_P)                   :: Ng=0      ! Number of ghost cells for boundary conditions handling.
  integer(I_P)                   :: Ns=0      ! Number of initial species.
  integer(I_P)                   :: Nc=0      ! Number of conservative variables, Ns+2.
  integer(I_P)                   :: Np=0      ! Number of primitive variables, Ns+4.
  real(R_P)                      :: Dx=0._R_P ! Space step.
  type(weno_interpolator_upwind) :: weno      ! WENO interpolator.
  real(R_P),    allocatable      :: U(:,:)    ! Integrand (state) variables, whole physical domain [1:Nc,1:Ni].
  real(R_P),    allocatable      :: cp0(:)    ! Specific heat cp of initial species [1:Ns].
  real(R_P),    allocatable      :: cv0(:)    ! Specific heat cv of initial species [1:Ns].
  character(:), allocatable      :: BC_L      ! Left boundary condition type.
  character(:), allocatable      :: BC_R      ! Right boundary condition type.
  integer(I_P)                   :: me=0      ! ID of this\_image().
  integer(I_P)                   :: we=0      ! Number of CAF images used.
  contains
    ! auxiliary methods
    procedure, pass(self), public :: init
    procedure, pass(self), public :: destroy
    procedure, pass(self), public :: output
    procedure, pass(self), public :: dt => compute_dt
    ! ADT integrand deferred methods
    procedure, pass(self), public :: t => dEuler_dt
    procedure, pass(lhs),  public :: local_error => euler_local_error
    procedure, pass(lhs),  public :: integrand_multiply_integrand => euler_multiply_euler
    procedure, pass(lhs),  public :: integrand_multiply_real => euler_multiply_real
    procedure, pass(rhs),  public :: real_multiply_integrand => real_multiply_euler
    procedure, pass(lhs),  public :: add => add_euler
    procedure, pass(lhs),  public :: sub => sub_euler
    procedure, pass(lhs),  public :: assign_integrand => euler_assign_euler
    procedure, pass(lhs),  public :: assign_real => euler_assign_real
    ! private methods
    procedure, pass(self), private :: primitive2conservative
    procedure, pass(self), private :: conservative2primitive
    procedure, pass(self), private :: synchronize
    procedure, pass(self), private :: impose_boundary_conditions
    procedure, pass(self), private :: reconstruct_interfaces_states
    procedure, pass(self), private :: riemann_solver
    final                          :: finalize
endtype euler_1D
\end{lstlisting}

\subsection{Euler 1D CAF Enabled API}\label{subsec:euler-1D-CAF-API}

\subsection{Euler 1D OpenMP Enabled API}\label{subsec:euler-1D-OpenMP-API}

### 2015-11-05 papers/announcement/src/parallel/caf.tex
improve manuscript

\subsection{CAF benchmark}\label{subsec:caf}

This subsection reports the parallel scaling analysis of Euler 1D test programs (with and without FOODIE) being parallelized by means of CoArrays Fortran (CAF) model. This parallel model is based on the concept of \emph{coarray} introduced into the Fortran 2008 standard: the array syntax is extended introducing the so called \emph{codimension} that is a new arrays indexing. Essentially, a CAF enabled code is designed to be replicated a certain number of times and all copies, conventionally named \emph{images}, are executed asynchronously. Each image has its own set of data (memory) and the codimension indexes are used to access to the (remote) memory of the other images. The CAF approach allows the implementation of Partitioned Global Address Space (PGAS) model following the SPMD (single program, multiple data) parallelization paradigm. The programmer must take care of defining the coarray variables and of synchronizing the images when necessary. This approach requires the refactoring of legacy serial codes, but it allows the exploitation of both shared and distributed memory architectures. Moreover, it is a standard feature of Fortran (2008), thus it is not chained to any particular compiler vendors extension.

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.80\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{caf_benchmark/euler-1D-caf/strong-scaling-comparison.png}
    \caption{Strong scaling, number of cells 240000}\label{fig:strong-scaling-caf}
  \end{subfigure}\\
  \begin{subfigure}[b]{0.80\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{caf_benchmark/euler-1D-caf/weak-scaling-comparison.png}
    \caption{Weak scaling, minimum number of cells 24000}\label{fig:weak-scaling-caf}
  \end{subfigure}\\
  \caption{Scaling efficiency with CAF programming model}\label{fig:scaling-caf}
\end{figure}

The benchmarks shows in this section have been done on a \emph{dual Intel(R) Xeon(R) CPU X5650} exacores workstation for a total of 12 physical cores, coupled with 24GB of RAM. In order to perform an accurate analysis 4 different codes have considered:

\begin{itemize}
  \item FOODIE-aware codes:
    \begin{itemize}
      \item serial code;
      \item CAF-enabled code;
      \end{itemize}
  \item procedural codes without using FOODIE library:
    \begin{itemize}
      \item serial code;
      \item CAF-enabled code;
      \end{itemize}
  \end{itemize}

These codes (see \ref{subsec:euler-1D-CAF-API} for the implementation details) have been compiled by means of the GNU gfortran compiler v5.2.0 coupled with OpenCoarrays v1.1.0\footnote{OpenCoarrays is an open-source software project for developing, porting and tuning transport layers that support coarray Fortran (CAF) compilers, see \cite{opencoarrays}.}.

The Euler conservation laws are integrated for 30 time steps by means of the TVD RK(5,4) solver: the measured CPU time used for computing the scaling efficiencies is the average of the 30 integrations, thus representing the mean CPU time for computing one time step integration.

For the strong scaling, the benchmark has been conducted with 240000 finite volumes. Figure \ref{fig:strong-scaling-caf} summarizes the strong scaling analysis: it shows that FOODIE-based code scales similarly to the baseline code without FOODIE.

For the weak scaling the minimum size is 24000 finite volumes and the size is scaled linearly with the CAF images, thus $N_{12} = 288000$ cells. Figure \ref{fig:weak-scaling-caf} summarizes the weak scaling analysis and it essentially confirms that FOODIE-based code scales similarly to the baseline code without FOODIE.

Both strong and weak scaling analysis point out that for the computing architecture considered the parallel scaling is reasonable up to 12 cores, the efficiency being always satisfactory.

To complete the comparison, the absolute CPU-time consumed by the two families of codes (with and without FOODIE) must be considered. Table \ref{tab:caf-results} summarizes the benchmarks results. As shown, procedural and FOODIE-aware codes consume a very similar CPU-time for both the strong and the weak benchmarks. The same results are shown in figure \ref{fig:cpu-time-caf}. These results prove that the abstraction of FOODIE environment does not degrade the computational efficiency.

\begin{table}[!ht]
  \centering
  \caption{caf benchmarks results\label{tab:caf-results}}
  \begin{subtable}[b]{0.80\textwidth}
    \centering
    \caption{Strong benchmarks, number of cells 240000\label{tab:caf-results-strong}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{ccccc}
      {\sc Number of caf threads} & \multicolumn{4}{c}{\sc CPU time for 1 time step integration}              \\
      \hline
                                     & FOODIE serial & FOODIE parallel & procedural serial & procedural parallel \\
      \cmidrule{2-5}
      1                              & 3.2970        & 3.3297          & 3.0049            & 3.0563              \\
      2                              & /             & 1.6536          & /                 & 1.5686              \\
      4                              & /             & 0.8515          & /                 & 1.8116              \\
      8                              & /             & 0.4296          & /                 & 0.4130              \\
      12                             & /             & 0.3094          & /                 & 0.2839              \\
      \hline
    \end{tabular}}
  \end{subtable}\\
  \begin{subtable}[b]{0.80\textwidth}
    \centering
    \caption{Weak benchmarks, minimum number of cells 24000\label{tab:caf-results-weak}}
    \resizebox{1.00\textwidth}{!}{%
    \begin{tabular}{cccccc}
      {\sc Number of caf threads} & {\sc Number of Cells} & \multicolumn{4}{c}{\sc CPU time for 1 time step integration} \\
      \hline
                                     &                       & FOODIE serial & FOODIE parallel & procedural serial & procedural parallel \\
      \cmidrule{3-6}
      1                              & 24000                 & 0.3105        & 0.3159          & 0.3089            & 0.3133              \\
      2                              & 48000                 & /             & 0.3209          & /                 & 0.3185              \\
      4                              & 96000                 & /             & 0.3384          & /                 & 0.3269              \\
      8                              & 192000                & /             & 0.3449          & /                 & 0.3369              \\
      12                             & 288000                & /             & 0.4291          & /                 & 0.3657              \\
      \hline
    \end{tabular}}
  \end{subtable}
\end{table}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.40\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{caf_benchmark/euler-1D-caf/strong-cpu-time-comparison.png}
    \caption{Strong benchmark, number of cells 240000}\label{fig:strong-cpu-time-caf}
  \end{subfigure}\quad%
  \begin{subfigure}[b]{0.40\textwidth}
    \centering
    \includegraphics[width=1.00\textwidth]{caf_benchmark/euler-1D-caf/weak-cpu-time-comparison.png}
    \caption{Weak benchmark, minimum number of cells 24000}\label{fig:weak-cpu-time-caf}
  \end{subfigure}\\
  \caption{CPU time consumed with caf programming model}\label{fig:cpu-time-caf}
\end{figure}

### 2015-11-05 papers/announcement/src/parallel/openmp.tex
improve manuscript

@@
 
-This subsection reports the parallel scaling analysis performed with OpenMP directives-based paradigm.
-These benchmarks have been done on a \emph{dual Intel(R) Xeon(R) CPU X5650} exacores workstation for a total of 12 physical cores, coupled with 24GB of RAM.
+This subsection reports the parallel scaling analysis of Euler 1D test programs (with and without FOODIE) being parallelized by means of OpenMP directives-based paradigm.
+This parallel model is based on the concept of \emph{threads}: an OpenMP enabled code start a single (master) threaded program and, at run-time, it is able to generate a team of (many) threads that work concurrently on the parallelized parts of the code, thus reducing the CPU time necessary for completing such parts.
+The parallelization is made by means of \emph{directives} explicitly inserted by the programmer: the communications between threads are automatically handled by the compiler (through the provided OpenMP library used as back-end).
+OpenMP parallel paradigm is not a standard feature of Fortran, rather it is an extension provided by the compiler vendors.
+This parallel paradigm constitutes an effective and easy approach for parallelizing legacy serial codes, however its usage is limited to shared memory architectures because all threads must have access to the same memory.
+
+\begin{figure}[!ht]
+  \centering
+  \begin{subfigure}[b]{0.80\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-scaling-comparison.png}
+    \caption{Strong scaling, number of cells 240000}\label{fig:strong-scaling-openmp}
+  \end{subfigure}\\
+  \begin{subfigure}[b]{0.80\textwidth}
+    \centering
+    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-scaling-comparison.png}
+    \caption{Weak scaling, minimum number of cells 24000}\label{fig:weak-scaling-openmp}
+  \end{subfigure}\\
+  \caption{Scaling efficiency with OpenMP programming model}\label{fig:scaling-openmp}
+\end{figure}
+
+The benchmarks shown in this section have been done on a \emph{dual Intel(R) Xeon(R) CPU X5650} exacores workstation for a total of 12 physical cores, coupled with 24GB of RAM.
 In order to perform an accurate analysis 4 different codes have considered:
@@
 
-These codes (see appendix \ref{sec:euler-1D-OpenMP-API} for the implementation details) have been compiled by means of the GNU gfortran compiler v5.2.0 with \emph{-O2 -fopenmp} compilation flags.
+These codes (see \ref{subsec:euler-1D-OpenMP-API} for the implementation details) have been compiled by means of the GNU gfortran compiler v5.2.0 with \emph{-O2 -fopenmp} compilation flags.
 
-The Euler conservation laws are integrated for 10 time steps by means of the TVD RK(5,4) solver: the measured CPU time used for computing the scaling efficiencies is the average of the 10 integrations, thus representing the mean CPU time for computing one time step integration.
+The Euler conservation laws are integrated for 30 time steps by means of the TVD RK(5,4) solver: the measured CPU time used for computing the scaling efficiencies is the average of the 30 integrations, thus representing the mean CPU time for computing one time step integration.
 
-\begin{figure}[!ht]
-  \centering
-  \begin{subfigure}[b]{0.85\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-scaling-comparison.png}
-    \caption{Strong scaling, number of cells 100000}\label{fig:strong-scaling-openmp}
-  \end{subfigure}\\
-  \begin{subfigure}[b]{0.85\textwidth}
-    \centering
-    \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-scaling-comparison.png}
-    \caption{Weak scaling, minimum number of cells 10000}\label{fig:weak-scaling-openmp}
-  \end{subfigure}\\
-  \caption{Scaling efficiency with OpenMP programming model}\label{fig:scaling-openmp}
-\end{figure}
-
-For the strong scaling, the benchmark has been conducted with 100000 finite volumes.
+For the strong scaling, the benchmark has been conducted with 240000 finite volumes.
 Figure \ref{fig:strong-scaling-openmp} summarizes the strong scaling analysis: it shows that FOODIE-based code scales similarly to the baseline code without FOODIE.
 
-For the weak scaling the minimum size is 10000 finite volumes and the size is scaled linearly with the OpenMP threads, thus $N_{12} = 120000$ cells.
+For the weak scaling the minimum size is 24000 finite volumes and the size is scaled linearly with the OpenMP threads, thus $N_{12} = 288000$ cells.
 Figure \ref{fig:weak-scaling-openmp} summarizes the weak scaling analysis and it essentially confirms that FOODIE-based code scales similarly to the baseline code without FOODIE.
 
-Both strong and weak scaling analysis point out that for the computing architecture considered the parallel scaling is reasonable up to 8 cores: using 12 cores the measured efficiencies become unsatisfactory, reducing below the 50\%.
+Both strong and weak scaling analysis point out that for the computing architecture considered the parallel scaling is reasonable up to 8 cores: using 12 cores the measured efficiencies become unsatisfactory, reducing below the 60\%.
 
@@
     \centering
-    \caption{Strong benchmarks, number of cells 100000\label{tab:openmp-results-strong}}
+    \caption{Strong benchmarks, number of cells 240000\label{tab:openmp-results-strong}}
     \resizebox{1.00\textwidth}{!}{%
@@
       \cmidrule{2-5}
-      1                              & 1.2941        & 1.2942          & 1.2820            & 1.2769              \\
-      2                              & /             & 0.7472          & /                 & 0.7784              \\
-      4                              & /             & 0.3988          & /                 & 0.4218              \\
-      8                              & /             & 0.2185          & /                 & 0.2559              \\
-      12                             & /             & 0.2352          & /                 & 0.2272              \\
+      1                              & 3.3466        & 3.3076          & 3.1252            & 3.0873              \\
+      2                              & /             & 1.8166          & /                 & 1.7765              \\
+      4                              & /             & 0.9798          & /                 & 1.0085              \\
+      8                              & /             & 0.5192          & /                 & 0.5055              \\
+      12                             & /             & 0.4847          & /                 & 0.4748              \\
       \hline
@@
     \centering
-    \caption{Weak benchmarks, minimum number of cells 10000\label{tab:openmp-results-weak}}
+    \caption{Weak benchmarks, minimum number of cells 24000\label{tab:openmp-results-weak}}
     \resizebox{1.00\textwidth}{!}{%
@@
       \cmidrule{3-6}
-      1                              & 10000                 & 0.1496        & 0.1419          & 1.2820            & 1.2769              \\
-      2                              & 20000                 & /             & 0.1643          & /                 & 0.7784              \\
-      4                              & 40000                 & /             & 0.1671          & /                 & 0.4218              \\
-      8                              & 80000                 & /             & 0.1752          & /                 & 0.2559              \\
-      12                             & 120000                & /             & 0.2883          & /                 & 0.2272              \\
+      1                              & 24000                 & 0.3171        & 0.3162          & 0.3089            & 0.3111              \\
+      2                              & 48000                 & /             & 0.3492          & /                 & 0.3854              \\
+      4                              & 96000                 & /             & 0.3666          & /                 & 0.4069              \\
+      8                              & 192000                & /             & 0.3862          & /                 & 0.4142              \\
+      12                             & 288000                & /             & 0.5727          & /                 & 0.6142              \\
       \hline
@@
   \centering
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.40\textwidth}
     \centering
     \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/strong-cpu-time-comparison.png}
-    \caption{Strong benchmark, number of cells 100000}\label{fig:strong-cpu-time-openmp}
+    \caption{Strong benchmark, number of cells 240000}\label{fig:strong-cpu-time-openmp}
   \end{subfigure}\quad%
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.40\textwidth}
     \centering
     \includegraphics[width=1.00\textwidth]{openmp_benchmark/euler-1D-openmp/weak-cpu-time-comparison.png}
-    \caption{Weak benchmark, minimum number of cells 10000}\label{fig:weak-cpu-time-openmp}
+    \caption{Weak benchmark, minimum number of cells 24000}\label{fig:weak-cpu-time-openmp}
   \end{subfigure}\\

### 2015-11-05 papers/announcement/src/parallel/parallel.tex
improve manuscript

@@
   \item OpenMP directive-based model, for only shared memory architectures;
-  \item Message Passing Interface (MPI) library-based model, for only distributed memory architectures;
   \end{itemize}
@@
 
-The same 1D Euler PDE problem is also solved by parallel-enabled codes that are not based on FOODIE: their solutions provide a reference for measuring the effect of FOODIE abstraction on the parallel scaling.
+The same 1D Euler PDEs problem is also solved by parallel-enabled codes that are not based on FOODIE: their solutions provide a reference for measuring the effect of FOODIE abstraction on the parallel scaling.
 
-\subsection{CAF benchmark}\label{subsec:caf}
-
-{\color{red} To be written.}
+\input{src/parallel/caf.tex}
 
 \input{src/parallel/openmp.tex}
-
-\subsection{MPI benchmark}\label{subsec:mpi}
-
-{\color{red} To be written.}

### 2015-11-05 papers/announcement/openmp_benchmark/openmp_benchmark.sh

@@ -51,7 +51,7 @@ function strong_scaling {
     exe='../tests-openmp/euler-1D-openmp-no-foodie'
     output='strong-scaling-no-foodie.dat'
   fi
-  echo "Strong scaling"
+  echo "Strong scaling, size $Ni"
   echo 'TITLE="Strong scaling analysis, cells '$Ni'"' > $output
   echo 'VARIABLES="OpenMP threads number" "CPU time"' >> $output
   echo "Benchmarking without OpenMP threads"

### 2017-05-05 papers/announcement/manuscript.tex
update paper

@@
 The present paper is the first \emph{manifesto} of FOODIE, a library aimed to numerically solve ODEs problems by means of a clear, concise and efficient \emph{abstract} interface.
-FOODIE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has manifolds aims: to provide a set to built-in numerical schemes that are accurate, robust, validated and efficient and to allow easy application of these schemes to (almost) all ODEs problems by means of an effective Abstract Calculus Pattern.
+FOODIE, meaning Fortran Object oriented Ordinary Differential equations Integration Environment, has manifolds aims: to provide a set of built-in numerical schemes that are accurate, robust, validated and efficient and to allow easy application of these schemes to (almost) all ODEs problems by means of an effective Abstract Calculus Pattern.
 The key idea is to allow the same solver-implementation to be applied to all ODEs problems thus avoiding the re-implementation of the ODEs solver for each different ODEs problem: code re-usability is consequently maximized, FOODIE being a general robust framework.
@@
 
+\input{src/appendix/solvers-details.tex}
+
 \input{src/appendix/euler-1D-parallel-tests-API.tex}

### 2017-05-05 papers/announcement/src/API/API.tex
update paper

@@
 
-As aforementioned, the programming language used is the Fortran 2008 standard, that is a minor revision of the previous Fortran 2003 standard.
-Such a new Fortran idioms provide (among other useful features) an almost complete support for OOP, in particular for ADT concept.
-Fortran 2003 has introduced the \emph{abstract derived type}: it is a derived type suitable to serve as \emph{contract} for concrete type-extensions that has not any actual implementations, rather it provides a well-defined set of type bound procedures interfaces, that in Fortran nomenclature are called \emph{deferred} procedures.
-Using such an abstract definition, we can implement algorithms operating on only this abstract type and on \emph{all its concrete extensions}.
-This is the key feature of FOODIE library: all the above described ODE solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax.
-In the meanwhile, client codes must implement their own IVPs extending only one simple abstract type.
+As aforementioned, the programming language used is the Fortran 2008 (or newer) standard, that is a minor revision of the previous Fortran 2003 standard.
+Such a new Fortran idioms provide (among other useful features) an almost complete support for OOP, in particular for Abstract Data Type (ADT) concept.
+Fortran 2003 has introduced the \emph{abstract derived type}: it is a derived type suitable to serve as \emph{contract} for concrete type-extensions that has not any actual implementations, rather it provides a well-defined set of methods (in Fortran jargon they are \emph{type bound procedures}) interfaces, that in Fortran nomenclature are called \emph{deferred} procedures.
+Using such an abstract definition, we can implement algorithms operating on only this abstract type \emph{and on all its concrete extensions}.
+This is the key feature of FOODIE library: all the above described ODEs solvers are implemented on the knowledge of \emph{only one abstract type}, allowing an implementation-style based on a very high-level syntax.
+In the meanwhile, client codes must implement their own IVPs extending only one simple abstract type with minimal requirements.
 
-In the subsection \ref{subsec:type_integrand} a review of the FOODIE main ADT, the \emph{integrand} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.
+In the subsection \ref{subsec:type_integrand} a review of the FOODIE main ADT, the \emph{integrand\_object} type, is provided, while subsections \ref{subsec:solver_euler}, \ref{subsec:solver_tvd_rk}, \ref{subsec:solver_ls_rk}, \ref{subsec:solver_ab} and \ref{subsec:solver_leapfrog} cover the API of the currently implemented solvers.
 
-It is worth noting that all FOODIE public \emph{entities} (ADT and solvers) must be accessed by the FOODIE module, see listing \ref{list:main_module} for an example on how access to all public FOODIE entities.
+It is worth noting that all FOODIE public \emph{entities} (ADT and solvers) must be accessed by the FOODIE module, see listing \ref{list:main_module} for an example on how to access to all public FOODIE entities.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={usage example importing all public entities of FOODIE main module},label={list:main_module}]
-use foodie, only: integrand, &
-                  adams_bashforth_integrator, &
-                  adams_moulton_integrator, &
-                  adams_bashforth_moulton_integrator, &
-                  euler_explicit_integrator, &
-                  leapfrog_integrator, &
-                  ls_runge_kutta_integrator, &
-                  tvd_runge_kutta_integrator
+use foodie, only : integrand_object,                   &
+                   integrator_object,                  &
+                   integrator_adams_bashforth,         &
+                   integrator_adams_bashforth_moulton, &
+                   integrator_adams_moulton,           &
+                   integrator_back_df,                 &
+                   integrator_euler_explicit,          &
+                   integrator_leapfrog,                &
+                   integrator_lmm_ssp,                 &
+                   integrator_lmm_ssp_vss,             &
+                   integrator_ms_runge_kutta_ssp,      &
+                   integrator_runge_kutta_emd,         &
+                   integrator_runge_kutta_ls,          &
+                   integrator_runge_kutta_lssp,        &
+                   integrator_runge_kutta_ssp
 ! or simply
@@
 
-\subsection{The main FOODIE Abstract Data Type: the integrand type}\label{subsec:type_integrand}
+\subsection{Main FOODIE Abstract Data Type: the integrand type}\label{subsec:type_integrand}
 
-The implemented ACP is based on one main ADT, the \emph{integrand} type, the definition of which is shown in listing~\ref{list:type_integrand}.
+The implemented ACP is based on one main ADT, the \emph{integrand\_object} type, the definition of which is shown in listing~\ref{list:type_integrand}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={integrand type definition},label={list:type_integrand}]
-type, abstract :: integrand
+type, abstract :: integrand_object
   !< Abstract type for building FOODIE ODE integrators.
@@
     ! public deferred procedures that concrete integrand-field must implement
-    procedure(time_derivative),       pass(self), deferred, public:: t
+    procedure(time_derivative), pass(self), deferred :: t
     ! operators
-    procedure(symmetric_operator),   pass(lhs), deferred, public:: integrand_multiply_integrand
-    procedure(integrand_op_real),    pass(lhs), deferred, public:: integrand_multiply_real
-    procedure(real_op_integrand),    pass(rhs), deferred, public:: real_multiply_integrand
-    procedure(symmetric_operator),   pass(lhs), deferred, public:: add
-    procedure(symmetric_operator),   pass(lhs), deferred, public:: sub
-    procedure(assignment_integrand), pass(lhs), deferred, public:: assign_integrand
-    ! operators overloading
-    generic, public:: operator(+) => add
-    generic, public:: operator(-) => sub
-    generic, public:: operator(*) => integrand_multiply_integrand, &
+    procedure(local_error_operator), pass(lhs), deferred :: local_error
+    generic :: operator(.lterror.) => local_error ! Estimate local truncation error.
+    procedure(symmetric_operator), pass(lhs), deferred :: integrand_add_integrand
+    procedure(integrand_op_real),  pass(lhs), deferred :: integrand_add_real
+    procedure(real_op_integrand),  pass(rhs), deferred :: real_add_integrand
+    generic :: operator(+) => integrand_add_integrand, &
+                              integrand_add_real,      &
+                              real_add_integrand
+    procedure(symmetric_operator),       pass(lhs), deferred :: integrand_multiply_integrand
+    procedure(integrand_op_real),        pass(lhs), deferred :: integrand_multiply_real
+    procedure(real_op_integrand),        pass(rhs), deferred :: real_multiply_integrand
+    procedure(integrand_op_real_scalar), pass(lhs), deferred :: integrand_multiply_real_scalar
+    procedure(real_scalar_op_integrand), pass(rhs), deferred :: real_scalar_multiply_integrand
+    generic :: operator(*) => integrand_multiply_integrand,   &
+                              integrand_multiply_real,        &
                               real_multiply_integrand,        &
-                                     integrand_multiply_real
-    generic, public:: assignment(=) => assign_integrand
-endtype integrand
+                              integrand_multiply_real_scalar, &
+                              real_scalar_multiply_integrand
+    procedure(symmetric_operator), pass(lhs), deferred :: integrand_sub_integrand
+    procedure(integrand_op_real),  pass(lhs), deferred :: integrand_sub_real
+    procedure(real_op_integrand),  pass(rhs), deferred :: real_sub_integrand
+    generic :: operator(-) => integrand_sub_integrand, &
+                              integrand_sub_real,      &
+                              real_sub_integrand
+    procedure(assignment_integrand), pass(lhs), deferred :: assign_integrand
+    procedure(assignment_real),      pass(lhs), deferred :: assign_real
+    generic :: assignment(=) => assign_integrand, assign_real
+endtype integrand_object
 \end{lstlisting}
 
-The \emph{integrand} type does not implement any actual integrand field, it being and abstract type.
+The \emph{integrand\_object} type does not implement any actual integrand field, it being and abstract type.
 It only specifies which deferred procedures are necessary for implementing an actual concrete integrand type that can use a FOODIE solver.
 
-As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 1 ODE-specific procedure plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects.
+As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is limited: essentially, there are 1 ODEs-specific procedure and some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects.
 These procedures are analyzed in the following paragraphs.
+
+It is worth noting that for keeping the computational efficiency high a compromise has been done in the TBPs API design: in order to avoid excessive overhead due to polymorphic entities allocation\footnote{Polymorphic allocation (and deallocation) can be very complex to optimize and it could easily generate memory leaks.} and to allows \emph{pure}\footnote{A TBP can be defined \emph{pure} when it do not have \emph{side-effects}.
+Purity is important for enabling compiler optimizations (otherwise prevented), for allowing easy parallelization and for imposing a clear and easy-to-debug style to the programmer.} operators, all the math operators are designed to return a real array rather than an allocatable class of \emph{integrand\_object}.
+The \emph{integrand\_object} is then \emph{restored} by the assignment operator that updates an instance of \emph{integrand\_object} by the real array resulting from the math operators.
+This is a clear restriction to the ADT abstraction level: it imposes a \emph{pattern} to the programmer that could be avoided, but, in the meanwhile, it greatly improves FOODIE performance retaining a reasonable level of abstraction.
 
@@
 function time_derivative(self, t) result(dState_dt)
-import :: integrand, R_P, I_P
-class(integrand),       intent(IN) :: self      !< Integrand field.
-real(R_P),    optional, intent(IN) :: t         !< Time.
-class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.
+import :: integrand_object, R_P
+class(integrand_object), intent(in) :: self         ! Integrand field.
+real(R_P),    optional,  intent(in) :: t            ! Time.
+real(R_P), allocatable              :: dState_dt(:) ! Result of the time derivative function of integrand field.
 endfunction time_derivative
@@
 
-This procedure-function takes two arguments, the first passed as a \emph{type bounded} argument, while the latter is optional, and it returns an integrand object.
-The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT.
+This procedure-function takes two arguments, the first passed as a \emph{type bounded} argument, while the latter is optional, and it returns a real array of kind \emph{R\_P} (that is a parametric constant of FOODIE, it being customizable at compile-time).
+The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand\_object} ADT.
 The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.
 
-Commonly, into the concrete implementation of this deferred abstract procedure clients embed the actual ODE equations being solved.
+Commonly, into the concrete implementation of this deferred abstract procedure clients embed the actual ODEs system being solved.
 As an example, for the Burgers equation, that is a Partial Differential Equations (PDE) system involving also a boundary value problem, this procedure embeds the spatial operator that convert the PDE to a system of algebraic ODE.
-As a consequence, the eventual concrete implementation of this procedure can be very complex and errors-prone.
+As a consequence, the eventual concrete implementation of this procedure can be of any level of complexity.
 Nevertheless, the FOODIE solvers are implemented only on the above abstract interface, thus emancipating the solvers implementation from any concrete complexity.
-
-{\color{red} Add citations to Burgers, Adams-Bashfort, leapfrog references.}
 
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={symmetric operator procedure interface},label={list:symmetric_operator}]
-function symmetric_operator(lhs, rhs) result(operator_result)
-import :: integrand
-class(integrand), intent(IN)  :: lhs             !< Left hand side.
-class(integrand), intent(IN)  :: rhs             !< Right hand side.
-class(integrand), allocatable :: operator_result !< Operator result.
+pure function symmetric_operator(lhs, rhs) result(operator_result)
+import :: integrand_object, R_P
+class(integrand_object), intent(in) :: lhs                ! Left hand side.
+class(integrand_object), intent(in) :: rhs                ! Right hand side.
+real(R_P), allocatable              :: operator_result(:) ! Operator result.
 endfunction symmetric_operator
@@
 
-This interface defines a class of procedures operating on 2 \emph{integrand} objects, namely it is used for the definition of the operators \emph{multiplication}, \emph{summation} and \emph{subtraction} of integrand objects.
+This interface defines a class of procedures operating on 2 \emph{integrand\_object} arguments, namely it is used for the definition of the operators \emph{multiplication}, \emph{summation} and \emph{subtraction} of integrand objects.
 These operators are used into the above described ODE solvers, for example see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
 The implementation details of such a procedures class are strictly dependent on the concrete extension of the integrand type.
-From the FOODIE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are two integrand object and the returned object is still an integrand one.
 
@@
 \begin{lstlisting}[firstnumber=1,style=code,caption={Integrand/real and real/integrand operators procedure interfaces},label={list:integrand_real_ops}]
-function integrand_op_real(lhs, rhs) result(operator_result)
-import :: integrand, R_P
-class(integrand), intent(IN)  :: lhs              !< Left hand side.
-real(R_P),        intent(IN)  :: rhs              !< Right hand side.
-class(integrand), allocatable :: operator_result  !< Operator result.
+pure function integrand_op_real(lhs, rhs) result(operator_result)
+import :: integrand_object, R_P
+class(integrand_object), intent(in) :: lhs                ! Left hand side.
+real(R_P),               intent(in) :: rhs(1:)            ! Right hand side.
+real(R_P), allocatable              :: operator_result(:) ! Operator result.
 endfunction integrand_op_real
 
-function real_op_integrand(lhs, rhs) result(operator_result)
-import :: integrand, R_P
-real(R_P),        intent(IN)  :: lhs             !< Left hand side.
-class(integrand), intent(IN)  :: rhs             !< Right hand side.
-class(integrand), allocatable :: operator_result !< Operator result.
+pure function real_op_integrand(lhs, rhs) result(operator_result)
+!< Asymmetric type operator `real.op.integrand`.
+import :: integrand_object, R_P
+class(integrand_object), intent(in) :: rhs                ! Right hand side.
+real(R_P),               intent(in) :: lhs(1:)            ! Left hand side.
+real(R_P), allocatable              :: operator_result(:) ! Operator result.
 endfunction real_op_integrand
+
+pure function integrand_op_real_scalar(lhs, rhs) result(operator_result)
+!< Asymmetric type operator `integrand.op.real`.
+import :: integrand_object, R_P
+class(integrand_object), intent(in) :: lhs                ! Left hand side.
+real(R_P),               intent(in) :: rhs                ! Right hand side.
+real(R_P), allocatable              :: operator_result(:) ! Operator result.
+endfunction integrand_op_real_scalar
+
+pure function real_scalar_op_integrand(lhs, rhs) result(operator_result)
+!< Asymmetric type operator `real.op.integrand`.
+import :: integrand_object, R_P
+real(R_P),               intent(in) :: lhs                ! Left hand side.
+class(integrand_object), intent(in) :: rhs                ! Right hand side.
+real(R_P), allocatable              :: operator_result(:) ! Operator result.
+endfunction real_scalar_op_integrand
 \end{lstlisting}
 
-These two interfaces are necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the multiplication of an integrand object for a real number, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
+These foure interfaces are necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the multiplication of an integrand object for a real number (scalar or array), circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
 The implementation details of these procedures are strictly dependent on the concrete extension of the integrand type.
-From the FOODIE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the operator, and the second argument, the right-hand-side of the operator, are an integrand object and real number of viceversa and the returned object is still an integrand one.
 
@@
 
-The abstract interface of \emph{integrand assignment} procedure is shown in listing \ref{list:integrand_assignment}.
+The abstract interfaces of \emph{integrand assignment} procedure is shown in listing \ref{list:integrand_assignment}.
 
 \begin{lstlisting}[firstnumber=1,style=code,caption={integrand assignment procedure interface},label={list:integrand_assignment}]
-subroutine assignment_integrand(lhs, rhs)
-import :: integrand
-class(integrand), intent(INOUT) :: lhs !< Left hand side.
-class(integrand), intent(IN)    :: rhs !< Right hand side.
+pure subroutine assignment_integrand(lhs, rhs)
+import :: integrand_object
+class(integrand_object), intent(inout) :: lhs ! Left hand side.
+class(integrand_object), intent(in)    :: rhs ! Right hand side.
 endsubroutine assignment_integrand
+
+pure subroutine assignment_real(lhs, rhs)
+!< Symmetric assignment integrand = integrand.
+import :: integrand_object, R_P
+class(integrand_object), intent(inout) :: lhs     ! Left hand side.
+real(R_P),               intent(in)    :: rhs(1:) ! Right hand side.
+endsubroutine assignment_real
 \end{lstlisting}
 
-The assignment statement is necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the assignment of an integrand object by another one, circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
+The assignment statement is necessary in order to complete the \emph{algebra} operating on the integrand object class, allowing the
+assignment of an integrand object by another one or the assignment of an integrand object and a real array (arising from the result
+of other operators), circumstance that happens in all solvers, see equations \ref{eq:solver-euler}, \ref{eq:RK}, \ref{eq:AB} or \ref{eq:leapfrog}.
 The implementation details of this assignment is strictly dependent on the concrete extension of the integrand type.
-From the FOODIE solvers point of view, we need to known only that first argument passed as bounded one, the left-hand-side of the assignment, and the second argument, the right-hand-side of the assignment, are two integrand objects.
 
@@
 
-\subsection{The explicit forward Euler solver}\label{subsec:solver_euler}
+\subsection{Explicit forward Euler solver API}\label{subsec:solver_euler}
 
@@
 
-\subsection{The explicit TVD/SSP Runge-Kutta class of solvers}\label{subsec:solver_tvd_rk}
+\subsection{Explicit Adams-Bashforth class API}\label{subsec:solver_ab}
+
+The explicit Adams-Bashforth class of solvers is exposed as a single derived type named \emph{adams\_bashforth\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
+  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time;
+  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
+  \end{itemize}
+
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_ab_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Adams-Bashforth integrator},label={list:define_ab_solver}]
+use FOODIE, only: adams_bashforth_integrator
+type(adams_bashforth_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ab_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Adams-Bashforth integrator},label={list:init_ab_solver}]
+call integrator%init(steps=4)
+\end{lstlisting}
+
+In the listing \ref{list:init_ab_solver} a 4-steps solver has been initialized.
+As a matter of facts, from the equation \ref{eq:AB} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen.
+The complete definition of the \emph{adams\_bashforth\_integrator} type is reported into listing \ref{list:ab_solver}.
+As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_bashforth\_integrator} type},label={list:ab_solver}]
+type :: adams_bashforth_integrator
+  private
+  integer(I_P)           :: steps=0 ! Number of time steps.
+  real(R_P), allocatable :: b(:)    ! \(b\) coefficients.
+  contains
+    procedure, pass(self), public :: destroy
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+    procedure, pass(self), public :: update_previous
+    final                         :: finalize
+endtype adams_bashforth_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ab_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
+real               :: times(1:4)
+type(my_integrand) :: my_field
+type(my_integrand) :: previous(1:4)
+call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.
+
+The complete implementation of the \emph{integrate} method of the explicit Adams-Bashforth class of solvers is reported in listing \ref{list:ab_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
+subroutine integrate(self, U, previous, Dt, t, autoupdate)
+class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
+class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
+class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
+real(R_P),                         intent(IN)    :: Dt           ! Time steps.
+real(R_P),                         intent(IN)    :: t(:)         ! Times.
+logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
+logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
+integer(I_P)                                     :: s            ! Steps counter.
+autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
+do s=1, self%steps
+  U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
+enddo
+if (autoupdate_) call self%update_previous(U=U, previous=previous)
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+
+It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
+This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.
+
+\subsection{Leapfrog solver}\label{subsec:solver_leapfrog}
+
+The explicit Leapfrog class of solvers is exposed as a single derived type named \emph{leapfrog\_integrator}. This type provides three methods:
+
+\begin{itemize}
+  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
+  \item \emph{integrate}: integrate integrand field one-step-over in time.
+  \end{itemize}
+
+As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_lf_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Leapfrog integrator},label={list:define_lf_solver}]
+use FOODIE, only: leapfrog_integrator
+type(leapfrog_integrator) :: integrator
+\end{lstlisting}
+
+Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_lf_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Leapfrog integrator},label={list:init_lf_solver}]
+! default coefficients nu=0.01, alpha=0.53
+call integrator%init()
+! custom coefficients
+call integrator%init(nu=0.015, alpha=0.6)
+\end{lstlisting}
+
+The complete definition of the \emph{leapfrog\_integrator} type is reported into listing \ref{list:lf_solver}.
+As shown, the filter coefficients are initialized to zero, suitable values are initialized by the \emph{init} method.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{leapfrog\_integrator} type},label={list:lf_solver}]
+type :: leapfrog_integrator
+  private
+  real(R_P) :: nu=0.01_R_P    ! Robert-Asselin filter coefficient.
+  real(R_P) :: alpha=0.53_R_P ! Robert-Asselin-Williams filter coefficient.
+  contains
+    procedure, pass(self), public :: init
+    procedure, pass(self), public :: integrate
+endtype leapfrog_integrator
+\end{lstlisting}
+
+After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_lf_solver}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a Leapfrog integrator},label={list:use_lf_solver}]
+real               :: times(1:2)
+type(my_integrand) :: filter_displacement
+type(my_integrand) :: my_field
+type(my_integrand) :: previous(1:2)
+call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, &
+                          t=times)
+\end{lstlisting}
+where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{previous} are the memory registers where previous time steps solutions are saved, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.
+
+The complete implementation of the \emph{integrate} method of the explicit Leapfrog class of solvers is reported in listing \ref{list:lf_integrate}.
+
+\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Leapfrog class},label={list:lf_integrate}]
+subroutine integrate(self, U, previous, Dt, t, filter)
+class(leapfrog_integrator), intent(IN)    :: self          ! LF integrator.
+class(integrand),           intent(INOUT) :: U             ! Field to be integrated.
+class(integrand),           intent(INOUT) :: previous(1:2) ! Previous time steps solutions.
+real(R_P),                  intent(in)    :: Dt            ! Time step.
+real(R_P),                  intent(IN)    :: t             ! Time.
+class(integrand), optional, intent(INOUT) :: filter        ! Filter field displacement.
+U = previous(1) + previous(2)%t(t=t) * (Dt * 2._R_P)
+if (present(filter)) then
+  filter = (previous(1) - previous(2) * 2._R_P + U) * self%nu * 0.5_R_P
+  previous(2) = previous(2) + filter * self%alpha
+  U = U + filter * (self%alpha - 1._R_P)
+endif
+previous(1) = previous(2)
+previous(2) = U
+return
+endsubroutine integrate
+\end{lstlisting}
+
+This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the optional filter-displacement-register, the fifth is the time step used and the sixth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
+The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
+It is worth noting that if the filter displacement argument is not passed, the solver reverts back to the standard unfiltered Leapfrog method.
+
+It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
+In particular, if the filter displacement argument is passed the method performs the RAW filtering.
+
+\subsection{Explicit SSP Linear Mutistep Methods class API}
+
+{\color{red}To be written.}
+
+\subsection{Explicit SSP Linear Mutistep Methods with variable step-size class API}
+
+{\color{red}To be written.}
+
+\subsection{Explicit SSP Linear Mutistep Methods Runge-Kutta class API}
+
+{\color{red}To be written.}
+
+\subsection{Explicit SSP Runge-Kutta class API}\label{subsec:solver_tvd_rk}
 
@@
 
-\subsection{The explicit low storage Runge-Kutta class of solvers}\label{subsec:solver_ls_rk}
+\subsection{Explicit low storage Runge-Kutta class API}\label{subsec:solver_ls_rk}
 
@@
 
-\subsection{The explicit Adams-Bashforth class of solvers}\label{subsec:solver_ab}
-
-The explicit Adams-Bashforth class of solvers is exposed as a single derived type named \emph{adams\_bashforth\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{destroy}: destroy the integrator previously initialized, eventually freeing the allocated dynamic memory registers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time;
-  \item \emph{update\_previous}: auto update (cyclically) previous time steps solutions.
-  \end{itemize}
-
-As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_ab_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Adams-Bashforth integrator},label={list:define_ab_solver}]
-use FOODIE, only: adams_bashforth_integrator
-type(adams_bashforth_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_ab_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Adams-Bashforth integrator},label={list:init_ab_solver}]
-call integrator%init(steps=4)
-\end{lstlisting}
-
-In the listing \ref{list:init_ab_solver} a 4-steps solver has been initialized.
-As a matter of facts, from the equation \ref{eq:AB} a solver belonging to this class is completely defined once the number of time steps adopted has been chosen.
-The complete definition of the \emph{adams\_bashforth\_integrator} type is reported into listing \ref{list:ab_solver}.
-As shown, the linear coefficients are stored as allocatable arrays the values of which are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{adams\_bashforth\_integrator} type},label={list:ab_solver}]
-type :: adams_bashforth_integrator
-  private
-  integer(I_P)           :: steps=0 ! Number of time steps.
-  real(R_P), allocatable :: b(:)    ! \(b\) coefficients.
-  contains
-    procedure, pass(self), public :: destroy
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-    procedure, pass(self), public :: update_previous
-    final                         :: finalize
-endtype adams_bashforth_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_ab_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
-real               :: times(1:4)
-type(my_integrand) :: my_field
-type(my_integrand) :: previous(1:4)
-call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.
-
-The complete implementation of the \emph{integrate} method of the explicit Adams-Bashforth class of solvers is reported in listing \ref{list:ab_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
-subroutine integrate(self, U, previous, Dt, t, autoupdate)
-class(adams_bashforth_integrator), intent(IN)    :: self         ! Actual AB integrator.
-class(integrand),                  intent(INOUT) :: U            ! Field to be integrated.
-class(integrand),                  intent(INOUT) :: previous(1:) ! Previous time steps solutions.
-real(R_P),                         intent(IN)    :: Dt           ! Time steps.
-real(R_P),                         intent(IN)    :: t(:)         ! Times.
-logical, optional,                 intent(IN)    :: autoupdate   ! Autoupdate previous time steps.
-logical                                          :: autoupdate_  ! autoupdate previous time steps, dummy var.
-integer(I_P)                                     :: s            ! Steps counter.
-autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
-do s=1, self%steps
-  U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
-enddo
-if (autoupdate_) call self%update_previous(U=U, previous=previous)
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes five arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the time step used, the fifth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time and the sixth is a logical flag for enabling/disabling the cyclic update of previous time steps solutions.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-
-It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
-This can be disable passing \emph{autoupdate=.false.}: it is useful in the framework of predictor-corrector solvers.
-
-\subsection{The implicit Adams-Moulton class of solvers}\label{subsec:solver_am}
+\subsection{Implicit Adams-Moulton class API}\label{subsec:solver_am}
 
@@
 
-\subsection{The predictor-corrector Adams-Bashforth-Moulton class of solvers}\label{subsec:solver_abm}
+\subsection{Predictor-corrector Adams-Bashforth-Moulton class API}\label{subsec:solver_abm}
 
 {\color{red} To be written.}
-
-\subsection{The leapfrog solver}\label{subsec:solver_leapfrog}
-
-The explicit Leapfrog class of solvers is exposed as a single derived type named \emph{leapfrog\_integrator}. This type provides three methods:
-
-\begin{itemize}
-  \item \emph{init}: initialize the integrator accordingly the possibilities offered by the class of solvers;
-  \item \emph{integrate}: integrate integrand field one-step-over in time.
-  \end{itemize}
-
-As common for FOODIE solvers, for using such a solver it must be previously defined as an instance of the exposed FOODIE integrator type, see listing \ref{list:define_lf_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of an explicit Leapfrog integrator},label={list:define_lf_solver}]
-use FOODIE, only: leapfrog_integrator
-type(leapfrog_integrator) :: integrator
-\end{lstlisting}
-
-Once an integrator of this type has been instantiated, it must be initialized before used, for example see listing \ref{list:init_lf_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of initialization of an explicit Leapfrog integrator},label={list:init_lf_solver}]
-! default coefficients nu=0.01, alpha=0.53
-call integrator%init()
-! custom coefficients
-call integrator%init(nu=0.015, alpha=0.6)
-\end{lstlisting}
-
-The complete definition of the \emph{leapfrog\_integrator} type is reported into listing \ref{list:lf_solver}.
-As shown, the filter coefficients are initialized to zero, suitable values are initialized by the \emph{init} method.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={definition of \emph{leapfrog\_integrator} type},label={list:lf_solver}]
-type :: leapfrog_integrator
-  private
-  real(R_P) :: nu=0.01_R_P    ! Robert-Asselin filter coefficient.
-  real(R_P) :: alpha=0.53_R_P ! Robert-Asselin-Williams filter coefficient.
-  contains
-    procedure, pass(self), public :: init
-    procedure, pass(self), public :: integrate
-endtype leapfrog_integrator
-\end{lstlisting}
-
-After the solver has been initialized it can be used for integrating an integrand field, as shown in listing \ref{list:use_lf_solver}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of a Leapfrog integrator},label={list:use_lf_solver}]
-real               :: times(1:2)
-type(my_integrand) :: filter_displacement
-type(my_integrand) :: my_field
-type(my_integrand) :: previous(1:2)
-call integrator%integrate(U=my_field, previous=previous, filter=filter_displacement, Dt=Dt, &
-                          t=times)
-\end{lstlisting}
-where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{previous} are the memory registers where previous time steps solutions are saved, \emph{filter\_displacement} is the register necessary for computing the eventual displacement of the applied filter and \emph{times} are the time at each 2 steps considered for the current one-step-over integration.
-
-The complete implementation of the \emph{integrate} method of the explicit Leapfrog class of solvers is reported in listing \ref{list:lf_integrate}.
-
-\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Leapfrog class},label={list:lf_integrate}]
-subroutine integrate(self, U, previous, Dt, t, filter)
-class(leapfrog_integrator), intent(IN)    :: self          ! LF integrator.
-class(integrand),           intent(INOUT) :: U             ! Field to be integrated.
-class(integrand),           intent(INOUT) :: previous(1:2) ! Previous time steps solutions.
-real(R_P),                  intent(in)    :: Dt            ! Time step.
-real(R_P),                  intent(IN)    :: t             ! Time.
-class(integrand), optional, intent(INOUT) :: filter        ! Filter field displacement.
-U = previous(1) + previous(2)%t(t=t) * (Dt * 2._R_P)
-if (present(filter)) then
-  filter = (previous(1) - previous(2) * 2._R_P + U) * self%nu * 0.5_R_P
-  previous(2) = previous(2) + filter * self%alpha
-  U = U + filter * (self%alpha - 1._R_P)
-endif
-previous(1) = previous(2)
-previous(2) = U
-return
-endsubroutine integrate
-\end{lstlisting}
-
-This method takes six arguments, the first argument is passed as bounded argument and it is the solver itself, the second is of an integrand class, it being the integrand field that must integrated one-step-over in time, the third are the previous time steps solutions, the fourth is the optional filter-displacement-register, the fifth is the time step used and the sixth is an array of the time values of the steps considered for the current one-step-over integration that are passed to the residuals function for taking into account the cases where the time derivative explicitly depends on time.
-The time step is not automatically computed (for example inspecting the passed integrand field), thus its value must be externally computed and passed to the \emph{integrate} method.
-It is worth noting that if the filter displacement argument is not passed, the solver reverts back to the standard unfiltered Leapfrog method.
-
-It is worth noting that the method also performs the cyclic update of the previous time steps solutions memory registers.
-In particular, if the filter displacement argument is passed the method performs the RAW filtering.
 

### 2017-05-05 papers/announcement/src/introduction/introduction.tex
update paper

@@
 Initial Value Problem (IVP, or Cauchy problem) constitutes a class of mathematical models of paramount relevance, it being applied to the modelling of a wide range of problems.
-Briefly, an IVP is an Ordinary Differential Equations system (ODE) coupled with specified initial values of the unknown state variables, the solution of which are searched at a given time after the initial time considered.
+An IVP is an Ordinary Differential Equation (ODE, or system of equations, ODEs) coupled with specified initial values of the unknown state variables, the solution of which are searched at a given time after the initial time considered.
 
@@
 
-FOODIE library must:
+FOODIE development arises from the following specifications:
 
@@
   \item be written in modern Fortran (standard 2008 or newer);
-  \item be written by means of Object Oriented Programming (OOP) paradigm;
+  \item be abstract, thus it is written by means of Object Oriented Programming (OOP) paradigm;
   \item be well documented;
@@
 
-FOODIE, meaning Fortran Object oriented Ordinary Differential Equations integration library, has been developed with the aim to satisfy the above specifications.
-The present paper is its first comprehensive presentation.
-
-The Fortran (Formula Translator) programming language is the \emph{de facto} standard into computer science field: it strongly facilitates the effective and efficient translation of (even complex) mathematical and numerical models into an operative software without compromise on computations speed and accuracy.
-Moreover, its simple syntax is suitable for scientific researchers that are interested (and skilled) in the physical aspects of the numerical computations rather than computer technicians.
+Fortran (Formula Translator) programming language is the \emph{de facto} standard into computer science field: it strongly facilitates the effective and efficient translation of (even complex) mathematical and numerical models into an operative software without compromise on computations speed and accuracy.
+Moreover, its simple syntax is suitable for scientific researchers that are interested (and skilled) in the physical aspects of the numerical computations rather than in the information technology aspects.
 Consequently, we develop FOODIE using Fortran language: FOODIE is written by research scientists for research scientists.
 
-One key-point of the FOODIE development is the \emph{problem generalization}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its actual definitions.
+One key-point of the FOODIE development is the \emph{problem abstraction}: the problem solved must be the IVP \ref{eq:IVP} rather than any of its real, concrete applications, e.g.
+the Lorentz system.
 Consequently, we must rely on a \emph{generic} implementation of the solvers.
 To this aim, OOP is very useful: it allows to express IVP \ref{eq:IVP} in a very concise and clear formulation that is really generic.
-In particular, our implementation is based on \emph{Abstract Calculus Pattern} (ACP) concept.
+In particular, our implementation is based on the \emph{Abstract Calculus Pattern} (ACP) concept.
 
@@
 Moreover, the FOODIE clients can use solvers being extremely robust: as a matter of facts, FOODIE solvers are expressed in a formulation very close to the mathematical one and are tested on an extremely varying family of problems.
-As shown in the following, such a great flexibility does not compromise the computational efficiency.
+As shown in the following, such a great flexibility does not compromise the computational efficiency: ACP introduces an overhead that is, however, very contained.
 

### 2017-05-05 papers/announcement/src/mathematical-numerical-models/mathematical-numerical-models.tex
update paper

@@
 On the contrary, a one-step scheme evolves the solution toward the next step using only the information coming from the current time approximation.
-In the framework of one-step schemes family an equivalent accurate approximation can be obtained by means of a multi-stage approach as the one due to Runge-Kutta.
-The current version of FOODIE provides schemes belonging to both these families.
+In the framework of one-step schemes family an equivalent accurate approximation can be obtained by means of a multi-stage approach as the one due to Runge and Kutta.
+FOODIE provides schemes belonging to both these families.
 
@@
 The implicit approach is of practical use for \emph{stiff} systems where the usage of explicit schemes could require an extremely small time step to evolve in a \emph{stable} way the solution.
-Mixing together explicit and implicit schemes it is possible to build a family of \emph{predictor-corrector} methods: using an explicit scheme to predict a guess for the next step approximation it is possible to use an implicit method for correcting this guess.
-Currently, FOODIE provides explicit solvers and predictor-correct ones.
+Mixing together explicit and implicit schemes it is possible to build families of \emph{predictor-corrector} methods: using an explicit scheme to predict a guess for the next step approximation it is possible to use an implicit method for correcting this guess.
+FOODIE provides explicit, implicit and predictor-correct solvers.
 
@@
 \begin{itemize}
-  \item one-step schemes:
+   \item explicit schemes:
    \begin{itemize}
-      \item explicit forward Euler scheme, it being $1^{st}$ order accurate;
-      \item explicit Runge-Kutta schemes:
+      \item forward Euler: $1^{st}$ order of accurate;
+      \item Adams-Bashforth: multi-step from $1^{st}$ to $16^{th}$ order of accuracy;
+      \item Leapfrog:
       \begin{itemize}
-          \item TVD/SSP Runge-Kutta schemes:
+         \item unfiltered leapfrog, $2^{nd}$ order accurate, mostly unstable;
+         \item Robert-Asselin filtered leapfrog, $1^{st}$ order accurate;
+         \item Robert-Asselin-Williams filtered leapfrog, $3^{rd}$ order accurate;
+      \end{itemize}
+      \item Strong Stability Preserving (SSP) Linear Multistep Methods:
       \begin{itemize}
-              \item 2-stages, it being $2^{nd}$ order accurate;
-              \item 3-stages, it being $3^{rd}$ order accurate;
-              \item 5-stages, it being $4^{th}$ order accurate;
+         \item 3 steps, $2^{nd}$ order accurate;
+         \item 4 steps, $3^{rd}$ order accurate;
+         \item 5 steps, $3^{rd}$ order accurate;
       \end{itemize}
-          \item low storage Runge-Kutta schemes:
+      \item Strong Stability Preserving (SSP) Linear Multistep Methods with variable step-size:
       \begin{itemize}
-              \item 5-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \item 6-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \item 7-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \item 12-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \item 13-stages 2N registers schemes, it being $4^{th}$ order accurate;
-              \item 14-stages 2N registers schemes, it being $4^{th}$ order accurate;
+         \item from 2 to 3 steps with $2^{nd}$ order of accuracy;
+         \item from 3 to 5 steps with $3^{rd}$ order of accuracy;
+      \end{itemize}
+      \item Strong Stability Preserving (SSP) Linear Multistep Runge-Kutta Methods:
+      \begin{itemize}
+         \item 2 steps, 2 stages, $3^{rd}$ order accurate;
+         \item 3 steps, 2 stages, $3^{rd}$ order accurate;
+         \item 4 steps, 5 stages, $8^{th}$ order accurate;
+      \end{itemize}
+      \item Runge-Kutta schemes:
+      \begin{itemize}
+         \item Linear Strong Stability Preserving (SSP) RK of any order:
+         \begin{itemize}
+            \item generic s-stages of order $(s-1)^{th}$;
+            \item generic s-stages of order $s^{th}$;
+         \end{itemize}
+         \item Strong Stability Preserving (SSP) Low Storage RK:
+         \begin{itemize}
+            \item 5, 6, 7, 12-14 stages, $4^{th}$ order accurate, 2 registers;
+         \end{itemize}
+         \item Strong Stability Preserving (SSP) RK:
+         \begin{itemize}
+            \item 2 stages, $2^{nd}$ order accurate;
+            \item 3 stages, $3^{rd}$ order accurate;
+            \item 5 stages, $4^{th}$ order accurate;
+         \end{itemize}
+         \item embedded (adaptive) RK:
+         \begin{itemize}
+            \item Heun-Euler, 2 stages, $2^{nd}$ order accurate;
+            \item Runge-Kutta-Cash-Karp, 6 stages, $5^{th}$ order accurate;
+            \item Prince-Dormand, 7 stages, $4^{th}$ order accurate;
+            \item Calvo, 9 stages, $6^{th}$ order accurate;
+            \item Feagin, 17 stages, $10^{th}$ order accurate;
          \end{itemize}
@@
    \end{itemize}
-  \item multi-step schemes:
+   \item implicit schemes:
    \begin{itemize}
-      \item explicit Adams-Bashforth schemes:
+      \item Adams-Moulton: multi-step from $1^{st}$ to $16^{th}$ order of accuracy;
+      \item Backward Differentiation Formula: multi-step from $2^{nd}$ to $6^{th}$ order of accuracy;
+   \end{itemize}
+   \item predictor-corrector schemes:
    \begin{itemize}
-          \item 2-steps, it being $2^{nd}$ order accurate;
-          \item 3-steps, it being $3^{rd}$ order accurate;
-          \item 4-steps, it being $4^{th}$ order accurate;
-          \end{itemize}
-      \item implicit Adams-Moulton schemes:
-        \begin{itemize}
-          \item 1-step, it being $2^{nd}$ order accurate;
-          \item 2-steps, it being $3^{rd}$ order accurate;
-          \item 3-steps, it being $4^{th}$ order accurate;
-          \end{itemize}
-      \item predictor-corrector Adams-Bashforth-Moulton schemes:
-        \begin{itemize}
-          \item 1-step, it being $2^{nd}$ order accurate;
-          \item 2-steps, it being $3^{rd}$ order accurate;
-          \item 3-steps, it being $4^{th}$ order accurate;
-          \end{itemize}
-      \item explicit Leapfrog schemes\footnote{Leapfrog method is a 2-steps solver being, in general $2^{nd}$ order accurate, but mostly \emph{unstable}.
-It is well suited for periodic/oscillatory problems, thus the error is generally categorized accordingly to the phase and amplitude errors: the unfiltered scheme provides a $2^{nd}$ order error on phase approximation while the amplitude error is null.
-However, the unfiltered scheme is mostly unstable, thus many filters have been devised.
-In general, the filters retain the accuracy on the phase error, while affect the amplitude one.}:
-        \begin{itemize}
-          \item 2-steps unfiltered, it being $2^{nd}$ order accurate, but mostly \emph{unstable};
-          \item 2-steps Robert-Asselin filtered, it being $1^{st}$ order accurate (on \emph{amplitude error});
-          \item 2-steps Robert-Asselin-Williams filtered, it being $3^{rd}$ order accurate (on \emph{amplitude error});
-          \end{itemize}
+      \item Adams-Bashforth-Moulton: multi-step from $1^{st}$ to $16^{th}$ order of accuracy;
    \end{itemize}
@@
 
-\subsection{The explicit forward Euler scheme}
+In the subsequent sections each of the above mentioned scheme is briefly described.
+
+It is worth noting that for multi-step methods like Adams-Bashforth, Adams-Moulton, Linear Multistep Methods, etc...
+classes the solvers are not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided, $N_s$ being the number of previous steps used.
+To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
+
+\subsection{Explicit forward Euler scheme}
 
@@
 \begin{equation}
-  U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left[t^n, U\left(t^n\right)\right]
+  U\left(t^{n+1}\right) = U\left(t^n\right) +\Delta t \cdot R\left(t^n, U\left(t^n\right)\right)
 \label{eq:solver-euler}
@@
 
-This scheme is Total Variation Diminishing (TVD), thus satisfies the maximum principle (or the equivalent positivity preserving property).
+This scheme is Total Variation Diminishing (TVD), thus satisfies the maximum principle or the equivalent positivity preserving property.
+In different fields this property is also indicated as Strong Stability Preserving.
 
@@
 
-\subsection{The explicit TVD/SSP Runge-Kutta class of schemes}
+\subsection{Explicit Adams-Bashforth class}
+
+Adams-Bashforth methods belong to the more general (linear) explicit multi-step family of schemes.
+This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
+Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
+
+In general, the Adams-Bashforth schemes provided by FOODIE library are written by means of the following algorithm (for only explicit schemes):
+
+\begin{equation}
+U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}
+\label{eq:AB}
+\end{equation}
+where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.
+
+Currently FOODIE provides schemes having $2^{nd}$ to $16^{th}$ formal order of accuracy. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.
+
+It is worth noting that FODDIE also provides a one-step AB solver that reverts back to the explicit forward Euler scheme: it can be used, for example, into a Recursive Order Reduction (ROR) framework that automatically checks some properties of the solution and, in case, reduces the order of the Runge-Kutta solver until those properties are obtained.
+
+{\color{red} Add citations.}
+
+\subsection{Leapfrog solver}
+
+\emph{Leapfrog} scheme belongs to the multi-step family, it being formally a centered second order approximation in time.
+The leapfrog method (in its original formulation) is mostly unstable, however it is well suited for periodic-oscillatory problems providing a null error on the amplitude value and a formal second order error on the phase one, under the satisfaction of the time-step size stable limit.
+Commonly, the leapfrog methods are said to provide a $2 \Delta t$ computational mode that can generate unphysical, unstable solutions.
+As consequence, the original leapfrog scheme is generally \emph{filtered} in order to suppress these computational modes.
+
+The unfiltered leapfrog scheme provided by FOODIE is:
+\begin{equation}
+  U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]
+\label{eq:leapfrog}
+\end{equation}
+
+FOODIE provides, in a \emph{seamless} API, also filtered leapfrog schemes.
+A widely used filter is due to Robert and Asselin, that suppress the computational modes at the cost of accuracy reduction resulting into a $1^{st}$ order error in amplitude value.
+A more accurate filter, able to provide a $3^{rd}$ order error on amplitude, is a modification of the Robert-Asselin filter due to Williams known as Robert-Asselin-Williams (RAW) filter, that filters the approximation of $U\left(t^{n+1}\right)$ and $U\left(t^{n+2}\right)$ by the following scalar coefficient:
+
+\begin{equation}
+  \begin{matrix}
+    U\left(t^{n+1}\right) = U\left(t^{n+1}\right) + \Delta * \alpha     \\
+    U\left(t^{n+2}\right) = U\left(t^{n+2}\right) + \Delta * (\alpha-1) \\
+    where \\
+    \Delta = \frac{\nu}{2}(U^{n} - 2 U^{n+1} + U^{n+2})
+  \end{matrix}
+\label{eq:leapfrog-RAW}
+\end{equation}
+
+The filter coefficients should be taken as $\nu \in (0,1]$ and $\alpha \in (0.5,1]$.
+If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude.
+The default values of the FOODIE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.
+
+{\color{red} Add citations.}
+
+\subsection{Explicit SSP Linear Mutistep Methods class}
+
+Explicit SSP Linear Mutistep Methods belong to multi-step schemes like the Adams-Bashforth ones, but they ensure the Strong Stability Preserving property.
+
+The algorithm implemented into FOODIE is:
+
+\begin{equation}
+U^{n+N_s} = \sum_{s=1}^{N_s}{\left[a_s U^{n+s-1} + \Delta t b_s \cdot R(t^{n+s-1}, U^{n+s-1}) \right]}
+\label{eq:lmm-ssp}
+\end{equation}
+where $N_s$ is the number of time steps considered and $a_s, \quad b_s$ are the linear coefficients selected.
+
+Currently FOODIE provides schemes having $2^{nd}$ to $3^{th}$ formal order of accuracy. The $a_s , \quad b_s$ coefficients are reported in table \ref{tab:lmm-ssp-coeff}.
+
+{\color{red} Add citations.}
+
+\subsection{Explicit SSP Linear Mutistep Methods with variable step-size class}
+
+Explicit SSP Linear Mutistep Methods with variable step-size belong to multi-step schemes like the Adams-Bashforth ones, but they ensure the Strong Stability Preserving property and they allow the time step-size to vary, as it happens for multi-stage scheme.
+
+Currently FOODIE provides schemes having $2^{nd}$ to $3^{th}$ formal order of accuracy. The $2^{nd}$ order algorithm implemented into FOODIE is:
+
+\begin{equation}
+U^{n+N_s} = \frac{1}{\Omega_{N_s-1}^2} U^n + \frac{\Omega_{N_s-1}^2 - 1}{\Omega_{N_s-1}^2} U^{n+N_s-1} + \frac{\Omega_{N_s-1} + 1}{\Omega_{N_s-1}} \Delta t^{n+N_s} R(U^{n+N_s-1})
+\label{eq:lmm-ssp-vss-2}
+\end{equation}
+while the $3^{rd}$ order algorithm is:
+\begin{equation}
+U^{n+N_s} = \frac{3 \Omega_{N_s-1} + 2}{\Omega_{N_s-1}^3} U^n +
+\frac{(\Omega_{N_s-1} + 1)^2(\Omega_{N_s-1} - 2)}{\Omega_{N_s-1}^3} U^{n+N_s-1} +
+\frac{\Omega_{N_s-1} + 1}{\Omega_{N_s-1}^2} \Delta t^{n+N_s} R(U^n) +
+\frac{(\Omega_{N_s-1} + 1)^2}{\Omega_{N_s-1}^2} \Delta t^{n+N_s} R(U^{n+N_s-1})
+\label{eq:lmm-ssp-vss-3}
+\end{equation}
+
+The coefficients are computed by recursive formula:
+
+$$ \Omega_s = \sum_{i=1}^s { \omega_i }\quad 1 \leq s \leq N_s $$
+$$ \omega_i = \frac{\Delta t^{n + s}}{\Delta t^{n + N_s}} $$
+where $N_s$ is the number of time steps considered.
+
+{\color{red} Add citations.}
+
+\subsection{Explicit SSP Linear Mutistep Methods Runge-Kutta class}
+
+Because SSP Runge-Kutta has an order barrier, see \cite{ketch}, namely they can be at most of $4^{th}$ order accurate, multi-step/multi-stage schemes have been devised.
+Such a class of schemes represent a good compromise between memory efficiency and accuracy.
+The schemes are not self-starting, being a multi-step class, but the same multi-step nature allows to overcome the order barrier of SSP multi-stage schemes.
+
+The algorithm implemented into FOODIE is:
+\begin{equation}
+\begin{matrix}
+y_1^n = u^n \\
+y_i^n = \sum_{l=1}^{k} d_{il} u^{n-k+l} + \Delta{t}\sum_{l=1}^{k-1} \hat{a}_{il} F(u^{n-k+l}) + \Delta{t}\sum_{j=1}^{i-1} a_{ij} F(y_j^n) \; \; \; \;  2 \leq i \leq N_s \\
+   u^{n+1} = \sum_{l=1}^{k} \theta_l u^{n-k+l} + \Delta{t}\sum_{l=1}^{k-1} \hat{b}_{l} F(u^{n-k+l}) + \Delta{t}\sum_{j=1}^{N_s} b_j F(y_j^n).
+\end{matrix}
+\label{eq:ms-rk-ssp}
+\end{equation}
+where $N_s$ is the number of time steps considered.
+
+{\color{red} Add citations.}
+
+\subsection{Explicit SSP Runge-Kutta class}
 
@@
 
-In general, the TVD/SSP Runge-Kutta schemes provided by FOODIE library are written by means of the following algorithm:
+In general, the SSP Runge-Kutta schemes provided by FOODIE library are written by means of the following algorithm:
 \begin{equation}
@@
 The nature of the scheme and the properties of the solutions obtained depend on the number of stages and on the value of the coefficients selected.
-Currently, FOODIE provides 3 Runge-Kutta schemes having TVD or Strong Stability Preserving (SSP) propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.
+Currently, FOODIE provides many Runge-Kutta schemes having Strong Stability Preserving propriety (thus they being suitable for ODE systems involving rapidly changing non linear dynamics) the Butcher's coefficients of which are reported in tables \ref{tab:RK-TVD-butcher-2}, \ref{tab:RK-SSP-butcher-3}, \ref{tab:RK-SSP-butcher-5}.
 
-\begin{table}[!ht]
-  \centering
-  \caption{Butcher's table of 2 stages, $2^{nd}$ order, Runge-Kutta TVD scheme\label{tab:RK-TVD-butcher-2}}
-  \resizebox{0.15\textwidth}{!}{%
-  \begin{tabular}{c|cc}
-   1 & 1 & 0       \\
-   \hline
-       & 1/2 & 1/2   \\
-  \end{tabular}}
-\end{table}
+The absolute stability locus depends on the coefficients selected, however, as a general principle, we can assume that greater is the stages number and wider is the stability locus on equal accuracy orders.
+Currently, FOODIE provides $2^{nd}$ to $4^{th}$ order accurate schemes.
+As demonstrated in \cite{ketch}, SSP RK are at most $4^{th}$ order accurate, namely the SSP property introduces an order barrier for single step methods.
+To increase the order it is necessary to adopt multi-step schemes: the hybrid multi-step RK schemes are devised to this aim.
 
-\begin{table}[!ht]
-  \centering
-  \caption{Butcher's table of 3 stages, $3^{rd}$ order, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-3}}
-  \resizebox{0.20\textwidth}{!}{%
-  \begin{tabular}{c|ccc}
-   1   & 1   &     &      \\
-   1/2 & 1/4 & 1/4 &      \\
-   \hline
-       & 1/6 & 1/6 & 2/3  \\
-  \end{tabular}}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \caption{Butcher's table of 5 stages, $4^{th}$ order accurate, Runge-Kutta SSP scheme\label{tab:RK-SSP-butcher-5}}
-  \resizebox{0.99\textwidth}{!}{%
-  \begin{tabular}{c|ccccc}
-   0.39175222700392 & 0.39175222700392 &                  &                  &                  &                  \\
-   0.58607968896779 & 0.21766909633821 & 0.36841059262959 &                  &                  &                  \\
-   0.47454236302687 & 0.08269208670950 & 0.13995850206999 & 0.25189177424738 &                  &                  \\
-   0.93501063100924 & 0.06796628370320 & 0.11503469844438 & 0.20703489864929 & 0.54497475021237 &                  \\
-   \hline
-                    & 0.14681187618661 & 0.24848290924556 & 0.10425883036650 & 0.27443890091960 & 0.22600748319395 \\
-  \end{tabular}}
-\end{table}
-The absolute stability locus depends on the coefficients selected, however, as a general principle, we can assume that greater is the stages number and wider is the stability locus on equal accuracy orders.
-
-It is worth noting that FODDiE also provides a one-stage TVD Runge-Kutta solver that reverts back to the explicit forward Euler scheme: it can be used, for example, into a Recursive Order Reduction (ROR) framework that automatically checks some properties of the solution and, in case, reduces the order of the Runge-Kutta solver until those properties are obtained.
+Similarly to the Adams-Bashforth class, the Runge-Kutta class also provides a fail-safe one-stage solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
 
@@
 
-\subsection{The explicit low storage Runge-Kutta class of schemes}
+\subsection{Explicit low storage Runge-Kutta class}
 
@@
 
-Currently FOODIE provides 5/6/7/12/13/14 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in table \ref{tab:RK-ls}.
+Currently FOODIE provides 5, 6, 7, 12, 13 and 14 stages, all $4^{th}$ order, 2N registers explicit schemes, the coefficients of which are listed in table \ref{tab:RK-ls}.
 
-\begin{table}[!ht]
-  \centering
-  \caption{Williamson's table of low storage Runge-Kutta schemes\label{tab:RK-ls}}
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{5 stages, $4^{th}$ order\label{tab:RK-ls-5}}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{cccc}
-      Stage & $A$                                    & $B$                                    & $C$                                   \\
-      \hline
-      1     & $ 0                                  $ & $\frac{1432997174477}{9575080441755 }$ & $ 0                                 $ \\
-      2     & $-\frac{567301805773 }{1357537059087}$ & $\frac{5161836677717}{13612068292357}$ & $\frac{1432997174477}{9575080441755}$ \\
-      3     & $-\frac{2404267990393}{2016746695238}$ & $\frac{1720146321549}{2090206949498 }$ & $\frac{2526269341429}{6820363962896}$ \\
-      4     & $-\frac{3550918686646}{2091501179385}$ & $\frac{3134564353537}{4481467310338 }$ & $\frac{2006345519317}{3224310063776}$ \\
-      5     & $-\frac{1275806237668}{842570457699 }$ & $\frac{2277821191437}{14882151754819}$ & $\frac{2802321613138}{2924317926251}$ \\
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{6 stages, $4^{th}$ order\label{tab:RK-ls-6}}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{cccc}
-      Stage & $A$               & $B$              & $C$              \\
-      \hline
-      1     & $ 0             $ & $0.122000000000$ & $0             $ \\
-      2     & $-0.691750960670$ & $0.477263056358$ & $0.122000000000$ \\
-      3     & $-1.727127405211$ & $0.381941220320$ & $0.269115878630$ \\
-      4     & $-0.694890150986$ & $0.447757195744$ & $0.447717183551$ \\
-      5     & $-1.039942756197$ & $0.498614246822$ & $0.749979795490$ \\
-      6     & $-1.531977447611$ & $0.186648570846$ & $0.898555413085$ \\
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{7 stages, $4^{th}$ order\label{tab:RK-ls-7}}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{cccc}
-      Stage & $A$               & $B$              & $C$              \\
-      \hline
-      1     & $ 0.000000000000$ & $0.117322146869$ & $0.000000000000$ \\
-      2     & $-0.647900745934$ & $0.503270262127$ & $0.117322146869$ \\
-      3     & $-2.704760863204$ & $0.233663281658$ & $0.294523230758$ \\
-      4     & $-0.460080550118$ & $0.283419634625$ & $0.305658622131$ \\
-      5     & $-0.500581787785$ & $0.540367414023$ & $0.582864148403$ \\
-      6     & $-1.906532255913$ & $0.371499414620$ & $0.858664273599$ \\
-      7     & $-1.450000000000$ & $0.136670099385$ & $0.868664273599$ \\
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{12 stages, $4^{th}$ order\label{tab:RK-ls-12}}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{cccc}
-      Stage & $A$                   & $B$                  & $C$                  \\
-      \hline
-      1     & $ 0.0000000000000000$ & $0.0650008435125904$ & $0.0000000000000000$ \\
-      2     & $-0.0923311242368072$ & $0.0161459902249842$ & $0.0650008435125904$ \\
-      3     & $-0.9441056581158819$ & $0.5758627178358159$ & $0.0796560563081853$ \\
-      4     & $-4.3271273247576394$ & $0.1649758848361671$ & $0.1620416710085376$ \\
-      5     & $-2.1557771329026072$ & $0.3934619494248182$ & $0.2248877362907778$ \\
-      6     & $-0.9770727190189062$ & $0.0443509641602719$ & $0.2952293985641261$ \\
-      7     & $-0.7581835342571139$ & $0.2074504268408778$ & $0.3318332506149405$ \\
-      8     & $-1.7977525470825499$ & $0.6914247433015102$ & $0.4094724050198658$ \\
-      9     & $-2.6915667972700770$ & $0.3766646883450449$ & $0.6356954475753369$ \\
-      10    & $-4.6466798960268143$ & $0.0757190350155483$ & $0.6806551557645497$ \\
-      11    & $-0.1539613783825189$ & $0.2027862031054088$ & $0.7143773712418350$ \\
-      12    & $-0.5943293901830616$ & $0.2167029365631842$ & $0.9032588871651854$ \\
-    \end{tabular}}
-  \end{subtable}\\
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{13 stages, $4^{th}$ order\label{tab:RK-ls-13}}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{cccc}
-      Stage & $A$                   & $B$                  & $C$                  \\
-      \hline
-      1     & $ 0.0000000000000000$ & $0.0271990297818803$ & $0.0000000000000000$ \\
-      2     & $-0.6160178650170565$ & $0.1772488819905108$ & $0.0271990297818803$ \\
-      3     & $-0.4449487060774118$ & $0.0378528418949694$ & $0.0952594339119365$ \\
-      4     & $-1.0952033345276178$ & $0.6086431830142991$ & $0.1266450286591127$ \\
-      5     & $-1.2256030785959187$ & $0.2154313974316100$ & $0.1825883045699772$ \\
-      6     & $-0.2740182222332805$ & $0.2066152563885843$ & $0.3737511439063931$ \\
-      7     & $-0.0411952089052647$ & $0.0415864076069797$ & $0.5301279418422206$ \\
-      8     & $-0.1797084899153560$ & $0.0219891884310925$ & $0.5704177433952291$ \\
-      9     & $-1.1771530652064288$ & $0.9893081222650993$ & $0.5885784947099155$ \\
-      10    & $-0.4078831463120878$ & $0.0063199019859826$ & $0.6160769826246714$ \\
-      11    & $-0.8295636426191777$ & $0.3749640721105318$ & $0.6223252334314046$ \\
-      12    & $-4.7895970584252288$ & $1.6080235151003195$ & $0.6897593128753419$ \\
-      13    & $-0.6606671432964504$ & $0.0961209123818189$ & $0.9126827615920843$ \\
-    \end{tabular}}
-  \end{subtable}\quad%
-  \begin{subtable}[b]{0.45\textwidth}
-    \centering
-    \caption{14 stages, $4^{th}$ order\label{tab:RK-ls-14}}
-    \resizebox{1.00\textwidth}{!}{%
-    \begin{tabular}{cccc}
-      Stage & $A$                   & $B$                  & $C$                  \\
-      \hline
-      1     & $ 0.0000000000000000$ & $0.0367762454319673$ & $0.0000000000000000$ \\
-      2     & $-0.7188012108672410$ & $0.3136296607553959$ & $0.0367762454319673$ \\
-      3     & $-0.7785331173421570$ & $0.1531848691869027$ & $0.1249685262725025$ \\
-      4     & $-0.0053282796654044$ & $0.0030097086818182$ & $0.2446177702277698$ \\
-      5     & $-0.8552979934029281$ & $0.3326293790646110$ & $0.2476149531070420$ \\
-      6     & $-3.9564138245774565$ & $0.2440251405350864$ & $0.2969311120382472$ \\
-      7     & $-1.5780575380587385$ & $0.3718879239592277$ & $0.3978149645802642$ \\
-      8     & $-2.0837094552574054$ & $0.6204126221582444$ & $0.5270854589440328$ \\
-      9     & $-0.7483334182761610$ & $0.1524043173028741$ & $0.6981269994175695$ \\
-      10    & $-0.7032861106563359$ & $0.0760894927419266$ & $0.8190890835352128$ \\
-      11    & $ 0.0013917096117681$ & $0.0077604214040978$ & $0.8527059887098624$ \\
-      12    & $-0.0932075369637460$ & $0.0024647284755382$ & $0.8604711817462826$ \\
-      13    & $-0.9514200470875948$ & $0.0780348340049386$ & $0.8627060376969976$ \\
-      14    & $-7.1151571693922548$ & $5.5059777270269628$ & $0.8734213127600976$ \\
-    \end{tabular}}
-  \end{subtable}\\
-\end{table}
-
-Similarly to the TVD/SSP Runge-Kutta class, the low storage class also provides a fail-safe one-stage solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
+Similarly to the SSP Runge-Kutta class, the low storage class also provides a fail-safe one-stage solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
 
@@
 
-\subsection{The explicit Adams-Bashforth class of schemes}
-
-Adams-Bashforth methods belong to the more general (linear) explicit multi-step family of schemes.
-This kind of schemes has been designed to achieve a more accurate solution than the $1^{st}$ Euler scheme using the information coming from the solutions already computed at previous time steps.
-Typically only one new residuals function $R$ evaluation is required at each time step, whereas Runge-Kutta schemes require many of them.
-
-In general, the Adams-Bashforth schemes provided by FOODIE library are written by means of the following algorithm (for only explicit schemes):
-
-\begin{equation}
-U\left(t^{n+N_s}\right) = U\left(t^{n+N_s-1}\right) +\Delta t \sum_{s=1}^{n+N_s}{ b_s \cdot R\left[t^{n+s-1}, U\left(t^{n+s-1}\right)\right]}
-\label{eq:AB}
-\end{equation}
-where $N_s$ is the number of time steps considered and $b_s$ are the linear coefficients selected.
-
-Currently FOODIE provides 2, 3, and 4 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AB-coeff}.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Explicit Adams-Bashforth coefficients\label{tab:AB-coeff}}
-  \resizebox{0.30\textwidth}{!}{%
-  \begin{tabular}{c|cccc}
-    $N_s$ & $b_1$           & $b_2$            & $b_3$            & $b_4$           \\
-    \hline
-    2     & $-\frac{1}{2}$  & $\frac{3}{2}$    &       /          &       /         \\
-    3     & $\frac{5}{12}$  & $-\frac{16}{12}$ & $\frac{23}{12}$  &       /         \\
-    4     & $-\frac{9}{24}$ & $\frac{37}{24}$  & $-\frac{59}{24}$ & $\frac{55}{24}$ \\
-  \end{tabular}}
-\end{table}
-
-Similarly to the Runge-Kutta classes, the Adams-Bashforth class also provides a fail-safe one-step solver reverting back to the explicit forward Euler solver, that is useful for ROR-like frameworks.
-
-It is worth noting that for $N_s>1$ the Adams-Bashforth class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided.
-To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
-
-{\color{red} Add citations.}
-
-\subsection{The implicit Adams-Moulton class of schemes}
+\subsection{Implicit Adams-Moulton class}
 
@@
 
-Currently FOODIE provides 1, 2, and 3 steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.
-
-\begin{table}[!ht]
-  \centering
-  \caption{Implicit Adams-Moulton coefficients\label{tab:AM-coeff}}
-  \resizebox{0.30\textwidth}{!}{%
-  \begin{tabular}{c|cccc}
-    $N_s$ & $b_0$           & $b_1$           & $b_2$           & $b_3$          \\
-    \hline
-    1     & $\frac{1}{2}$   & $\frac{1}{2}$   &       /         &       /        \\
-    2     & $-\frac{1}{12}$ & $\frac{8}{12}$  & $\frac{5}{12}$  &       /        \\
-    3     & $\frac{1}{24}$  & $-\frac{5}{24}$ & $\frac{19}{24}$ & $\frac{9}{24}$ \\
-  \end{tabular}}
-\end{table}
+Currently FOODIE provides schemes having $2^{nd}$ to $16^{th}$ formal order of accuracy. The $b_s$ coefficients are reported in table \ref{tab:AM-coeff}.
 
@@
 
-It is worth noting that for $N_s>1$ the Adams-Moulton class of solvers is not \emph{self-starting}: the values of $U\left(t^{1}\right)$, $U\left(t^{2}\right)$, \dots, $U\left(t^{N_s-1}\right)$ must be provided.
-To this aim, a lower order multi-step scheme or an equivalent order one-step multi-stage scheme can be used.
-
 {\color{red} Add citations.}
 
-\subsection{The predictor-corrector Adams-Bashforth-Moulton class of schemes}
+\subsection{Predictor-corrector Adams-Bashforth-Moulton class}
 
@@
 
-Currently FOODIE provides $N_s^c=1, 2, 3 \rightarrow N_s^c=2, 3, 4$ steps schemes having $2^{nd}$, $3^{rd}$ and $4^{th}$ formal order of accuracy, respectively.
-The $b_s^{p,c}$ coefficients are those reported in tables \ref{tab:AB-coeff} and \ref{tab:AM-coeff}.
+Currently FOODIE provides schemes having $2^{nd}$ to $16^{th}$ formal order of accuracy. The $b_s^{p,c}$ coefficients are those reported in tables \ref{tab:AB-coeff} and \ref{tab:AM-coeff}.
 
 {\color{red} Add citations.}
-
-\subsection{The leapfrog solver}
-
-The \emph{leapfrog} scheme belongs to the multi-step family, it being formally a centered second order approximation in time.
-The leapfrog method (in its original formulation) is mostly unstable, however it is well suited for periodic-oscillatory problems providing a null error on the amplitude value and a formal second order error on the phase one, under the satisfaction of the time-step size stable limit.
-Commonly, the leapfrog methods are said to provide a $2 \Delta t$ computational mode that can generate unphysical, unstable solutions.
-As consequence, the original leapfrog scheme is generally \emph{filtered} in order to suppress these computational modes.
-
-The unfiltered leapfrog scheme provided by FOODIE is:
-\begin{equation}
-  U\left(t^{n+2}\right) = U\left(t^{n}\right) + 2\Delta t \cdot R\left[t^{n+1}, U\left(t^{n+1}\right)\right]
-\label{eq:leapfrog}
-\end{equation}
-
-FOODIE provides, in a \emph{seamless} API, also filtered leapfrog schemes.
-A widely used filter is due to Robert and Asselin, that suppress the computational modes at the cost of accuracy reduction resulting into a $1^{st}$ order error in amplitude value.
-A more accurate filter, able to provide a $3^{rd}$ order error on amplitude, is a modification of the Robert-Asselin filter due to Williams known as Robert-Asselin-Williams (RAW) filter, that filters the approximation of $U\left(t^{n+1}\right)$ and $U\left(t^{n+2}\right)$ by the following scalar coefficient:
-
-\begin{equation}
-  \begin{matrix}
-    U\left(t^{n+1}\right) = U\left(t^{n+1}\right) + \Delta * \alpha     \\
-    U\left(t^{n+2}\right) = U\left(t^{n+2}\right) + \Delta * (\alpha-1) \\
-    where \\
-    \Delta = \frac{\nu}{2}(U^{n} - 2 U^{n+1} + U^{n+2})
-  \end{matrix}
-\label{eq:leapfrog-RAW}
-\end{equation}
-
-The filter coefficients should be taken as $\nu \in (0,1]$ and $\alpha \in (0.5,1]$.
-If $\alpha=0.5$ the filters of time $t^{n+1}$ and $t^{n+2}$ have the same amplitude and opposite sign thus allowing to the optimal $3^{rd}$ order error on amplitude.
-The default values of the FOODIE provided scheme are $\nu=0.01$ $\alpha=0.53$, but they can be customized at runtime.
-
-{\color{red} Add citations.}

### 2017-05-05 src/foodie_integrator_adams_bashforth.f90
Travis CI autocommit from travis build 121

@@ -351,7 +351,7 @@ contains
 
   autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
   do s=1, self%steps
-    U = U + (previous(s)%t(t=t(s)) * (Dt * self%b(s)))
+    U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
   enddo
   if (autoupdate_) call self%update_previous(U=U, previous=previous)
   endsubroutine integrate

### 2017-05-05 src/foodie_integrator_adams_moulton.f90

@@ -355,20 +355,20 @@ contains
       allocate(delta, mold=U)
       delta = previous(self%steps)
       do s=0, self%steps - 1
-        delta = delta + (previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s)))
+        delta = delta + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
       enddo
       do s=1, iterations
-        U = delta + (U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps)))
+        U = delta + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
       enddo
     else
-      U = previous(self%steps) + (U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps)))
+      U = previous(self%steps) + U%t(t=t(self%steps) + Dt) * (Dt * self%b(self%steps))
       do s=0, self%steps - 1
-        U = U + (previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s)))
+        U = U + previous(s+1)%t(t=t(s+1)) * (Dt * self%b(s))
       enddo
     endif
     if (autoupdate_) call self%update_previous(U=U, previous=previous)
   else
-    U = U + (U%t(t=t(1)) * (Dt * self%b(0)))
+    U = U + U%t(t=t(1)) * (Dt * self%b(0))
   endif
   endsubroutine integrate
 

### 2017-05-05 src/foodie_integrator_backward_differentiation_formula.f90

@@ -227,10 +227,10 @@ contains
   allocate(delta, mold=U)
   delta = previous(self%steps) * (-self%a(self%steps))
   do s=1, self%steps - 1
-    delta = delta + (previous(s) * (-self%a(s)))
+    delta = delta + previous(s) * (-self%a(s))
   enddo
   do s=1, iterations
-    U = delta + (U%t(t=t(self%steps) + Dt) * (Dt * self%b))
+    U = delta + U%t(t=t(self%steps) + Dt) * (Dt * self%b)
   enddo
   if (autoupdate_) call self%update_previous(U=U, previous=previous)
   endsubroutine integrate

### 2017-05-05 src/foodie_integrator_euler_explicit.f90

@@ -108,6 +108,6 @@ contains
   real(R_P),               intent(in)    :: Dt !< Time step.
   real(R_P), optional,     intent(in)    :: t  !< Time.
 
-  U = U + (U%t(t=t) * Dt)
+  U = U + U%t(t=t) * Dt
   endsubroutine integrate
 endmodule foodie_integrator_euler_explicit

### 2017-05-05 src/foodie_integrator_leapfrog.f90

@@ -169,11 +169,11 @@ contains
   real(R_P),                         intent(in)    :: t             !< Time.
   class(integrand_object), optional, intent(inout) :: filter        !< Filter field displacement.
 
-  U = previous(1) + (previous(2)%t(t=t) * (Dt * 2._R_P))
+  U = previous(1) + previous(2)%t(t=t) * (Dt * 2._R_P)
   if (present(filter)) then
-    filter = (previous(1) - (previous(2) * 2._R_P) + U) * self%nu * 0.5_R_P
-    previous(2) = previous(2) + (filter * self%alpha)
-    U = U + (filter * (self%alpha - 1._R_P))
+    filter = (previous(1) - previous(2) * 2._R_P + U) * self%nu * 0.5_R_P
+    previous(2) = previous(2) + filter * self%alpha
+    U = U + filter * (self%alpha - 1._R_P)
   endif
   previous(1) = previous(2)
   previous(2) = U

### 2017-05-05 src/foodie_integrator_lmm_ssp.f90

@@ -214,8 +214,8 @@ contains
   autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
   U = U * 0._R_P
   do s=1, self%steps
-    if (self%a(s) /= 0._R_P) U = U + (previous(s) * self%a(s))
-    if (self%b(s) /= 0._R_P) U = U + (previous(s)%t(t=t(s)) * (Dt * self%b(s)))
+    if (self%a(s) /= 0._R_P) U = U + previous(s) * self%a(s)
+    if (self%b(s) /= 0._R_P) U = U + previous(s)%t(t=t(s)) * (Dt * self%b(s))
   enddo
   if (autoupdate_) call self%update_previous(U=U, previous=previous)
   endsubroutine integrate

### 2017-05-05 src/foodie_integrator_lmm_ssp_vss.f90

@@ -245,10 +245,10 @@ contains
   real(R_P)                                    :: omega_sq      !< Square of omega coefficient.
 
   autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
-  omega_ = omega(Dt=Dt, s=self%steps-1)
+  omega_= omega(Dt=Dt, s=self%steps-1)
   omega_sq = omega_ * omega_
-  U = (previous(1) * (1._R_P / omega_sq)) + (previous(self%steps) * ((omega_sq - 1._R_P) / omega_sq)) + &
-      (previous(self%steps)%t(t=t(self%steps)) * (Dt(self%steps) * (omega_ + 1._R_P) / omega_))
+  U = previous(1) * (1._R_P / omega_sq) + previous(self%steps) * ((omega_sq - 1._R_P) / omega_sq) + &
+      previous(self%steps)%t(t=t(self%steps)) * (Dt(self%steps) * (omega_ + 1._R_P) / omega_)
   if (autoupdate_) call self%update_previous(U=U, previous=previous, Dt=Dt)
   endsubroutine integrate_order_2
 
@@ -265,10 +265,10 @@ contains
 
   autoupdate_ = .true. ; if (present(autoupdate)) autoupdate_ = autoupdate
   omega_= omega(Dt=Dt, s=self%steps-1)
-  U = (previous(1) * ((3._R_P * omega_ + 2._R_P) / omega_ ** 3 )) +                           &
-      (previous(self%steps) * (((omega_ + 1._R_P) ** 2) * (omega_ - 2._R_P) / omega_ ** 3)) + &
-      (previous(1)%t(t=t(1)) * (Dt(self%steps) * (omega_ + 1._R_P) / omega_ ** 2)) +          &
-      (previous(self%steps)%t(t=t(self%steps)) * (Dt(self%steps) * (omega_ + 1._R_P) ** 2 / omega_ ** 2))
+  U = previous(1) * ((3._R_P * omega_ + 2._R_P) / omega_ ** 3 ) +                           &
+      previous(self%steps) * (((omega_ + 1._R_P) ** 2) * (omega_ - 2._R_P) / omega_ ** 3) + &
+      previous(1)%t(t=t(1)) * (Dt(self%steps) * (omega_ + 1._R_P) / omega_ ** 2) +          &
[…]

### 2017-05-05 src/foodie_integrator_ms_runge_kutta_ssp.f90

@@ -327,25 +327,25 @@ contains
   do k=2, self%stages
     stage(k) = U * 0._R_P
     do s=1, self%steps
-      if (self%D(k, s) /= 0._R_P) stage(k) = stage(k) + (previous(s) * self%D(k, s))
+      if (self%D(k, s) /= 0._R_P) stage(k) = stage(k) + previous(s) * self%D(k, s)
     enddo
     do s=1, self%steps - 1
-      if (self%Ahat(k, s) /= 0._R_P) stage(k) = stage(k) + (previous(s)%t(t=t(s)) * (Dt * self%Ahat(k, s)))
+      if (self%Ahat(k, s) /= 0._R_P) stage(k) = stage(k) + previous(s)%t(t=t(s)) * (Dt * self%Ahat(k, s))
     enddo
     do kk=1, k - 1
-      if (self%A(k, kk) /= 0._R_P) stage(k) = stage(k) + (stage(kk)%t(t=t(self%steps)) * (Dt * self%A(k, kk)))
+      if (self%A(k, kk) /= 0._R_P) stage(k) = stage(k) + stage(kk)%t(t=t(self%steps)) * (Dt * self%A(k, kk))
     enddo
   enddo
   ! computing new time step
   U = U * 0._R_P
   do s=1, self%steps
-    if (self%T(s) /= 0._R_P) U = U + (previous(s) * self%T(s))
+    if (self%T(s) /= 0._R_P) U = U + previous(s) * self%T(s)
   enddo
   do s=1, self%steps - 1
-    if (self%Bhat(s) /= 0._R_P) U = U + (previous(s)%t(t=t(s)) * (Dt * self%Bhat(s)))
+    if (self%Bhat(s) /= 0._R_P) U = U + previous(s)%t(t=t(s)) * (Dt * self%Bhat(s))
   enddo
   do k=1, self%stages
-    if (self%B(k) /= 0._R_P) U = U + (stage(k)%t(t=t(self%steps)) * (Dt * self%B(k)))
+    if (self%B(k) /= 0._R_P) U = U + stage(k)%t(t=t(self%steps)) * (Dt * self%B(k))
   enddo
   if (autoupdate_) call self%update_previous(U=U, previous=previous)
   endsubroutine integrate

### 2017-05-05 src/foodie_integrator_runge_kutta_embedded.f90

@@ -687,7 +687,7 @@ contains
     do s=1, self%stages
       stage(s) = U
       do ss=1, s - 1
-        stage(s) = stage(s) + (stage(ss) * (Dt * self%alph(s, ss)))
+        stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
       enddo
       stage(s) = stage(s)%t(t=t + self%gamm(s) * Dt)
     enddo
@@ -695,8 +695,8 @@ contains
     U1 = U
     U2 = U
     do s=1, self%stages
-      U1 = U1 + (stage(s) * (Dt * self%beta(s, 1)))
-      U2 = U2 + (stage(s) * (Dt * self%beta(s, 2)))
+      U1 = U1 +  stage(s) * (Dt * self%beta(s, 1))
+      U2 = U2 +  stage(s) * (Dt * self%beta(s, 2))
     enddo
     error = U2.lterror.U1
     call self%new_Dt(error=error, Dt=Dt)

### 2017-05-05 src/foodie_integrator_runge_kutta_low_storage.f90

@@ -414,8 +414,8 @@ contains
   stage(1) = U
   stage(2) = U * 0._R_P
   do s=1, self%stages
-    stage(2) = (stage(2) * self%A(s)) + (stage(1)%t(t=t + self%C(s) * Dt) * Dt)
-    stage(1) = stage(1) + (stage(2) * self%B(s))
+    stage(2) = stage(2) * self%A(s) + stage(1)%t(t=t + self%C(s) * Dt) * Dt
+    stage(1) = stage(1) + stage(2) * self%B(s)
   enddo
   U = stage(1)
   endsubroutine integrate

### 2017-05-05 src/foodie_integrator_runge_kutta_lssp.f90

@@ -298,13 +298,13 @@ contains
   ! computing stages
   stage(1) = U
   do s=2, self%stages
-    stage(s) = stage(s-1) + (stage(s-1)%t(t=t) * (Dt * 0.5_R_P))
+    stage(s) = stage(s-1) + stage(s-1)%t(t=t) * (Dt * 0.5_R_P)
   enddo
-  stage(self%stages) = stage(self%stages) + (stage(self%stages)%t(t=t) * (Dt * 0.5_R_P))
+  stage(self%stages) = stage(self%stages) + stage(self%stages)%t(t=t) * (Dt * 0.5_R_P)
   ! computing new time step
   U = U * 0._R_P
   do s=1, self%stages
-    U = U + (stage(s) * self%alpha(s))
+    U = U + stage(s) * self%alpha(s)
   enddo
   endsubroutine integrate_order_s_1
 
@@ -322,13 +322,13 @@ contains
   ! computing stages
   stage(1) = U
   do s=2, self%stages
-    stage(s) = stage(s-1) + (stage(s-1)%t(t=t) * Dt)
+    stage(s) = stage(s-1) + stage(s-1)%t(t=t) * Dt
   enddo
-  stage(self%stages) = stage(self%stages) + (stage(self%stages)%t(t=t) * Dt)
+  stage(self%stages) = stage(self%stages) + stage(self%stages)%t(t=t) * Dt
   ! computing new time step
   U = U * 0._R_P
   do s=1, self%stages
-    U = U + (stage(s) * self%alpha(s))
+    U = U + stage(s) * self%alpha(s)
   enddo
   endsubroutine integrate_order_s
 endmodule foodie_integrator_runge_kutta_lssp

### 2017-05-05 src/foodie_integrator_runge_kutta_ssp.f90

@@ -307,13 +307,13 @@ contains
   do s=1, self%stages
     stage(s) = U
     do ss=1, s - 1
-      stage(s) = stage(s) + (stage(ss) * (Dt * self%alph(s, ss)))
+      stage(s) = stage(s) + stage(ss) * (Dt * self%alph(s, ss))
     enddo
     stage(s) = stage(s)%t(t=t + self%gamm(s) * Dt)
   enddo
   ! computing new time step
   do s=1, self%stages
-    U = U + (stage(s) * (Dt * self%beta(s)))
+    U = U + stage(s) * (Dt * self%beta(s))
   enddo
   endsubroutine integrate
 endmodule foodie_integrator_runge_kutta_ssp
